### 2019-10-14 draft/paper.tex
stuff

\documentclass{llncs}

\usepackage{proof}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage[utf8]{inputenc}
\usepackage{todonotes}
\input{abbrevs.tex}

\begin{document}

\title{Generalizations of Quotient Inductive-Inductive Types\thanks{This work
    was supported by EFOP-3.6.3-VEKOP-16-2017-00002 grant and COST
    Action EUTypes CA15123.}}

\author{Ambrus Kaposi \and Andr{\'a}s Kov{\'a}cs}
\institute{E{\"o}tv{\"o}s Lor{\'a}nd University, Budapest, Hungary \\ \{akaposi$|$kovacsandras\}@inf.elte.hu}

\maketitle

\begin{abstract}
Quotient inductive-inductive types (QIITs) are generalized inductive types which
allow sorts to be indexed over previously declared sorts, and allow usage of
equality constructors. QIITs are especially useful for algebraic descriptions of
type theories and constructive definitions of real, ordinal and surreal
numbers. We develop new metatheory for large QIITs, large elimination, recursive
equations, infinitary constructors and equations between sorts.
%% First, we develop tools which allow precise and convenient handling
%% of universe levels, working internally in an extensional type theory with
%% cumulative universes.
As in prior work, we describe QIITs using a type theory where each context
represents a QIIT signature. However, in our case the theory of signatures can
also describe its own signature. We use self-description to bootstrap a model
theory for the theory of signatures without using preterms or assuming a
pre-existing internal syntax for a type theory. We give initial algebra
semantics for described QIITs, and we show the equivalence of initiality and
induction. We present two extensions of a previous term model construction. The
first one constructs all large infinitary QIITs without sort equations from the
QIIT of the theory of signatures. The second one constructs all large finitary
QIITs with sort equations, from the same syntax. This separation is required
because handling infinitary constructors in the term model requires showing a
strong form of invariance under algebra isomorphisms, which is violated by sort
equations.

\end{abstract}

\section{Introduction}

\todo[inline]{Citations}

The aim of this work is to provide theoretical underpinning to a general notion
of inductive types, called quotient inductive-inductive types (QIITs). QIITs are
of interest because there are many commonly used mathematical structures, which
can be conveniently described as QIITs in type theory, but cannot be defined as
less general inductive types, or doing so incurs large encoding overhead.

Categories are a good example. Signatures for QIITs allow having multiple sorts,
with later ones indexed over previous ones, and equations as well. We need both
features in order to write down the signature of categories:
\begin{alignat*}{5}
  & Obj && : \Set \\
  & Mor && : Obj \ra Obj \ra \Set \\
  & id  && : Mor\,i\,j \\
  & \blank\circ\blank && : Mor\,j\,k\ra Mor\,i\,j \ra Mor\,i\,k \\
  & idl && : id \circ f = f \\
  & idr && : f \circ id = f \\
  & ass && : f \circ (g \circ h) = (f \circ g) \circ h
\end{alignat*}
The benefit of having a QIIT signature is getting a model theory ``for free'',
from the metatheory of QIITs. This model theory includes a category of algebras
which has an initial object and also some additional structure. For the
signature of categories, we get the empty category as the initial object, but it
is common to consider categories with more structure, which have more
interesting initial models.

Algebraic notions of models of type theories are examples for this. Here,
initial models represent syntax, and initiality corresponds to induction on
syntax. A number of different notions have been used, from contextual categories
and comprehension categories to categories with families, but all of these are
categories with extra structure.

The main motivation of the current paper is to extend QIITs so that it
accommodates all algebraic notions of type theories which have been used in
previous works. As a beneficial side effect of fulfilling this goal, infinitary
QIITs such as real numbers will be covered as well.

We generalize previous notions of QIITs in the following ways:
\begin{enumerate}
  \item
  \textbf{Large constructors, large elimination} and models at different
  universe levels. This feature is routinely used in the metatheory of type
  theory, but it has not been presented explicitly in previous works about
  QIITs. In order to handle universe levels gracefully, we develop a set of
  tools, working inside an extensional type theory with \emph{algebraic
    cumulativity}.
  \item
  \textbf{Infinitary constructors}. This enables defining Cauchy real numbers
  and surreal numbers as QIITs. Also, the theory of QIIT signatures is itself
  large and infinitary, thus it can ``eat itself'', i.e.\ include its own
  signature and provide its own metatheory.  We use this self-representation to
  bootstrap the model theory of signatures, without having to assume any
  pre-existing internal syntax.
  \item
  \textbf{Recursive equations}, i.e. equations appearing as assumptions
  of constructors. These have occurred previously in syntaxes of cubical type
  theories, as boundary conditions.
  \item
  \textbf{Sort equations}, or equations between type constructors. They have
  been used recently to conveniently represent type-theoretic universes, for
  example Russell-style universes or cumulative hierarchies.
\end{enumerate}

We also develop semantics. We show that for each signature, there is a cwf
(category with families) of algebras, extended with $\Sigma$-types, an
extensional identity type and constant families. This additional structure
corresponds to a type-theoretic flavor of finite limits, and it was shown
in \cite{TODO} that the category of such cwfs is biequivalent to the category
of finitely complete categories.

As to the existence of initial algebras, we present two different term model
constructions, yielding the following statements:

\begin{enumerate}
\item
  All large infinitary QIITs with recursive equations, but \emph{without
  sort equations}, are reducible to the theory of signatures, i.e.\
  their initial algebras are constructible from the initial algebra for
  the theory of signatures.
\item
  All large finitary QIITs with sort equations, but no recursive equations,
  are reducible to the theory of signatures.
\end{enumerate}

The reason for the two separate constructions is that sort equations are not as
well-behaved as the other extensions: they are modelled as strict equalities of
sets in algebras, and so they do not respect isomorphism of sets. We make
essential use of invariance under set isomorphism in the first term model
construction, so we cannot throw sort equations into that mix.

\bibliography{references}

\end{document}

### 2019-10-14 draft/paper.tex
stuff

@@
 \input{abbrevs.tex}
+
+%% abbrevs
+%% ------------------------------------------------------------
+
+\newcommand{\Lift}{\Uparrow}
+
+%% ------------------------------------------------------------
 
@@
 \section{Introduction}
+\label{sec:intro}
 
@@
 
+In Section \ref{sec:levels} we set up cumulative universes which we use in later
+sections. In Section \ref{sec:tos} we introduce the theory of QIIT
+signatures. In Section \ref{sec:cwfmodel} we develop initial algebra semantics,
+also covering the theory of signatures itself. In Sections
+\ref{sec:infinitary_term}-\ref{sec:sort_term} we present the two term model
+constructions. We discuss related work and conclusions in Sections
+\ref{sec:related}-\ref{sec:conclusion}.
+
+\section{Cumulative Extensional Type Theory}
+\label{sec:levels}
+
+In the following sections, we will consider QIIT algebras at arbitrary finite
+levels, along with large eliminations, where the initial algebra can be at a
+different level than the target algebra. For a simple example, consider natural
+number algebras at a given level $i$, given as the $\Sigma$-type $NatAlg_i :\equiv(Nat :
+\Set_i)\times Nat\times(Nat \ra Nat)$. The initial such algebra is the set of
+natural numbers, which is at level $0$, but in type theory we often want to
+eliminate into larger $\Set_i$, for example when computing a $Nat$-indexed family
+of types.
+
+Universe levels are usually viewed as a bureaucratic and tedious part of type
+theory, and are often omitted or mentioned only in passing. Here we intend to
+model universe levels in a precise way, because large eliminations are an
+essential part of the practical usage of inductive types. Hence, we need a way
+to conveniently reason about levels without overmuch administrative
+burden. Unfortunately, simple universe setups are insufficient for this
+purpose. We need two features: cumulativity and a way to quantify over finite
+levels.
+
+\subsection{Cumulativity}
+
+Cumulativity makes handling of algebras and constructions at different levels
+much easier. For example, given an algebra $\gamma : NatAlg_i$, cumulativity
+allows us to also have $\gamma : NatAlg_{i+j}$. In a non-cumulative setting such
+as Agda, we may only have for each type $A : \Set_i$ a type $\mathsf{Lift}_{i\,j}\,A :
+\Set_{i+j}$ which is isomorphic to $A$ up to definitional equality.
+
+In the most basic form, cumulativity requires that whenever $A : \Set_i$, then
+also $A : \Set_{i+j+1}$. However, the $NatAlg$ example above also required a form
+of cumulativity for $\Sigma$ and $\Pi$ types, and we need this in general for
+QIIT algebras.
+
+We use Sterling's cumulative algebraic type theory \cite{TODO} as the general
+setting for the rest of the paper. The reason for this is twofold. First, it
+supports the right kind of cumulativity for our purposes. Second, it is itself a
+finitary QIIT with sort equations, hence its model theory is covered in this
+paper\footnote{Which is, of course, somewhat circular, but not more circular than
+  e.g.\ the study of models of ZFC in ZFC.}.
+
+We extend the base theory with an extensional identity type and
+$\Sigma$-types. In \cite{TODO} a proof of canonicity is provided for the base
+theory, which also includes a standard set-theoretic model. It is
+straightforward to extend the canonicity proof to cover our additional type
+formers.
+
+From now on, we refer to this theory as cETT (cumulative extensional type
+theory). When working in cETT, we use conventional type-theoretic notation. We
+have Russell-style universes $\Set_i$ indexed by natural numbers, dependent
+functions as $(x : A)\ra B$, and dependent pairs as $(x : A)\times B$ with
+projections $\proj1$ and $\proj2$. We also have a \emph{lifting} operation on
+types, which introduces cumulativity. The most important rules for lifting are
+the following:
+
+\begin{figure}
+
+Universe formation
+\[
+  \infer{\Gamma \vdash \Set_i : \Set_{i+j+1}}{}
+\]
+
+Type lifting
+\[
+%% \begin{gathered}
+  \infer{\Gamma \vdash\,\Lift_j A : \Set_{i+j+1}}
+        {\Gamma \vdash A : \Set_i}
+%% \end{gathered}
+\]
+Universe lifting
+\[
+  \Lift_j Set_i = Set_{i+j}
+\]
+Term lifting
+\[
+\{t\,|\,\Gamma \vdash t : A\} = \{t\,|\,\Gamma \vdash t :\,\,\Lift_j A\}
+\]
+Context lifting
+\[
+(\Gamma,\,x : A) = (\Gamma,\,x :\,\Lift_j A)
+\]
+Function lifting
+\[
+\Lift_j((x : A)\ra B) = (x :\,\Lift_j A)\ra\,\Lift_j B
+\]
+Pair lifting
+\[
+\Lift_j((x : A)\times B) = (x :\,\Lift_j A)\,\times\Lift_j B
+\]
+
+\caption{Some of the rules for lifting}
+\label{lifting}
+\end{figure}
+
+In Figure \ref{lifting}, we include an excerpt of Sterling's rules, with a
+slight change of presentation. We use $i+j+1$ levels instead of assuming $i < j$
+and lifting to $\Set_j$. We also omit rules for naturality and substitution of
+lifting. In general, we can expect that lifting never impedes constructions,
+because it appropriately computes out of the way.
+
+Of special note is the \emph{term lifting rule}. It is a sort equation, an
+equation between sets of terms, expressing that lifted types have exactly the
+same terms as unlifted ones. This allows us to have $t :\,\,\Lift_j A$ whenever $t
+: A$. Together with the universe lifting rule, this implies $A : Set_{i+j+1}$
+whenever $A : Set_i$. Similarly, the lifting rules for functions and pairs give
+us cumulativity for $NatAlg$.
+
+\section{Theory of Signatures}
+\label{sec:tos}
+
+\subsection{Bootstrapping}
+
+\section{Categories with Families of Algebras}
+\label{sec:cwfmodel}
+
+\section{Infinitary Term Models}
+\label{sec:infinitary_term}
+
+\subsection{Preservation of Isomorphisms}
+
+\section{Term Models for Sort Equations}
+\label{sec:sort_term}
+
+\section{Related Work}
+\label{sec:related}
+
+\section{Conclusion}
+\label{sec:conclusion}
+
 \bibliography{references}

### 2019-10-15 draft/paper.tex
progress

@@
 
-\usepackage{proof}
-\usepackage{amssymb}
-\usepackage{amsmath}
 \usepackage[utf8]{inputenc}
 \usepackage{todonotes}
-\input{abbrevs.tex}
+\presetkeys{todonotes}{inline}{}
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{amsfonts}
+\usepackage{mathpartir}
+\usepackage{scalerel}
+
+%% Definition env with non-italic body text
+\spnewtheorem{nidefinition}[definition]{Definition}{\bfseries}{\textnormal}
 
@@
 
+\input{abbrevs.tex}
 \newcommand{\Lift}{\Uparrow}
+\newcommand{\ToS}{\mathsf{ToS}}
+\newcommand{\ext}{\triangleright}
+\newcommand{\emptycon}{\scaleobj{.75}\bullet}
+\newcommand{\Pii}{\mathsf{\Pi}}
+\newcommand{\appi}{\mathsf{app}}
+\newcommand{\lami}{\mathsf{lam}}
+\newcommand{\Pie}{\mathsf{\hat{\Pi}}}
+\newcommand{\appe}{\mathsf{ap\hat{p}}}
+\newcommand{\lame}{\mathsf{la\hat{m}}}
+\newcommand{\Piinf}{\mathsf{\tilde{\Pi}}}
+\newcommand{\appinf}{\mathsf{ap\tilde{p}}}
+\newcommand{\laminf}{\mathsf{la\tilde{m}}}
+
+\newcommand{\appitt}{\mathop{{\scriptstyle @}}}
+
+\newcommand{\Refl}{\mathsf{Refl}}
+
+\newcommand{\IdU}{\mathsf{IdU}}
+\newcommand{\ReflU}{\mathsf{ReflU}}
+\newcommand{\Sig}{\mathsf{Sig}}
 
@@
 features in order to write down the signature of categories:
-\begin{alignat*}{5}
+\begin{alignat*}{3}
   & Obj && : \Set \\
@@
 previous works. As a beneficial side effect of fulfilling this goal, infinitary
-QIITs such as real numbers will be covered as well.
+QIITs such as Cauchy real numbers will be covered as well.
 
@@
   theory, but it has not been presented explicitly in previous works about
-  QIITs. In order to handle universe levels gracefully, we develop a set of
-  tools, working inside an extensional type theory with \emph{algebraic
-    cumulativity}.
+  QIITs. In order to handle universe levels gracefully, we work in an
+  extensional type theory extended with cumulative universes.
   \item
@@
 We also develop semantics. We show that for each signature, there is a cwf
-(category with families) of algebras, extended with $\Sigma$-types, an
-extensional identity type and constant families. This additional structure
-corresponds to a type-theoretic flavor of finite limits, and it was shown
-in \cite{TODO} that the category of such cwfs is biequivalent to the category
-of finitely complete categories.
+(category with families) of algebras, extended with $\Sigma$-types, extensional
+identity, and constant families. This additional structure corresponds to a
+type-theoretic flavor of finite limits, and it was shown in \cite{TODO} that the
+category of such cwfs is biequivalent to the category of finitely complete
+categories.
 
 As to the existence of initial algebras, we present two different term model
-constructions, yielding the following statements:
+constructions, yielding the following results:
 
@@
 well-behaved as the other extensions: they are modelled as strict equalities of
-sets in algebras, and so they do not respect isomorphism of sets. We make
+sets in algebras, hence they do not respect isomorphism of sets. We make
 essential use of invariance under set isomorphism in the first term model
@@
 essential part of the practical usage of inductive types. Hence, we need a way
-to conveniently reason about levels without overmuch administrative
+to conveniently reason about levels without too much administrative
 burden. Unfortunately, simple universe setups are insufficient for this
@@
 In the most basic form, cumulativity requires that whenever $A : \Set_i$, then
-also $A : \Set_{i+j+1}$. However, the $NatAlg$ example above also required a form
-of cumulativity for $\Sigma$ and $\Pi$ types, and we need this in general for
-QIIT algebras.
+also $A : \Set_{i+j+1}$. However, the $NatAlg$ example above also required a
+form of cumulativity for $\Sigma$ and $\Pi$ types, which we also need in general
+for QIIT algebras.
 
@@
 supports the right kind of cumulativity for our purposes. Second, it is itself a
-finitary QIIT with sort equations, hence its model theory is covered in this
-paper\footnote{Which is, of course, somewhat circular, but not more circular than
+finitary QIIT with sort equations, hence this paper provides a model theory for
+it\footnote{Which is, of course, somewhat circular, but not more circular than
   e.g.\ the study of models of ZFC in ZFC.}.
 
-We extend the base theory with an extensional identity type and
-$\Sigma$-types. In \cite{TODO} a proof of canonicity is provided for the base
+We extend the base theory with an extensional identity type, $\Sigma$-types and
+the unity type. In \cite{TODO} a proof of canonicity is provided for the base
 theory, which also includes a standard set-theoretic model. It is
@@
 From now on, we refer to this theory as cETT (cumulative extensional type
-theory). When working in cETT, we use conventional type-theoretic notation. We
+theory). When working in cETT, we use the following notation. We
 have Russell-style universes $\Set_i$ indexed by natural numbers, dependent
 functions as $(x : A)\ra B$, and dependent pairs as $(x : A)\times B$ with
-projections $\proj1$ and $\proj2$. We also have a \emph{lifting} operation on
-types, which introduces cumulativity. The most important rules for lifting are
-the following:
+projections $\proj1$ and $\proj2$. We also use a field projection notation
+for iterated pairs. For example, for $t : (A : \Set_i) \times (B : \Set_i)
+\times (f : A \ra B)$, we use $t.B$ to denote the projection of the second
+component.
+
+We write propositional equality as $t = u$, with $\refl_t$ for reflexivity. We
+have equality reflection for $\blank=\blank$. The unit type is $\top$, with
+inhabitant $\tt$.
+
+We also have a lifting operation on types, which introduces cumulativity.
 
 \begin{figure}
+\begin{mathpar}
+  \inferrule*[lab=universe formation]
+             {\\}
+             {\Gamma \vdash \Set_i : \Set_{i+j+1}}
 
-Universe formation
-\[
-  \infer{\Gamma \vdash \Set_i : \Set_{i+j+1}}{}
-\]
+  \inferrule*[lab=type lifting]
+             {\Gamma \vdash A : \Set_i}
+             {\Gamma \vdash\,\Lift_j A : \Set_{i+j+1}}
 
-Type lifting
-\[
-%% \begin{gathered}
-  \infer{\Gamma \vdash\,\Lift_j A : \Set_{i+j+1}}
-        {\Gamma \vdash A : \Set_i}
-%% \end{gathered}
-\]
-Universe lifting
-\[
-  \Lift_j Set_i = Set_{i+j}
-\]
-Term lifting
-\[
-\{t\,|\,\Gamma \vdash t : A\} = \{t\,|\,\Gamma \vdash t :\,\,\Lift_j A\}
-\]
-Context lifting
-\[
-(\Gamma,\,x : A) = (\Gamma,\,x :\,\Lift_j A)
-\]
-Function lifting
-\[
-\Lift_j((x : A)\ra B) = (x :\,\Lift_j A)\ra\,\Lift_j B
-\]
-Pair lifting
-\[
-\Lift_j((x : A)\times B) = (x :\,\Lift_j A)\,\times\Lift_j B
-\]
+  \inferrule*[lab=universe lifting]{}{\Lift_j \Set_i = \Set_{i+j}}
 
-\caption{Some of the rules for lifting}
+  \inferrule*[lab=term lifting]
+             {}
+             {\{t\,|\,\Gamma \vdash t : A\} = \{t\,|\,\Gamma \vdash t :\,\,\Lift_j A\}}
+
+  \inferrule*[lab=context lifting]
+             {}
+             {(\Gamma,\,x : A) = (\Gamma,\,x :\,\,\Lift_j A)}
+
+  \inferrule*[lab=function lifting]
+             {}
+             {\Lift_j((x : A)\ra B) = (x :\,\Lift_j A)\ra\,\Lift_j B}
+
+  \inferrule*[lab=pair lifting]
+             {}
+             {\Lift_j((x : A)\times B) = (x :\,\Lift_j A)\,\times\Lift_j B}
+\end{mathpar}
+
+\caption{Some of the rules for lifting.}
 \label{lifting}
@@
 In Figure \ref{lifting}, we include an excerpt of Sterling's rules, with a
-slight change of presentation. We use $i+j+1$ levels instead of assuming $i < j$
-and lifting to $\Set_j$. We also omit rules for naturality and substitution of
-lifting. In general, we can expect that lifting never impedes constructions,
-because it appropriately computes out of the way.
+slight change of presentation. We lift from $i$ to $i+j+1$ levels instead of
+assuming $i < j$ and lifting to $j$. We also omit rules for naturality and
+substitution of lifting. In general, we can expect that lifting never impedes
+constructions, because it appropriately computes out of the way.
 
@@
 whenever $A : Set_i$. Similarly, the lifting rules for functions and pairs give
-us cumulativity for $NatAlg$.
+us cumulativity for $NatAlg$. We derive a notion of subtyping:
 
-\section{Theory of Signatures}
+\begin{nidefinition}{Cumulative subtyping.}
+  For $A : \Set_i$ and $B : \Set_{i+j+1}$, we define $A \prec B : \Set_{i+j+1}$
+  as $B =\,\,\Lift_j A$.  It follows from term lifting and equality reflection
+  that whenever $A \prec B$ and $t : A$, then also $t : B$.
+\end{nidefinition}
+
+\subsection{Universe Polymorphism}
+\label{sec:universe_polymorphism}
+
+We also need to quantify over universe levels. cETT does not support this, and
+we leave it like that. Instead, for the sake of simplicity, we quantify over
+levels in an unspecified metatheory \emph{outside} cETT. Hence, a universe
+polymorphic cETT term is understood as a $\mathbb{N}$-indexed family of
+terms. We use the $\forall$ symbol for this, e.g.\ as $\forall\,i.\,\Set_i \ra
+\Set_i$. This is sufficient, because we do not need to internalize
+level-polymorphic constructions in cETT.
+
+\section{Signatures}
 \label{sec:tos}
 
-\subsection{Bootstrapping}
+In this section, we define signatures for QIITs which support all the mentioned
+generalizations. Signatures are given as contexts in a certain type theory,
+called the theory of signatures. We shall abbreviate it as ToS. However, ToS
+turns out to be a large infinitary QIIT itself, and we would like to define ToS
+and a notion of signature without referring to QIITs, only using features
+present in cETT. As a first step, we define a notion of model.
+
+\begin{nidefinition}{Notion of model for ToS.}\label{def:tos}
+For levels $i$ and $j$, there is a cETT type $\ToS_{i,j} : \Set_{\max(i,\,j)+1}$, whose elements are
+ToS models (or ToS-algebras). $\ToS_{i,j}$ is an iterated $\Sigma$-type, containing
+all of the following components.
+\begin{enumerate}
+\item
+  A category with families (cwf), where all four underlying sets (of
+  objects, morphisms, types and terms) are in $\Set_i$. Following notation in
+  \cite{TODO}, we denote these respectively as $\Con : \Set_i$, $\Sub : \Con \ra
+  \Con \ra \Set_i$, $\Ty : \Con \ra \Set_i$ and $\Tm : (\Gamma : \Con) \ra
+  \Ty\,\Gamma \ra \Set_i$. We denote the empty context as $\emptycon : \Con$,
+  context extension as $\blank\ext\blank : (\Gamma : \Con)\ra \Ty\,\Gamma \ra \Con$.
+  Substitution on types and terms is written as $\blank[\blank]$.
+\item
+  A universe $\U : \Ty\,\Gamma$ with decoding $\El : (a : \Tm\,\Gamma\,\U) \ra
+  \Ty\,\Gamma$.
+\item
+  Inductive function space, with $\Pii : (a : \Tm\,\Gamma\,\U) \ra
+  \Ty\,(\Gamma\,\ext\,\El\,a) \ra \Ty\,\Gamma$, application as $\appi :
+  \Tm\,\Gamma\,(\Pii\,a\,B)\ra \Tm\,(\Gamma\,\ext\,\El\,a)\,B$, and its inverse
+  $\lami$.
+\item
+  External function space with $\Set_j$ domain: $\Pie : (A : \Set_j)\ra(A \ra
+  \Ty\,\Gamma)\ra \Ty\,\Gamma$, with $\appe : \Tm\,\Gamma\,(\Pie\,A\,B)\ra(x :
+  A)\ra \Tm\,\Gamma\,(B\,x)$, and likewise $\lame$ as inverse.
+\item
+  Infinitary function space, with $\Piinf : (A : \Set_j)\ra(A \ra
+  \Tm\,\Gamma\,\U)\ra \Tm\,\Gamma\,\U$, $\appinf :
+  \Tm\,\Gamma\,(\El\,(\Pie\,A\,b))\ra(x : A)\ra \Tm\,\Gamma\,(\El\,(b\,x))$, and
+  $\laminf$.
+\item
+  An identity type $\Id : (a : \Tm\,\Gamma\,\U)\ra
+  \Tm\,\Gamma\,(\El\,a)\ra\Tm\,\Gamma\,(\El\,a)\ra \Tm\,\Gamma\,\U$, with
+  $\Refl : (T : \Tm\,\Gamma\,(\El\,a))\ra \Id\,a\,t\,t$ and equality reflection.
+\item
+  An identity type $\IdU : \Tm\,\Gamma\,\U \ra \Tm\,\Gamma\,\U \ra \Ty\,\Gamma$, with
+  evident $\ReflU$ and equality reflection.
+\end{enumerate}
+\end{nidefinition}
+In the above listing, we omit most cwf components and equations for substitution and
+$\beta\eta$-conversion, but these should be understood to be also part of $\ToS_{i,j}$.
+
+%% \paragraph{Notational conventions.}
+\emph{Notational conventions.} Cwf components by default support de Bruijn
+indices, which are not easily readable. We use instead a nameful notation for
+binders in context extension, $\Pii$ and $\lami$, e.g.\ as $(\emptycon \ext a :
+U \ext t : \El\,a)$. We also define a type-theoretic version of $\appi$ for
+convenience:
+\begin{alignat*}{3}
+  & \blank\appitt\blank && :
+      \Tm\,\Gamma\,(\Pii\,a\,B)\ra
+      (u : \Tm\,\Gamma\,(\El\,a)) \ra \Tm\,\Gamma\,(B[id, u])\\
+  & t\appitt u && :\equiv (\appi\,t)[id, u]
+\end{alignat*}
+
+\begin{nidefinition}{Notion of signature.}
+A QIIT signature at level $j$ is a context in an arbitrary $M : \ToS_{i,j}$
+model. We define the type of such signatures as follows:
+\[
+  \Sig_j :\equiv \forall i.\,(M : \ToS_{i,j})\ra M.Con
+\]
+
+Note that $\Sig_j$ is defined as a universe-polymorphic type, so it does not
+have a type internally to cETT. A concrete signature $sig$ at level $j$ has type
+$\forall i.\,(M : \ToS_{i,j})\ra M.Con$, so it builds up a context from the
+components of an arbitrary $\ToS$ model.
+
+\begin{example}
+The signature for natural
+numbers is the following, where $\emptycon$, $\blank\ext\blank$, $\U$, $\El$ and
+$\Pii$ all refer to components of $M$:
+\begin{alignat*}{3}
+  & NatSig : \Sig_0 \\
+  & NatSig :\equiv \lambda (M : \ToS_{i,0}). (\emptycon\,\ext\, N : \U \,\ext\, zero : \El\,N \ext suc : \Pii\,(n : N)\,(\El\,N))
+\end{alignat*}
+\end{example}
+
+The $j$ level in $\Sig_j$ specifies the level of the external types appearing in
+a signatures, since the $\Pie$ and $\Piinf$ function spaces have domains in $\Set_j$.
+In the case of naturals numbers, no external types appear, so the level of $NatSig$
+can be chosen as $0$.
+
+With this, we are able to specify QIITs, and we can also interpret each
+signature in an arbitrary ToS model, just by applying a signature to a model.
+$\Sig_j$ can be viewed as a precursor to a Church-encoding for the theory of
+signatures, but we only need contexts encoded in this way, and not other ToS
+components. In functional programming, this encoding is sometimes called
+``finally tagless''\cite{TODO}, and it is used for defining and interpreting
+embedded languages.
+\end{nidefinition}
+
+\subsection{Example Signatures}
+
+Compared to signatures in the previous work \cite{TODO}, we additionally support
+$\Piinf$ and $\IdU$, and $\Id$ has a more liberal type which allows recursive
+equations. Let us see examples for these three in order. For brevity, we shall
+only write contexts, and assume that all ToS operations come from an arbitrary
+model.
+
+\begin{example}{Infinitary constructors}. The universe $\U$ is closed under
+the $\Piinf$ function type, which makes it possible to write such functions
+in the domains of $\Pii$ types. This allows a signature for $W$-types, given
+arbitrary $S : \Set_i$ and $P : S \ra \Set_i$ types for shapes and positions.
+\[
+WSig :\equiv
+\emptycon \ext W : \U \ext sup : \Pie\,S\,(\lambda s.\,\Pii\,(\Piinf\,(P\,s)\,\lambda \_.\,W) \,(\El\,W))
+\]
+\end{example}
+
+\begin{example}{Sort equations}. An interesting example is the term lifting
+rule for cETT (omitting everything except term lifting):
+\begin{alignat*}{3}
+  & cETTSig :\equiv \\
+  & ...\,\ext\,Tm{\Lift} : \Pii\,(\Gamma : Con)\,(\Pii\,(A : Ty{\appitt}\Gamma)((\IdU\,(Tm{\appitt}\Gamma{\appitt}A)\,(Tm{\appitt}\Gamma{\appitt}(\Lift_j{\appitt}A))))
+\end{alignat*}
+\end{example}
+
+\begin{example}{Recursive equations}. Our definition of $\Id$ returns in $\U$,
+which, similarly to the case of infinitary functions, allows us to write equations
+in $\Pii$ domains. A minimal example:
+\[
+RecEqSig :\equiv \emptycon \ext A : \U \ext a : \El\,A \ext f : \Pii\,(x : A)\,(\Pii\,(\Id\,A\,x\,a)\,(\El\,A))
+\]
+There more complicated and interesting examples, e.g.\ boundary conditions in
+various cubical type theories\cite{TODO}. Note that our $\Id$ allows iterated
+equations as well, but these are all made trivial in the semantics, where we
+assume uniqueness of identity proofs.
+\end{example}
+
+\subsection{A Signature for the Theory of Signatures}
+
+\todo{put this in infinitary term models beginning}
+
+Starting from Section \ref{sec:infinitary_term}, we need to do proofs and
+constructions by induction on signatures. However, the current signatures do not
+support induction, analogously to how induction is not available for Church
+encodings. We would like to have two things: a notion of signature induction,
+and a model of ToS which supports induction. Then, we can use contexts in the
+syntax of ToS (i.e.\ in the inductive model) as signatures. We provide a notion
+of induction on signatures by
+\begin{enumerate}
+  \item Showing that there is a signature which describes ToS.
+  \item Building categorical semantics, which for each signature provides a notion of induction.
+\end{enumerate}
+On the other hand, in this paper we will only \emph{assume} the existence of a
+syntax for ToS starting from Section \ref{sec:infinitary_term}. We do not provide a
+construction of the syntax from simpler type formers, although we do show that
+lower-level ToS syntaxes are constructible from higher-level ones.
 

### 2019-10-16 draft/paper.tex
progress

@@
 \newcommand{\Sig}{\mathsf{Sig}}
+
+\newcommand{\Sg}{\mathsf{\Sigma}}
+\newcommand{\flCwF}{\mathsf{flCwF}}
+\newcommand{\Kfam}{\mathsf{K}}
 
@@
 
-We also develop semantics. We show that for each signature, there is a cwf
+We also develop semantics. We show that for each signature, there is a CwF
 (category with families) of algebras, extended with $\Sigma$-types, extensional
@@
 type-theoretic flavor of finite limits, and it was shown in \cite{TODO} that the
-category of such cwfs is biequivalent to the category of finitely complete
+category of such CwFs is biequivalent to the category of finitely complete
 categories.
@@
 sections. In Section \ref{sec:tos} we introduce the theory of QIIT
-signatures. In Section \ref{sec:cwfmodel} we develop initial algebra semantics,
+signatures. In Section \ref{sec:categorical_semantics} we develop initial algebra semantics,
 also covering the theory of signatures itself. In Sections
-\ref{sec:infinitary_term}-\ref{sec:sort_term} we present the two term model
+\ref{sec:infinitary_term_model}-\ref{sec:sort_term} we present the two term model
 constructions. We discuss related work and conclusions in Sections
@@
 \times (f : A \ra B)$, we use $t.B$ to denote the projection of the second
-component.
+component. Sometimes we omit the $t.$ prefix if it is clear from context.
 
@@
 terms. We use the $\forall$ symbol for this, e.g.\ as $\forall\,i.\,\Set_i \ra
-\Set_i$. This is sufficient, because we do not need to internalize
+\Set_i$. This is sufficient in this paper, because we do not need to internalize
 level-polymorphic constructions in cETT.
@@
 \item
-  A category with families (cwf), where all four underlying sets (of
+  A category with families (CwF), where all four underlying sets (of
   objects, morphisms, types and terms) are in $\Set_i$. Following notation in
@@
   \Tm\,\Gamma\,(\El\,a)\ra\Tm\,\Gamma\,(\El\,a)\ra \Tm\,\Gamma\,\U$, with
-  $\Refl : (T : \Tm\,\Gamma\,(\El\,a))\ra \Id\,a\,t\,t$ and equality reflection.
+  $\Refl : (T : \Tm\,\Gamma\,(\El\,a))\ra \Id\,a\,t\,t$, equality reflection
+  and uniqueness of identity proofs.
 \item
   An identity type $\IdU : \Tm\,\Gamma\,\U \ra \Tm\,\Gamma\,\U \ra \Ty\,\Gamma$, with
-  evident $\ReflU$ and equality reflection.
+  evident $\ReflU$, equality reflection and uniqueness of identity proofs.
 \end{enumerate}
 \end{nidefinition}
-In the above listing, we omit most cwf components and equations for substitution and
+In the above listing, we omit most CwF components and equations for substitution and
 $\beta\eta$-conversion, but these should be understood to be also part of $\ToS_{i,j}$.
@@
 %% \paragraph{Notational conventions.}
-\emph{Notational conventions.} Cwf components by default support de Bruijn
+\emph{Notational conventions.} CwF components by default support de Bruijn
 indices, which are not easily readable. We use instead a nameful notation for
@@
 numbers is the following, where $\emptycon$, $\blank\ext\blank$, $\U$, $\El$ and
-$\Pii$ all refer to components of $M$:
+$\Pii$ all refer to components of the $\lambda$-bound $M$:
 \begin{alignat*}{3}
@@
 \]
-There more complicated and interesting examples, e.g.\ boundary conditions in
+There are more complicated and interesting examples, e.g.\ boundary conditions in
 various cubical type theories\cite{TODO}. Note that our $\Id$ allows iterated
@@
 
-\subsection{A Signature for the Theory of Signatures}
+\section{Categorical Semantics}
+\label{sec:categorical_semantics}
 
-\todo{put this in infinitary term models beginning}
+For each signature, we would like to have at least
+\begin{enumerate}
+  \item A category of algebras, with homomorphisms as morphisms.
+  \item A notion of induction, which requires a notion of dependent algebras.
+  \item A proof that for algebras, initiality is equivalent to supporting induction.
+\end{enumerate}
 
-Starting from Section \ref{sec:infinitary_term}, we need to do proofs and
+Following \cite{TODO}, we do this by creating a model of ToS, where contexts are
+categories supporting the above requirements and substitutions are strictly
+structure-preserving functors. Then, each signature can applied to this model,
+yielding an interpretation of the signature as a structured category of
+algebras. First, we fix our notion of such structured category.
+
+\begin{nidefinition}{Finite limit CwFs.}
+For a level $i$, there is a cETT type $\flCwF_i : \Set_{i+1}$, which is an
+iterated $\Sigma$-type with the following components:
+\begin{enumerate}
+  \item A CwF with underlying sets all in $\Set_i$.
+  \item $\Sigma$-type $\Sg : (A : \Ty\,\Gamma)\ra \Ty\,(\Gamma \ext A)
+    \ra \Ty\,\Gamma$, with usual projections and pairing.
+  \item Identity type $\Id : (A : \Ty\,\Gamma)\ra \Tm\,\Gamma\,A\ra
+    \Tm\,\Gamma\,A\ra \Ty\,\Gamma$, with equality reflection and uniqueness
+    of identity proofs.
+  \item Constant families. This includes a type former $\Kfam : \Con \ra
+    \Ty\,\Gamma$, where $\Gamma$ is implicitly quantified, and an isomorphism
+    between $\Sub\,\Gamma\,\Delta$ and $\Tm\,\Gamma\,(\Kfam\,\Delta)$ which is
+    natural in $\Gamma$. The idea is that $\Kfam\,\Delta$ is a representation
+    of $\Delta$ as a type in any context. Dybjer and Clairambault called constant
+    families ``democracy'' in \cite{TODO}.
+\end{enumerate}
+\end{nidefinition}
+
+We use $\Ty$ in an flCwF to represent ``dependent objects'', and $\Tm$
+for ``dependent morphisms''. We use these to define induction.
+
+\begin{nidefinition} Notion of induction internally to $C : \flCwF_i$:
+\begin{alignat*}{3}
+  & Inductive : C.\Con \ra \Set_i \\
+  & Inductive\,\Gamma :\equiv (A : C.\Ty\,\Gamma)\ra C.\Tm\,\Gamma\,A
+\end{alignat*}
+\end{nidefinition}
+
+In the categorical semantics for a concrete signature, this informally means
+that an algebra supports induction if for any dependent algebra over it (which
+is a bundle of induction motives and methods), there is a dependent morphism
+into it (which is a bundle of eliminator functions and their $\beta$-rules).
+
+\begin{theorem}{Equivalence of initiality and induction.}
+An object $\Gamma : C.\Con$ supports induction if and only if it is
+initial. Moreover, induction and initiality are both proof irrelevant
+predicates on objects.
+\end{theorem}
+\begin{proof} Shown by brief internal flCwF reasoning in \cite{TODO}.\qed
+\end{proof}
+
+The reason for the ``finite limit CwF'' naming is the following: Dybjer and
+Clairambault showed\cite{TODO} that the 2-category of flCwFs is biequivalent to
+the 2-category of finitely complete (or ``lex'') categories. It can be also seen
+that the categorical product of $\Gamma$ and $\Delta$ in an flCwF can be given
+as $\Gamma \ext \Kfam\, \Delta$, and equalizers can be given using the identity
+type.
+
+So, why do we use flCwF instead of lex categories? After all, the former
+structure has more components and has more immediate complexity than the
+latter. The reason is that flCwFs allow us to define induction in strictly the
+same way as one would write it down in type theory, since we have $\Ty$ and
+$\Tm$ for a primitive notion of dependent objects and morphisms. In contrast,
+dependent objects in lex categories is a derived notion, and the induction
+principles we get are what we want only up to isomorphisms. This issue is
+perhaps not very relevant to those with a purely categorical perspective, but we
+are concerned with eventually implementing QIITs in proof assistants, so we
+prefer if our semantics computes notions strictly.
+
+\begin{nidefinition}{flCwF model for ToS.} For $i$ and $j$ levels, there is
+an $M : \ToS_{1+\max(i+1,j),j}$ such that $M.\Con = \flCwF_{\max(i+1,j)}$.
+
+\todo{flip level indices in ToS, maybe shorten cETT part, add more explanation
+      on role of i,j in ToS(i,j)}
+
+\end{nidefinition}
+
+\section{Infinitary Term Models}
+\label{sec:infinitary_term_models}
+
+\subsection{Induction on Signatures}
+
+\todo{sync}
+
+Starting from Section \ref{sec:infinitary_term_models}, we need to do proofs and
 constructions by induction on signatures. However, the current signatures do not
@@
 On the other hand, in this paper we will only \emph{assume} the existence of a
-syntax for ToS starting from Section \ref{sec:infinitary_term}. We do not provide a
+syntax for ToS starting from Section \ref{sec:infinitary_term_models}. We do not provide a
 construction of the syntax from simpler type formers, although we do show that
@@
 
-\section{Categories with Families of Algebras}
-\label{sec:cwfmodel}
-
-\section{Infinitary Term Models}
-\label{sec:infinitary_term}
-
 \subsection{Preservation of Isomorphisms}
 
-\section{Term Models for Sort Equations}
+\section{Term Models with Sort Equations}
 \label{sec:sort_term}

### 2019-10-17 draft/paper.tex
progress

@@
 different level than the target algebra. For a simple example, consider natural
-number algebras at a given level $i$, given as the $\Sigma$-type $NatAlg_i :\equiv(Nat :
-\Set_i)\times Nat\times(Nat \ra Nat)$. The initial such algebra is the set of
-natural numbers, which is at level $0$, but in type theory we often want to
-eliminate into larger $\Set_i$, for example when computing a $Nat$-indexed family
-of types.
+number algebras at a given level $i$, given as the $\Sigma$-type $NatAlg_i
+:\equiv(Nat : \Set_i)\times Nat\times(Nat \ra Nat)$. The initial such algebra is
+the set of natural numbers, which is at level $0$, but in type theory we often
+want to eliminate into larger $\Set_i$, for example when computing a
+$Nat$-indexed family of types. To support convenient reasoning about levels in
+this paper, we need two features: cumulativity and a way to quantify over finite
+levels.
 
-Universe levels are usually viewed as a bureaucratic and tedious part of type
-theory, and are often omitted or mentioned only in passing. Here we intend to
-model universe levels in a precise way, because large eliminations are an
-essential part of the practical usage of inductive types. Hence, we need a way
-to conveniently reason about levels without too much administrative
-burden. Unfortunately, simple universe setups are insufficient for this
-purpose. We need two features: cumulativity and a way to quantify over finite
-levels.
+%% Universe levels are usually viewed as a bureaucratic and tedious part of type
+%% theory, and are often omitted or mentioned only in passing. Here we intend to
+%% model universe levels in a precise way, because large eliminations are an
+%% essential part of the practical usage of inductive types. Hence, we need a way
+%% to conveniently reason about levels without too much administrative
+%% burden. Unfortunately, simple universe setups are insufficient for this
+%% purpose. We need two features: cumulativity and a way to quantify over finite
+%% levels.
 
@@
 
-Cumulativity makes handling of algebras and constructions at different levels
-much easier. For example, given an algebra $\gamma : NatAlg_i$, cumulativity
-allows us to also have $\gamma : NatAlg_{i+j}$. In a non-cumulative setting such
-as Agda, we may only have for each type $A : \Set_i$ a type $\mathsf{Lift}_{i\,j}\,A :
-\Set_{i+j}$ which is isomorphic to $A$ up to definitional equality.
+%% Cumulativity makes handling of algebras and constructions at different levels
+%% much easier. For example, given an algebra $\gamma : NatAlg_i$, cumulativity
+%% allows us to also have $\gamma : NatAlg_{i+j}$. In a non-cumulative setting such
+%% as Agda, we may only have for each type $A : \Set_i$ a type $\mathsf{Lift}_{i\,j}\,A :
+%% \Set_{i+j}$ which is isomorphic to $A$ up to definitional equality.
 
-In the most basic form, cumulativity requires that whenever $A : \Set_i$, then
-also $A : \Set_{i+j+1}$. However, the $NatAlg$ example above also required a
-form of cumulativity for $\Sigma$ and $\Pi$ types, which we also need in general
-for QIIT algebras.
+We need cumulativity to reduce administrative overhead. In the most basic form,
+cumulativity requires that whenever $A : \Set_i$, and $i < j$ then also $A :
+\Set_j$. We also need cumulativity for $\Sigma$ and function types, in order to
+have cumulativity for algebras in general. For example, we want to have $\gamma
+: NatAlg_j$ whenever $\gamma : NatAlg_i$ and $i < j$.
+
+%% However, the $NatAlg$ example above also required a form of
+%% cumulativity for $\Sigma$ and $\Pi$ types, which we also need in general for
+%% QIIT algebras.
 
@@
 We extend the base theory with an extensional identity type, $\Sigma$-types and
-the unity type. In \cite{TODO} a proof of canonicity is provided for the base
+the unit type. In \cite{TODO} a proof of canonicity is provided for the base
 theory, which also includes a standard set-theoretic model. It is
@@
 We also have a lifting operation on types, which introduces cumulativity.
-
+\vspace{-1.5em}
 \begin{figure}
@@
 
+  \inferrule*[lab=function formation]
+             {\Gamma \vdash A : \U_i \\ \Gamma,\,x : A \vdash B : \U_j}
+             {\Gamma \vdash (x : A)\ra B : \U_{\max(i,j)} }
+
+  \inferrule*[lab=$\Sigma$ formation]
+             {\Gamma \vdash A : \U_i \\ \Gamma,\,x : A \vdash B : \U_j}
+             {\Gamma \vdash (x : A)\times B : \U_{\max(i,j)} }
+
+  \inferrule*[lab=unit formation]
+             {\\}
+             {\Gamma \vdash \top : \Set_i}
+
   \inferrule*[lab=type lifting]
-             {\Gamma \vdash A : \Set_i}
-             {\Gamma \vdash\,\Lift_j A : \Set_{i+j+1}}
+             {\Gamma \vdash A : \Set_i \\ i < j}
+             {\Gamma \vdash\,\Lift_j A : \Set_{j}}
 
-  \inferrule*[lab=universe lifting]{}{\Lift_j \Set_i = \Set_{i+j}}
+  \inferrule*[lab=lift composition]{}{\Lift_j (\Lift_i A) =\,\,\Lift_{j} A}
 
@@
              {}
-             {\{t\,|\,\Gamma \vdash t : A\} = \{t\,|\,\Gamma \vdash t :\,\,\Lift_j A\}}
+             {\{t\,|\,\Gamma \vdash t : A\} = \{t\,|\,\Gamma \vdash t :\,\,\Lift_i A\}}
 
@@
              {}
-             {(\Gamma,\,x : A) = (\Gamma,\,x :\,\,\Lift_j A)}
+             {(\Gamma,\,x : A) = (\Gamma,\,x :\,\,\Lift_i A)}
+
+  \inferrule*[lab=universe lifting]{}{\Lift_j \Set_i = \Set_i}
 
@@
              {}
-             {\Lift_j((x : A)\ra B) = (x :\,\Lift_j A)\ra\,\Lift_j B}
+             {\Lift_i((x : A)\ra B) = (x :\,\Lift_i A)\ra\,\Lift_i B}
 
@@
              {}
-             {\Lift_j((x : A)\times B) = (x :\,\Lift_j A)\,\times\Lift_j B}
+             {\Lift_i((x : A)\times B) = (x :\,\Lift_i A)\,\times\Lift_i B}
+
+  \inferrule*[lab=unit lifting]
+             {}
+             {\Lift_i\top = \top}
 \end{mathpar}
@@
 
-In Figure \ref{lifting}, we include an excerpt of Sterling's rules, with a
-slight change of presentation. We lift from $i$ to $i+j+1$ levels instead of
-assuming $i < j$ and lifting to $j$. We also omit rules for naturality and
-substitution of lifting. In general, we can expect that lifting never impedes
-constructions, because it appropriately computes out of the way.
+In Figure \ref{lifting}, we include an excerpt of cETT's rules. We change
+presentation slightly from Sterling. We lift from $i$ to $i+j+1$ levels instead
+of assuming $i < j$ and lifting to $j$, and we use a $\U_{\max(i,j)}$ return
+type in functions and pairs, instead of having both $A$, $B$ types in the same
+$\Set_i$. These changed rules are all derivable from the original ones. In
+general, we can expect that lifting never impedes constructions, because it
+appropriately computes out of the way.
 
@@
 polymorphic cETT term is understood as a $\mathbb{N}$-indexed family of
-terms. We use the $\forall$ symbol for this, e.g.\ as $\forall\,i.\,\Set_i \ra
+terms. We use the $\forall$ symbol for this, e.g.\ as $t : \forall\,i.\,\Set_i \ra
 \Set_i$. This is sufficient in this paper, because we do not need to internalize
@@
   A category with families (CwF), where all four underlying sets (of
-  objects, morphisms, types and terms) are in $\Set_i$. Following notation in
-  \cite{TODO}, we denote these respectively as $\Con : \Set_i$, $\Sub : \Con \ra
-  \Con \ra \Set_i$, $\Ty : \Con \ra \Set_i$ and $\Tm : (\Gamma : \Con) \ra
-  \Ty\,\Gamma \ra \Set_i$. We denote the empty context as $\emptycon : \Con$,
+  objects, morphisms, types and terms) are in $\Set_j$. Following notation in
+  \cite{TODO}, we denote these respectively as $\Con : \Set_j$, $\Sub : \Con \ra
+  \Con \ra \Set_j$, $\Ty : \Con \ra \Set_j$ and $\Tm : (\Gamma : \Con) \ra
+  \Ty\,\Gamma \ra \Set_j$. We denote the empty context as $\emptycon : \Con$,
   context extension as $\blank\ext\blank : (\Gamma : \Con)\ra \Ty\,\Gamma \ra \Con$.
@@
   Inductive function space, with $\Pii : (a : \Tm\,\Gamma\,\U) \ra
-  \Ty\,(\Gamma\,\ext\,\El\,a) \ra \Ty\,\Gamma$, application as $\appi :
-  \Tm\,\Gamma\,(\Pii\,a\,B)\ra \Tm\,(\Gamma\,\ext\,\El\,a)\,B$, and its inverse
-  $\lami$.
+  \Ty\,(\Gamma\,\ext\,\El\,a) \ra \Ty\,\Gamma$ and application as $\appi :
+  \Tm\,\Gamma\,(\Pii\,a\,B)\ra \Tm\,(\Gamma\,\ext\,\El\,a)\,B$. We do not
+  support $\lami$ here and in other function types, as they are not essential
+  for QIIT signatures.
 \item
-  External function space with $\Set_j$ domain: $\Pie : (A : \Set_j)\ra(A \ra
+  External function space with $\Set_i$ domain: $\Pie : (A : \Set_i)\ra(A \ra
   \Ty\,\Gamma)\ra \Ty\,\Gamma$, with $\appe : \Tm\,\Gamma\,(\Pie\,A\,B)\ra(x :
-  A)\ra \Tm\,\Gamma\,(B\,x)$, and likewise $\lame$ as inverse.
+  A)\ra \Tm\,\Gamma\,(B\,x)$.
 \item
-  Infinitary function space, with $\Piinf : (A : \Set_j)\ra(A \ra
+  Infinitary function space, with $\Piinf : (A : \Set_i)\ra(A \ra
   \Tm\,\Gamma\,\U)\ra \Tm\,\Gamma\,\U$, $\appinf :
-  \Tm\,\Gamma\,(\El\,(\Pie\,A\,b))\ra(x : A)\ra \Tm\,\Gamma\,(\El\,(b\,x))$, and
-  $\laminf$.
+  \Tm\,\Gamma\,(\El\,(\Pie\,A\,b))\ra(x : A)\ra \Tm\,\Gamma\,(\El\,(b\,x))$.
 \item
@@
 
-%% \paragraph{Notational conventions.}
 \emph{Notational conventions.} CwF components by default support de Bruijn
@@
 \begin{nidefinition}{Notion of signature.}
-A QIIT signature at level $j$ is a context in an arbitrary $M : \ToS_{i,j}$
+A QIIT signature at level $i$ is a context in an arbitrary $M : \ToS_{i,j}$
 model. We define the type of such signatures as follows:
 \[
-  \Sig_j :\equiv \forall i.\,(M : \ToS_{i,j})\ra M.Con
+  \Sig_i :\equiv \forall j.\,(M : \ToS_{i,j})\ra M.Con
 \]
 
-Note that $\Sig_j$ is defined as a universe-polymorphic type, so it does not
-have a type internally to cETT. A concrete signature $sig$ at level $j$ has type
-$\forall i.\,(M : \ToS_{i,j})\ra M.Con$, so it builds up a context from the
-components of an arbitrary $\ToS$ model.
+Note that $\Sig_i$ is defined as a universe-polymorphic type, so it does not
+have a type internally to cETT.
 
@@
   & NatSig : \Sig_0 \\
-  & NatSig :\equiv \lambda (M : \ToS_{i,0}). (\emptycon\,\ext\, N : \U \,\ext\, zero : \El\,N \ext suc : \Pii\,(n : N)\,(\El\,N))
+  & NatSig :\equiv \lambda (M : \ToS_{0,j}).\,(\emptycon\,\ext\, N : \U \,\ext\, zero : \El\,N \ext suc : \Pii\,(n : N)\,(\El\,N))
 \end{alignat*}
@@
 
-The $j$ level in $\Sig_j$ specifies the level of the external types appearing in
-a signatures, since the $\Pie$ and $\Piinf$ function spaces have domains in $\Set_j$.
+The $i$ level in $\Sig_i$ specifies the level of the external types appearing in
+a signatures, since the $\Pie$ and $\Piinf$ function spaces have domains in $\Set_i$.
 In the case of naturals numbers, no external types appear, so the level of $NatSig$
@@
 With this, we are able to specify QIITs, and we can also interpret each
-signature in an arbitrary ToS model, just by applying a signature to a model.
-$\Sig_j$ can be viewed as a precursor to a Church-encoding for the theory of
+signature in an arbitrary ToS model, by applying a signature to a model.
+$\Sig_i$ can be viewed as a precursor to a Church-encoding for the theory of
 signatures, but we only need contexts encoded in this way, and not other ToS
@@
 \begin{example}{Infinitary constructors}. The universe $\U$ is closed under
-the $\Piinf$ function type, which makes it possible to write such functions
-in the domains of $\Pii$ types. This allows a signature for $W$-types, given
-arbitrary $S : \Set_i$ and $P : S \ra \Set_i$ types for shapes and positions.
+the $\Piinf$ function type, which makes it possible to write such functions in
+the domains of $\Pii$ types. This allows a signature for trees branching with
+arbitrary small sets. This is a signature at level 1, since we have $\Set_0$ as
+domain type.
 \[
-WSig :\equiv
-\emptycon \ext W : \U \ext sup : \Pie\,S\,(\lambda s.\,\Pii\,(\Piinf\,(P\,s)\,\lambda \_.\,W) \,(\El\,W))
+TreeSig :\equiv
+\emptycon \ext Tree : \U \ext node : \Pie\,\Set_0\,\lambda A.\,\Pii\,(\Piinf\,A\,\lambda \_.\,Tree)\,(\El\,Tree)
 \]
@@
 principles we get are what we want only up to isomorphisms. This issue is
-perhaps not very relevant to those with a purely categorical perspective, but we
-are concerned with eventually implementing QIITs in proof assistants, so we
-prefer if our semantics computes notions strictly.
+perhaps not relevant from a purely categorical perspective, but we are concerned
+with eventually implementing QIITs in proof assistants, so we prefer if our
+semantics computes strictly.
 
-\begin{nidefinition}{flCwF model for ToS.} For $i$ and $j$ levels, there is
-an $M : \ToS_{1+\max(i+1,j),j}$ such that $M.\Con = \flCwF_{\max(i+1,j)}$.
+\begin{nidefinition}{flCwF model for ToS.} We have for each $i$ and $j$ levels
+an $M : \ToS_{i, 1+\max(i,j+1)}$ such that $M.\Con = \flCwF_{\max(i,j+1)}$. This
+model is an extension to the CwF model of signatures in \cite{TODO}, with
+$\Sigma$-types in the semantic CwF, the new type formers in signatures, and the
+treatment of universe levels. We summarize the model below.
+\begin{itemize}
+\item As noted, $M.\Con = \flCwF_{\max(i,j+1)}$. The level $i$ marks the level of
+      all external sets in function domains, and $j$ marks the level of all inductive
+      sets in an algebra. Hence, every algebra, i.e.\ object of the flCwF, has
+      level $\max(i,j+1)$. The bump is only needed for $j$, since algebras only contain
+      \emph{elements} of $T : \Set_j$ types, while inductive sets are themselves elements
+      of $\Set_i$. For example, $NatAlg_i : \Set_{\max(0, i+1)}$.
+\item $\sigma : \Sub\,\Gamma\,\Delta$ substitutions are strictly structure-preserving flCwF
+      functors.
+    \item $A : \Ty\,\Gamma$ types are displayed flCwFs over $\Gamma$. This is an
+      flCwF where every underlying set is indexed over underlying sets in
+      $\Gamma$, and it is an extension of the displayed categories of Ahrens and
+      Lumsdaine\cite{TODO}.  It contains equivalent data to an flCwF slice
+      $\delta : \Sub\,\Delta\,\Gamma$. We prefer it to slices because it yields
+      expected notions of induction strictly.
+\item $t : \Tm\,\Gamma\,A$ terms are sections of the $A$ displayed flCwF. This can be viewed
+      as a dependent functor, which maps structure in $\Gamma$ to corresponding displayed
+      structure in $A$, e.g.\ it maps objects to displayed objects, morphisms to displayed
+      morphisms, and so on.
+\item $\emptycon : \Con$ is the terminal flCwF, whose underlying category is the terminal
+      category with one object.
 
-\todo{flip level indices in ToS, maybe shorten cETT part, add more explanation
-      on role of i,j in ToS(i,j)}
+\item $\Gamma\,\ext\,A$ is the total flCwF of the $A$ displayed flCwF, where every underlying
+      set is given by packing up indexed underlying sets of $A$ in $\Sigma$-s.
+
+\item $\U : \Ty\,\Gamma$ is a displayed flCwF which is constantly (i.e. it does
+      not depend on $\Gamma$) the flCwF of sets in $\Set_j$, whose underlying category
+      is the usual category of sets.
+\item
+  $\Pii\,a\,B : \Ty\,\Gamma$ is a displayed flCwF where objects are
+  dependent functions from the set of objects of $a$ to the set of objects of
+  $B$. It is possible to build such flCwF because the $a : \Tm\,\Gamma\,\U$
+  domain is discrete. A general $\Pii$ type is not possible because of variance
+  issues; this is an extension of there being no $\Pii$ types in the category
+  of categories\cite{elephant}
+\item
+  $\Pie\,A\,B : \Ty\,\Gamma$ is the $A$-indexed direct product of displayed flCwFs
+  given by $B : A \ra \Ty\,\Gamma$.
+
+\end{itemize}
+
+%%   For $i$ and $j$ levels, there is
+%% an $M : \ToS_{i, 1+\max(i,j+1)}$ such that $M.\Con = \flCwF_{\max(i,j+1)}$.
 

### 2019-10-30 draft/paper.tex
stuff

@@
 
-\todo[inline]{Citations}
+ISSUE. We actually need flCwF pseudomorphisms for Sub in infinitary semantics.
+$a : \Tm\,\Gamma\,\U$ is \emph{not} a discrete displayed flCwF, only $\El\,a$ is!
+$\Piinf\,A\,b : \Tm\,\Gamma,\U$ does not preserve (displayed) object in $\Gamma$
+strictly.
+
+\begin{itemize}
+\item New semantics with pseudomorphisms for Sub and Tm. Is it enough to
+      have weak preservation of comprehension, and we get all other preservation
+      for free, see ``gluing for type theory''?
+\item Check term model and eliminators to see if change needed.
+\item Check if the isomorphism for K in flCwF can be made strict, i.e.
+      $\Sub\,\Gamma\,\Delta = \Tm\,\Gamma\,(\Kfam\,\Delta)$.
+\end{itemize}
 
@@
 
-The main motivation of the current paper is to extend QIITs so that it
-accommodates all algebraic notions of type theories which have been used in
-previous works. As a beneficial side effect of fulfilling this goal, infinitary
-QIITs such as Cauchy real numbers will be covered as well.
+The main motivation of the current paper is to extend the notion of QIIT so that
+it accommodates all algebraic notions of type theories which have been used in
+previous works. As a side effect of fulfilling this goal, infinitary QIITs such
+as Cauchy real numbers are covered as well.
 
@@
   theory, but it has not been presented explicitly in previous works about
-  QIITs. In order to handle universe levels gracefully, we work in an
-  extensional type theory extended with cumulative universes.
+  QIITs.
   \item
-  \textbf{Infinitary constructors}. This enables defining Cauchy real numbers
-  and surreal numbers as QIITs. Also, the theory of QIIT signatures is itself
-  large and infinitary, thus it can ``eat itself'', i.e.\ include its own
-  signature and provide its own metatheory.  We use this self-representation to
+  \textbf{Infinitary constructors}. This allows infinitely branching
+  trees. Also, the theory of QIIT signatures is itself large and infinitary,
+  thus now it can ``eat itself'', i.e.\ include its own signature and provide
+  its own metatheory. This was previously not possible in \cite{TODO}, where
+  only finitary QIITs were described. We use this self-representation to
   bootstrap the model theory of signatures, without having to assume any
@@
   of constructors. These have occurred previously in syntaxes of cubical type
-  theories, as boundary conditions.
+  theories, as boundary conditions\cite{TODO}.
   \item
@@
   been used recently to conveniently represent type-theoretic universes, for
-  example Russell-style universes or cumulative hierarchies.
+  example Russell-style universes\cite{TODO} or cumulative hierarchies\cite{TODO}.
 \end{enumerate}
@@
 different level than the target algebra. For a simple example, consider natural
-number algebras at a given level $i$, given as the $\Sigma$-type $NatAlg_i
-:\equiv(Nat : \Set_i)\times Nat\times(Nat \ra Nat)$. The initial such algebra is
-the set of natural numbers, which is at level $0$, but in type theory we often
-want to eliminate into larger $\Set_i$, for example when computing a
-$Nat$-indexed family of types. To support convenient reasoning about levels in
-this paper, we need two features: cumulativity and a way to quantify over finite
-levels.
+number algebras at level $i$, given as the $\Sigma$-type $NatAlg_i :\equiv(Nat :
+\Set_i)\times Nat\times(Nat \ra Nat)$. The initial such algebra is the set of
+natural numbers, which is at level $0$, but in type theory we often want to
+eliminate into larger $\Set_i$, for example when computing a $Nat$-indexed
+family of types. To support convenient reasoning about levels in this paper, we
+need two features: cumulativity and a way to quantify over finite levels.
 
@@
 
-We need cumulativity to reduce administrative overhead. In the most basic form,
+We need cumulativity to reduce bureaucratic overhead. In its most basic form,
 cumulativity requires that whenever $A : \Set_i$, and $i < j$ then also $A :
@@
 have cumulativity for algebras in general. For example, we want to have $\gamma
-: NatAlg_j$ whenever $\gamma : NatAlg_i$ and $i < j$.
+: NatAlg_j$ whenever $\gamma : NatAlg_i$ and $i < j$. We have found that
+non-cumulative algebras induce excessive ``lifting'' noise, which we would like
+to avoid.
 
@@
 From now on, we refer to this theory as cETT (cumulative extensional type
-theory). When working in cETT, we use the following notation. We
-have Russell-style universes $\Set_i$ indexed by natural numbers, dependent
-functions as $(x : A)\ra B$, and dependent pairs as $(x : A)\times B$ with
-projections $\proj1$ and $\proj2$. We also use a field projection notation
-for iterated pairs. For example, for $t : (A : \Set_i) \times (B : \Set_i)
-\times (f : A \ra B)$, we use $t.B$ to denote the projection of the second
-component. Sometimes we omit the $t.$ prefix if it is clear from context.
+theory). When working in cETT, we use the following notation. We have
+Russell-style universes $\Set_i$ indexed by natural numbers, dependent functions
+as $(x : A)\ra B$, and dependent pairs as $(x : A)\times B$ with projections
+$\proj1$ and $\proj2$. We also use subscripts as a field projection notation for
+iterated pairs. For example, for $t : (A : \Set_i) \times (B : \Set_i) \times (f
+: A \ra B)$, we use $B_t$ to denote the projection of the second
+component. Sometimes we omit the subscript if it is clear from context.
 
 We write propositional equality as $t = u$, with $\refl_t$ for reflexivity. We
-have equality reflection for $\blank=\blank$. The unit type is $\top$, with
-inhabitant $\tt$.
+have equality reflection and uniqueness of identity proofs (UIP) for
+$\blank=\blank$. The unit type is $\top$, with inhabitant $\tt$.
 
@@
   \inferrule*[lab=universe formation]
-             {\\}
-             {\Gamma \vdash \Set_i : \Set_{i+j+1}}
+             {i < j}
+             {\Gamma \vdash \Set_i : \Set_{j}}
 
@@
 In Figure \ref{lifting}, we include an excerpt of cETT's rules. We change
-presentation slightly from Sterling. We lift from $i$ to $i+j+1$ levels instead
-of assuming $i < j$ and lifting to $j$, and we use a $\U_{\max(i,j)}$ return
-type in functions and pairs, instead of having both $A$, $B$ types in the same
+presentation slightly from Sterling: we use a $\U_{\max(i,j)}$ return type in
+functions and pairs, instead of having both $A$ and $B$ types in the same
 $\Set_i$. These changed rules are all derivable from the original ones. In
@@
 
-\begin{nidefinition}{Cumulative subtyping.}
+\begin{nidefinition}[Cumulative subtyping]
   For $A : \Set_i$ and $B : \Set_{i+j+1}$, we define $A \prec B : \Set_{i+j+1}$
@@
 polymorphic cETT term is understood as a $\mathbb{N}$-indexed family of
-terms. We use the $\forall$ symbol for this, e.g.\ as $t : \forall\,i.\,\Set_i \ra
-\Set_i$. This is sufficient in this paper, because we do not need to internalize
-level-polymorphic constructions in cETT.
+terms. We reuse the notation of cETT functions for level-polymorphism, e.g.\
+as in the following function:
+\[
+\lambda i.\,\Set_i\ra\Set_i : (i : \mathbb{N})\ra \Set_{i+1}
+\]
+In this paper, we do not need to internalize level-polymorphic constructions,
+so this setup is sufficient.
 
@@
 
-\begin{nidefinition}{Notion of model for ToS.}\label{def:tos}
+\begin{nidefinition}[Notion of model for the theory of signatures]\label{def:tos}
 For levels $i$ and $j$, there is a cETT type $\ToS_{i,j} : \Set_{\max(i,\,j)+1}$, whose elements are
@@
 \item
-  A category with families (CwF), where all four underlying sets (of
-  objects, morphisms, types and terms) are in $\Set_j$. Following notation in
+  A category with families (CwF), where all four underlying sets (of objects,
+  morphisms, types and terms) are in $\Set_j$. Following notation in
   \cite{TODO}, we denote these respectively as $\Con : \Set_j$, $\Sub : \Con \ra
@@
   \Ty\,\Gamma \ra \Set_j$. We denote the empty context as $\emptycon : \Con$,
-  context extension as $\blank\ext\blank : (\Gamma : \Con)\ra \Ty\,\Gamma \ra \Con$.
-  Substitution on types and terms is written as $\blank[\blank]$.
+  context extension as $\blank\ext\blank : (\Gamma : \Con)\ra \Ty\,\Gamma \ra
+  \Con$.  Substitution on types and terms is written as $\blank[\blank]$, with
+  $\id$ and $\blank\circ\blank$ for identity and composition, and $\blank,\blank
+  : (\sigma : \Sub\,\Gamma\,\Delta)\ra \Tm\,\Gamma\,A[\sigma]\ra
+  \Sub\,\Gamma\,(\Delta\ext A)$ for substitution extension.
 \item
@@
   \Ty\,(\Gamma\,\ext\,\El\,a) \ra \Ty\,\Gamma$ and application as $\appi :
-  \Tm\,\Gamma\,(\Pii\,a\,B)\ra \Tm\,(\Gamma\,\ext\,\El\,a)\,B$. We do not
-  support $\lami$ here and in other function types, as they are not essential
-  for QIIT signatures.
+  \Tm\,\Gamma\,(\Pii\,a\,B)\ra \Tm\,(\Gamma\,\ext\,\El\,a)\,B$ and its
+  inverse $\lami$.
 \item
   External function space with $\Set_i$ domain: $\Pie : (A : \Set_i)\ra(A \ra
-  \Ty\,\Gamma)\ra \Ty\,\Gamma$, with $\appe : \Tm\,\Gamma\,(\Pie\,A\,B)\ra(x :
-  A)\ra \Tm\,\Gamma\,(B\,x)$.
+  \Ty\,\Gamma)\ra \Ty\,\Gamma$, with $\appe : \Tm\,\Gamma\,(\Pie\,A\,B)\ra((x :
+  A)\ra \Tm\,\Gamma\,(B\,x))$ and its inverse $\lame$.
 \item
@@
   \Tm\,\Gamma\,\U)\ra \Tm\,\Gamma\,\U$, $\appinf :
-  \Tm\,\Gamma\,(\El\,(\Pie\,A\,b))\ra(x : A)\ra \Tm\,\Gamma\,(\El\,(b\,x))$.
+  \Tm\,\Gamma\,(\El\,(\Pie\,A\,b))\ra((x : A)\ra \Tm\,\Gamma\,(\El\,(b\,x)))$
+  and $\laminf$.
 \item
@@
   $\Refl : (T : \Tm\,\Gamma\,(\El\,a))\ra \Id\,a\,t\,t$, equality reflection
-  and uniqueness of identity proofs.
+  and UIP.
 \item
-  An identity type $\IdU : \Tm\,\Gamma\,\U \ra \Tm\,\Gamma\,\U \ra \Ty\,\Gamma$, with
-  evident $\ReflU$, equality reflection and uniqueness of identity proofs.
+  An identity type $\IdU : \Tm\,\Gamma\,\U \ra \Tm\,\Gamma\,\U \ra \Ty\,\Gamma$,
+  with reflexivity as $\ReflU$, equality reflection and UIP.
 \end{enumerate}
@@
       \Tm\,\Gamma\,(\Pii\,a\,B)\ra
-      (u : \Tm\,\Gamma\,(\El\,a)) \ra \Tm\,\Gamma\,(B[id, u])\\
-  & t\appitt u && :\equiv (\appi\,t)[id, u]
+      (u : \Tm\,\Gamma\,(\El\,a)) \ra \Tm\,\Gamma\,(B[\id, u])\\
+  & t\appitt u && :\equiv (\appi\,t)[\id, u]
 \end{alignat*}
 
-\begin{nidefinition}{Notion of signature.}
+\begin{nidefinition}[Notion of signature]
 A QIIT signature at level $i$ is a context in an arbitrary $M : \ToS_{i,j}$
@@
 \[
-  \Sig_i :\equiv \forall j.\,(M : \ToS_{i,j})\ra M.Con
+  \Sig_i :\equiv (j : \mathbb{N})(M : \ToS_{i,j})\ra M.Con
 \]
 
-Note that $\Sig_i$ is defined as a universe-polymorphic type, so it does not
-have a type internally to cETT.
+Here, $i$ refers to the level of external types appearing in the signature, in
+the domains of $\Pie$ and $\Piinf$ functions, while the quantified $j$ level is
+required to allow interpreting a signature in arbitrary-sized ToS models. Note
+that $\Sig_i$ is universe-polymorphic, so it does not have a type internally to
+cETT.
 
-\begin{example}
-The signature for natural
-numbers is the following, where $\emptycon$, $\blank\ext\blank$, $\U$, $\El$ and
-$\Pii$ all refer to components of the $\lambda$-bound $M$:
+\begin{example}{
+    Signature for natural numbers. Here, no external types appear, so the level
+    of $NatSig$ can be chosen as $0$.}
 \begin{alignat*}{3}
   & NatSig : \Sig_0 \\
-  & NatSig :\equiv \lambda (M : \ToS_{0,j}).\,(\emptycon\,\ext\, N : \U \,\ext\, zero : \El\,N \ext suc : \Pii\,(n : N)\,(\El\,N))
+  & NatSig :\equiv \lambda (j :\mathbb{N})(M : \ToS_{0,j}).\\
+  & \hspace{2em}(\emptycon_M\,\ext_M\, (N : \U_M) \,\ext_M\, (zero : \El_M\,N) \ext_M (suc : \Pii_M\,N\,(\El_M\,N)))
 \end{alignat*}
 \end{example}
-
-The $i$ level in $\Sig_i$ specifies the level of the external types appearing in
-a signatures, since the $\Pie$ and $\Piinf$ function spaces have domains in $\Set_i$.
-In the case of naturals numbers, no external types appear, so the level of $NatSig$
-can be chosen as $0$.
 
@@
 arbitrary small sets. This is a signature at level 1, since we have $\Set_0$ as
-domain type.
+a $\Pie$ domain type.
 \[
 TreeSig :\equiv
-\emptycon \ext Tree : \U \ext node : \Pie\,\Set_0\,\lambda A.\,\Pii\,(\Piinf\,A\,\lambda \_.\,Tree)\,(\El\,Tree)
+\emptycon \ext (Tree : \U) \ext (node : \Pie\,\Set_0\,\lambda A.\,\Pii\,(\Piinf\,A\,\lambda \_.\,Tree)\,(\El\,Tree))
 \]
@@
 \[
-RecEqSig :\equiv \emptycon \ext A : \U \ext a : \El\,A \ext f : \Pii\,(x : A)\,(\Pii\,(\Id\,A\,x\,a)\,(\El\,A))
+RecEqSig :\equiv \emptycon \ext (A : \U) \ext (a : \El\,A) \ext (f : \Pii\,(x : A)\,(\Pii\,(\Id\,A\,x\,a)\,(\El\,A)))
 \]
-There are more complicated and interesting examples, e.g.\ boundary conditions in
-various cubical type theories\cite{TODO}. Note that our $\Id$ allows iterated
-equations as well, but these are all made trivial in the semantics, where we
-assume uniqueness of identity proofs.
+More interesting (and complicated) examples are boundary conditions in various
+cubical type theories\cite{TODO}. Note that our $\Id$ allows iterated equations
+as well, but these are all made trivial in the semantics, where we assume UIP.
 \end{example}
@@
 
-\begin{nidefinition}{Finite limit CwFs.}
+\begin{nidefinition}[Finite limit CwFs]\label{def:flCwF}
 For a level $i$, there is a cETT type $\flCwF_i : \Set_{i+1}$, which is an
@@
   \item Identity type $\Id : (A : \Ty\,\Gamma)\ra \Tm\,\Gamma\,A\ra
-    \Tm\,\Gamma\,A\ra \Ty\,\Gamma$, with equality reflection and uniqueness
-    of identity proofs.
+    \Tm\,\Gamma\,A\ra \Ty\,\Gamma$, with equality reflection and UIP.
   \item Constant families. This includes a type former $\Kfam : \Con \ra
@@
 
-\begin{nidefinition} Notion of induction internally to $C : \flCwF_i$:
+\begin{nidefinition}[Notion of induction in an flCwF]
+Given $C : \flCwF_i$, we have the following predicate on objects:
 \begin{alignat*}{3}
-  & Inductive : C.\Con \ra \Set_i \\
-  & Inductive\,\Gamma :\equiv (A : C.\Ty\,\Gamma)\ra C.\Tm\,\Gamma\,A
+  & Inductive : \Con_C \ra \Set_i \\
+  & Inductive\,\Gamma :\equiv (A : \Ty_C\,\Gamma)\ra \Tm_C\,\Gamma\,A
 \end{alignat*}
@@
 
-\begin{theorem}{Equivalence of initiality and induction.}
-An object $\Gamma : C.\Con$ supports induction if and only if it is
+\begin{theorem}[Equivalence of initiality and induction]
+An object $\Gamma : \Con_C$ supports induction if and only if it is
 initial. Moreover, induction and initiality are both proof irrelevant
-predicates on objects.
+predicates.
 \end{theorem}
@@
 
-So, why do we use flCwF instead of lex categories? After all, the former
-structure has more components and has more immediate complexity than the
-latter. The reason is that flCwFs allow us to define induction in strictly the
-same way as one would write it down in type theory, since we have $\Ty$ and
-$\Tm$ for a primitive notion of dependent objects and morphisms. In contrast,
-dependent objects in lex categories is a derived notion, and the induction
-principles we get are what we want only up to isomorphisms. This issue is
-perhaps not relevant from a purely categorical perspective, but we are concerned
-with eventually implementing QIITs in proof assistants, so we prefer if our
-semantics computes strictly.
+So, why do we use flCwFs instead of lex categories? After all, the former has
+more components and has more immediate complexity than the latter. The reason is
+that flCwFs allow us to define induction in strictly the same way as one would
+write it down in type theory, since we have $\Ty$ and $\Tm$ for a primitive
+notion of dependent objects and morphisms. In contrast, dependent objects in lex
+categories is a derived notion, and the induction principles we get are what we
+want only up to isomorphisms. This issue is perhaps not relevant from a purely
+categorical perspective, but we are concerned with eventually implementing QIITs
+in proof assistants, so we prefer if our semantics computes strictly.
 
-\begin{nidefinition}{flCwF model for ToS.} We have for each $i$ and $j$ levels
-an $M : \ToS_{i, 1+\max(i,j+1)}$ such that $M.\Con = \flCwF_{\max(i,j+1)}$. This
-model is an extension to the CwF model of signatures in \cite{TODO}, with
-$\Sigma$-types in the semantic CwF, the new type formers in signatures, and the
-treatment of universe levels. We summarize the model below.
-\begin{itemize}
-\item As noted, $M.\Con = \flCwF_{\max(i,j+1)}$. The level $i$ marks the level of
-      all external sets in function domains, and $j$ marks the level of all inductive
-      sets in an algebra. Hence, every algebra, i.e.\ object of the flCwF, has
-      level $\max(i,j+1)$. The bump is only needed for $j$, since algebras only contain
-      \emph{elements} of $T : \Set_j$ types, while inductive sets are themselves elements
-      of $\Set_i$. For example, $NatAlg_i : \Set_{\max(0, i+1)}$.
-\item $\sigma : \Sub\,\Gamma\,\Delta$ substitutions are strictly structure-preserving flCwF
-      functors.
-    \item $A : \Ty\,\Gamma$ types are displayed flCwFs over $\Gamma$. This is an
+\begin{nidefinition}[flCwF model for ToS] We have for each $i$ and $j$ levels
+an $M : \ToS_{i, 1+\max(i,j+1)}$ such that $\Con_M = \flCwF_{\max(i,j+1)}$. This
+model is an extension of the CwF model of signatures in \cite{TODO}. The new
+additions are: $\Sigma$-types in the semantic CwF, the new type formers in
+signatures, and the treatment of universe levels. We summarize the model below.
+
+As noted, $\Con$ is given as $\flCwF_{\max(i,j+1)}$. The level $i$ marks the
+level of all external sets in function domains, and $j$ marks the level of all
+internal sorts in an algebra. Hence, every algebra, i.e.\ object of the flCwF,
+has level $\max(i,j+1)$. The bump is only needed for $j$, since algebras only
+contain elements of $T : \Set_j$ types, while inductive sets are themselves
+elements of $\Set_i$. For example, $NatAlg_i : \Set_{\max(0, i+1)}$.
+
+$\sigma : \Sub\,\Gamma\,\Delta$ substitutions are functors preserving all flCwF
+structure strictly.
+
+$A : \Ty\,\Gamma$ types are displayed flCwFs over $\Gamma$. This is an
 flCwF where every underlying set is indexed over underlying sets in
@@
 expected notions of induction strictly.
-\item $t : \Tm\,\Gamma\,A$ terms are sections of the $A$ displayed flCwF. This can be viewed
-      as a dependent functor, which maps structure in $\Gamma$ to corresponding displayed
-      structure in $A$, e.g.\ it maps objects to displayed objects, morphisms to displayed
-      morphisms, and so on.
-\item $\emptycon : \Con$ is the terminal flCwF, whose underlying category is the terminal
-      category with one object.
 
-\item $\Gamma\,\ext\,A$ is the total flCwF of the $A$ displayed flCwF, where every underlying
-      set is given by packing up indexed underlying sets of $A$ in $\Sigma$-s.
+$t : \Tm\,\Gamma\,A$ terms are sections of the displayed flCwF $A$. This can
+be viewed as a dependent functor, which maps structure in $\Gamma$ to
+corresponding displayed structure in $A$, e.g.\ it maps objects to displayed
+objects, morphisms to displayed morphisms, and so on.
 
-\item $\U : \Ty\,\Gamma$ is a displayed flCwF which is constantly (i.e. it does
-      not depend on $\Gamma$) the flCwF of sets in $\Set_j$, whose underlying category
-      is the usual category of sets.
-\item
-  $\Pii\,a\,B : \Ty\,\Gamma$ is a displayed flCwF where objects are
-  dependent functions from the set of objects of $a$ to the set of objects of
-  $B$. It is possible to build such flCwF because the $a : \Tm\,\Gamma\,\U$
-  domain is discrete. A general $\Pii$ type is not possible because of variance
-  issues; this is an extension of there being no $\Pii$ types in the category
-  of categories\cite{elephant}
-\item
-  $\Pie\,A\,B : \Ty\,\Gamma$ is the $A$-indexed direct product of displayed flCwFs
-  given by $B : A \ra \Ty\,\Gamma$.
+$\emptycon : \Con$ is the terminal flCwF, whose underlying category is the
+terminal category with one object.
 
-\end{itemize}
+$\Gamma\,\ext\,A$ is the total flCwF of the $A$ displayed flCwF, where every
+underlying set is given by packing up indexed underlying sets of $A$ in
+$\Sigma$-s.
+
+$\U : \Ty\,\Gamma$ is a displayed flCwF which is constantly (i.e. it does not
+depend on $\Gamma$) the flCwF of sets in $\Set_j$, whose underlying category
+is the usual category of sets, and which has evident finite limit structure
+($\Sg$, $\Id$, $\K$ from Definition \ref{def:flCwF}). Hence, elements of
+$\Tm\,\Gamma\,\U$ are functors from $\Gamma$ to \textbf{Set}. $\El :
+\Tm\,\Gamma\,\U\ra\Ty\,\Gamma$ interprets such functors as discrete displayed
+flCwFs, where all morphisms are identities.
+
+$\Pii\,a\,B : \Ty\,\Gamma$ is a displayed flCwF where objects are dependent
+functions from the set of objects of $a$ to the set of objects of $B$, and
+morphisms are pointwise morphisms. This is possible to build because the $a :
+\Tm\,\Gamma\,\U$ domain is discrete. A general $\Pii$ type is not possible
+because of variance issues; this follows from the fact that the category of
+categories does not support $\Pii$ types\cite{elephant}. Finite limit structure
+is given pointwise, and is inherited from the domain $B$.
+
+$\Pie\,A\,B : \Ty\,\Gamma$ is the $A$-indexed direct product of the family of
+displayed flCwFs given by $B : A \ra \Ty\,\Gamma$.
+
+$\Piinf\,A\,b : \Tm\,\Gamma\,\U$ is likewise the $A$-indexed direct product of
+the family of discrete displayed flCwFs given by $b : A \ra \Tm\,\Gamma\,\U$.
+
+$\Id\,a\,t\,u : \Tm\,\Gamma\,\U$ maps every object $\gamma$ of $\Gamma$ to the
+set $t\,\gamma = u\,\gamma$, and every morphism $\sigma :
+\Sub_\Gamma\,\gamma\,\gamma'$ to a function with type $t\,\gamma = u\,\gamma
+\ra t\,\gamma' = u\,\gamma'$, which can be given by the functoriality of $t$
+and $u$. In short, $\Id\,a\,t\,u$ expresses pointwise equality of $t$ and $u$
+dependent functors.
+
+$\IdU\,a\,b : \Ty\,\Gamma$ is a displayed flCwF where the set of objects over
+$\gamma : \Con_\Gamma$ is $a\,\gamma = b\,\gamma$, that is, a strict equation
+of types in $\Set_j$. Displayed morphisms are trivial (i.e.\ constantly
+$\top$) because of UIP in cETT.
+
+%% \begin{itemize}
+%% \item
+%%   As noted, $\Con = \flCwF_{\max(i,j+1)}$. The level $i$ marks the level of
+%%   all external sets in function domains, and $j$ marks the level of all
+%%   inductive sets in an algebra. Hence, every algebra, i.e.\ object of the flCwF,
+%%   has level $\max(i,j+1)$. The bump is only needed for $j$, since algebras only
+%%   contain elements of $T : \Set_j$ types, while inductive sets are themselves
+%%   elements of $\Set_i$. For example, $NatAlg_i : \Set_{\max(0, i+1)}$.
+%% \item
+%%   $\sigma : \Sub\,\Gamma\,\Delta$ substitutions are functors preserving all flCwF
+%%   structure strictly.
+%% \item
+%%   $A : \Ty\,\Gamma$ types are displayed flCwFs over $\Gamma$. This is an
+%%   flCwF where every underlying set is indexed over underlying sets in
+%%   $\Gamma$, and it is an extension of the displayed categories of Ahrens and
+%%   Lumsdaine\cite{TODO}.  It contains equivalent data to an flCwF slice
+%%   $\delta : \Sub\,\Delta\,\Gamma$. We prefer it to slices because it yields
+%%   expected notions of induction strictly.
+%% \item
+%%   $t : \Tm\,\Gamma\,A$ terms are sections of the $A$ displayed flCwF. This can
+%%   be viewed as a dependent functor, which maps structure in $\Gamma$ to
+%%   corresponding displayed structure in $A$, e.g.\ it maps objects to displayed
+%%   objects, morphisms to displayed morphisms, and so on.
+%% \item
+%%   $\emptycon : \Con$ is the terminal flCwF, whose underlying category is the
+%%   terminal category with one object.
+%% \item
+%%   $\Gamma\,\ext\,A$ is the total flCwF of the $A$ displayed flCwF, where every
+%%   underlying set is given by packing up indexed underlying sets of $A$ in
+%%   $\Sigma$-s.
+%% \item
+%%   $\U : \Ty\,\Gamma$ is a displayed flCwF which is constantly (i.e. it does not
+%%   depend on $\Gamma$) the flCwF of sets in $\Set_j$, whose underlying category
+%%   is the usual category of sets, and which has evident finite limit structure
+%%   ($\Sg$, $\Id$, $\K$ from Definition \ref{def:flCwF}). Hence, elements of
+%%   $\Tm\,\Gamma\,\U$ are functors from $\Gamma$ to \textbf{Set}. $\El :
+%%   \Tm\,\Gamma\,\U\ra\Ty\,\Gamma$ interprets such functors as discrete displayed
+%%   flCwFs, where all morphisms are identities.
+%% \item
+%%   $\Pii\,a\,B : \Ty\,\Gamma$ is a displayed flCwF where objects are dependent
+%%   functions from the set of objects of $a$ to the set of objects of $B$, and
+%%   morphisms are pointwise morphisms. This is possible to build because the $a :
+%%   \Tm\,\Gamma\,\U$ domain is discrete. A general $\Pii$ type is not possible
+%%   because of variance issues; this follows from the fact that the category of
+%%   categories does not support $\Pii$ types\cite{elephant}.
+%% \item
+%%   $\Pie\,A\,B : \Ty\,\Gamma$ is the $A$-indexed direct product of the family of
+%%   displayed flCwFs given by $B : A \ra \Ty\,\Gamma$.
+%% \item
+%%   $\Piinf\,A\,b : \Tm\,\Gamma\,\U$ is likewise the $A$-indexed direct product of
+%%   the family of discrete displayed flCwFs given by $b : A \ra \Tm\,\Gamma\,\U$.
+%% \item
+%%   $\Id\,a\,t\,u : \Tm\,\Gamma\,\U$ maps every object $\gamma$ of $\Gamma$ to the
+%%   set $t\,\gamma = u\,\gamma$, and every morphism $\sigma :
+%%   \Sub_\Gamma\,\gamma\,\gamma'$ to a function with type $t\,\gamma = u\,\gamma
+%%   \ra t\,\gamma' = u\,\gamma'$, which can be given by the functoriality of $t$
+%%   and $u$. In short, $\Id\,a\,t\,u$ expresses pointwise equality of $t$ and $u$
+%%   dependent functors.
+%% \item
+%%   $\IdU\,a\,b : \Ty\,\Gamma$ is a displayed flCwF where the set of objects over
+%%   $\gamma : \Con_\Gamma$ is $a\,\gamma = b\,\gamma$, that is, a strict equation
+%%   of types in $\Set_j$. Displayed morphisms are trivial (i.e.\ constantly
+%%   $\top$) because of UIP in cETT.
+%% \end{itemize}
+
+\begin{example}{Algebras for pointed sets.}
+
+\end{example}
 

### 2019-12-26 draft/old.tex
work on paper

--- 2019-10-30 draft/paper.tex
@@
 \newcommand{\laminf}{\mathsf{la\tilde{m}}}
-
 \newcommand{\appitt}{\mathop{{\scriptstyle @}}}
-
 \newcommand{\Refl}{\mathsf{Refl}}
-
 \newcommand{\IdU}{\mathsf{IdU}}
@@
 \newcommand{\Sig}{\mathsf{Sig}}
-
 \newcommand{\Sg}{\mathsf{\Sigma}}
@@
 
-\title{Generalizations of Quotient Inductive-Inductive Types\thanks{This work
+\title{Large and Infinitary Quotient Inductive-Inductive Types\thanks{This work
     was supported by EFOP-3.6.3-VEKOP-16-2017-00002 grant and COST
@@
 numbers. We develop new metatheory for large QIITs, large elimination, recursive
-equations, infinitary constructors and equations between sorts.
-%% First, we develop tools which allow precise and convenient handling
-%% of universe levels, working internally in an extensional type theory with
-%% cumulative universes.
-As in prior work, we describe QIITs using a type theory where each context
-represents a QIIT signature. However, in our case the theory of signatures can
-also describe its own signature. We use self-description to bootstrap a model
-theory for the theory of signatures without using preterms or assuming a
-pre-existing internal syntax for a type theory. We give initial algebra
-semantics for described QIITs, and we show the equivalence of initiality and
-induction. We present two extensions of a previous term model construction. The
-first one constructs all large infinitary QIITs without sort equations from the
-QIIT of the theory of signatures. The second one constructs all large finitary
-QIITs with sort equations, from the same syntax. This separation is required
-because handling infinitary constructors in the term model requires showing a
-strong form of invariance under algebra isomorphisms, which is violated by sort
-equations.
-
+equations and infinitary constructors. As in prior work, we describe QIITs using
+a type theory where each context represents a QIIT signature. However, in our
+case the theory of signatures can also describe its own signature, modulo
+universe sizes. We bootstrap the model theory of signatures using
+self-description and a Church-coded notion of signature, without using
+complicated raw syntax or assuming an existing internal QIIT of signatures. We
+give semantics to described QIITs by modeling each signature as a finitely
+complete CwF (category with families) of algebras. Compared to the case of
+finitary QIITs, we additionally need to show invariance under algebra
+isomorphisms in the semantics. We do this by modeling signature types as
+isofibrations. We show by a term model construction that every QIIT is
+constructible from the syntax of the theory of signatures.
 \end{abstract}
@@
 
-ISSUE. We actually need flCwF pseudomorphisms for Sub in infinitary semantics.
-$a : \Tm\,\Gamma\,\U$ is \emph{not} a discrete displayed flCwF, only $\El\,a$ is!
-$\Piinf\,A\,b : \Tm\,\Gamma,\U$ does not preserve (displayed) object in $\Gamma$
-strictly.
+%% ISSUE. We actually need flCwF pseudomorphisms for Sub in infinitary semantics.
+%% $a : \Tm\,\Gamma\,\U$ is \emph{not} a discrete displayed flCwF, only $\El\,a$ is!
+%% $\Piinf\,A\,b : \Tm\,\Gamma,\U$ does not preserve (displayed) object in $\Gamma$
+%% strictly.
 
-\begin{itemize}
-\item New semantics with pseudomorphisms for Sub and Tm. Is it enough to
-      have weak preservation of comprehension, and we get all other preservation
-      for free, see ``gluing for type theory''?
-\item Check term model and eliminators to see if change needed.
-\item Check if the isomorphism for K in flCwF can be made strict, i.e.
-      $\Sub\,\Gamma\,\Delta = \Tm\,\Gamma\,(\Kfam\,\Delta)$.
-\end{itemize}
+%% \begin{itemize}
+%% \item New semantics with pseudomorphisms for Sub and Tm. Is it enough to
+%%       have weak preservation of comprehension, and we get all other preservation
+%%       for free, see ``gluing for type theory''?
+%% \item Check term model and eliminators to see if change needed.
+%% \item Check if the isomorphism for K in flCwF can be made strict, i.e.
+%%       $\Sub\,\Gamma\,\Delta = \Tm\,\Gamma\,(\Kfam\,\Delta)$.
+%% \end{itemize}
 
@@
 The main motivation of the current paper is to extend the notion of QIIT so that
-it accommodates all algebraic notions of type theories which have been used in
-previous works. As a side effect of fulfilling this goal, infinitary QIITs such
-as Cauchy real numbers are covered as well.
+it accommodates more algebraic notions of type theories than previously, and
+more aspects of their practical metatheory. As a side effect of fulfilling this
+goal, infinitary QIITs such as Cauchy real numbers are covered as well.
 
@@
   \textbf{Large constructors, large elimination} and models at different
-  universe levels. This feature is routinely used in the metatheory of type
+  universe levels. Large elimination is routinely used in the metatheory of type
   theory, but it has not been presented explicitly in previous works about
@@
   thus now it can ``eat itself'', i.e.\ include its own signature and provide
-  its own metatheory. This was previously not possible in \cite{TODO}, where
-  only finitary QIITs were described. We use this self-representation to
-  bootstrap the model theory of signatures, without having to assume any
-  pre-existing internal syntax.
+  its own metatheory. This was not possible previously in
+  \cite{kaposi2019constructing}, where only finitary QIITs were described. We
+  exploit self-representation to bootstrap the model theory of signatures,
+  without having to assume any pre-existing internal syntax.
   \item
@@
   theories, as boundary conditions\cite{TODO}.
-  \item
-  \textbf{Sort equations}, or equations between type constructors. They have
-  been used recently to conveniently represent type-theoretic universes, for
-  example Russell-style universes\cite{TODO} or cumulative hierarchies\cite{TODO}.
 \end{enumerate}

### 2019-12-26 draft/paper.tex
work on paper

@@
-\documentclass{llncs}
+%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
+\documentclass[sigplan,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+%% For double-blind review submission, w/ CCS and ACM Reference
+%\documentclass[sigplan,review,anonymous]{acmart}\settopmatter{printfolios=true}
+%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
+%\documentclass[sigplan,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+%% For single-blind review submission, w/ CCS and ACM Reference
+%\documentclass[sigplan,review]{acmart}\settopmatter{printfolios=true}
+%% For final camera-ready submission, w/ required CCS and ACM Reference
+%\documentclass[sigplan]{acmart}\settopmatter{}
 
+%% Conference information
+%% Supplied to authors by publisher for camera-ready submission;
+%% use defaults for review submission.
+\acmConference[PL'18]{ACM SIGPLAN Conference on Programming Languages}{January 01--03, 2018}{New York, NY, USA}
+\acmYear{2018}
+\acmISBN{} % \acmISBN{978-x-xxxx-xxxx-x/YY/MM}
+\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
+\startPage{1}
+
+%% Copyright information
+%% Supplied to authors (based on authors' rights management selection;
+%% see authors.acm.org) by publisher for camera-ready submission;
+%% use 'none' for review submission.
+\setcopyright{none}
+%\setcopyright{acmcopyright}
+%\setcopyright{acmlicensed}
+%\setcopyright{rightsretained}
+%\copyrightyear{2018}           %% If different from \acmYear
+
+%% Bibliography style
+\bibliographystyle{ACM-Reference-Format}
+%% Citation style
+%\citestyle{acmauthoryear}  %% For author/year citations
+%\citestyle{acmnumeric}     %% For numeric citations
+%\setcitestyle{nosort}      %% With 'acmnumeric', to disable automatic
+                            %% sorting of references within a single citation;
+                            %% e.g., \cite{Smith99,Carpenter05,Baker12}
+                            %% rendered as [14,5,2] rather than [2,5,14].
+%\setcitesyle{nocompress}   %% With 'acmnumeric', to disable automatic
+                            %% compression of sequential references within a
+                            %% single citation;
+                            %% e.g., \cite{Baker12,Baker14,Baker16}
+                            %% rendered as [2,3,4] rather than [2-4].
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Note: Authors migrating a paper from traditional SIGPLAN
+%% proceedings format to PACMPL format must update the
+%% '\documentclass' and topmatter commands above; see
+%% 'acmart-pacmpl-template.tex'.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%% Some recommended packages.
+\usepackage{booktabs}   %% For formal tables:
+                        %% http://ctan.org/pkg/booktabs
+\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
+                        %% http://ctan.org/pkg/subcaption
+
+%% ------------------------------------------------------------
 \usepackage[utf8]{inputenc}
@@
 \usepackage{amsmath}
+\usepackage{amsthm}
 \usepackage{amssymb}
@@
 \usepackage{scalerel}
-
-%% Definition env with non-italic body text
-\spnewtheorem{nidefinition}[definition]{Definition}{\bfseries}{\textnormal}
-
-%% abbrevs
-%% ------------------------------------------------------------
 
@@
 \newcommand{\emptycon}{\scaleobj{.75}\bullet}
-\newcommand{\Pii}{\mathsf{\Pi}}
+\newcommand{\Pii}{\Pi}
 \newcommand{\appi}{\mathsf{app}}
 \newcommand{\lami}{\mathsf{lam}}
-\newcommand{\Pie}{\mathsf{\hat{\Pi}}}
+\newcommand{\Pie}{\hat{\Pi}}
 \newcommand{\appe}{\mathsf{ap\hat{p}}}
 \newcommand{\lame}{\mathsf{la\hat{m}}}
-\newcommand{\Piinf}{\mathsf{\tilde{\Pi}}}
+\newcommand{\Piinf}{\tilde{\Pi}}
 \newcommand{\appinf}{\mathsf{ap\tilde{p}}}
 \newcommand{\laminf}{\mathsf{la\tilde{m}}}
-
 \newcommand{\appitt}{\mathop{{\scriptstyle @}}}
-
 \newcommand{\Refl}{\mathsf{Refl}}
-
 \newcommand{\IdU}{\mathsf{IdU}}
@@
 \newcommand{\Sig}{\mathsf{Sig}}
-
-\newcommand{\Sg}{\mathsf{\Sigma}}
+\newcommand{\Sg}{\Sigma}
 \newcommand{\flCwF}{\mathsf{flCwF}}
 \newcommand{\Kfam}{\mathsf{K}}
+\newcommand{\p}{\mathsf{p}}
+\newcommand{\q}{\mathsf{q}}
 
+\newtheorem{abbreviation}{Abbreviation}
 %% ------------------------------------------------------------
@@
 
-\title{Generalizations of Quotient Inductive-Inductive Types\thanks{This work
-    was supported by EFOP-3.6.3-VEKOP-16-2017-00002 grant and COST
-    Action EUTypes CA15123.}}
+%% Title information
+\title{Large and Infinitary Quotient Inductive-Inductive Types} %% [Short Title] is optional;
+                                        %% when present, will be used in
+                                        %% header instead of Full Title.
+%% \titlenote{with title note}             %% \titlenote is optional;
+%%                                         %% can be repeated if necessary;
+%%                                         %% contents suppressed with 'anonymous'
+%% \subtitle{Subtitle}                     %% \subtitle is optional
+%% \subtitlenote{with subtitle note}       %% \subtitlenote is optional;
+%%                                         %% can be repeated if necessary;
+%%                                         %% contents suppressed with 'anonymous'
 
-\author{Ambrus Kaposi \and Andr{\'a}s Kov{\'a}cs}
-\institute{E{\"o}tv{\"o}s Lor{\'a}nd University, Budapest, Hungary \\ \{akaposi$|$kovacsandras\}@inf.elte.hu}
+%% Author information
+%% Contents and number of authors suppressed with 'anonymous'.
+%% Each author should be introduced by \author, followed by
+%% \authornote (optional), \orcid (optional), \affiliation, and
+%% \email.
+%% An author may have multiple affiliations and/or emails; repeat the
+%% appropriate command.
+%% Many elements are not rendered, but should be provided for metadata
+%% extraction tools.
 
-\maketitle
+\author{Andr{\'a}s Kov{\'a}cs}
+\orcid{0000-0002-6375-9781}
+\affiliation{
+  \department{Department of Programming Languages and Compilers}
+  \institution{E{\"o}tv{\"o}s Lor{\'a}nd University}
+  \city{Budapest}
+  \country{Hungary}                   %% \country is recommended
+}
+\email{kovacsandras@inf.elte.hu}         %% \email is recommended
 
+%% Abstract
+%% Note: \begin{abstract}...\end{abstract} environment must come
+%% before \maketitle command
 \begin{abstract}
@@
 numbers. We develop new metatheory for large QIITs, large elimination, recursive
-equations, infinitary constructors and equations between sorts.
-%% First, we develop tools which allow precise and convenient handling
-%% of universe levels, working internally in an extensional type theory with
-%% cumulative universes.
-As in prior work, we describe QIITs using a type theory where each context
-represents a QIIT signature. However, in our case the theory of signatures can
-also describe its own signature. We use self-description to bootstrap a model
-theory for the theory of signatures without using preterms or assuming a
-pre-existing internal syntax for a type theory. We give initial algebra
-semantics for described QIITs, and we show the equivalence of initiality and
-induction. We present two extensions of a previous term model construction. The
-first one constructs all large infinitary QIITs without sort equations from the
-QIIT of the theory of signatures. The second one constructs all large finitary
-QIITs with sort equations, from the same syntax. This separation is required
-because handling infinitary constructors in the term model requires showing a
-strong form of invariance under algebra isomorphisms, which is violated by sort
-equations.
-
+equations and infinitary constructors. As in prior work, we describe QIITs using
+a type theory where each context represents a QIIT signature. However, in our
+case the theory of signatures can also describe its own signature, modulo
+universe sizes. We bootstrap the model theory of signatures using
+self-description and a Church-coded notion of signature, without using
+complicated raw syntax or assuming an existing internal QIIT of signatures. We
+give semantics to described QIITs by modeling each signature as a finitely
+complete CwF (category with families) of algebras. Compared to the case of
+finitary QIITs, we additionally need to show invariance under algebra
+isomorphisms in the semantics. We do this by modeling signature types as
+isofibrations. Finally, we show by a term model construction that every QIIT is
+constructible from the syntax of the theory of signatures.
 \end{abstract}
 
+%% 2012 ACM Computing Classification System (CSS) concepts
+%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
+\begin{CCSXML}
+<ccs2012>
+<concept>
+<concept_id>10003752.10003790.10011740</concept_id>
+<concept_desc>Theory of computation~Type theory</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+</ccs2012>
+\end{CCSXML}
+
+\ccsdesc[500]{Theory of computation~Type theory}
+%% End of generated code
+
+%% Keywords
+%% comma separated list
+\keywords{inductive types, quotient inductive-inductive types}  %% \keywords are mandatory in final camera-ready submission
+
+%% \maketitle
+%% Note: \maketitle command must come after title commands, author
+%% commands, abstract environment, Computing Classification System
+%% environment and commands, and keywords command.
+\maketitle
+
 \section{Introduction}
-\label{sec:intro}
-
-ISSUE. We actually need flCwF pseudomorphisms for Sub in infinitary semantics.
-$a : \Tm\,\Gamma\,\U$ is \emph{not} a discrete displayed flCwF, only $\El\,a$ is!
-$\Piinf\,A\,b : \Tm\,\Gamma,\U$ does not preserve (displayed) object in $\Gamma$
-strictly.
-
-\begin{itemize}
-\item New semantics with pseudomorphisms for Sub and Tm. Is it enough to
-      have weak preservation of comprehension, and we get all other preservation
-      for free, see ``gluing for type theory''?
-\item Check term model and eliminators to see if change needed.
-\item Check if the isomorphism for K in flCwF can be made strict, i.e.
-      $\Sub\,\Gamma\,\Delta = \Tm\,\Gamma\,(\Kfam\,\Delta)$.
-\end{itemize}
 
@@
 with later ones indexed over previous ones, and equations as well. We need both
-features in order to write down the signature of categories:
-\begin{alignat*}{3}
-  & Obj && : \Set \\
-  & Mor && : Obj \ra Obj \ra \Set \\
-  & id  && : Mor\,i\,j \\
-  & \blank\circ\blank && : Mor\,j\,k\ra Mor\,i\,j \ra Mor\,i\,k \\
-  & idl && : id \circ f = f \\
-  & idr && : f \circ id = f \\
-  & ass && : f \circ (g \circ h) = (f \circ g) \circ h
-\end{alignat*}
+features in order to write down the signature of categories.
+
 The benefit of having a QIIT signature is getting a model theory ``for free'',
@@
 initial models represent syntax, and initiality corresponds to induction on
-syntax. A number of different notions have been used, from contextual categories
-and comprehension categories to categories with families, but all of these are
+syntax. Several variants have been used, from contextual categories \cite{TODO} and
+comprehension categories \cite{TODO} to categories with families \cite{TODO}, but all of these are
 categories with extra structure.
 
-The main motivation of the current paper is to extend the notion of QIIT so that
-it accommodates all algebraic notions of type theories which have been used in
-previous works. As a side effect of fulfilling this goal, infinitary QIITs such
-as Cauchy real numbers are covered as well.
+The main motivation of the current paper is to generalize previously formalized
+\cite{kaposi2019constructing} QIITs so that they accommodate more algebraic
+formulations of type theories, and more aspects of their practical
+metatheory. As a side effect of fulfilling this goal, infinitary QIITs such as
+Cauchy real numbers \cite{TODO} are covered as well.
 
-We generalize previous notions of QIITs in the following ways:
+\subsection{Contributions}
+
+We add the following features to QIITs:
 \begin{enumerate}
   \item
-  \textbf{Large constructors, large elimination} and models at different
-  universe levels. This feature is routinely used in the metatheory of type
+  \textbf{Large constructors, large elimination} and algebras at different
+  universe levels. Large elimination is routinely used in the metatheory of type
   theory, but it has not been presented explicitly in previous works about
@@
   trees. Also, the theory of QIIT signatures is itself large and infinitary,
-  thus now it can ``eat itself'', i.e.\ include its own signature and provide
-  its own metatheory. This was previously not possible in \cite{TODO}, where
-  only finitary QIITs were described. We use this self-representation to
-  bootstrap the model theory of signatures, without having to assume any
-  pre-existing internal syntax.
+  thus it can ``eat itself'', i.e.\ include its own signature and provide
+  its own metatheory. This was not possible previously in
+  \cite{kaposi2019constructing}, where only finitary QIITs were described. We
+  exploit self-representation to bootstrap the model theory of signatures,
+  without having to assume any pre-existing internal syntax.
   \item
@@
   theories, as boundary conditions\cite{TODO}.
-  \item
-  \textbf{Sort equations}, or equations between type constructors. They have
-  been used recently to conveniently represent type-theoretic universes, for
-  example Russell-style universes\cite{TODO} or cumulative hierarchies\cite{TODO}.
 \end{enumerate}
 
-We also develop semantics. We show that for each signature, there is a CwF
-(category with families) of algebras, extended with $\Sigma$-types, extensional
-identity, and constant families. This additional structure corresponds to a
-type-theoretic flavor of finite limits, and it was shown in \cite{TODO} that the
-category of such CwFs is biequivalent to the category of finitely complete
-categories.
+To provide semantics, we show that for each signature, there is a CwF (category
+with families) of algebras, extended with $\Sigma$-types, extensional identity,
+and constant families. This additional structure corresponds to a type-theoretic
+flavor of finite limits, as it was shown in \cite{clairambault2014biequivalence}
+that the category of such CwFs is biequivalent to the category of finitely
+complete categories.
 
-As to the existence of initial algebras, we present two different term model
-constructions, yielding the following results:
+Compared to the case of finitary QIITs, the addition of infinitary constructors
+and recursive equations requires a significant change in semantics: instead of
+strict CwF morphisms, we need to consider weak ones, and instead of modeling
+types as displayed CwFs, we need to model them as CwF isofibrations. In a
+nutshell, the new semantics must be given mutually with a form of a ``structure
+identity principle''\cite{TODO}, which says that signature extension respects algebra
+isomorphisms.
 
-\begin{enumerate}
-\item
-  All large infinitary QIITs with recursive equations, but \emph{without
-  sort equations}, are reducible to the theory of signatures, i.e.\
-  their initial algebras are constructible from the initial algebra for
-  the theory of signatures.
-\item
-  All large finitary QIITs with sort equations, but no recursive equations,
-  are reducible to the theory of signatures.
-\end{enumerate}
+We also show, by a term model construction, that all QIITs are reducible to the
+syntax of signatures. This construction also essentially relies on invariance under
+isomorphisms.
 
-The reason for the two separate constructions is that sort equations are not as
-well-behaved as the other extensions: they are modelled as strict equalities of
-sets in algebras, hence they do not respect isomorphism of sets. We make
-essential use of invariance under set isomorphism in the first term model
-construction, so we cannot throw sort equations into that mix.
+\subsection{Outline of the Paper}
 
-In Section \ref{sec:levels} we set up cumulative universes which we use in later
-sections. In Section \ref{sec:tos} we introduce the theory of QIIT
-signatures. In Section \ref{sec:categorical_semantics} we develop initial algebra semantics,
-also covering the theory of signatures itself. In Sections
-\ref{sec:infinitary_term_model}-\ref{sec:sort_term} we present the two term model
-constructions. We discuss related work and conclusions in Sections
-\ref{sec:related}-\ref{sec:conclusion}.
+In Section \ref{sec:metatheory}, we describe a cumulative variant of extensional type
+theory which is used as metatheory in the rest of the paper. In Section \ref{sec:tos},
+we introduce the theory of QIIT signatures. TODO
 
-\section{Cumulative Extensional Type Theory}
-\label{sec:levels}
+\section{Metatheory}
 
-In the following sections, we will consider QIIT algebras at arbitrary finite
-levels, along with large eliminations, where the initial algebra can be at a
-different level than the target algebra. For a simple example, consider natural
-number algebras at level $i$, given as the $\Sigma$-type $NatAlg_i :\equiv(Nat :
-\Set_i)\times Nat\times(Nat \ra Nat)$. The initial such algebra is the set of
-natural numbers, which is at level $0$, but in type theory we often want to
-eliminate into larger $\Set_i$, for example when computing a $Nat$-indexed
-family of types. To support convenient reasoning about levels in this paper, we
-need two features: cumulativity and a way to quantify over finite levels.
-
-%% Universe levels are usually viewed as a bureaucratic and tedious part of type
-%% theory, and are often omitted or mentioned only in passing. Here we intend to
-%% model universe levels in a precise way, because large eliminations are an
-%% essential part of the practical usage of inductive types. Hence, we need a way
-%% to conveniently reason about levels without too much administrative
-%% burden. Unfortunately, simple universe setups are insufficient for this
-%% purpose. We need two features: cumulativity and a way to quantify over finite
-%% levels.
+In this paper we consider QII algebras at arbitrary finite levels, along with
+large eliminations, where the initial algebra can be at a different level than
+the target algebra. For a simple example, consider natural number algebras at
+level $i$, given as the $\Sigma$-type $NatAlg_i :\equiv(Nat : \Set_i)\times
+Nat\times(Nat \ra Nat)$. The initial such algebra is the set of natural numbers,
+which is at level $0$, but in type theory we often want to eliminate into larger
+$\Set_i$, for example when computing a $Nat$-indexed family of types. To support
+convenient reasoning about levels in this paper, we need two features:
+cumulativity and a way to quantify over finite levels. The former is essential
+to reduce bureaucratic overhead, while the latter is required to talk about
+arbitrary levels.
 
@@
 
-We need cumulativity to reduce bureaucratic overhead. In its most basic form,
-cumulativity requires that whenever $A : \Set_i$, and $i < j$ then also $A :
-\Set_j$. We also need cumulativity for $\Sigma$ and function types, in order to
-have cumulativity for algebras in general. For example, we want to have $\gamma
-: NatAlg_j$ whenever $\gamma : NatAlg_i$ and $i < j$. We have found that
-non-cumulative algebras induce excessive ``lifting'' noise, which we would like
-to avoid.
+In its most basic form, cumulativity requires that whenever $A : \Set_i$, and $i
+< j$ then also $A : \Set_j$. We also need cumulativity for $\Sigma$ and function
+types, in order to have cumulativity for QII algebras in general. For example,
+we want to have $\gamma : NatAlg_j$ whenever $\gamma : NatAlg_i$ and $i < j$.
 
-%% However, the $NatAlg$ example above also required a form of
-%% cumulativity for $\Sigma$ and $\Pi$ types, which we also need in general for
-%% QIIT algebras.
+We use Sterling's cumulative algebraic type theory \cite{sterling2019algebraic} as basis for our metatheory. \todo{REASONS}
 
-We use Sterling's cumulative algebraic type theory \cite{TODO} as the general
-setting for the rest of the paper. The reason for this is twofold. First, it
-supports the right kind of cumulativity for our purposes. Second, it is itself a
-finitary QIIT with sort equations, hence this paper provides a model theory for
-it\footnote{Which is, of course, somewhat circular, but not more circular than
-  e.g.\ the study of models of ZFC in ZFC.}.
+%% The reason is that it is the only rigorous
+%% algebraic definition of cumulativity that we are aware of, and it also
+
+%% The reason for this is twofold. First, it
+%% supports the right kind of cumulativity for our purposes. Second, it is itself a
+%% finitary QIIT with sort equations, hence this paper provides a model theory for
+%% it
 
 We extend the base theory with an extensional identity type, $\Sigma$-types and
-the unit type. In \cite{TODO} a proof of canonicity is provided for the base
+the unit type. In \cite{sterling2019algebraic} a proof of canonicity is provided for the base
 theory, which also includes a standard set-theoretic model. It is
@@
 
-We also have a lifting operation on types, which introduces cumulativity.
-\vspace{-1.5em}
 \begin{figure}
@@
 
+  \inferrule*[lab=equality formation]
+             {\Gamma \vdash A : \Set_i \\ \Gamma \vdash t : A \\ \Gamma \vdash u : A}
+             {\Gamma \vdash t = u : \Set_i}
+
   \inferrule*[lab=type lifting]
-             {\Gamma \vdash A : \Set_i \\ i < j}
+             {\Gamma \vdash A : \Set_i \\ i \leq j}
              {\Gamma \vdash\,\Lift_j A : \Set_{j}}
 
-  \inferrule*[lab=lift composition]{}{\Lift_j (\Lift_i A) =\,\,\Lift_{j} A}
+  \inferrule*[lab=lift composition]{}{\Lift_j (\Lift_i A) \equiv\,\,\Lift_{j} A}
+
+  \inferrule*[lab=identity lift]{\Gamma \vdash A : \Set_i}{(\Lift_i A) \equiv\,A}
 
@@
              {}
-             {\{t\,|\,\Gamma \vdash t : A\} = \{t\,|\,\Gamma \vdash t :\,\,\Lift_i A\}}
+             {\{t\,|\,\Gamma \vdash t : A\} \equiv \{t\,|\,\Gamma \vdash t :\,\,\Lift_i A\}}
 
@@
              {}
-             {(\Gamma,\,x : A) = (\Gamma,\,x :\,\,\Lift_i A)}
+             {(\Gamma,\,x : A) \equiv (\Gamma,\,x :\,\,\Lift_i A)}
 
-  \inferrule*[lab=universe lifting]{}{\Lift_j \Set_i = \Set_i}
+  \inferrule*[lab=universe lifting]{}{\Lift_j \Set_i \equiv \Set_i}
 
@@
              {}
-             {\Lift_i((x : A)\ra B) = (x :\,\Lift_i A)\ra\,\Lift_i B}
+             {\Lift_i((x : A)\ra B) \equiv (x :\,\Lift_i A)\ra\,\Lift_i B}
 
@@
              {}
-             {\Lift_i((x : A)\times B) = (x :\,\Lift_i A)\,\times\Lift_i B}
+             {\Lift_i((x : A)\times B) \equiv (x :\,\Lift_i A)\,\times\Lift_i B}
 
@@
              {}
-             {\Lift_i\top = \top}
+             {\Lift_i\top \equiv \top}
+
+  \inferrule*[lab=equality lifting]
+             {}
+             {\Lift_i(t = u) \equiv (t = u)}
 \end{mathpar}
@@
 
-In Figure \ref{lifting}, we include an excerpt of cETT's rules. We change
+We also have a lifting operation on types, which introduces cumulativity. In
+Figure \ref{lifting}, we include an excerpt of cETT's rules. We change
 presentation slightly from Sterling: we use a $\U_{\max(i,j)}$ return type in
 functions and pairs, instead of having both $A$ and $B$ types in the same
-$\Set_i$. These changed rules are all derivable from the original ones. In
-general, we can expect that lifting never impedes constructions, because it
-appropriately computes out of the way.
+$\Set_i$, and we define type lifting with $i \leq j$ instead of $i < j$. These
+changed rules are all derivable from the original ones. In general, we can
+expect that lifting never impedes constructions, because it appropriately
+computes out of the way.
 
@@
 same terms as unlifted ones. This allows us to have $t :\,\,\Lift_j A$ whenever $t
-: A$. Together with the universe lifting rule, this implies $A : Set_{i+j+1}$
-whenever $A : Set_i$. Similarly, the lifting rules for functions and pairs give
+: A$. Together with the universe lifting rule, this implies $A : \Set_{i+j}$
+whenever $A : \Set_i$. Similarly, the lifting rules for functions and pairs give
 us cumulativity for $NatAlg$. We derive a notion of subtyping:
 
-\begin{nidefinition}[Cumulative subtyping]
-  For $A : \Set_i$ and $B : \Set_{i+j+1}$, we define $A \prec B : \Set_{i+j+1}$
+\begin{definition}[Cumulative subtyping]
+  For $A : \Set_i$, $i \leq j$ and $B : \Set_{j}$, we define $A \leq B : \Set_{j}$
   as $B =\,\,\Lift_j A$.  It follows from term lifting and equality reflection
-  that whenever $A \prec B$ and $t : A$, then also $t : B$.
-\end{nidefinition}
+  that whenever $A \leq B$ and $t : A$, then also $t : B$.
+\end{definition}
 
@@
 
-We also need to quantify over universe levels. cETT does not support this, and
-we leave it like that. Instead, for the sake of simplicity, we quantify over
-levels in an unspecified metatheory \emph{outside} cETT. Hence, a universe
-polymorphic cETT term is understood as a $\mathbb{N}$-indexed family of
-terms. We reuse the notation of cETT functions for level-polymorphism, e.g.\
-as in the following function:
+For the sake of simplicity, we do not add this to cETT directly. Instead, we
+quantify over levels in an unspecified metatheory \emph{outside} cETT. Hence, a
+universe polymorphic cETT term is understood as a $\mathbb{N}$-indexed family of
+terms. We reuse the notation of cETT functions for universe polymorphism, e.g.\ as
+in the following function:
 \[
-\lambda i.\,\Set_i\ra\Set_i : (i : \mathbb{N})\ra \Set_{i+1}
+\lambda\,i.\,\Set_i : (i : \mathbb{N})\ra \Set_{i+1}
 \]
@@
 
-\section{Signatures}
+\section{QIIT Signatures}
 \label{sec:tos}
 
-In this section, we define signatures for QIITs which support all the mentioned
-generalizations. Signatures are given as contexts in a certain type theory,
-called the theory of signatures. We shall abbreviate it as ToS. However, ToS
-turns out to be a large infinitary QIIT itself, and we would like to define ToS
-and a notion of signature without referring to QIITs, only using features
-present in cETT. As a first step, we define a notion of model.
+Signatures are given as contexts in a certain type theory, the theory of
+signatures. We shall abbreviate it as ToS. However, ToS turns out to be a large
+infinitary QIIT itself, and we would like to define ToS and a notion of
+signature without referring to QIITs, only using features present in cETT. As a
+first step, we define a notion of model. $\Pi$
 
-\begin{nidefinition}[Notion of model for the theory of signatures]\label{def:tos}
+\begin{definition}[Notion of model for the theory of signatures]\label{def:tos}
 For levels $i$ and $j$, there is a cETT type $\ToS_{i,j} : \Set_{\max(i,\,j)+1}$, whose elements are
@@
   morphisms, types and terms) are in $\Set_j$. Following notation in
-  \cite{TODO}, we denote these respectively as $\Con : \Set_j$, $\Sub : \Con \ra
+  \cite{kaposi2019constructing}, we denote these respectively as $\Con : \Set_j$, $\Sub : \Con \ra
   \Con \ra \Set_j$, $\Ty : \Con \ra \Set_j$ and $\Tm : (\Gamma : \Con) \ra
@@
   and UIP.
-\item
-  An identity type $\IdU : \Tm\,\Gamma\,\U \ra \Tm\,\Gamma\,\U \ra \Ty\,\Gamma$,
-  with reflexivity as $\ReflU$, equality reflection and UIP.
 \end{enumerate}
-\end{nidefinition}
+\end{definition}
 In the above listing, we omit most CwF components and equations for substitution and
@@
 
-\emph{Notational conventions.} CwF components by default support de Bruijn
-indices, which are not easily readable. We use instead a nameful notation for
-binders in context extension, $\Pii$ and $\lami$, e.g.\ as $(\emptycon \ext a :
-U \ext t : \El\,a)$. We also define a type-theoretic version of $\appi$ for
-convenience:
+\emph{Notational conventions.} We name elements of $\Con$ as $\Gamma$, $\Delta$,
+$\Theta$, elements of $\Sub\,\Gamma\,\Delta$ as $\sigma$, $\delta$, $\nu$,
+elements of $\Ty\,\Gamma$ as $A$, $B$, $C$, and elements of $\Tm\,\Gamma\,A$ as
+$t$, $u$, $v$. CwF components by default support de Bruijn indices, which are
+not easily readable. We use instead a nameful notation for binders in context
+extension, $\Pii$ and $\lami$, e.g.\ as $(\emptycon\,\ext (a : U) \ext (t :
+\El\,a))$. We also define a type-theoretic flavor of $\appi$ for convenience:
 \begin{alignat*}{3}
@@
 
-\begin{nidefinition}[Notion of signature]
+\begin{definition}[Notion of signature]
 A QIIT signature at level $i$ is a context in an arbitrary $M : \ToS_{i,j}$
@@
 \[
-  \Sig_i :\equiv (j : \mathbb{N})(M : \ToS_{i,j})\ra M.Con
+  \Sig_i :\equiv (j : \mathbb{N})\ra(M : \ToS_{i,j})\ra Con_M
 \]
@@
 required to allow interpreting a signature in arbitrary-sized ToS models. Note
-that $\Sig_i$ is universe-polymorphic, so it does not have a type internally to
-cETT.
+that $\Sig_i$ is universe-polymorphic, so it is a family of cETT types and it is
+not in any cETT universe.
 
@@
     Signature for natural numbers. Here, no external types appear, so the level
-    of $NatSig$ can be chosen as $0$.}
+    can be chosen as $0$.}
 \begin{alignat*}{3}
@@
   & NatSig :\equiv \lambda (j :\mathbb{N})(M : \ToS_{0,j}).\\
-  & \hspace{2em}(\emptycon_M\,\ext_M\, (N : \U_M) \,\ext_M\, (zero : \El_M\,N) \ext_M (suc : \Pii_M\,N\,(\El_M\,N)))
+  & \hspace{1em}(\emptycon_M\,\ext_M\, (N : \U_M) \,\ext_M\,(zero : \El_M\,N)\\
+  & \hspace{2.67em}\ext_M (suc : \Pii_M\,N\,(\El_M\,N)))
 \end{alignat*}
@@
 signatures, but we only need contexts encoded in this way, and not other ToS
-components. In functional programming, this encoding is sometimes called
-``finally tagless''\cite{TODO}, and it is used for defining and interpreting
-embedded languages.
-\end{nidefinition}
-
-\subsection{Example Signatures}
-
-Compared to signatures in the previous work \cite{TODO}, we additionally support
-$\Piinf$ and $\IdU$, and $\Id$ has a more liberal type which allows recursive
-equations. Let us see examples for these three in order. For brevity, we shall
-only write contexts, and assume that all ToS operations come from an arbitrary
-model.
+components. In functional programming, this is sometimes called
+``finally tagless''\cite{carette2007finally}, and it is used for defining and
+interpreting embedded languages.
+\end{definition}
 
 \begin{example}{Infinitary constructors}. The universe $\U$ is closed under
-the $\Piinf$ function type, which makes it possible to write such functions in
-the domains of $\Pii$ types. This allows a signature for trees branching with
-arbitrary small sets. This is a signature at level 1, since we have $\Set_0$ as
-a $\Pie$ domain type.
+the $\Piinf$ function type, which allows such functions to appear in the domains
+of $\Pii$ types. This allows a signature for trees branching with arbitrary
+small sets. This is a signature at level 1, since we have $\Set_0$ as a $\Pie$
+domain type.
 \[
@@
 
-\begin{example}{Sort equations}. An interesting example is the term lifting
-rule for cETT (omitting everything except term lifting):
-\begin{alignat*}{3}
-  & cETTSig :\equiv \\
-  & ...\,\ext\,Tm{\Lift} : \Pii\,(\Gamma : Con)\,(\Pii\,(A : Ty{\appitt}\Gamma)((\IdU\,(Tm{\appitt}\Gamma{\appitt}A)\,(Tm{\appitt}\Gamma{\appitt}(\Lift_j{\appitt}A))))
-\end{alignat*}
-\end{example}
+%% \begin{example}{Sort equations}. An interesting example is the term lifting
+%% rule for cETT (omitting everything except term lifting):
+%% \begin{alignat*}{3}
+%%   & cETTSig :\equiv \\
+%%   & ...\,\ext\,Tm{\Lift} : \Pii\,(\Gamma : Con)\,(\Pii\,(A : Ty{\appitt}\Gamma)((\IdU\,(Tm{\appitt}\Gamma{\appitt}A)\,(Tm{\appitt}\Gamma{\appitt}(\Lift_j{\appitt}A))))
+%% \end{alignat*}
+%% \end{example}
 
-\begin{example}{Recursive equations}. Our definition of $\Id$ returns in $\U$,
-which, similarly to the case of infinitary functions, allows us to write equations
+\begin{example}{Recursive equations}. Again, the universe is closed under $\Id$, which allows us to write equations
 in $\Pii$ domains. A minimal example:
@@
 cubical type theories\cite{TODO}. Note that our $\Id$ allows iterated equations
-as well, but these are all made trivial in the semantics, where we assume UIP.
+as well, but these are all trivial in the semantics, where we assume UIP.
 \end{example}
 
-\section{Categorical Semantics}
-\label{sec:categorical_semantics}
+\section{Semantics}
+\label{sec:semantics}
+
+\subsection{Overview}
 
@@
 
-Following \cite{TODO}, we do this by creating a model of ToS, where contexts are
-categories supporting the above requirements and substitutions are strictly
-structure-preserving functors. Then, each signature can applied to this model,
-yielding an interpretation of the signature as a structured category of
-algebras. First, we fix our notion of such structured category.
+Following \cite{kaposi2019constructing}, we do this by creating a model of ToS,
+where contexts are categories supporting the above requirements and
+substitutions are appropriate structure-preserving functors. Then, each
+signature can applied to this model, yielding an interpretation of the signature
+as a structured category of algebras.
 
-\begin{nidefinition}[Finite limit CwFs]\label{def:flCwF}
+Our semantics has a ``type-theoretic'' flavor, which is inspired by the cubical
+set model of Martin-Löf type theory by Bezem et al.\cite{cubical}. The core idea
+is to avoid strictness issues by starting from basic ingredients which are already
+strict enough. Hence, instead of modeling types as certain slices and
+substitution by pullback, we model types as displayed categories with extra
+structure, which naturally support strict reindexing.
+
+We make a similar choice in the interpretation of signatures themselves: we
+use structured CwFs instead of lex categories to model notions of induction. The
+reason here is that CwFs allow us to define induction in strictly the same way as
+one would write it down in type theory, since we have $\Ty$ and $\Tm$ for a
+primitive notion of dependent objects and morphisms. In contrast, dependent
+objects in lex categories is a derived notion, and the induction principles we
+get are only up to isomorphisms. This issue is perhaps not relevant from a
+purely categorical perspective, but we are concerned with eventually
+implementing QIITs in proof assistants, so we prefer if our semantics computes
+strictly.
+
+In the following, we fix $i$ and $j$ levels, and define a model $M : \ToS_{i,
+  1+\max(i,j+1)}$, such than $\Con_M$ is a type of structured categories (of
+algebras). The level $i$ marks the level of all external sets in function
+domains, and $j$ marks the level of all internal sorts in an algebra. Hence,
+every algebra has level $\max(i,j+1)$. The bump is only needed for $j$, since
+algebras only contain elements of $T : \Set_j$ types, while inductive sets are
+themselves elements of $\Set_i$. For example, $NatAlg_i : \Set_{\max(0, i+1)}$.
+
+We present the components of the model in order. In the following, we use
+\textbf{bold} font to disambiguate components of the model from internal
+structures. For example, we use $\boldsymbol{\sigma : \Sub\,\Gamma\,\Delta}$
+to denote a substitution in the model.
+
+\subsection{Contexts}
+
+We define $\boldsymbol{\Con} : \Set_{1+\max(i,j+1)}$ as $\flCwF_{\max(i,j+1)}$. We define
+$\flCwF$ as follows.
+
+\begin{definition}[Finite limit CwFs]\label{def:flCwF}
 For a level $i$, there is a cETT type $\flCwF_i : \Set_{i+1}$, which is an
@@
     natural in $\Gamma$. The idea is that $\Kfam\,\Delta$ is a representation
-    of $\Delta$ as a type in any context. Dybjer and Clairambault called constant
-    families ``democracy'' in \cite{TODO}.
+    of $\Delta$ as a type in any context. Clairambault and Dybjer called constant
+    families ``democracy'' in \cite{clairambault2014biequivalence}.
 \end{enumerate}
-\end{nidefinition}
+\end{definition}
 
-We use $\Ty$ in an flCwF to represent ``dependent objects'', and $\Tm$
-for ``dependent morphisms''. We use these to define induction.
+\begin{definition} We abbreviate the additional structure on CwFs consisting of $\Sigma$, $\Id$ and
+$\Kfam$ as \emph{fl-structure}.
+\end{definition}
 
-\begin{nidefinition}[Notion of induction in an flCwF]
+%% We abbreviate the additional structure on CwFs consisting of $\Sigma$, $\Id$ and
+%% $\Kfam$ as fl-structure. We use $\Ty$ in an flCwF to represent ``dependent
+%% objects'', and $\Tm$ for ``dependent morphisms''. We use these to define
+%% induction.
+
+\begin{definition}[Notion of induction in an flCwF]
 Given $C : \flCwF_i$, we have the following predicate on objects:
@@
 \end{alignat*}
-\end{nidefinition}
+\end{definition}
 
@@
 \end{theorem}
-\begin{proof} Shown by brief internal flCwF reasoning in \cite{TODO}.\qed
+\begin{proof} By brief internal flCwF reasoning in \cite{kaposi2019constructing}.
 \end{proof}
 
-The reason for the ``finite limit CwF'' naming is the following: Dybjer and
-Clairambault showed\cite{TODO} that the 2-category of flCwFs is biequivalent to
-the 2-category of finitely complete (or ``lex'') categories. It can be also seen
-that the categorical product of $\Gamma$ and $\Delta$ in an flCwF can be given
-as $\Gamma \ext \Kfam\, \Delta$, and equalizers can be given using the identity
-type.
+The reason for the ``finite limit CwF'' naming is the following: Clairambault
+and Dybjer showed that the 2-category of flCwFs is biequivalent to the
+2-category of finitely complete (or ``lex'') categories
+\cite{clairambault2014biequivalence}. It can be also seen that the categorical
+product of $\Gamma$ and $\Delta$ in an flCwF can be given as $\Gamma \ext
+\Kfam\, \Delta$, and equalizers can be given using the identity type.
 
-So, why do we use flCwFs instead of lex categories? After all, the former has
-more components and has more immediate complexity than the latter. The reason is
-that flCwFs allow us to define induction in strictly the same way as one would
-write it down in type theory, since we have $\Ty$ and $\Tm$ for a primitive
-notion of dependent objects and morphisms. In contrast, dependent objects in lex
-categories is a derived notion, and the induction principles we get are what we
-want only up to isomorphisms. This issue is perhaps not relevant from a purely
-categorical perspective, but we are concerned with eventually implementing QIITs
-in proof assistants, so we prefer if our semantics computes strictly.
+\todo{segue text?}
 
-\begin{nidefinition}[flCwF model for ToS] We have for each $i$ and $j$ levels
-an $M : \ToS_{i, 1+\max(i,j+1)}$ such that $\Con_M = \flCwF_{\max(i,j+1)}$. This
-model is an extension of the CwF model of signatures in \cite{TODO}. The new
-additions are: $\Sigma$-types in the semantic CwF, the new type formers in
-signatures, and the treatment of universe levels. We summarize the model below.
+\begin{definition} A \emph{context isomorphism} is an invertible morphism $\sigma : \Sub\,\Gamma\,\Delta$.
+We note the inverse as $\sigma^{-1}$.
+We also use the notation $\sigma : \Gamma \simeq \Delta$.
+\end{definition}
 
-As noted, $\Con$ is given as $\flCwF_{\max(i,j+1)}$. The level $i$ marks the
-level of all external sets in function domains, and $j$ marks the level of all
-internal sorts in an algebra. Hence, every algebra, i.e.\ object of the flCwF,
-has level $\max(i,j+1)$. The bump is only needed for $j$, since algebras only
-contain elements of $T : \Set_j$ types, while inductive sets are themselves
-elements of $\Set_i$. For example, $NatAlg_i : \Set_{\max(0, i+1)}$.
+\begin{definition}[Type categories, c.f.\ \cite{clairambault2014biequivalence}] For each $\Gamma : \Con$, there is a category
+whose objects are types $A : \Ty\,\Gamma$, and morphisms from $A$ to $B$ are
+terms $t : \Tm\,(\Gamma\,\ext\,A)\,B[\p]$. Identity morphisms are given by $\q :
+\Tm\,(\Gamma\,\ext\,A)\,A[\p]$, and composition $t \circ u$ by $t[\p, u]$. The
+assignment of type categories to contexts extends to an indexed category. For
+each $\sigma : \Sub\,\Gamma\,\Delta$, there is a functor from $\Ty\,\Delta$ to
+$\Ty\,\Gamma$, which sends $A$ to $A[\sigma]$ and $t :
+\Tm\,(\Gamma\,\ext\,A)\,B[\p]$ to $t[\sigma\circ \p, \q]$.
+\end{definition}
 
-$\sigma : \Sub\,\Gamma\,\Delta$ substitutions are functors preserving all flCwF
-structure strictly.
+\begin{definition} A \emph{type isomorphism}, notated $t : A \simeq B$ is an isomorphism in a type category. We note the inverse as $t^{-1}$.
+\end{definition}
 
-$A : \Ty\,\Gamma$ types are displayed flCwFs over $\Gamma$. This is an
-flCwF where every underlying set is indexed over underlying sets in
-$\Gamma$, and it is an extension of the displayed categories of Ahrens and
-Lumsdaine\cite{TODO}.  It contains equivalent data to an flCwF slice
-$\delta : \Sub\,\Delta\,\Gamma$. We prefer it to slices because it yields
-expected notions of induction strictly.
+\subsection{Substitutions}
 
-$t : \Tm\,\Gamma\,A$ terms are sections of the displayed flCwF $A$. This can
-be viewed as a dependent functor, which maps structure in $\Gamma$ to
-corresponding displayed structure in $A$, e.g.\ it maps objects to displayed
-objects, morphisms to displayed morphisms, and so on.
+%% Acknowledgments
+\begin{acks}
+This work was supported by the European Union, co-financed by the
+European Social Fund (EFOP-3.6.3-VEKOP-16-2017-00002) and COST Action
+EUTypes CA15123.
+\end{acks}
 
-$\emptycon : \Con$ is the terminal flCwF, whose underlying category is the
-terminal category with one object.
-
-$\Gamma\,\ext\,A$ is the total flCwF of the $A$ displayed flCwF, where every
-underlying set is given by packing up indexed underlying sets of $A$ in
-$\Sigma$-s.
-
-$\U : \Ty\,\Gamma$ is a displayed flCwF which is constantly (i.e. it does not
-depend on $\Gamma$) the flCwF of sets in $\Set_j$, whose underlying category
-is the usual category of sets, and which has evident finite limit structure
-($\Sg$, $\Id$, $\K$ from Definition \ref{def:flCwF}). Hence, elements of
-$\Tm\,\Gamma\,\U$ are functors from $\Gamma$ to \textbf{Set}. $\El :
-\Tm\,\Gamma\,\U\ra\Ty\,\Gamma$ interprets such functors as discrete displayed
-flCwFs, where all morphisms are identities.
-
-$\Pii\,a\,B : \Ty\,\Gamma$ is a displayed flCwF where objects are dependent
-functions from the set of objects of $a$ to the set of objects of $B$, and
-morphisms are pointwise morphisms. This is possible to build because the $a :
-\Tm\,\Gamma\,\U$ domain is discrete. A general $\Pii$ type is not possible
-because of variance issues; this follows from the fact that the category of
-categories does not support $\Pii$ types\cite{elephant}. Finite limit structure
-is given pointwise, and is inherited from the domain $B$.
-
-$\Pie\,A\,B : \Ty\,\Gamma$ is the $A$-indexed direct product of the family of
-displayed flCwFs given by $B : A \ra \Ty\,\Gamma$.
-
-$\Piinf\,A\,b : \Tm\,\Gamma\,\U$ is likewise the $A$-indexed direct product of
-the family of discrete displayed flCwFs given by $b : A \ra \Tm\,\Gamma\,\U$.
-
-$\Id\,a\,t\,u : \Tm\,\Gamma\,\U$ maps every object $\gamma$ of $\Gamma$ to the
-set $t\,\gamma = u\,\gamma$, and every morphism $\sigma :
-\Sub_\Gamma\,\gamma\,\gamma'$ to a function with type $t\,\gamma = u\,\gamma
-\ra t\,\gamma' = u\,\gamma'$, which can be given by the functoriality of $t$
-and $u$. In short, $\Id\,a\,t\,u$ expresses pointwise equality of $t$ and $u$
-dependent functors.
-
-$\IdU\,a\,b : \Ty\,\Gamma$ is a displayed flCwF where the set of objects over
-$\gamma : \Con_\Gamma$ is $a\,\gamma = b\,\gamma$, that is, a strict equation
-of types in $\Set_j$. Displayed morphisms are trivial (i.e.\ constantly
-$\top$) because of UIP in cETT.
-
-%% \begin{itemize}
-%% \item
-%%   As noted, $\Con = \flCwF_{\max(i,j+1)}$. The level $i$ marks the level of
-%%   all external sets in function domains, and $j$ marks the level of all
-%%   inductive sets in an algebra. Hence, every algebra, i.e.\ object of the flCwF,
-%%   has level $\max(i,j+1)$. The bump is only needed for $j$, since algebras only
-%%   contain elements of $T : \Set_j$ types, while inductive sets are themselves
-%%   elements of $\Set_i$. For example, $NatAlg_i : \Set_{\max(0, i+1)}$.
-%% \item
-%%   $\sigma : \Sub\,\Gamma\,\Delta$ substitutions are functors preserving all flCwF
-%%   structure strictly.
-%% \item
-%%   $A : \Ty\,\Gamma$ types are displayed flCwFs over $\Gamma$. This is an
-%%   flCwF where every underlying set is indexed over underlying sets in
-%%   $\Gamma$, and it is an extension of the displayed categories of Ahrens and
-%%   Lumsdaine\cite{TODO}.  It contains equivalent data to an flCwF slice
-%%   $\delta : \Sub\,\Delta\,\Gamma$. We prefer it to slices because it yields
-%%   expected notions of induction strictly.
-%% \item
-%%   $t : \Tm\,\Gamma\,A$ terms are sections of the $A$ displayed flCwF. This can
-%%   be viewed as a dependent functor, which maps structure in $\Gamma$ to
-%%   corresponding displayed structure in $A$, e.g.\ it maps objects to displayed
-%%   objects, morphisms to displayed morphisms, and so on.
-%% \item
-%%   $\emptycon : \Con$ is the terminal flCwF, whose underlying category is the
-%%   terminal category with one object.
-%% \item
-%%   $\Gamma\,\ext\,A$ is the total flCwF of the $A$ displayed flCwF, where every
-%%   underlying set is given by packing up indexed underlying sets of $A$ in
-%%   $\Sigma$-s.
-%% \item
-%%   $\U : \Ty\,\Gamma$ is a displayed flCwF which is constantly (i.e. it does not
-%%   depend on $\Gamma$) the flCwF of sets in $\Set_j$, whose underlying category
-%%   is the usual category of sets, and which has evident finite limit structure
-%%   ($\Sg$, $\Id$, $\K$ from Definition \ref{def:flCwF}). Hence, elements of
-%%   $\Tm\,\Gamma\,\U$ are functors from $\Gamma$ to \textbf{Set}. $\El :
-%%   \Tm\,\Gamma\,\U\ra\Ty\,\Gamma$ interprets such functors as discrete displayed
-%%   flCwFs, where all morphisms are identities.
-%% \item
-%%   $\Pii\,a\,B : \Ty\,\Gamma$ is a displayed flCwF where objects are dependent
-%%   functions from the set of objects of $a$ to the set of objects of $B$, and
-%%   morphisms are pointwise morphisms. This is possible to build because the $a :
-%%   \Tm\,\Gamma\,\U$ domain is discrete. A general $\Pii$ type is not possible
-%%   because of variance issues; this follows from the fact that the category of
-%%   categories does not support $\Pii$ types\cite{elephant}.
-%% \item
-%%   $\Pie\,A\,B : \Ty\,\Gamma$ is the $A$-indexed direct product of the family of
-%%   displayed flCwFs given by $B : A \ra \Ty\,\Gamma$.
-%% \item
-%%   $\Piinf\,A\,b : \Tm\,\Gamma\,\U$ is likewise the $A$-indexed direct product of
-%%   the family of discrete displayed flCwFs given by $b : A \ra \Tm\,\Gamma\,\U$.
-%% \item
-%%   $\Id\,a\,t\,u : \Tm\,\Gamma\,\U$ maps every object $\gamma$ of $\Gamma$ to the
-%%   set $t\,\gamma = u\,\gamma$, and every morphism $\sigma :
-%%   \Sub_\Gamma\,\gamma\,\gamma'$ to a function with type $t\,\gamma = u\,\gamma
-%%   \ra t\,\gamma' = u\,\gamma'$, which can be given by the functoriality of $t$
-%%   and $u$. In short, $\Id\,a\,t\,u$ expresses pointwise equality of $t$ and $u$
-%%   dependent functors.
-%% \item
-%%   $\IdU\,a\,b : \Ty\,\Gamma$ is a displayed flCwF where the set of objects over
-%%   $\gamma : \Con_\Gamma$ is $a\,\gamma = b\,\gamma$, that is, a strict equation
-%%   of types in $\Set_j$. Displayed morphisms are trivial (i.e.\ constantly
-%%   $\top$) because of UIP in cETT.
-%% \end{itemize}
-
-\begin{example}{Algebras for pointed sets.}
-
-\end{example}
-
-%%   For $i$ and $j$ levels, there is
-%% an $M : \ToS_{i, 1+\max(i,j+1)}$ such that $M.\Con = \flCwF_{\max(i,j+1)}$.
-
-\end{nidefinition}
-
-\section{Infinitary Term Models}
-\label{sec:infinitary_term_models}
-
-\subsection{Induction on Signatures}
-
-\todo{sync}
-
-Starting from Section \ref{sec:infinitary_term_models}, we need to do proofs and
-constructions by induction on signatures. However, the current signatures do not
-support induction, analogously to how induction is not available for Church
-encodings. We would like to have two things: a notion of signature induction,
-and a model of ToS which supports induction. Then, we can use contexts in the
-syntax of ToS (i.e.\ in the inductive model) as signatures. We provide a notion
-of induction on signatures by
-\begin{enumerate}
-  \item Showing that there is a signature which describes ToS.
-  \item Building categorical semantics, which for each signature provides a notion of induction.
-\end{enumerate}
-On the other hand, in this paper we will only \emph{assume} the existence of a
-syntax for ToS starting from Section \ref{sec:infinitary_term_models}. We do not provide a
-construction of the syntax from simpler type formers, although we do show that
-lower-level ToS syntaxes are constructible from higher-level ones.
-
-\subsection{Preservation of Isomorphisms}
-
-\section{Term Models with Sort Equations}
-\label{sec:sort_term}
-
-\section{Related Work}
-\label{sec:related}
-
-\section{Conclusion}
-\label{sec:conclusion}
-
+%% Bibliography
 \bibliography{references}

### 2019-12-28 draft/paper.tex
paper

@@
 \newcommand{\q}{\mathsf{q}}
+\newcommand{\lamK}{\mathsf{lam}^{\K}}
+\newcommand{\appK}{\mathsf{app}^{\K}}
 
-\newtheorem{abbreviation}{Abbreviation}
+\newcommand{\bCon}{\boldsymbol{\Con}}
+\newcommand{\bTy}{\boldsymbol{\Ty}}
+\newcommand{\bSub}{\boldsymbol{\Sub}}
+\newcommand{\bTm}{\boldsymbol{\Tm}}
+\newcommand{\bGamma}{\boldsymbol{\Gamma}}
+\newcommand{\bDelta}{\boldsymbol{\Delta}}
+\newcommand{\bsigma}{\boldsymbol{\sigma}}
+\newcommand{\bdelta}{\boldsymbol{\delta}}
+\newcommand{\bt}{\boldsymbol{t}}
+\newcommand{\bu}{\boldsymbol{u}}
+\newcommand{\bA}{\boldsymbol{A}}
+\newcommand{\bB}{\boldsymbol{B}}
+\newcommand{\bid}{\boldsymbol{\id}}
+
+%% \newcommand{\bDelta}{\boldsymbol{\Delta}}
+
 %% ------------------------------------------------------------
@@
 \section{Introduction}
+
+\todo{P and Q in cwf, level lifts, think about revising cETT or compressing}
 
@@
 \begin{definition}[Notion of model for the theory of signatures]\label{def:tos}
-For levels $i$ and $j$, there is a cETT type $\ToS_{i,j} : \Set_{\max(i,\,j)+1}$, whose elements are
-ToS models (or ToS-algebras). $\ToS_{i,j}$ is an iterated $\Sigma$-type, containing
-all of the following components.
-\begin{enumerate}
-\item
-  A category with families (CwF), where all four underlying sets (of objects,
-  morphisms, types and terms) are in $\Set_j$. Following notation in
-  \cite{kaposi2019constructing}, we denote these respectively as $\Con : \Set_j$, $\Sub : \Con \ra
-  \Con \ra \Set_j$, $\Ty : \Con \ra \Set_j$ and $\Tm : (\Gamma : \Con) \ra
-  \Ty\,\Gamma \ra \Set_j$. We denote the empty context as $\emptycon : \Con$,
-  context extension as $\blank\ext\blank : (\Gamma : \Con)\ra \Ty\,\Gamma \ra
-  \Con$.  Substitution on types and terms is written as $\blank[\blank]$, with
-  $\id$ and $\blank\circ\blank$ for identity and composition, and $\blank,\blank
-  : (\sigma : \Sub\,\Gamma\,\Delta)\ra \Tm\,\Gamma\,A[\sigma]\ra
-  \Sub\,\Gamma\,(\Delta\ext A)$ for substitution extension.
-\item
-  A universe $\U : \Ty\,\Gamma$ with decoding $\El : (a : \Tm\,\Gamma\,\U) \ra
-  \Ty\,\Gamma$.
-\item
-  Inductive function space, with $\Pii : (a : \Tm\,\Gamma\,\U) \ra
+For levels $i$ and $j$, $\ToS_{i,j} : \Set_{\max(i,\,j)+1}$ is a cETT type whose
+elements are ToS models (or ToS-algebras). $\ToS_{i,j}$ is an iterated
+$\Sigma$-type, containing all of the following components.
+
+A \textbf{category with families} (CwF), where all four underlying sets (of
+objects, morphisms, types and terms) are in $\Set_j$. Following notation in
+\cite{kaposi2019constructing}, we denote these respectively as $\Con : \Set_j$,
+$\Sub : \Con \ra \Con \ra \Set_j$, $\Ty : \Con \ra \Set_j$ and $\Tm : (\Gamma :
+\Con) \ra \Ty\,\Gamma \ra \Set_j$. We use $\id$ and $\blank\circ\blank$ to
+denote identity and composition for substitution. We denote the empty context as
+$\emptycon : \Con$, and the unique substitution into the empty context as
+$\epsilon : \Sub\,\Gamma\,\emptycon$. Context extension is $\blank\ext\blank :
+(\Gamma : \Con)\ra \Ty\,\Gamma \ra \Con$. Substitution on types and terms is
+written as $\blank[\blank]$. Projections are noted as $\p :
+\Sub\,(\Gamma\,\ext\,A)\,\Gamma$ and $\q : \Tm\,(\Gamma\,\ext\,A)\,(A[\p])$, and
+substitution extension is $\blank,\blank : (\sigma : \Sub\,\Gamma\,\Delta)\ra
+\Tm\,\Gamma\,A[\sigma]\ra \Sub\,\Gamma\,(\Delta\ext A)$.
+
+A \textbf{universe} $\U : \Ty\,\Gamma$ with decoding $\El : (a :
+\Tm\,\Gamma\,\U) \ra \Ty\,\Gamma$.
+
+\textbf{Inductive function space}, with $\Pii : (a : \Tm\,\Gamma\,\U) \ra
 \Ty\,(\Gamma\,\ext\,\El\,a) \ra \Ty\,\Gamma$ and application as $\appi :
@@
 inverse $\lami$.
-\item
-  External function space with $\Set_i$ domain: $\Pie : (A : \Set_i)\ra(A \ra
-  \Ty\,\Gamma)\ra \Ty\,\Gamma$, with $\appe : \Tm\,\Gamma\,(\Pie\,A\,B)\ra((x :
-  A)\ra \Tm\,\Gamma\,(B\,x))$ and its inverse $\lame$.
-\item
-  Infinitary function space, with $\Piinf : (A : \Set_i)\ra(A \ra
+
+\textbf{External function space} with $\Set_i$ domain: $\Pie : (A : \Set_i)\ra(A
+\ra \Ty\,\Gamma)\ra \Ty\,\Gamma$, with $\appe : \Tm\,\Gamma\,(\Pie\,A\,B)\ra((x
+: A)\ra \Tm\,\Gamma\,(B\,x))$ and its inverse $\lame$.
+
+\textbf{Infinitary function space}, with $\Piinf : (A : \Set_i)\ra(A \ra
 \Tm\,\Gamma\,\U)\ra \Tm\,\Gamma\,\U$, $\appinf :
-  \Tm\,\Gamma\,(\El\,(\Pie\,A\,b))\ra((x : A)\ra \Tm\,\Gamma\,(\El\,(b\,x)))$
-  and $\laminf$.
-\item
-  An identity type $\Id : (a : \Tm\,\Gamma\,\U)\ra
-  \Tm\,\Gamma\,(\El\,a)\ra\Tm\,\Gamma\,(\El\,a)\ra \Tm\,\Gamma\,\U$, with
-  $\Refl : (T : \Tm\,\Gamma\,(\El\,a))\ra \Id\,a\,t\,t$, equality reflection
-  and UIP.
-\end{enumerate}
+\Tm\,\Gamma\,(\El\,(\Pie\,A\,b))\ra((x : A)\ra \Tm\,\Gamma\,(\El\,(b\,x)))$ and
+$\laminf$.
+
+\textbf{An identity type} $\Id : (a : \Tm\,\Gamma\,\U)\ra
+\Tm\,\Gamma\,(\El\,a)\ra\Tm\,\Gamma\,(\El\,a)\ra \Tm\,\Gamma\,\U$, with $\Refl :
+(T : \Tm\,\Gamma\,(\El\,a))\ra \Id\,a\,t\,t$, equality reflection and UIP.
+
 \end{definition}
-In the above listing, we omit most CwF components and equations for substitution and
-$\beta\eta$-conversion, but these should be understood to be also part of $\ToS_{i,j}$.
+In the above listing, we omit equations for substitution and
+$\beta\eta$-conversion, but these should be understood to be also part of
+$\ToS_{i,j}$.
 
@@
 We present the components of the model in order. In the following, we use
-\textbf{bold} font to disambiguate components of the model from internal
+\textbf{bold} font to disambiguate components of this model from internal
 structures. For example, we use $\boldsymbol{\sigma : \Sub\,\Gamma\,\Delta}$
@@
 
-We define $\boldsymbol{\Con} : \Set_{1+\max(i,j+1)}$ as $\flCwF_{\max(i,j+1)}$. We define
+We define $\bCon : \Set_{1+\max(i,j+1)}$ as $\flCwF_{\max(i,j+1)}$. We define
 $\flCwF$ as follows.
@@
 \begin{enumerate}
-  \item A CwF with underlying sets all in $\Set_i$.
+  \item A CwF with underlying sets all in $\Set_i$. We reuse the notation
+    from definition \ref{def:tos}.
   \item $\Sigma$-type $\Sg : (A : \Ty\,\Gamma)\ra \Ty\,(\Gamma \ext A)
-    \ra \Ty\,\Gamma$, with usual projections and pairing.
+    \ra \Ty\,\Gamma$, with term formers $\proj1$, $\proj2$ and $\blank,\blank$.
   \item Identity type $\Id : (A : \Ty\,\Gamma)\ra \Tm\,\Gamma\,A\ra
@@
     between $\Sub\,\Gamma\,\Delta$ and $\Tm\,\Gamma\,(\Kfam\,\Delta)$ which is
-    natural in $\Gamma$. The idea is that $\Kfam\,\Delta$ is a representation
-    of $\Delta$ as a type in any context. Clairambault and Dybjer called constant
-    families ``democracy'' in \cite{clairambault2014biequivalence}.
+    natural in $\Gamma$. We have $\lamK : \Sub\,\Gamma\,\Delta \ra
+    \Tm\,\Gamma\,(\Kfam\,\Delta)$ and $\appK : \Tm\,\Gamma\,(\Kfam\,\Delta)
+    \ra \Sub\,\Gamma\,\Delta$ as its components.  The idea is that
+    $\Kfam\,\Delta$ is a representation of $\Delta$ as a type in any
+    context. Clairambault and Dybjer called constant families ``democracy'' in
+    \cite{clairambault2014biequivalence}.
 \end{enumerate}
@@
 \begin{definition}[Notion of induction in an flCwF]
-Given $C : \flCwF_i$, we have the following predicate on objects:
+Given $\bGamma : \flCwF_i$, we have the following predicate on contexts:
 \begin{alignat*}{3}
-  & Inductive : \Con_C \ra \Set_i \\
-  & Inductive\,\Gamma :\equiv (A : \Ty_C\,\Gamma)\ra \Tm_C\,\Gamma\,A
+  & Inductive : \Con_{\bGamma} \ra \Set_i \\
+  & Inductive\,\Gamma :\equiv (A : \Ty_{\bGamma}\,\Gamma)\ra \Tm_{\bGamma}\,\Gamma\,A
 \end{alignat*}
@@
 \begin{theorem}[Equivalence of initiality and induction]
-An object $\Gamma : \Con_C$ supports induction if and only if it is
+An object $\Gamma : \Con_{\bGamma}$ supports induction if and only if it is
 initial. Moreover, induction and initiality are both proof irrelevant
@@
 
+\todo{we have a level lift here, decide how to handle}
+
+We define $\boldsymbol{\Sub\,\Gamma\,\Delta} : \Set_{1 + \max(i, j+1)}$ as the type
+of weak flCwF morphisms from $\boldsymbol{\Gamma}$ to $\boldsymbol{\Delta}$.
+
+\begin{definition} A weak flCwF morphism $\boldsymbol{\sigma : \Sub\,\Gamma\,\Delta}$ is a functor between underlying categories, which additionally maps types to types and terms to terms, and satisfies the following:
+  \begin{enumerate}
+    \item $\bsigma\,(A[\sigma]) = (\bsigma\,A)\,[\bsigma\,\sigma]$
+    \item $\bsigma\,(t[\sigma]) = (\bsigma\,t)\,[\bsigma\,\sigma]$
+    \item The unique map $\epsilon : \Sub\,(\bsigma\,\emptycon)\,\emptycon$ has a retraction.
+    \item Each $(\bsigma\,\p,\,\bsigma\,\q) : \Sub\,(\bsigma\,(\Gamma\,\ext\,A))\,(\bsigma\,\Gamma\,\ext\,\bsigma\,A)$ has an inverse.
+  \end{enumerate}
+\end{definition}
+
+In short, $\bsigma$ preserves substitution strictly and preserves
+empty context and context extension up to isomorphism. We notate the evident
+isomorphisms as $\bsigma_{\emptycon} :
+\bsigma\,\emptycon \simeq \emptycon$ and $\bsigma_{\ext}
+:
+\bsigma\,(\Gamma\,\ext\,A)\,\simeq\,\bsigma\,\Gamma\,\ext\,\bsigma\,A$.
+
+%% We do not require preservation of fl-structure, because the above conditions already insure that.
+
+\begin{theorem} Every $\boldsymbol{\sigma : \Sub\,\Gamma\,\Delta}$ preserves fl-structure up to type isomorphism. That is, we have isomorphisms
+\begin{alignat*}{3}
+  & \bsigma_{\Sigma} : \bsigma\,(\Sigma\,A\,B) \simeq \Sigma\,(\bsigma\,A)\,((\bsigma\,B)[\bsigma_{\ext}^{-1}]) \\
+  & \bsigma_{\K} : \bsigma\,(\K\,\Delta) \simeq \K\,(\bsigma\,\Delta) \\
+  & \bsigma_{\Id} : \bsigma\,(\Id\,t\,u) \simeq \Id\,(\bsigma\,t)\,(\bsigma\,u)
+\end{alignat*}
+Moreover, $\bsigma$ preserves all term and substitution formers in the
+fl-structure. For example, $\bsigma\,(\proj1\,t) = \proj1\,
+(\bsigma_{\Sigma}[\id, \bsigma\,t])$.
+\end{theorem}
+\begin{proof}
+For $\bsigma_{\Sigma}$, we construct the following context isomorphism:
+\begin{alignat*}{3}
+& (\bsigma\,\Gamma\,\ext\,\bsigma\,(\Sigma\,A\,B)) \simeq
+  (\bsigma\,\Gamma\,\ext\,\bsigma\,A\,\ext\,(\Sigma\,B)[\bsigma_{\ext}^{-1}]) \\
+& \simeq (\bsigma\,\Gamma\,\ext\,\Sigma\,(\bsigma\,A)\,((\Sigma\,B)[\bsigma_{\ext}^{-1}]))
+\end{alignat*}
+This isomorphism is the identity on $\bsigma\,\Gamma$, hence we can extract the
+desired $\bsigma_{\Sigma} : \bsigma\,\Sigma\,A\,B) \simeq
+\Sigma\,(\bsigma\,A)\,((\bsigma\,B)[\bsigma_{\ext}^{-1}])$ from it.
+
+For $\bsigma_{\K}$, note the following:
+\begin{alignat*}{3}
+  & (\emptycon\,\ext\,\bsigma\,(\K\,\Delta)) \simeq
+    (\bsigma\,\emptycon\,\ext\,\bsigma\,(\K\,\Delta)) \simeq
+    \bsigma\,(\emptycon\,\ext\,\K\,\Delta)\\
+  & \simeq \bsigma\,\Delta \simeq (\emptycon\,\ext\,K\,(\bsigma\,\Delta))
+\end{alignat*}
+This yields a type isomorphism $\bsigma\,(\K\,\Delta)) \simeq
+\K\,(\bsigma\,\Delta)$ in the empty context, and we use the functorial action of
+$\epsilon : \Sub\,\Gamma\,\emptycon$ to weaken it to any $\Gamma$ context.
+
+For $\bsigma_{\Id}$, both component morphisms can be constructed by $\refl$ and
+equality reflection, and the morphisms are inverses by UIP. We omit here the
+verification that $\bsigma$ preserves term and substitution formers.
+\end{proof}
+
+\subsection{Identity and Composition}
+
+%% $\bid : \bSub\,\bGamma\,\bGamma$ is defined in the obvious way, with $\bid
+%% For composition
+
 %% Acknowledgments

### 2019-12-28 draft/paper.tex
paper

@@
 
+\usepackage{tikz-cd}
+
+\newtheorem{notation}{Notation}
+
 \input{abbrevs.tex}
@@
 \newcommand{\bdelta}{\boldsymbol{\delta}}
+\newcommand{\bepsilon}{\boldsymbol{\epsilon}}
 \newcommand{\bt}{\boldsymbol{t}}
@@
 \newcommand{\bid}{\boldsymbol{\id}}
+\newcommand{\bemptycon}{\boldsymbol{\emptycon}}
+
+\newcommand{\ul}[1]{\underline{#1}}
+\newcommand{\ulGamma}{\ul{\Gamma}}
+\newcommand{\ulDelta}{\ul{\Delta}}
+\newcommand{\ulsigma}{\ul{\sigma}}
+\newcommand{\uldelta}{\ul{\delta}}
+\newcommand{\ulepsilon}{\ul{\epsilon}}
+\newcommand{\ult}{\ul{t}}
+\newcommand{\ulu}{\ul{u}}
+\newcommand{\ulA}{\ul{A}}
+\newcommand{\ulB}{\ul{B}}
+
+\newcommand{\coe}{\mathsf{coe}}
+\newcommand{\coh}{\mathsf{coh}}
 
@@
 
-We use Sterling's cumulative algebraic type theory \cite{sterling2019algebraic} as basis for our metatheory. \todo{REASONS}
+We use Sterling's cumulative algebraic type theory \cite{sterling2019algebraic} as basis for our metatheory. \todo{REASONS, revamp}
 
@@
 as $(x : A)\ra B$, and dependent pairs as $(x : A)\times B$ with projections
-$\proj1$ and $\proj2$. We also use subscripts as a field projection notation for
-iterated pairs. For example, for $t : (A : \Set_i) \times (B : \Set_i) \times (f
-: A \ra B)$, we use $B_t$ to denote the projection of the second
-component. Sometimes we omit the subscript if it is clear from context.
+$\proj1$ and $\proj2$. We sometimes leave parameters implicit in dependent
+function types, e.g.\ write $\id : A \ra A$ instead of $\id : (A : \Set_i)\ra A
+\ra A$. We also use subscripts as a field projection notation for iterated
+pairs. For example, for $t : (A : \Set_i) \times (B : \Set_i) \times (f : A \ra
+B)$, we use $B_t$ to denote the projection of the second component. Sometimes we
+omit the subscript if it is clear from context. When we write ``exists'' in this
+paper, we always mean chosen structure given by a $\Sigma$-type.
 
@@
 $\blank=\blank$. The unit type is $\top$, with inhabitant $\tt$.
+
+\todo{compactify cumulativity}
 
@@
   \item Identity type $\Id : (A : \Ty\,\Gamma)\ra \Tm\,\Gamma\,A\ra
-    \Tm\,\Gamma\,A\ra \Ty\,\Gamma$, with equality reflection and UIP.
+    \Tm\,\Gamma\,A\ra \Ty\,\Gamma$, with $\refl$, equality reflection and UIP.
   \item Constant families. This includes a type former $\Kfam : \Con \ra
-    \Ty\,\Gamma$, where $\Gamma$ is implicitly quantified, and an isomorphism
-    between $\Sub\,\Gamma\,\Delta$ and $\Tm\,\Gamma\,(\Kfam\,\Delta)$ which is
-    natural in $\Gamma$. We have $\lamK : \Sub\,\Gamma\,\Delta \ra
-    \Tm\,\Gamma\,(\Kfam\,\Delta)$ and $\appK : \Tm\,\Gamma\,(\Kfam\,\Delta)
-    \ra \Sub\,\Gamma\,\Delta$ as its components.  The idea is that
-    $\Kfam\,\Delta$ is a representation of $\Delta$ as a type in any
-    context. Clairambault and Dybjer called constant families ``democracy'' in
-    \cite{clairambault2014biequivalence}.
+    \Ty\,\Gamma$, where $\Gamma$ is implicitly quantified, together with $\lamK
+    : \Sub\,\Gamma\,\Delta \ra \Tm\,\Gamma\,(\Kfam\,\Delta)$ and its inverse
+    $\appK$. The idea is that $\Kfam\,\Delta$ is a representation of $\Delta$ as
+    a type in any context. Clairambault and Dybjer called constant families
+    ``democracy'' in \cite{clairambault2014biequivalence}.
 \end{enumerate}
@@
 and Dybjer showed that the 2-category of flCwFs is biequivalent to the
-2-category of finitely complete (or ``lex'') categories
-\cite{clairambault2014biequivalence}. It can be also seen that the categorical
-product of $\Gamma$ and $\Delta$ in an flCwF can be given as $\Gamma \ext
-\Kfam\, \Delta$, and equalizers can be given using the identity type.
+2-category of finitely complete categories
+\cite{clairambault2014biequivalence}. In particular, in an flCwF the categorical
+product of $\Gamma$ and $\Delta$ can be given as $\Gamma \ext \Kfam\, \Delta$,
+and the equalizer of $\sigma$ and $\delta$ as $\Gamma\ext
+\Id\,(\K\,\Delta)\,(\lamK\,\sigma)\,(\lamK\,\delta)$.
 
@@
 
-\begin{definition}[Type categories, c.f.\ \cite{clairambault2014biequivalence}] For each $\Gamma : \Con$, there is a category
+\begin{definition}[Type categories, c.f.\ \cite{clairambault2014biequivalence}]\label{def:type_categories} For each $\Gamma : \Con$, there is a category
 whose objects are types $A : \Ty\,\Gamma$, and morphisms from $A$ to $B$ are
@@
 \Tm\,(\Gamma\,\ext\,A)\,A[\p]$, and composition $t \circ u$ by $t[\p, u]$. The
-assignment of type categories to contexts extends to an indexed category. For
+assignment of type categories to contexts extends to a split indexed category. For
 each $\sigma : \Sub\,\Gamma\,\Delta$, there is a functor from $\Ty\,\Delta$ to
@@
 
-\begin{definition} A weak flCwF morphism $\boldsymbol{\sigma : \Sub\,\Gamma\,\Delta}$ is a functor between underlying categories, which additionally maps types to types and terms to terms, and satisfies the following:
+\begin{definition} A \emph{weak flCwF morphism} $\boldsymbol{\sigma : \Sub\,\Gamma\,\Delta}$ is a functor between underlying categories, which additionally maps types to types and terms to terms, and satisfies the following:
   \begin{enumerate}
@@
 :
-\bsigma\,(\Gamma\,\ext\,A)\,\simeq\,\bsigma\,\Gamma\,\ext\,\bsigma\,A$.
+\bsigma\,(\Gamma\,\ext\,A)\,\simeq\,\bsigma\,\Gamma\,\ext\,\bsigma\,A$. Our notion of weak morphism is the same as in \cite{birkedal2018modal}, when restricted to CwFs.
 
@@
 
-\begin{theorem} Every $\boldsymbol{\sigma : \Sub\,\Gamma\,\Delta}$ preserves fl-structure up to type isomorphism. That is, we have isomorphisms
+\begin{theorem} Every $\boldsymbol{\sigma : \Sub\,\Gamma\,\Delta}$ preserves fl-structure up to type isomorphism. That is, we have
 \begin{alignat*}{3}
@@
 equality reflection, and the morphisms are inverses by UIP. We omit here the
-verification that $\bsigma$ preserves term and substitution formers.
+verification that $\bsigma$ preserves term and substitution formers in the
+fl-structure.
 \end{proof}
@@
 
-%% $\bid : \bSub\,\bGamma\,\bGamma$ is defined in the obvious way, with $\bid
-%% For composition
+$\bid : \bSub\,\bGamma\,\bGamma$ is defined in the obvious way, with identities for
+underlying functions and for preservation morphisms.
+
+For $\boldsymbol{\sigma \circ \delta}$, the underlying functions are given by
+function composition, and the preservation morphisms are given as follows:
+\begin{alignat*}{3}
+  & (\boldsymbol{\sigma \circ \delta})_{\emptycon}^{-1} :\equiv
+    \bsigma\,\bdelta_{\emptycon}^{-1} \circ \bdelta_{\emptycon}^{-1} \\
+  & (\boldsymbol{\sigma \circ \delta})_{\ext}^{-1} :\equiv
+    \bsigma\,\bdelta_{\ext}^{-1} \circ \bdelta_{\ext}^{-1}
+\end{alignat*}
+
+\begin{lemma} The derived preservation isomorphisms for the fl-structure can be decomposed analogously; all derived isomorphisms in $\bid$ are identities, and we have
+\begin{alignat*}{3}
+  & (\boldsymbol{\sigma \circ \delta})_{\Sigma} =
+  \bsigma\,\bdelta_{\Sigma} \circ \bdelta_{\Sigma}\\
+  & (\boldsymbol{\sigma \circ \delta})_{\K} =
+  \bsigma\,\bdelta_{\K} \circ \bdelta_{\K}\\
+  & (\boldsymbol{\sigma \circ \delta})_{\Id} =
+  \bsigma\,\bdelta_{\Id} \circ \bdelta_{\Id}
+\end{alignat*}
+Here, $\blank\circ\blank$ denotes composition of type morphisms.
+\end{lemma}
+\begin{proof} In the case of $\Id$, the equations hold immediately by UIP. For $\Sigma$ and $\K$, we prove by flCwF computation and straightforward unfolding of definitions.
+\end{proof}
+
+\subsection{Empty Context}
+The empty context $\boldsymbol{\emptycon : \Con}$ is the terminal flCwF, which
+has all underlying sets defined as $\top$ (or constantly $\top$), with an
+evident unique $\boldsymbol{\epsilon : \Sub\,\Gamma\,\emptycon}$. Since
+$\bepsilon$ is a strict flCwF morphism, $\bepsilon_{\emptycon}^{-1}$ and
+$\bepsilon_{\ext}^{-1}$ are both identity morphisms.
+
+\subsection{Types}
+
+We define $\boldsymbol{\Ty\,\Gamma} : \Set_{1 + \max(i, j+1)}$ as the type of
+split flCwF-isofibrations over $\bGamma$. We extend Ahrens' and Lumsdaine's
+displayed categories and their definition of isofibrations \cite{displayedcats}.
+We first define displayed flCwFs, then specify iso-cleaving as additional structure
+on top of that.
+
+\begin{definition}[Displayed flCwF] The type of displayed flCwFs is given as the logical predicate interpretation \cite{TODO} of $\flCwF_i$.
+For each flCwF component in $\bGamma$, there is a component in a displayed flCwF which ``lies over'' it.
+
+In situations where we need to refer to both ``base'' and displayed things, we give
+\ul{underlined} names to contexts, substitutions, types and terms in a base
+flCwF. For example, we may have $\ulGamma : \Con_{\bGamma}$ living in
+$\boldsymbol{\Gamma : \Con}$, and $\Gamma : \Con_{\bA}\,\ulGamma$ living in
+$\boldsymbol{A : \Ty\,\Gamma}$. We only use underlining on cETT variable names,
+and overload flCwF component names for displayed counterparts.
+
+Concretely, $\boldsymbol{A : \Ty\,\Gamma}$ has the following underlying sets, which we call displayed contexts, substitutions, types and terms respectively.
+\begin{alignat*}{3}
+  & \Con_{\bA} && : \Con_{\bGamma}\ra \Set_i\\
+  & \Sub_{\bA} && : \Con_{\bA}\,\ulGamma \ra \Con_{\bA}\,\ulDelta \ra \Sub_{\bGamma}\,\ulGamma\,\ulDelta \ra \Set_i \\
+  & \Ty_{\bA}  && : \Con_{\bA}\,\ulGamma \ra \Ty_{\bGamma}\,\ulGamma \ra \Set_i\\
+  & \Tm_{\bA}  && : (\Gamma : \Con_{\bA}\,\ulGamma)\ra \Ty_{\bA}\,\Gamma\,\ulA \ra \Tm_{\bGamma}\,\ulGamma\,\ulA \ra \Set_i
+\end{alignat*}
+
+Above, we implicitly quantify over $\ulGamma$, $\ulDelta$ and $\ulA$ base
+parameters. We also have the following components for empty context, context
+extension and substitution. We omit listing other components here.
+\begin{alignat*}{3}
+  & \emptycon_{\bA} && : \Con_{\bA}\,\emptycon_{\bGamma}\\
+  & \ext_{\bA}      && : (\Gamma : \Con_{\bA}\,\ulGamma)\ra \Ty_{\bA}\,\Gamma\,\ulA \ra
+                     \Con_{\bA}\,\Gamma\,(\ulGamma \ext_{\bGamma} \ulA)\\
+  & \blank[\blank]_{\bA} && : \Ty_{\bA}\,\Delta\,\ulA \ra \Sub_{\bA}\,\Gamma\,\Delta\,\ulsigma
+                     \ra \Ty_{\bA}\,\Gamma\, (\ulA[\ulsigma]_{\bGamma})\\
+  & \blank[\blank]_{\bA} && : \Tm_{\bA}\,\Delta\,A\,\ult \ra (\sigma : \Sub_{\bA}\,\Gamma\,\Delta\,\ulsigma)\\
+  & && \hspace{0.5em}\ra \Tm_{\bA}\,\Gamma\, (A[\sigma]_{\bA})\,(\ult[\ulsigma]_{\bGamma})
+\end{alignat*}
+\end{definition}
+
+In the following we will often omit $_{\bGamma}$ and $_{\bA}$ subscripts on
+components; for example, in the type $\Con_{\bA}\,\emptycon$, the $\emptycon$ is clearly
+a base component in $\bGamma$.
+
+We also need displayed counterparts to the previously defined derived notions on flCwFs;
+these are again given as logical predicate interpretations of the non-displayed definitions.
+
+\begin{definition}[Displayed type categories] For each $\Gamma : \Con_{\bA}\,\ulGamma$, there
+is a displayed category over the type category $\Ty_{\bGamma}\,\ulGamma$, whose objects
+over $\ulA : \Ty_{\bGamma}\,\ulGamma$ are elements of $\Ty_{\bA}\,\Gamma\,\ulA$, and displayed morphisms over $\ult : \Tm_{\bGamma}\,(\ulGamma \ext \ulA)\,(\ulB[\p])$ are elements of $\Tm_{\bA}\,(\Gamma \ext A)\,(B[\p])\,\ult$.
+The identity morphism is given by $\q_{\bA}$, and the composition of $t$ and $u$ is $t[\p_{\bA},u]$.
+Analogously to Definition \ref{def:type_categories}, this extends to a displayed split indexed category.
+\end{definition}
+
+\begin{definition}[Displayed isomorphisms] A \emph{displayed context isomorphism} over $\ulsigma : \ulGamma \simeq \ulDelta$, notated $\sigma : \Gamma \simeq_{\ulsigma} \Delta$, is an invertible displayed morphism $\sigma : \Sub_{\bA}\,\Gamma\,\Delta\,\ulsigma$, with inverse $\sigma^{-1} : \Sub_{\bA}\,\Delta\,\Gamma\,\ulsigma^{-1}$.
+A \emph{displayed type isomorphism} over $\ult : \ulA \simeq \ulB$, notated $t : A \simeq_{\ult} B$, is an isomorphism in a displayed type category.
+\end{definition}
+
+\begin{definition} A \emph{vertical morphism} lies over an identity morphism. We use this definition for context and type morphisms as well.
+\end{definition}
+
+In contrast to \cite{kaposi2019constructing}, it is not sufficient to model
+types with displayed flCwFs. In ibid.\ substitutions are modeled as strict
+morphisms, which we cannot do because our infinitary function types and identity
+types cannot be interpreted strictly, as we will see in Section
+\ref{TODO}. Strict preservation is expressed with metatheoretic equality, which
+is respected by all cETT constructions, but isomorphisms are not necessarily
+respected. Hence, we need to add additional structure to displayed flCwFs which
+expresses preservation of base isomorphisms.
+
+\begin{definition}[Context iso-cleaving] This lifts a base context isomorphism to a displyed one. It consists of
+\begin{alignat*}{3}
+  & \coe : \ulGamma \simeq \ulDelta \ra \Con_{\bA}\,\ulGamma \ra \Con_{\bA}\,\ulDelta\\
+  & \coh : (\ulsigma : \ulGamma \simeq \ulDelta)(\Gamma : \Con_{\bA}\,\ulGamma)
+           \ra \Gamma \simeq_{\ulsigma} \coe\,\ulsigma\,\Gamma
+\end{alignat*}
+such that
+\begin{alignat*}{3}
+  & \coe\,\id\,\Gamma && = \Gamma\\
+  & \coe\,(\ulsigma\circ\uldelta)\,\Gamma && = \coe\,\ulsigma\,(\coe\,\uldelta\,\Gamma)\\
+  & \coh\,\id\,\Gamma && = \id\\
+  & \coh\,(\ulsigma\circ\uldelta)\,\Gamma && = \coh\,\ulsigma\,(\coe\,\uldelta\,\Gamma)
+          \circ \coh\,\uldelta\,\Gamma
+\end{alignat*}
+Here, $\coe$ and $\coh$ abbreviate ``coercion'' and ``coherence'' respectively.
+\end{definition}
+
+\begin{definition}[Type iso-cleaving] This consists of
+\begin{alignat*}{3}
+  & \coe : \ulA \simeq \ulB \ra \Ty_{\bA}\,\Gamma\,\ulA \ra \Ty_{\bA}\,\Gamma\,\ulB\\
+  & \coh : (\ult : \ulA \simeq \ulB)(A : \Ty_{\bA}\,\Gamma\,\ulA)
+           \ra A \simeq_{\ult} \coe\,\ult\,A
+\end{alignat*}
+such that
+\begin{alignat*}{3}
+  & \coe\,\id\,A && = A\\
+  & \coe\,\ult\,(\coe\,\uldelta\,A) && = \coe\,(\ult\circ\uldelta)\,A\\
+  & \coh\,\id\,A && = \id\\
+  & \coh\,(\ult\circ\uldelta)\,A && = \coh\,\ult\,(\coe\,\uldelta\,A)
+          \circ \coh\,\uldelta\,A
+\end{alignat*}
+Also, whenever $\sigma : \Sub_{\bA}\,\Gamma\,\Delta\,\ulsigma$, we have
+\begin{alignat*}{3}
+  & \coe\,(\ult[\ulsigma])\,(A[\sigma]) && = (\coe\,\ult\,A)[\sigma]\\
+  & \coh\,(\ult[\ulsigma\circ \p,\q])\,(A[\sigma]) && = (\coh\,\ult\,A)[\sigma]
+\end{alignat*}
+
+\end{definition}
+
+\begin{definition} A \emph{split flCwF isofibration} is a displayed flCwF equipped with iso-cleaving for contexts and types.
+\end{definition}
+
+\subsection{Type Substitution}
+
+We seek to define $\boldsymbol{\blank[\blank] : \Ty\,\Delta \ra \Sub\,\Gamma\,\Delta \ra \Ty\,\Gamma}$.
 

### 2019-12-29 draft/paper.tex
paper

@@
 
-%% \newcommand{\bDelta}{\boldsymbol{\Delta}}
-
 %% ------------------------------------------------------------
@@
 
-\begin{definition} A \emph{weak flCwF morphism} $\boldsymbol{\sigma : \Sub\,\Gamma\,\Delta}$ is a functor between underlying categories, which additionally maps types to types and terms to terms, and satisfies the following:
+\begin{definition}\label{def:weakmorphism} A \emph{weak flCwF morphism} $\boldsymbol{\sigma : \Sub\,\Gamma\,\Delta}$ is a functor between underlying categories, which additionally maps types to types and terms to terms, and satisfies the following:
   \begin{enumerate}
@@
 
-%% We do not require preservation of fl-structure, because the above conditions already insure that.
-
-\begin{theorem} Every $\boldsymbol{\sigma : \Sub\,\Gamma\,\Delta}$ preserves fl-structure up to type isomorphism. That is, we have
+\begin{theorem}\label{thm:flpres} Every $\boldsymbol{\sigma : \Sub\,\Gamma\,\Delta}$ preserves fl-structure up to type isomorphism. That is, we have
 \begin{alignat*}{3}
@@
 \end{alignat*}
+These are all natural in the following sense: for $\sigma :
+\Sub_{\bGamma}\,\Gamma\,\Delta$, the functorial action of $\bsigma\,\sigma :
+\Sub_{\bDelta}\,(\bsigma\,\Gamma)\,(\bsigma\,\Delta)$ on $\bsigma_{\Sigma}$ (in
+the $\bsigma\,\Gamma$ context) is equal to $\bsigma_{\Sigma}$ (in
+$\bsigma\,\Delta$), and similarly for $\bsigma_{\K}$ and $\bsigma_{\Id}$.
+
 Moreover, $\bsigma$ preserves all term and substitution formers in the
@@
 equality reflection, and the morphisms are inverses by UIP. We omit here the
-verification that $\bsigma$ preserves term and substitution formers in the
-fl-structure.
+verification of naturality and that $\bsigma$ preserves term and substitution
+formers in the fl-structure.
 \end{proof}
@@
 
-\begin{lemma} The derived preservation isomorphisms for the fl-structure can be decomposed analogously; all derived isomorphisms in $\bid$ are identities, and we have
+It is easy to verify the left and right identity laws and associativity for $\boldsymbol{\blank\circ\blank}$.
+
+\begin{lemma}\label{lem:idcomppres} The derived preservation isomorphisms for the fl-structure can be decomposed analogously; all derived isomorphisms in $\bid$ are identities, and we have
 \begin{alignat*}{3}
@@
 \end{alignat*}
-Here, $\blank\circ\blank$ denotes composition of type morphisms.
+On the right sides, $\blank\circ\blank$ refers to composition of type morphisms.
 \end{lemma}
@@
 
-\begin{definition}[Displayed flCwF] The type of displayed flCwFs is given as the logical predicate interpretation \cite{TODO} of $\flCwF_i$.
+\begin{definition}[Displayed flCwF] The type of displayed flCwFs at level $i$ is given as the logical predicate interpretation \cite{TODO} of $\flCwF_i$.
 For each flCwF component in $\bGamma$, there is a component in a displayed flCwF which ``lies over'' it.
@@
 
-Concretely, $\boldsymbol{A : \Ty\,\Gamma}$ has the following underlying sets, which we call displayed contexts, substitutions, types and terms respectively.
+Concretely, a displayed flCwF $\bA$ over $\bGamma$ has the following underlying sets, which we call displayed contexts, substitutions, types and terms respectively.
 \begin{alignat*}{3}
@@
 
-\begin{definition}[Context iso-cleaving] This lifts a base context isomorphism to a displyed one. It consists of
-\begin{alignat*}{3}
-  & \coe : \ulGamma \simeq \ulDelta \ra \Con_{\bA}\,\ulGamma \ra \Con_{\bA}\,\ulDelta\\
-  & \coh : (\ulsigma : \ulGamma \simeq \ulDelta)(\Gamma : \Con_{\bA}\,\ulGamma)
-           \ra \Gamma \simeq_{\ulsigma} \coe\,\ulsigma\,\Gamma
-\end{alignat*}
-such that
-\begin{alignat*}{3}
-  & \coe\,\id\,\Gamma && = \Gamma\\
-  & \coe\,(\ulsigma\circ\uldelta)\,\Gamma && = \coe\,\ulsigma\,(\coe\,\uldelta\,\Gamma)\\
-  & \coh\,\id\,\Gamma && = \id\\
-  & \coh\,(\ulsigma\circ\uldelta)\,\Gamma && = \coh\,\ulsigma\,(\coe\,\uldelta\,\Gamma)
+\begin{definition}[Context iso-cleaving] This lifts a base context isomorphism to a displayed one. It consists of
+\begin{alignat*}{3}
+  & \coe &&: \ulGamma \simeq \ulDelta \ra \Con_{\bA}\,\ulGamma \ra \Con_{\bA}\,\ulDelta\\
+  & \coh &&: (\ulsigma : \ulGamma \simeq \ulDelta)(\Gamma : \Con_{\bA}\,\ulGamma)
+           \ra \Gamma \simeq_{\ulsigma} \coe\,\ulsigma\,\Gamma\\
+  & \coe_{\id} && : \coe\,\id\,\Gamma = \Gamma\\
+  & \coe_{\circ} && : \coe\,(\ulsigma\circ\uldelta)\,\Gamma = \coe\,\ulsigma\,(\coe\,\uldelta\,\Gamma)\\
+  & \coh_{\id} && : \coh\,\id\,\Gamma = \id\\
+  & \coh_{\circ} && : \coh\,(\ulsigma\circ\uldelta)\,\Gamma = \coh\,\ulsigma\,(\coe\,\uldelta\,\Gamma)
           \circ \coh\,\uldelta\,\Gamma
@@
 \begin{alignat*}{3}
-  & \coe : \ulA \simeq \ulB \ra \Ty_{\bA}\,\Gamma\,\ulA \ra \Ty_{\bA}\,\Gamma\,\ulB\\
-  & \coh : (\ult : \ulA \simeq \ulB)(A : \Ty_{\bA}\,\Gamma\,\ulA)
-           \ra A \simeq_{\ult} \coe\,\ult\,A
-\end{alignat*}
-such that
-\begin{alignat*}{3}
-  & \coe\,\id\,A && = A\\
-  & \coe\,\ult\,(\coe\,\uldelta\,A) && = \coe\,(\ult\circ\uldelta)\,A\\
-  & \coh\,\id\,A && = \id\\
-  & \coh\,(\ult\circ\uldelta)\,A && = \coh\,\ult\,(\coe\,\uldelta\,A)
+  & \coe &&: \ulA \simeq \ulB \ra \Ty_{\bA}\,\Gamma\,\ulA \ra \Ty_{\bA}\,\Gamma\,\ulB\\
+  & \coh &&: (\ult : \ulA \simeq \ulB)(A : \Ty_{\bA}\,\Gamma\,\ulA)
+           \ra A \simeq_{\ult} \coe\,\ult\,A\\
+  & \coe_{\id} && : \coe\,\id\,A = A\\
+  & \coe_{\circ} && : \coe\,\ult\,(\coe\,\uldelta\,A) = \coe\,(\ult\circ\uldelta)\,A\\
+  & \coh_{\id} &&: \coh\,\id\,A = \id\\
+  & \coh_{\circ} &&: \coh\,(\ult\circ\uldelta)\,A = \coh\,\ult\,(\coe\,\uldelta\,A)
           \circ \coh\,\uldelta\,A
 \end{alignat*}
-Also, whenever $\sigma : \Sub_{\bA}\,\Gamma\,\Delta\,\ulsigma$, we have
-\begin{alignat*}{3}
-  & \coe\,(\ult[\ulsigma])\,(A[\sigma]) && = (\coe\,\ult\,A)[\sigma]\\
-  & \coh\,(\ult[\ulsigma\circ \p,\q])\,(A[\sigma]) && = (\coh\,\ult\,A)[\sigma]
+Additionally, for $\sigma : \Sub_{\bA}\,\Gamma\,\Delta\,\ulsigma$, we have
+\begin{alignat*}{3}
+  & \coe_{[]} &&: \coe\,(\ult[\ulsigma])\,(A[\sigma]) = (\coe\,\ult\,A)[\sigma]\\
+  & \coh_{[]} &&: \coh\,(\ult[\ulsigma\circ \p,\q])\,(A[\sigma]) = (\coh\,\ult\,A)[\sigma]
 \end{alignat*}
@@
 \subsection{Type Substitution}
-
-We seek to define $\boldsymbol{\blank[\blank] : \Ty\,\Delta \ra \Sub\,\Gamma\,\Delta \ra \Ty\,\Gamma}$.
+We aim to define $\boldsymbol{\blank[\blank] : \Ty\,\Delta \ra
+  \Sub\,\Gamma\,\Delta \ra \Ty\,\Gamma}$, such that $\boldsymbol{A[\id]} = \bA$
+and $\boldsymbol{A[\sigma\circ\delta]} = \boldsymbol{A[\sigma][\delta]}$. This
+is rather laborious to fully write out, so we omit most details and focus on the
+main points.
+
+The underlying sets are given by simple composition:
+\begin{alignat*}{3}
+  & \Con_{\boldsymbol{A[\sigma]}}\,\ulGamma && :\equiv \Con_{\bA}\,(\bsigma\,\ulGamma)\\
+  & \Sub_{\boldsymbol{A[\sigma]}}\,\Gamma\,\Delta\,\ulsigma && :\equiv
+    \Sub_{\bA}\,\Gamma\,\Delta\,(\bsigma\,\ulsigma)\\
+  & \Ty_{\boldsymbol{A[\sigma]}}\,\Gamma\,\ulA && :\equiv
+      \Ty_{\bA}\,\Gamma\,(\bsigma\,\ulA)\\
+  & \Tm_{\boldsymbol{A[\sigma]}}\,\Gamma\,A\,\ult && :\equiv
+      \Tm_{\bA}\,\Gamma\,A\,(\bsigma\,\ult)
+\end{alignat*}
+
+Moreover, $\id_{\boldsymbol{A[\sigma]}} :\equiv \id_{\bA}$,
+$\sigma\circ_{\boldsymbol{A[\sigma]}}\delta :\equiv \sigma\circ_{\bA}\delta$,
+and likewise substitution components are given by corresponding substitution
+components in $\bA$.
+
+Context and type formers are given by coercing $\bA$
+structures along $\bsigma$ preservation isomorphisms. For example:
+\begin{alignat*}{3}
+  &\emptycon_{\boldsymbol{A[\sigma]}} && :\equiv
+    \coe\,\bsigma_{\emptycon}^{-1}\,\emptycon_{\bA}\\
+  &\Gamma\ext_{\boldsymbol{A[\sigma]}}A && :\equiv
+    \coe\,\bsigma_{\ext}^{-1}\,(\Gamma\ext_{\bA} A)\\
+  &\Id_{\boldsymbol{A[\sigma]}}\,t\,u && :\equiv
+    \coe\,\bsigma_{\Id}^{-1}\,(\Id_{\bA}\,t\,u)
+\end{alignat*}
+
+Term and substitution formers are given by composing $\coh$-lifted
+isomorphisms with term and substitution formers from $\bA$. For example:
+\begin{alignat*}{3}
+  & \epsilon_{\boldsymbol{A[\sigma]}} && :\equiv
+    \coh\,\bsigma_{\emptycon}^{-1}\,\emptycon_{\bA} \circ \epsilon_{\bA}\\
+  & \p_{\boldsymbol{A[\sigma]}} && :\equiv
+    \p_{\bA} \circ (\coh\,\bsigma_{\ext}^{-1}\,(\Gamma\ext A))^{-1}\\
+  & \appK_{\boldsymbol{A[\sigma]}}\,t && :\equiv
+    \appK_{\bA}\,((\coh\,\bsigma_{\K}\,(\K\,\Delta))^{-1}\circ t)
+\end{alignat*}
+Equations for term and type substitution follow from naturality of preservation
+isomorphisms in $\bsigma$, $\coe[]$, $\coh[]$ and substitution equations in
+$\bA$.
+
+\todo{iso-cleaving}
+
+Functoriality of type substitution, i.e.\ $\boldsymbol{A[\id]} = \bA$ and
+$\boldsymbol{A[\sigma\circ\delta]} = \boldsymbol{A[\sigma][\delta]}$, follow
+from Lemma \ref{lem:idcomppres} and split cleaving given by $\coe_{\id}$,
+$\coe_{\circ}$, $\coh_{\id}$ and $\coh_{\circ}$ in $\bA$.
+
+\subsection{Terms}
+
+$\boldsymbol{\Tm\,\Gamma\,A} : \Set_{1 + \max(i, j)}$ is defined as the type of \emph{weak flCwF sections} of $\bA$.
+
+\begin{definition}[Weak flCwF section] Given $\bA : \bTy\,\bGamma$, the underlying functions of $\bt : \bTm\,\bGamma\,\bA$ are as follows:
+\begin{alignat*}{3}
+  & \bt : (\ulGamma : \Con_{\bGamma}) \ra \Con_{\bA}\,\ulGamma\\
+  & \bt : (\ulsigma : \Sub_{\bGamma}\,\ulGamma\,\ulDelta)
+         \ra \Sub_{\bA}\,(\bt\,\ulGamma)\,(\bt\,\ulDelta)\,\ulsigma\\
+  & \bt : (\ulA : \Ty_{\bGamma}) \ra \Ty_{\bA}\,(\bt\,\ulGamma)\,\ulA\\
+  & \bt : (\ult : \Tm_{\bGamma}\,\ulGamma\,\ulA) \ra
+          \Tm_{\bA}\,(\bt\,\ulGamma)\,(\bt\,\ulA)\,\ult
+\end{alignat*}
+Such that
+\begin{enumerate}
+  \item $\bsigma\,(\ulA[\ulsigma]) = (\bsigma\,\ulA)\,[\bsigma\,\ulsigma]$
+  \item $\bsigma\,(\ult[\ulsigma]) = (\bsigma\,\ult)\,[\bsigma\,\ulsigma]$
+  \item The unique map $\epsilon_{\bA} : \Sub\,(\bsigma\,\emptycon)\,\emptycon_{\bGamma}\,\id$ has a vertical retraction.
+  \item Each $(\bsigma\,\p,\,\bsigma\,\q) : \Sub\,(\bsigma\,(\ulGamma\,\ext\,\ulA))\,(\bsigma\,\ulGamma\,\ext\,\bsigma\,\ulA)\,\id$ has a vertical inverse.
+\end{enumerate}
+Similarly to Definition \ref{def:weakmorphism}, we denote the evident preservation
+isomorphisms as $\bt_{\emptycon} : \bt\,\emptycon \simeq_{\id} \emptycon$ and
+$\bt_{\ext} : \bt\,(\ulGamma\ext \ulA) \simeq_{\id} \bt\,\ulGamma \ext
+\bt\,\ulA$. In short, weak section is a dependently typed analogue of weak
+morphism, with dependent underlying functions and displayed preservation
+isomorphisms. We also have the derived fl-preservation isomorphisms.
+\end{definition}
+
+\begin{theorem} A weak section $\boldsymbol{t : \Tm\,\Gamma\,A}$ preserves fl-structure up to vertical type isomorphisms, that is, the following are derivable:
+\begin{alignat*}{3}
+  & \bt_{\Sigma} : \bt\,(\Sigma\,\ulA\,\ulB) \simeq_{\id} \Sigma\,(\bt\,\ulA)\,((\bt\,\ulB)[\bt_{\ext}^{-1}]) \\
+  & \bt_{\K} : \bt\,(\K\,\ulDelta) \simeq_{\id} \K\,(\bt\,\ulDelta) \\
+  & \bt_{\Id} : \bt\,(\Id\,\ult\,\ulu) \simeq_{\id} \Id\,(\bt\,\ult)\,(\bt\,\ulu)
+\end{alignat*}
+Also, the above isomorphisms are natural in the sense of Theorem
+\ref{thm:flpres}, and $\bt$ preserves type and substitution formers in the
+fl-structure.
+\end{theorem}
+\begin{proof} The construction of isomorphisms is the same as in Theorem \ref{thm:flpres}. Indeed, every construction there has a displayed counterpart which we can use here.
+\end{proof}
+
+We note though that the move from Theorem \ref{thm:flpres} to here is not simply a
+logical predicate translation, because we are only lifting the codomain of a
+weak morphism to a displayed version, and we leave the domain non-displayed. We
+leave to future work the investigation of such asymmetrical (or
+``modal'') logical predicate translations.
+
+\subsection{Term Substitution}
+
+$\boldsymbol{\blank[\blank] : \Tm\,\Delta\,A \ra (\sigma : \Sub\,\Gamma\,\Delta)
+  \ra \Tm\,\Gamma\,(A[\sigma])}$ is given the similarly to
+$\boldsymbol{\blank\circ\blank}$ in Section \ref{TODO}. Underlying functions
+are given by function composition, and preservation morphisms are also similar:
+\begin{alignat*}{3}
+  & (\boldsymbol{t[\sigma]})_{\emptycon}^{-1} :\equiv
+    \bt\,\bsigma_{\emptycon}^{-1} \circ \bt_{\emptycon}^{-1} \\
+  & (\boldsymbol{t[\sigma]})_{\ext}^{-1} :\equiv
+    \bt\,\bsigma_{\ext}^{-1} \circ \bt_{\ext}^{-1}
+\end{alignat*}
+
+We also have the same decomposition of derived isomorphisms as in Lemma
+\ref{lem:idcomppres}. We do not have to show functoriality of term substitution
+here, since that is derivable in any CwF \cite{TODO}.
+
+\subsection{Context Extension and Comprehension}
+
+$\boldsymbol{\Gamma \ext A : \Con}$ is defined as the \emph{total flCwF} of
+$\bA$. This is given by bundling together all displayed flCwF components in
+$\bA$ with corresponding base components in $\bGamma$, using metatheoretic
+$\Sigma$-types. It is a straightforward extension of total categories in
+\cite{displayedcats}.
+
+$\boldsymbol{\p : \Sub\,(\Gamma\ext A)\,\Gamma}$ is a strict morphism given by
+taking a first projection for each component. $\boldsymbol{\q : \Tm\,(\Gamma\ext
+  A)\,(A[\p])}$ is likewise a strict flCwF section given by second projections.
 

### 2019-12-30 draft/paper.tex
paper

@@
 \newcommand{\lami}{\mathsf{lam}}
-\newcommand{\Pie}{\hat{\Pi}}
-\newcommand{\appe}{\mathsf{ap\hat{p}}}
-\newcommand{\lame}{\mathsf{la\hat{m}}}
-\newcommand{\Piinf}{\tilde{\Pi}}
-\newcommand{\appinf}{\mathsf{ap\tilde{p}}}
-\newcommand{\laminf}{\mathsf{la\tilde{m}}}
+\newcommand{\Pie}{\Pi^{\mathsf{ext}}}
+\newcommand{\appe}{\mathsf{app^{ext}}}
+\newcommand{\lame}{\mathsf{lam^{ext}}}
+\newcommand{\Piinf}{\Pi^{\mathsf{inf}}}
+\newcommand{\appinf}{\mathsf{app^{inf}}}
+\newcommand{\laminf}{\mathsf{lam^{inf}}}
 \newcommand{\appitt}{\mathop{{\scriptstyle @}}}
@@
 
+\newcommand{\arri}{\Rightarrow}
+\newcommand{\arre}{\Rightarrow^{\mathsf{ext}}}
+\newcommand{\arrinf}{\Rightarrow^{\mathsf{inf}}}
+
 \newcommand{\bCon}{\boldsymbol{\Con}}
@@
 \newcommand{\bA}{\boldsymbol{A}}
+\newcommand{\ba}{\boldsymbol{a}}
 \newcommand{\bB}{\boldsymbol{B}}
@@
 \newcommand{\bemptycon}{\boldsymbol{\emptycon}}
+\newcommand{\bSet}{\boldsymbol{\Set}}
+\newcommand{\bU}{\boldsymbol{\U}}
+\newcommand{\bEl}{\boldsymbol{\El}}
+\newcommand{\bPii}{\boldsymbol{\Pi}}
+\newcommand{\bappi}{\boldsymbol{\mathsf{app}}}
+\newcommand{\blami}{\boldsymbol{\mathsf{lam}}}
 
@@
 signature without referring to QIITs, only using features present in cETT. As a
-first step, we define a notion of model. $\Pi$
+first step, we define a notion of model.
 
@@
 substitution extension is $\blank,\blank : (\sigma : \Sub\,\Gamma\,\Delta)\ra
-\Tm\,\Gamma\,A[\sigma]\ra \Sub\,\Gamma\,(\Delta\ext A)$.
+\Tm\,\Gamma\,(A[\sigma])\ra \Sub\,\Gamma\,(\Delta\ext A)$.
 
@@
 
-\textbf{Inductive function space}, with $\Pii : (a : \Tm\,\Gamma\,\U) \ra
-\Ty\,(\Gamma\,\ext\,\El\,a) \ra \Ty\,\Gamma$ and application as $\appi :
+\textbf{Inductive function space} $\Pii : (a : \Tm\,\Gamma\,\U) \ra
+\Ty\,(\Gamma\,\ext\,\El\,a) \ra \Ty\,\Gamma$, with application as $\appi :
 \Tm\,\Gamma\,(\Pii\,a\,B)\ra \Tm\,(\Gamma\,\ext\,\El\,a)\,B$ and its
@@
 
-\textbf{External function space} with $\Set_i$ domain: $\Pie : (A : \Set_i)\ra(A
-\ra \Ty\,\Gamma)\ra \Ty\,\Gamma$, with $\appe : \Tm\,\Gamma\,(\Pie\,A\,B)\ra((x
-: A)\ra \Tm\,\Gamma\,(B\,x))$ and its inverse $\lame$.
-
-\textbf{Infinitary function space}, with $\Piinf : (A : \Set_i)\ra(A \ra
-\Tm\,\Gamma\,\U)\ra \Tm\,\Gamma\,\U$, $\appinf :
+\textbf{External function space} $\Pie : (A : \Set_i)\ra(A \ra \Ty\,\Gamma)\ra
+\Ty\,\Gamma$, with $\appe : \Tm\,\Gamma\,(\Pie\,A\,B)\ra((x : A)\ra
+\Tm\,\Gamma\,(B\,x))$ and its inverse $\lame$.
+
+\textbf{Infinitary function space} $\Piinf : (A : \Set_i)\ra(A \ra
+\Tm\,\Gamma\,\U)\ra \Tm\,\Gamma\,\U$, with $\appinf :
 \Tm\,\Gamma\,(\El\,(\Pie\,A\,b))\ra((x : A)\ra \Tm\,\Gamma\,(\El\,(b\,x)))$ and
-$\laminf$.
+its inverse $\laminf$.
 
@@
 \Tm\,\Gamma\,(\El\,a)\ra\Tm\,\Gamma\,(\El\,a)\ra \Tm\,\Gamma\,\U$, with $\Refl :
-(T : \Tm\,\Gamma\,(\El\,a))\ra \Id\,a\,t\,t$, equality reflection and UIP.
+(t : \Tm\,\Gamma\,(\El\,a))\ra \Id\,a\,t\,t$, equality reflection and UIP.
 
@@
 \end{alignat*}
+We abbreviate non-dependent inductive $\Pii$ as $\blank\arri\blank$, and likewise we use $\blank\arre\blank$ and $\blank\arrinf\blank$.
+
+%% We use abbreviations for non-dependent function types:
+%% \begin{alignat*}{3}
+%%   & A\arri B && :\equiv \Pi\,A\,(B[\p])\\
+%%   & A \arre\,B &&:\equiv \Pie\,A\,(\lambda\,\_.\,B) \\
+%%   & A \arrinf\,B &&:\equiv \Piinf\,A\,(\lambda\,\_.\,B)
+%% $A\arri B :\equiv \Pi\,A\,(B[\p])$, and $A \arre\,B :\equiv \Pie\,A\,(\lambda\,\_.\,B)$, and likewise
+%% $A \arrinf\,B :\equiv \Piinf\,A\,(\lambda\,\_.\,B)$.
 
@@
   & \hspace{1em}(\emptycon_M\,\ext_M\, (N : \U_M) \,\ext_M\,(zero : \El_M\,N)\\
-  & \hspace{2.67em}\ext_M (suc : \Pii_M\,N\,(\El_M\,N)))
+  & \hspace{2.67em}\ext_M (suc : N\arri_M\El_M\,N))
 \end{alignat*}
@@
 
+In the following examples, we leave the abstracted $M : \ToS_{0,j}$ implicit.
+
 \begin{example}{Infinitary constructors}. The universe $\U$ is closed under
 the $\Piinf$ function type, which allows such functions to appear in the domains
-of $\Pii$ types. This allows a signature for trees branching with arbitrary
-small sets. This is a signature at level 1, since we have $\Set_0$ as a $\Pie$
-domain type.
-\[
-TreeSig :\equiv
-\emptycon \ext (Tree : \U) \ext (node : \Pie\,\Set_0\,\lambda A.\,\Pii\,(\Piinf\,A\,\lambda \_.\,Tree)\,(\El\,Tree))
-\]
+of $\Pii$ types. This allows, for example, a signature for trees branching with
+arbitrary small sets. This is a signature at level 1, since we have $\Set_0$ as
+a $\Pie$ domain type.
+\begin{alignat*}{3}
+& TreeSig :\equiv \\
+& \hspace{1em}\emptycon \ext (Tree : \U)\\
+& \hspace{1.35em}\ext (node : \Pie\,\Set_0\,(\lambda A.\,(A\,\arrinf\,Tree)\arri\El\,Tree))
+\end{alignat*}
 \end{example}
@@
 in $\Pii$ domains. A minimal example:
-\[
-RecEqSig :\equiv \emptycon \ext (A : \U) \ext (a : \El\,A) \ext (f : \Pii\,(x : A)\,(\Pii\,(\Id\,A\,x\,a)\,(\El\,A)))
-\]
+\begin{alignat*}{3}
+  & RecEqSig :\equiv\\
+  & \hspace{0.5em}\emptycon\,\ext (A : \U) \ext (a : \El\,A)\ext (f : \Pii\,(x : A)\,(\Id\,A\,x\,a\arri\El\,A))
+\end{alignat*}
 More interesting (and complicated) examples are boundary conditions in various
@@
 
-We make a similar choice in the interpretation of signatures themselves: we
-use structured CwFs instead of lex categories to model notions of induction. The
-reason here is that CwFs allow us to define induction in strictly the same way as
-one would write it down in type theory, since we have $\Ty$ and $\Tm$ for a
-primitive notion of dependent objects and morphisms. In contrast, dependent
-objects in lex categories is a derived notion, and the induction principles we
-get are only up to isomorphisms. This issue is perhaps not relevant from a
-purely categorical perspective, but we are concerned with eventually
-implementing QIITs in proof assistants, so we prefer if our semantics computes
-strictly.
+We make a similar choice in the interpretation of signatures themselves: we use
+structured CwFs instead of lex categories to model notions of induction. The
+reason here is that CwFs allow us to compute induction principles in strictly
+the same way as one would write them down in type theory, since we have $\Ty$
+and $\Tm$ for a primitive notion of dependent objects and morphisms. In
+contrast, dependent objects in lex categories is a derived notion, and the
+induction principles we get are only up to isomorphism. This issue is perhaps
+not relevant from a purely categorical perspective, but we are concerned with
+eventually implementing QIITs in proof assistants, so we prefer if our semantics
+computes strictly.
 
@@
 
+The model involves a large amount of technical detail, most of which we omit,
+and only present the most salient parts.
+
 \subsection{Contexts}
 
-We define $\bCon : \Set_{1+\max(i,j+1)}$ as $\flCwF_{\max(i,j+1)}$. We define
-$\flCwF$ as follows.
+We define $\bCon : \Set_{1+\max(i,j+1)}$ as $\flCwF_{\max(i,j+1)}$.
+%% We define
+%% $\flCwF$ as follows.
 
@@
 \begin{enumerate}
-  \item A CwF with underlying sets all in $\Set_i$. We reuse the notation
-    from definition \ref{def:tos}.
+  \item A CwF with underlying sets all in $\Set_i$. We reuse the component
+    notations from Definition \ref{def:tos}.
   \item $\Sigma$-type $\Sg : (A : \Ty\,\Gamma)\ra \Ty\,(\Gamma \ext A)
@@
 
-We define $\boldsymbol{\Sub\,\Gamma\,\Delta} : \Set_{1 + \max(i, j+1)}$ as the type
-of weak flCwF morphisms from $\boldsymbol{\Gamma}$ to $\boldsymbol{\Delta}$.
+%% We define $\boldsymbol{\Sub\,\Gamma\,\Delta} : \Set_{1 + \max(i, j+1)}$ as the type
+%% of weak flCwF morphisms from $\boldsymbol{\Gamma}$ to $\boldsymbol{\Delta}$.
 
@@
   \Sub\,\Gamma\,\Delta \ra \Ty\,\Gamma}$, such that $\boldsymbol{A[\id]} = \bA$
-and $\boldsymbol{A[\sigma\circ\delta]} = \boldsymbol{A[\sigma][\delta]}$. This
-is rather laborious to fully write out, so we omit most details and focus on the
-main points.
-
-The underlying sets are given by simple composition:
+and $\boldsymbol{A[\sigma\circ\delta]} = \boldsymbol{A[\sigma][\delta]}$. The
+underlying sets are given by simple composition:
 \begin{alignat*}{3}
@@
 Functoriality of type substitution, i.e.\ $\boldsymbol{A[\id]} = \bA$ and
-$\boldsymbol{A[\sigma\circ\delta]} = \boldsymbol{A[\sigma][\delta]}$, follow
+$\boldsymbol{A[\sigma\circ\delta]} = \boldsymbol{A[\sigma][\delta]}$, follows
 from Lemma \ref{lem:idcomppres} and split cleaving given by $\coe_{\id}$,
@@
 
-$\boldsymbol{\Tm\,\Gamma\,A} : \Set_{1 + \max(i, j)}$ is defined as the type of \emph{weak flCwF sections} of $\bA$.
-
-\begin{definition}[Weak flCwF section] Given $\bA : \bTy\,\bGamma$, the underlying functions of $\bt : \bTm\,\bGamma\,\bA$ are as follows:
+\begin{definition} $\boldsymbol{\Tm\,\Gamma\,A} : \Set_{1 + \max(i, j)}$ is defined as the type of
+\emph{weak flCwF sections} of $\bA$. The underlying functions of $\bt :
+\bTm\,\bGamma\,\bA$ are as follows:
 \begin{alignat*}{3}
@@
 \begin{enumerate}
-  \item $\bsigma\,(\ulA[\ulsigma]) = (\bsigma\,\ulA)\,[\bsigma\,\ulsigma]$
-  \item $\bsigma\,(\ult[\ulsigma]) = (\bsigma\,\ult)\,[\bsigma\,\ulsigma]$
-  \item The unique map $\epsilon_{\bA} : \Sub\,(\bsigma\,\emptycon)\,\emptycon_{\bGamma}\,\id$ has a vertical retraction.
-  \item Each $(\bsigma\,\p,\,\bsigma\,\q) : \Sub\,(\bsigma\,(\ulGamma\,\ext\,\ulA))\,(\bsigma\,\ulGamma\,\ext\,\bsigma\,\ulA)\,\id$ has a vertical inverse.
+  \item $\bt\,(\ulA[\ulsigma]) = (\bt\,\ulA)\,[\bt\,\ulsigma]$
+  \item $\bt\,(\ult[\ulsigma]) = (\bt\,\ult)\,[\bt\,\ulsigma]$
+  \item The unique map $\epsilon_{\bA} : \Sub\,(\bt\,\emptycon)\,\emptycon\,\id$ has a vertical retraction.
+  \item Each $(\bt\,\p,\,\bt\,\q) : \Sub\,(\bt\,(\ulGamma\,\ext\,\ulA))\,(\bt\,\ulGamma\,\ext\,\bt\,\ulA)\,\id$ has a vertical inverse.
 \end{enumerate}
+\end{definition}
 Similarly to Definition \ref{def:weakmorphism}, we denote the evident preservation
@@
 isomorphisms. We also have the derived fl-preservation isomorphisms.
-\end{definition}
 
@@
 $\boldsymbol{\blank[\blank] : \Tm\,\Delta\,A \ra (\sigma : \Sub\,\Gamma\,\Delta)
-  \ra \Tm\,\Gamma\,(A[\sigma])}$ is given the similarly to
+  \ra \Tm\,\Gamma\,(A[\sigma])}$ is given similarly to
 $\boldsymbol{\blank\circ\blank}$ in Section \ref{TODO}. Underlying functions
@@
 $\bA$. This is given by bundling together all displayed flCwF components in
-$\bA$ with corresponding base components in $\bGamma$, using metatheoretic
-$\Sigma$-types. It is a straightforward extension of total categories in
+$\bA$ with corresponding base components in $\bGamma$, using the metatheoretic
+$\Sigma$-type. It is a straightforward extension of total categories in
 \cite{displayedcats}.
@@
   A)\,(A[\p])}$ is likewise a strict flCwF section given by second projections.
+Substitution extension $\boldsymbol{\sigma,\,t}$ is given by pointwise combining $\bsigma$ and $\bt$ with metatheoretic $\Sigma$ pairing, e.g.\ $\Con_{\boldsymbol{(\sigma,t)}}\,\ulGamma :\equiv (\bsigma\,\ulGamma,\,\bt\,\ulGamma)$.
+
+\subsection{Universe}
+
+\begin{definition} For a level $i$, we write $\bSet_i$ for the flCwF of sets where $\Con_{\bSet_i} :\equiv \Set_i$ and $\Sub_{\bSet_i}\,\Gamma\,\Delta :\equiv \Gamma \ra \Delta$.
+\end{definition}
+
+We define $\boldsymbol{\U : \Ty\,\Gamma}$ as the isofibration which is
+constantly $\bSet_j$. A constant isofibration does not actually depend on
+the base flCwF, and has trivial iso-cleaving where $\coe$-s are identity
+functions. Hence, we have $\Con_{\bU}\,\ulGamma :\equiv \Set_j$ and $\Sub_{\bU}\,\Gamma\,\Delta\,\ulsigma :\equiv \Gamma \ra \Delta$.
+
+\emph{Remark.} The type $\bTm\,\bGamma\,\bU$ is strictly equal to $\bSub\,\bGamma\,\bSet_j$, so in the following we will think about semantic elements of the universe as weak morphisms from $\bGamma$ to $\bSet_j$.
+
+\subsection{Elements of the Universe}
+
+\begin{definition}
+We define $\boldsymbol{\El : \Tm\,\Gamma\,\U \ra \Ty\,\Gamma}$ as discrete
+isofibration formation. For $\boldsymbol{a : \Tm\,\Gamma\,\U}$, the underlying
+sets of $\bEl\,\ba$ are the following:
+\begin{alignat*}{3}
+  & \Con_{\bEl\,\ba}\,\ulGamma && :\equiv \ba\,\ulGamma\\
+  & \Sub_{\bEl\,\ba}\,\Gamma\,\Delta\,\ulsigma && :\equiv \ba\,\ulsigma\,\Gamma = \Delta\\
+  & \Ty_{\bEl\,\ba}\,\Gamma\,\ulA && :\equiv \ba\,\ulA\,\Gamma\\
+  & \Tm_{\bEl\,\ba}\,\Gamma\,A\,\ult && :\equiv \ba\,\ult\,\Gamma = A
+\end{alignat*}
+Hence, in $\bEl\,\ba$, $\Sub$ and $\Tm$ are propositional. We use the
+isomorphisms $\ba_{\emptycon} : \ba\,\emptycon \simeq \top$ and $\ba_{\ext} :
+\ba\,(\ulGamma\ext\ulA) \simeq (\Gamma :
+\ba\,\ulGamma)\times(\ba\,\ulA\,\Gamma)$ to define empty context and context
+extension:
+\begin{alignat*}{3}
+  & \emptycon_{\bEl\,\ba} && :\equiv \ba_{\emptycon}^{-1}\,\tt\\
+  & (\Gamma\ext_{\bEl\,\ba} A) && :\equiv \ba_{\ext}^{-1}\,(\Gamma,\,A)
+\end{alignat*}
+We likewise use preservation isomorphisms to define $\K$, $\Id$ and $\Sigma$.
+Context coercion is $\coe\,\ulsigma\,\Gamma :\equiv \ba\,\ulsigma\,\Gamma$. Type coercion, for $A : \ba\,\ulA\,\Gamma$ is given as $\coe\,\ult\,A :\equiv \ba\,\ult\,(\ba_{\ext}^{-1}\,(\Gamma,\,A))$.
+\end{definition}
+
+\subsection{Inductive Function Space}
+
+For $\boldsymbol{a : \Tm\,\Gamma\,\U}$ and $\boldsymbol{B :
+  \Ty\,(\Gamma\ext\El\,a)}$, we aim to define $\boldsymbol{\Pi\,a\,B}
+\boldsymbol{:} \bTy\,\bGamma$. We define this as a dependent product of
+isofibrations, indexed by a discrete domain. $\bPii$ would not be definable with
+general $\bA \boldsymbol{:} \bTy\,\bGamma$ domain, since variance issues already
+prevent $\Pi$-types in the CwF of categories, see e.g. \cite{TODO}.
 

### 2019-12-31 draft/paper.tex
paper

@@
 \newcommand{\ba}{\boldsymbol{a}}
+\newcommand{\bb}{\boldsymbol{b}}
 \newcommand{\bB}{\boldsymbol{B}}
@@
 \newcommand{\bPii}{\boldsymbol{\Pi}}
+\newcommand{\bPie}{\boldsymbol{\Pie}}
+\newcommand{\bPiinf}{\boldsymbol{\Piinf}}
 \newcommand{\bappi}{\boldsymbol{\mathsf{app}}}
@@
 \newcommand{\ulDelta}{\ul{\Delta}}
+\newcommand{\ulgamma}{\ul{\gamma}}
+\newcommand{\uldelta}{\ul{\delta}}
 \newcommand{\ulsigma}{\ul{\sigma}}
-\newcommand{\uldelta}{\ul{\delta}}
 \newcommand{\ulepsilon}{\ul{\epsilon}}
@@
 
-In the following examples, we leave the abstracted $M : \ToS_{0,j}$ implicit.
+In the following examples, we leave the abstracted $M : \ToS_{i,j}$ implicit.
 
@@
 We make a similar choice in the interpretation of signatures themselves: we use
-structured CwFs instead of lex categories to model notions of induction. The
-reason here is that CwFs allow us to compute induction principles in strictly
-the same way as one would write them down in type theory, since we have $\Ty$
-and $\Tm$ for a primitive notion of dependent objects and morphisms. In
-contrast, dependent objects in lex categories is a derived notion, and the
-induction principles we get are only up to isomorphism. This issue is perhaps
-not relevant from a purely categorical perspective, but we are concerned with
-eventually implementing QIITs in proof assistants, so we prefer if our semantics
-computes strictly.
+structured CwFs instead of lex categories. The reason here is that CwFs allow us
+to compute induction principles in strictly the same way as one would write them
+down in type theory, since we have $\Ty$ and $\Tm$ for a primitive notion of
+dependent objects and morphisms. In contrast, dependent objects in lex
+categories is a derived notion, and the induction principles we get are only up
+to isomorphism. This issue is perhaps not relevant from a purely categorical
+perspective, but we are concerned with eventually implementing QIITs in proof
+assistants, so we prefer if our semantics computes strictly.
 
 In the following, we fix $i$ and $j$ levels, and define a model $M : \ToS_{i,
-  1+\max(i,j+1)}$, such than $\Con_M$ is a type of structured categories (of
+  1+\max(i,1+j)}$, such than $\Con_M$ is a type of structured categories (of
 algebras). The level $i$ marks the level of all external sets in function
 domains, and $j$ marks the level of all internal sorts in an algebra. Hence,
-every algebra has level $\max(i,j+1)$. The bump is only needed for $j$, since
+every algebra has level $\max(i,1+j)$. The bump is only needed for $j$, since
 algebras only contain elements of $T : \Set_j$ types, while inductive sets are
-themselves elements of $\Set_i$. For example, $NatAlg_i : \Set_{\max(0, i+1)}$.
+themselves elements of $\Set_i$. For example, $NatAlg_i : \Set_{\max(0, 1+i)}$.
 
@@
 
-We define $\bCon : \Set_{1+\max(i,j+1)}$ as $\flCwF_{\max(i,j+1)}$.
+We define $\bCon : \Set_{1+\max(i,1+j)}$ as $\flCwF_{\max(i,1+j)}$.
 %% We define
@@
 \begin{definition}[Finite limit CwFs]\label{def:flCwF}
-For a level $i$, there is a cETT type $\flCwF_i : \Set_{i+1}$, which is an
+For a level $i$, there is a cETT type $\flCwF_i : \Set_{1+i}$, which is an
 iterated $\Sigma$-type with the following components:
@@
 
-\begin{definition} A \emph{vertical morphism} lies over an identity morphism. We use this definition for context and type morphisms as well.
+\begin{definition} A \emph{vertical morphism} lies over an identity morphism. We use this definition for context morphisms (substitutions) and type morphisms as well.
 \end{definition}
@@
            \ra \Gamma \simeq_{\ulsigma} \coe\,\ulsigma\,\Gamma\\
-  & \coe_{\id} && : \coe\,\id\,\Gamma = \Gamma\\
-  & \coe_{\circ} && : \coe\,(\ulsigma\circ\uldelta)\,\Gamma = \coe\,\ulsigma\,(\coe\,\uldelta\,\Gamma)\\
-  & \coh_{\id} && : \coh\,\id\,\Gamma = \id\\
-  & \coh_{\circ} && : \coh\,(\ulsigma\circ\uldelta)\,\Gamma = \coh\,\ulsigma\,(\coe\,\uldelta\,\Gamma)
+  & \coe^{\id} && : \coe\,\id\,\Gamma = \Gamma\\
+  & \coe^{\circ} && : \coe\,(\ulsigma\circ\uldelta)\,\Gamma = \coe\,\ulsigma\,(\coe\,\uldelta\,\Gamma)\\
+  & \coh^{\id} && : \coh\,\id\,\Gamma = \id\\
+  & \coh^{\circ} && : \coh\,(\ulsigma\circ\uldelta)\,\Gamma = \coh\,\ulsigma\,(\coe\,\uldelta\,\Gamma)
           \circ \coh\,\uldelta\,\Gamma
@@
            \ra A \simeq_{\ult} \coe\,\ult\,A\\
-  & \coe_{\id} && : \coe\,\id\,A = A\\
-  & \coe_{\circ} && : \coe\,\ult\,(\coe\,\uldelta\,A) = \coe\,(\ult\circ\uldelta)\,A\\
-  & \coh_{\id} &&: \coh\,\id\,A = \id\\
-  & \coh_{\circ} &&: \coh\,(\ult\circ\uldelta)\,A = \coh\,\ult\,(\coe\,\uldelta\,A)
+  & \coe^{\id} && : \coe\,\id\,A = A\\
+  & \coe^{\circ} && : \coe\,\ult\,(\coe\,\uldelta\,A) = \coe\,(\ult\circ\uldelta)\,A\\
+  & \coh^{\id} &&: \coh\,\id\,A = \id\\
+  & \coh^{\circ} &&: \coh\,(\ult\circ\uldelta)\,A = \coh\,\ult\,(\coe\,\uldelta\,A)
           \circ \coh\,\uldelta\,A
@@
 \begin{alignat*}{3}
-  & \coe_{[]} &&: \coe\,(\ult[\ulsigma])\,(A[\sigma]) = (\coe\,\ult\,A)[\sigma]\\
-  & \coh_{[]} &&: \coh\,(\ult[\ulsigma\circ \p,\q])\,(A[\sigma]) = (\coh\,\ult\,A)[\sigma]
+  & \coe[] &&: \coe\,(\ult[\ulsigma])\,(A[\sigma]) = (\coe\,\ult\,A)[\sigma]\\
+  & \coh[] &&: \coh\,(\ult[\ulsigma\circ \p,\q])\,(A[\sigma]) = (\coh\,\ult\,A)[\sigma]
 \end{alignat*}
@@
 \end{definition}
+
+\emph{Remark.} It is not possible to model types as fibrations or opfibrations, because
+we have no restriction on the variance of ToS types. For example, the type which extends
+a pointed set to a $Nat$-signature is invariant in the base signature.
 
@@
 
-\begin{definition} $\boldsymbol{\Tm\,\Gamma\,A} : \Set_{1 + \max(i, j)}$ is defined as the type of
+$\boldsymbol{\Tm\,\Gamma\,A} : \Set_{1 + \max(i, 1+j)}$ is defined as the type of
 \emph{weak flCwF sections} of $\bA$. The underlying functions of $\bt :
@@
 \end{enumerate}
-\end{definition}
+
 Similarly to Definition \ref{def:weakmorphism}, we denote the evident preservation
@@
 
-\begin{definition}
 We define $\boldsymbol{\El : \Tm\,\Gamma\,\U \ra \Ty\,\Gamma}$ as discrete
@@
 Context coercion is $\coe\,\ulsigma\,\Gamma :\equiv \ba\,\ulsigma\,\Gamma$. Type coercion, for $A : \ba\,\ulA\,\Gamma$ is given as $\coe\,\ult\,A :\equiv \ba\,\ult\,(\ba_{\ext}^{-1}\,(\Gamma,\,A))$.
-\end{definition}
 
@@
 \boldsymbol{:} \bTy\,\bGamma$. We define this as a dependent product of
-isofibrations, indexed by a discrete domain. $\bPii$ would not be definable with
-general $\bA \boldsymbol{:} \bTy\,\bGamma$ domain, since variance issues already
-prevent $\Pi$-types in the CwF of categories, see e.g. \cite{TODO}.
+isofibrations, indexed by a discrete domain. With a general $\bA \boldsymbol{:}
+\bTy\,\bGamma$ domain, $\bPii$ would not be definable, since variance issues
+preclude $\Pi$-types in the CwF of categories \cite{TODO}.
+
+%% _{(\bPii\,\ba\,\bB)}
+
+Contexts are products of $\bB$-contexts, and types are products of $\bB$-types,
+indexed respectively by contexts and types of $\bEl\,\ba$.
+\begin{alignat*}{3}
+  & \Con_{(\bPii\,\ba\,\bB)}\,\ulGamma &&:\equiv (\gamma : \ba\,\ulGamma)\ra \Con_{\bB}\,(\ulGamma, \gamma)\\
+  & \Ty_{(\bPii\,\ba\,\bB)}\,\Gamma\,\ulA &&:\equiv (\gamma : \ba\,\ulGamma)(a : \ba\,\ulA\,\gamma)\ra \Ty_{\bB}\,(\Gamma\,\gamma)\,(\ulA, a)
+\end{alignat*}
+Note that since $\bB$ is over the total $\boldsymbol{(\Gamma\ext\El\,a)}$,
+$\Con_{\bB}$ has a $\Sigma$-typed argument, and likewise the last argument
+of every $\bB$ component.  We could define substitutions similarly, as
+products of substitutions:
+\begin{alignat*}{3}
+  & \Sub_{(\bPii\,\ba\,\bB)}\,\Gamma\,\Delta\,\ulsigma :\equiv
+  (\gamma : \ba\,\ulGamma)(\delta : \ba\,\ulDelta)
+  (\sigma : \Sub_{(\bEl\,\ba)}\,\gamma\,\delta\,\ulsigma)\\
+  &\hspace{6.5em}\ra \Sub_{\bB}\,(\Gamma\,\gamma)\,(\Delta\,\delta)\,(\ulsigma, \sigma)
+\end{alignat*}
+This would work, but we know that $\Sub_{(\bEl\,\ba)}\,\gamma\,\delta\,\ulsigma$ is defined
+as $\ba\,\ulsigma\,\gamma = \delta$, so we can eliminate $\sigma$ by singleton contraction,
+and use the following equivalent definition:
+\begin{alignat*}{3}
+  & \Sub_{(\bPii\,\ba\,\bB)}\,\Gamma\,\Delta\,\ulsigma :\equiv
+  (\gamma : \ba\,\ulGamma)\ra\Sub_{\bB}\,(\Gamma\,\gamma)\,(\Delta\,(\ba\,\ulsigma\,\gamma)\,(\ulsigma, \refl)
+\end{alignat*}
+The benefit of the contracted definition is that it computes preservation laws
+in algebra homomorphisms strictly as expected, while the non-contracted
+definition computes homomorphisms as functional logical relations.
+
+Terms are also given as a singleton-contracted version of products of terms. In
+$\bPii\,\ba\,\bB$, all other structure is given pointwise by $\bB$-structure.
+
+Iso-cleaving is given by transporting indices backwards in $\bEl\,\ba$ and outputs forwards
+in $\bB$:
+\begin{alignat*}{3}
+  & \coe\,\ulsigma\,\Gamma &&:\equiv
+    \lambda\,\gamma.\,\coe_{\bB}\,(\ulsigma,\refl)\,(\Gamma\,(\ba\,(\ulsigma^{-1})\,\gamma))\\
+  & \coe\,\ult\,A &&:\equiv
+    \lambda\,\gamma\,a.\,\coe_{\bB}\,(\ult,\refl)\,(A\,(\ba\,(\ult^{-1})\,(\ba_{\ext}^{-1}(\gamma,a))))
+\end{alignat*}
+Likewise, $\coh$-s are given by backwards-forwards $\coh$-s.
+
+\subsection{External Function Space}
+
+For $A : \Set_i$ and $\bB : A \ra \bTy\,\bGamma$, we define $\bPie\,A\,\bB \boldsymbol{:} \bTy\,\bGamma$ as
+the evident $A$-indexed direct product of $\bB$.
+
+\subsection{Infinitary Function Space}
+
+For $A : \Set_i$ and $\bb : A \ra \bTm\,\bGamma\,\bU$, we aim to define
+$\bPiinf\,A\,\bB \boldsymbol{:} \bTm\,\bGamma\,\bU$.
 

### 2019-12-31 draft/paper.tex
paper

@@
 \newcommand{\blami}{\boldsymbol{\mathsf{lam}}}
+\newcommand{\bId}{\boldsymbol{\Id}}
 
@@
 
-\todo{P and Q in cwf, level lifts, think about revising cETT or compressing}
+\todo{revise cETT (compress), more mention of interp of term constructors in model, include missing iso-cleaving to type subst}
 
@@
 
-For $A : \Set_i$ and $\bB : A \ra \bTy\,\bGamma$, we define $\bPie\,A\,\bB \boldsymbol{:} \bTy\,\bGamma$ as
-the evident $A$-indexed direct product of $\bB$.
+For $A : \Set_i$ and $\bB : A \ra \bTy\,\bGamma$, we define $\bPie\,A\,\bB
+\boldsymbol{:} \bTy\,\bGamma$ as the $A$-indexed direct product of $\bB$. Since
+the indexing is given by a metatheoretic function, every component is given in the
+evident pointwise way.
 
@@
 For $A : \Set_i$ and $\bb : A \ra \bTm\,\bGamma\,\bU$, we aim to define
-$\bPiinf\,A\,\bB \boldsymbol{:} \bTm\,\bGamma\,\bU$.
+$\bPiinf\,A\,\bb \boldsymbol{:} \bTm\,\bGamma\,\bU$. The underlying functions
+are:
+\begin{alignat*}{3}
+  & (\bPiinf\,A\,\bb)\,\ulGamma    &&:\equiv (a : A)\ra \bb\,a\,\ulGamma\\
+  & (\bPiinf\,A\,\bb)\,\ulsigma    &&:\equiv \lambda\,a.\, \bb\,a\,\ulsigma\\
+  & (\bPiinf\,A\,\bb)\,\ulA\       &&:\equiv \lambda\,\Gamma.\,(a : A)\ra \bb\,a\,\ulA\,(\Gamma\, a)\\
+  & (\bPiinf\,A\,\bb)\,\ult        &&:\equiv \lambda\,a.\, \bb\,a\,\ult
+\end{alignat*}
+The preservation morphisms are as follows. Note that $\emptycon_{\bU} = \top$ and $\ext_{\bU}$ is
+metatheoretic $\Sigma$.
+\begin{alignat*}{3}
+  &(\bPiinf\,A\,\bb)_{\emptycon}^{-1} && : \top\ra (\bPiinf\,A\,\bb)\,\emptycon\\
+  %% &                               && : \top\ra (a : A) \ra \bb\,a\,\emptycon\\
+  &(\bPiinf\,A\,\bb)_{\emptycon}^{-1} && :\equiv \lambda\,\_\,a.\,(\bb\,a)_{\emptycon}^{-1}\,\tt\\
+  &(\bPiinf\,A\,\bb)_{\ext}^{-1} && : (\Gamma : (\bPiinf\,A\,\bb)\,\ulGamma)\times((\bPiinf\,A\,\bb)\,\ulA\,\Gamma)\\
+  & && \hspace{0.5em}\ra (\bPiinf\,A\,\bb)\,(\ulGamma \ext \ulA)\\
+  & (\bPiinf\,A\,\bb)_{\ext}^{-1} && :\equiv \lambda\,(\Gamma,A)\,a.\,(\bb\,a)_{\ext}^{-1}(\Gamma\,a,\,A\,a)
+\end{alignat*}
+
+The preservation of $\emptycon$ and $\blank\ext\blank$ here is in fact the main
+point of divergence from \cite{kaposi2019constructing}. In ibid., substitutions
+and terms are modeled as strict morphisms and types as displayed CwFs (with no
+iso-cleaving). However, it is not the case that $(\bPiinf\,A\,\bb)\,\emptycon =
+\top$, which is the statement of strict $\emptycon$-preservation. The left side
+reduces to $(a : A)\ra \bb\,a\,\emptycon$, which is isomorphic to $\top$ but not
+strictly equal to it. Likewise for $\ext$-preservation.
+
+Hence, we are forced to
+interpret terms as weak sections, which in turn forces us to interpret
+types as isofibrations, since type substitution requires iso-cleaving.
+
+\subsection{Identity}
+For $\bt$ and $\bu$ in $\bTm\,\bGamma\,(\bEl\,\ba)$, we define $\bId\,\bt\,\bu
+\boldsymbol{:} \bTm\,\bGamma\,\bU$ as expressing pointwise equality of weak
+sections.
+\begin{alignat*}{3}
+& (\bId\,\bt\,\bu)\,\ulGamma &&:\equiv \bt\,\ulGamma = \bu\,\ulGamma\\
+& (\bId\,\bt\,\bu)\,\ulA     && :\equiv \lambda\,e.\, \bt\,\ulA = \bu\,\ulA
+\end{alignat*}
+Above, $\bt\,\ulA = \bu\,\ulA$ is well-typed because of $e :
+\bt\,\ulGamma = \bu\,\ulGamma$. For substitutions, we have to complete a square
+of equalities:
+\begin{alignat*}{3}
+  (\bId\,\bt\,\bu)\,(\ulsigma : \Sub\,\ulGamma\,\ulDelta) : \bt\,\ulGamma = \bt\,\ulGamma \ra
+       \bt\,\ulDelta = \bt\,\ulDelta
+\end{alignat*}
+This can be given by $\bt\,\ulsigma : \ba\,\ulsigma\,(\bt\,\ulGamma) =
+\bt\,\ulDelta$ and $\bu\,\ulsigma : \ba\,\ulsigma\,(\bu\,\ulGamma) =
+\bu\,\ulDelta$. The action on terms is analogous. We omit preservation morphisms
+here as they are straightforward. Like $\bPiinf$, $\bId$ also does not support
+strict preservation of $\emptycon$ and $\ext$. Equality reflection and $\boldsymbol{\refl :}
+\bId\,\bt\,\bt$ are also evident.
+
+\section{Model Theory of the Theory of Signatures}
 

### 2020-01-01 draft/paper.tex
paper

@@
 \newcommand{\Sig}{\mathsf{Sig}}
+\newcommand{\ToSSig}{\mathsf{ToSSig}}
+
 \newcommand{\Sg}{\Sigma}
@@
 \newcommand{\bId}{\boldsymbol{\Id}}
+\newcommand{\bM}{\boldsymbol{\mathsf{M}}}
 
@@
 
-\begin{definition}[Notion of signature]
-A QIIT signature at level $i$ is a context in an arbitrary $M : \ToS_{i,j}$
+\begin{definition}[Notion of signature]\label{def:signature}
+A QIIT signature at level $i$ is a context in an arbitrary p$M : \ToS_{i,j}$
 model. We define the type of such signatures as follows:
 \[
-  \Sig_i :\equiv (j : \mathbb{N})\ra(M : \ToS_{i,j})\ra Con_M
+  \Sig_i :\equiv (j : \mathbb{N})\ra(M : \ToS_{i,j})\ra \Con_M
 \]
@@
 
-\subsection{Overview}
+\subsection{Overview}\label{sec:overview}
 
@@
 
-In the following, we fix $i$ and $j$ levels, and define a model $M : \ToS_{i,
-  1+\max(i,1+j)}$, such than $\Con_M$ is a type of structured categories (of
+In the following, for $i$ and $j$ levels we define a model $\bM_{i,j} : \ToS_{i,
+  1+\max(i,1+j)}$, such than $\Con_{\bM_{i,j}}$ is a type of structured categories (of
 algebras). The level $i$ marks the level of all external sets in function
@@
 algebras only contain elements of $T : \Set_j$ types, while inductive sets are
-themselves elements of $\Set_i$. For example, $NatAlg_i : \Set_{\max(0, 1+i)}$.
+themselves elements of $\Set_i$. For example, $NatAlg_i : \Set_{\max(0, 1+i)} :
+\Set_{1+\max(0, 1+i)}$,
 
 We present the components of the model in order. In the following, we use
-\textbf{bold} font to disambiguate components of this model from internal
-structures. For example, we use $\boldsymbol{\sigma : \Sub\,\Gamma\,\Delta}$
-to denote a substitution in the model.
-
-The model involves a large amount of technical detail, most of which we omit,
-and only present the most salient parts.
+\textbf{bold} font to disambiguate components of $\bM_{i,j}$ from components of other
+structures. For example, we use $\boldsymbol{\sigma : \Sub\,\Gamma\,\Delta}$ to
+denote a substitution in $\bM_{i,j}$.
+
+The model involves a large amount of technical detail; we omit a significant
+part of this, and only present the most salient parts.
 
@@
 \begin{definition}[Finite limit CwFs]\label{def:flCwF}
-For a level $i$, there is a cETT type $\flCwF_i : \Set_{1+i}$, which is an
-iterated $\Sigma$-type with the following components:
+For each level $i$ we define $\flCwF_i : \Set_{1+i}$ as an iterated $\Sigma$-type
+with the following components:
 \begin{enumerate}
@@
 
-\begin{definition}[Notion of induction in an flCwF]
+\begin{definition}[Notion of induction in an flCwF]\label{def:induction}
 Given $\bGamma : \flCwF_i$, we have the following predicate on contexts:
@@
 
+$\boldsymbol{\appi : \Tm\,\Gamma\,(\Pii\,a\,B)\ra
+  \Tm\,(\Gamma\,\ext\,\El\,a)\,B}$ can be defined as currying of the underlying
+functions, and $\boldsymbol{\lami}$ as uncurrying.
+
 \subsection{External Function Space}
@@
 \begin{alignat*}{3}
-  (\bId\,\bt\,\bu)\,(\ulsigma : \Sub\,\ulGamma\,\ulDelta) : \bt\,\ulGamma = \bt\,\ulGamma \ra
-       \bt\,\ulDelta = \bt\,\ulDelta
+  (\bId\,\bt\,\bu)\,(\ulsigma : \Sub\,\ulGamma\,\ulDelta) : \bt\,\ulGamma = \bu\,\ulGamma \ra
+       \bt\,\ulDelta = \bu\,\ulDelta
 \end{alignat*}
@@
 
+With this, we have defined the $\bM_{i,j} : \ToS_{i, 1+\max(i,1+j)}$ model that we set out to define
+in Section \ref{sec:overview}.
+
 \section{Model Theory of the Theory of Signatures}
+
+At this point, we only have a notion of algebra for ToS, from Definition
+\ref{def:tos}. In the following sections, we would also like to talk about
+initial ToS-algebras and ToS-induction. We get these notions by giving a QIIT
+signature for ToS, and interpreting it in the $\bM$ model from the previous
+section.
+
+\begin{definition}[Signature for ToS] For each level $i$,
+we define $\ToSSig_i : \Sig_{1+i}$, as the signature for the theory of
+signatures with external sets in $\Set_i$. This is a large and infinitary QIIT
+signature, as we have $\Pie$ and $\Piinf$ abstracting over $A : \Set_i$ and
+branching with $A \ra \Ty\,\Gamma$ and $A \ra \Tm\,\Gamma\,\U$ respectively. We
+present an excerpt from $\ToSSig_i$ below.
+\begin{alignat*}{3}
+  & \emptycon && \ext (Con : \U)\\
+  &           && \ext (Sub : Con\arri Con \arri \U)\\
+  &           && \ext (Ty  : Con\arri\U)\\
+  &           && \ext (Tm  : \Pii(\Gamma : Con)(\Ty\,\Gamma \arri \U))\\
+  &           && ...\\
+  &           && \ext (\Piinf : \Pii(\Gamma : Con)\\
+  &           && \hspace{4.2em}(\Pie \Set_i (\lambda\,A.\,(A \arrinf \Ty\,\Gamma)\arri\El(\Ty\,\Gamma))))\\
+  &           && ...
+\end{alignat*}
+
+Now, for each $j$, the interpretation of $\ToSSig_i$ in $\bM_{1+i,j}$ yields an
+flCwF $\bGamma$ such that $\Con_{\bGamma} = \ToS_{i,j}$. In short, we can
+recover ToS algebras from $\ToSSig$. This follows by computation of the
+interpretation and the fact that $\ToSSig$ is the internal representation of
+$\ToS$. Also, since we have an flCwF of $\ToS_{i,j}$-algebras, we can use
+Definition \ref{def:induction} for the notion of $\ToS$-induction.
+
+\emph{Remark.} By the definition of $\bemptycon$ and
+$\boldsymbol{\blank\ext\blank}$, the types of algebras computed by $\bM$ are
+always left-nested iterated $\Sigma$-types which start with $\top$. Hence, we
+need to require that Definition \ref{def:tos} is similarly left-nested and
+starts with $\top$, in order to make the match strict.
+
+\section{Term Models of QIITs}
+
+%% \begin{theorem}  For each $j$, the interpretation of $\ToSSig_i$ in $\bM_{1+i,j}$
+%% yields an flCwF $\bGamma$ such that $\Con_{\bGamma} = \ToS_{i,j}$. In short, we
+%% can recover ToS algebras from $\ToSSig$.
+%% \end{theorem}
+%% \begin{proof}
+%% \end{proof}
+
+%% Overall, this is just rephrasing Definition \ref{def:tos}
+%% internally to an arbitrary $M : \ToS_{1+i,j}$; recall from Definition
+%% \ref{def:signature} that $\Sig_{1+i} = (j : \mathbb{N})(M : \ToS_{1+i,j})\ra
+%% p\Con_M$. For example, the underlying sets are specified as
+
+\end{definition}
 

### 2020-01-02 draft/paper.tex
paper

@@
 \newcommand{\ulB}{\ul{B}}
+\newcommand{\tos}{\mathsf{tos}}
 
@@
 \end{alignat*}
+\end{definition}
 
@@
 flCwF $\bGamma$ such that $\Con_{\bGamma} = \ToS_{i,j}$. In short, we can
-recover ToS algebras from $\ToSSig$. This follows by computation of the
-interpretation and the fact that $\ToSSig$ is the internal representation of
-$\ToS$. Also, since we have an flCwF of $\ToS_{i,j}$-algebras, we can use
-Definition \ref{def:induction} for the notion of $\ToS$-induction.
+recover ToS algebras from the semantics of $\ToSSig$. This follows by
+computation of the interpretation and the fact that $\ToSSig$ is precisely the
+internal representation of $\ToS$. Also, since we have an flCwF of
+$\ToS_{i,j}$-algebras, we can use Definition \ref{def:induction} for the notion
+of $\ToS$-induction.
 
@@
 
-%% \begin{theorem}  For each $j$, the interpretation of $\ToSSig_i$ in $\bM_{1+i,j}$
-%% yields an flCwF $\bGamma$ such that $\Con_{\bGamma} = \ToS_{i,j}$. In short, we
-%% can recover ToS algebras from $\ToSSig$.
-%% \end{theorem}
-%% \begin{proof}
-%% \end{proof}
-
-%% Overall, this is just rephrasing Definition \ref{def:tos}
-%% internally to an arbitrary $M : \ToS_{1+i,j}$; recall from Definition
-%% \ref{def:signature} that $\Sig_{1+i} = (j : \mathbb{N})(M : \ToS_{1+i,j})\ra
-%% p\Con_M$. For example, the underlying sets are specified as
-
-\end{definition}
+In this section we construct QIITs from initial ToS-algebras. From now on, we assume
+the existence of such algebras.
+
+\todo{assumption in environment}
+
+\textbf{For each $i$, we assume that $\tos_i : \ToS_{i, 1+i}$ exists}. The reason for the
+$1 + i$ level of $\ToS$ sorts is that $\tos_i$ may contain $A : \Set_i$ types in
+$\Ty$, and since we intend $\tos_i$ to be a freely generated algebra, the sorts
+must be large enough to avoid size paradoxes.
+
+\begin{lemma}[Cumulativity of $\ToS$] If $i \leq i'$ and $j \leq j'$,
+  then $\ToS_{i,j} \leq \ToS_{i', j'}$.
+\end{lemma}
+\begin{proof} $\ToS_{i,j}$ is uniformly defined in $i$, $j$, hence cumulative subtyping extends
+  through dependent pairs and functions to $\ToS$.
+\end{proof}
+
+%% \todo{assumption as environment}
+
+%% \emph{Assumption.}
 

### 2020-01-07 draft/paper.tex
type of refl
paper

@@
 \newcommand{\ToSSig}{\mathsf{ToSSig}}
+\newcommand{\Inductive}{\mathsf{Inductive}}
 
@@
 
-\todo{revise cETT (compress), more mention of interp of term constructors in model, include missing iso-cleaving to type subst}
+\todo{revise metatheory intro, use Timany cumulativity}
+\todo{recheck universe levels, prettify levels}
+\todo{explain the limitations and design choices in univ levels}
+\todo{write the rest}
+\todo{see if we can cut back on flCwF, add more examples instead}
 
@@
 
-In Section \ref{sec:metatheory}, we describe a cumulative variant of extensional type
-theory which is used as metatheory in the rest of the paper. In Section \ref{sec:tos},
-we introduce the theory of QIIT signatures. TODO
+In Section \ref{sec:metatheory}, we describe the metatheory used in the rest of
+the paper.  In Section \ref{sec:tos}, we introduce the theory of QIIT
+signatures. TODO
 
@@
 
-In this paper we consider QII algebras at arbitrary finite levels, along with
-large eliminations, where the initial algebra can be at a different level than
-the target algebra. For a simple example, consider natural number algebras at
-level $i$, given as the $\Sigma$-type $NatAlg_i :\equiv(Nat : \Set_i)\times
-Nat\times(Nat \ra Nat)$. The initial such algebra is the set of natural numbers,
-which is at level $0$, but in type theory we often want to eliminate into larger
-$\Set_i$, for example when computing a $Nat$-indexed family of types. To support
-convenient reasoning about levels in this paper, we need two features:
-cumulativity and a way to quantify over finite levels. The former is essential
-to reduce bureaucratic overhead, while the latter is required to talk about
-arbitrary levels.
-
-\subsection{Cumulativity}
-
-%% Cumulativity makes handling of algebras and constructions at different levels
-%% much easier. For example, given an algebra $\gamma : NatAlg_i$, cumulativity
-%% allows us to also have $\gamma : NatAlg_{i+j}$. In a non-cumulative setting such
-%% as Agda, we may only have for each type $A : \Set_i$ a type $\mathsf{Lift}_{i\,j}\,A :
-%% \Set_{i+j}$ which is isomorphic to $A$ up to definitional equality.
-
-In its most basic form, cumulativity requires that whenever $A : \Set_i$, and $i
-< j$ then also $A : \Set_j$. We also need cumulativity for $\Sigma$ and function
-types, in order to have cumulativity for QII algebras in general. For example,
-we want to have $\gamma : NatAlg_j$ whenever $\gamma : NatAlg_i$ and $i < j$.
-
-We use Sterling's cumulative algebraic type theory \cite{sterling2019algebraic} as basis for our metatheory. \todo{REASONS, revamp}
-
-%% The reason is that it is the only rigorous
-%% algebraic definition of cumulativity that we are aware of, and it also
-
-%% The reason for this is twofold. First, it
-%% supports the right kind of cumulativity for our purposes. Second, it is itself a
-%% finitary QIIT with sort equations, hence this paper provides a model theory for
-%% it
-
-We extend the base theory with an extensional identity type, $\Sigma$-types and
-the unit type. In \cite{sterling2019algebraic} a proof of canonicity is provided for the base
-theory, which also includes a standard set-theoretic model. It is
-straightforward to extend the canonicity proof to cover our additional type
-formers.
-
-From now on, we refer to this theory as cETT (cumulative extensional type
-theory). When working in cETT, we use the following notation. We have
-Russell-style universes $\Set_i$ indexed by natural numbers, dependent functions
-as $(x : A)\ra B$, and dependent pairs as $(x : A)\times B$ with projections
-$\proj1$ and $\proj2$. We sometimes leave parameters implicit in dependent
-function types, e.g.\ write $\id : A \ra A$ instead of $\id : (A : \Set_i)\ra A
-\ra A$. We also use subscripts as a field projection notation for iterated
-pairs. For example, for $t : (A : \Set_i) \times (B : \Set_i) \times (f : A \ra
-B)$, we use $B_t$ to denote the projection of the second component. Sometimes we
-omit the subscript if it is clear from context. When we write ``exists'' in this
-paper, we always mean chosen structure given by a $\Sigma$-type.
-
-We write propositional equality as $t = u$, with $\refl_t$ for reflexivity. We
-have equality reflection and uniqueness of identity proofs (UIP) for
-$\blank=\blank$. The unit type is $\top$, with inhabitant $\tt$.
-
-\todo{compactify cumulativity}
-
-\begin{figure}
-\begin{mathpar}
-  \inferrule*[lab=universe formation]
-             {i < j}
-             {\Gamma \vdash \Set_i : \Set_{j}}
-
-  \inferrule*[lab=function formation]
-             {\Gamma \vdash A : \U_i \\ \Gamma,\,x : A \vdash B : \U_j}
-             {\Gamma \vdash (x : A)\ra B : \U_{\max(i,j)} }
-
-  \inferrule*[lab=$\Sigma$ formation]
-             {\Gamma \vdash A : \U_i \\ \Gamma,\,x : A \vdash B : \U_j}
-             {\Gamma \vdash (x : A)\times B : \U_{\max(i,j)} }
-
-  \inferrule*[lab=unit formation]
-             {\\}
-             {\Gamma \vdash \top : \Set_i}
-
-  \inferrule*[lab=equality formation]
-             {\Gamma \vdash A : \Set_i \\ \Gamma \vdash t : A \\ \Gamma \vdash u : A}
-             {\Gamma \vdash t = u : \Set_i}
-
-  \inferrule*[lab=type lifting]
-             {\Gamma \vdash A : \Set_i \\ i \leq j}
-             {\Gamma \vdash\,\Lift_j A : \Set_{j}}
-
-  \inferrule*[lab=lift composition]{}{\Lift_j (\Lift_i A) \equiv\,\,\Lift_{j} A}
-
-  \inferrule*[lab=identity lift]{\Gamma \vdash A : \Set_i}{(\Lift_i A) \equiv\,A}
-
-  \inferrule*[lab=term lifting]
-             {}
-             {\{t\,|\,\Gamma \vdash t : A\} \equiv \{t\,|\,\Gamma \vdash t :\,\,\Lift_i A\}}
-
-  \inferrule*[lab=context lifting]
-             {}
-             {(\Gamma,\,x : A) \equiv (\Gamma,\,x :\,\,\Lift_i A)}
-
-  \inferrule*[lab=universe lifting]{}{\Lift_j \Set_i \equiv \Set_i}
-
-  \inferrule*[lab=function lifting]
-             {}
-             {\Lift_i((x : A)\ra B) \equiv (x :\,\Lift_i A)\ra\,\Lift_i B}
-
-  \inferrule*[lab=pair lifting]
-             {}
-             {\Lift_i((x : A)\times B) \equiv (x :\,\Lift_i A)\,\times\Lift_i B}
-
-  \inferrule*[lab=unit lifting]
-             {}
-             {\Lift_i\top \equiv \top}
-
-  \inferrule*[lab=equality lifting]
-             {}
-             {\Lift_i(t = u) \equiv (t = u)}
-\end{mathpar}
-
-\caption{Some of the rules for lifting.}
-\label{lifting}
-\end{figure}
-
-We also have a lifting operation on types, which introduces cumulativity. In
-Figure \ref{lifting}, we include an excerpt of cETT's rules. We change
-presentation slightly from Sterling: we use a $\U_{\max(i,j)}$ return type in
-functions and pairs, instead of having both $A$ and $B$ types in the same
-$\Set_i$, and we define type lifting with $i \leq j$ instead of $i < j$. These
-changed rules are all derivable from the original ones. In general, we can
-expect that lifting never impedes constructions, because it appropriately
-computes out of the way.
-
-Of special note is the \emph{term lifting rule}. It is a sort equation, an
-equation between sets of terms, expressing that lifted types have exactly the
-same terms as unlifted ones. This allows us to have $t :\,\,\Lift_j A$ whenever $t
-: A$. Together with the universe lifting rule, this implies $A : \Set_{i+j}$
-whenever $A : \Set_i$. Similarly, the lifting rules for functions and pairs give
-us cumulativity for $NatAlg$. We derive a notion of subtyping:
-
-\begin{definition}[Cumulative subtyping]
-  For $A : \Set_i$, $i \leq j$ and $B : \Set_{j}$, we define $A \leq B : \Set_{j}$
-  as $B =\,\,\Lift_j A$.  It follows from term lifting and equality reflection
-  that whenever $A \leq B$ and $t : A$, then also $t : B$.
-\end{definition}
-
-\subsection{Universe Polymorphism}
-\label{sec:universe_polymorphism}
-
-For the sake of simplicity, we do not add this to cETT directly. Instead, we
-quantify over levels in an unspecified metatheory \emph{outside} cETT. Hence, a
-universe polymorphic cETT term is understood as a $\mathbb{N}$-indexed family of
-terms. We reuse the notation of cETT functions for universe polymorphism, e.g.\ as
-in the following function:
-\[
-\lambda\,i.\,\Set_i : (i : \mathbb{N})\ra \Set_{i+1}
-\]
-In this paper, we do not need to internalize level-polymorphic constructions,
-so this setup is sufficient.
+%% We use extensional type theory as metatheory, extended with two features:
+%% cumulative universes and a limited form of universe polymorphism.
+
+%% The extensions are used to conveniently talk about large elimination and
+%% algebras at different universe levels. For example, consider natural number
+%% algebras at level $i$, given as the $\Sigma$-type $NatAlg_i :\equiv(Nat :
+%% \Set_i)\times Nat\times(Nat \ra Nat)$. The initial such algebra is the set of
+%% natural numbers, which is at level $0$, but in type theory we often want to
+%% eliminate into larger $\Set_i$, for example when computing a $Nat$-indexed
+%% family of types. Hence, we would like to
+
+%% Cumulativity is used to reduce bureaucratic overhead of
+%% handling algebras at different universe levels. For a simple example, consider
+%% natural number algebras at level $i$, given as the $\Sigma$-type $NatAlg_i
+%% :\equiv(Nat : \Set_i)\times Nat\times(Nat \ra Nat)$. The initial such algebra is
+%% the set of natural numbers, which is at level $0$, but in type theory we often
+%% want to eliminate into larger $\Set_i$, for example when computing a
+%% $Nat$-indexed family of types.
+
+%% We are working in a variant of extensional type theory with cumulative
+%% universes, abbreviated as cETT.
+
+%% In this paper we consider QII algebras at arbitrary finite levels, along with
+%% large eliminations, where the initial algebra can be at a different level than
+%% the target algebra. For a simple example, consider natural number algebras at
+%% level $i$, given as the $\Sigma$-type $NatAlg_i :\equiv(Nat : \Set_i)\times
+%% Nat\times(Nat \ra Nat)$. The initial such algebra is the set of natural numbers,
+%% which is at level $0$, but in type theory we often want to eliminate into larger
+%% $\Set_i$, for example when computing a $Nat$-indexed family of types. To support
+%% convenient reasoning about levels in this paper, we need two features:
+%% cumulativity and a way to quantify over finite levels. The former is essential
+%% to reduce bureaucratic overhead, while the latter is required to talk about
+%% arbitrary levels.
+
+%% \subsection{Cumulativity}
+
+%% %% Cumulativity makes handling of algebras and constructions at different levels
+%% %% much easier. For example, given an algebra $\gamma : NatAlg_i$, cumulativity
+%% %% allows us to also have $\gamma : NatAlg_{i+j}$. In a non-cumulative setting such
+%% %% as Agda, we may only have for each type $A : \Set_i$ a type $\mathsf{Lift}_{i\,j}\,A :
+%% %% \Set_{i+j}$ which is isomorphic to $A$ up to definitional equality.
+
+%% In its most basic form, cumulativity requires that whenever $A : \Set_i$, and $i
+%% < j$ then also $A : \Set_j$. We also need cumulativity for $\Sigma$ and function
+%% types, in order to have cumulativity for QII algebras in general. For example,
+%% we want to have $\gamma : NatAlg_j$ whenever $\gamma : NatAlg_i$ and $i < j$.
+
+%% We use Sterling's cumulative algebraic type theory \cite{sterling2019algebraic} as basis for our metatheory. \todo{REASONS, revamp}
+
+%% %% The reason is that it is the only rigorous
+%% %% algebraic definition of cumulativity that we are aware of, and it also
+
+%% %% The reason for this is twofold. First, it
+%% %% supports the right kind of cumulativity for our purposes. Second, it is itself a
+%% %% finitary QIIT with sort equations, hence this paper provides a model theory for
+%% %% it
+
+%% We extend the base theory with an extensional identity type, $\Sigma$-types and
+%% the unit type. In \cite{sterling2019algebraic} a proof of canonicity is provided for the base
+%% theory, which also includes a standard set-theoretic model. It is
+%% straightforward to extend the canonicity proof to cover our additional type
+%% formers.
+
+%% From now on, we refer to this theory as cETT (cumulative extensional type
+%% theory). When working in cETT, we use the following notation. We have
+%% Russell-style universes $\Set_i$ indexed by natural numbers, dependent functions
+%% as $(x : A)\ra B$, and dependent pairs as $(x : A)\times B$ with projections
+%% $\proj1$ and $\proj2$. We sometimes leave parameters implicit in dependent
+%% function types, e.g.\ write $\id : A \ra A$ instead of $\id : (A : \Set_i)\ra A
+%% \ra A$. We also use subscripts as a field projection notation for iterated
+%% pairs. For example, for $t : (A : \Set_i) \times (B : \Set_i) \times (f : A \ra
+%% B)$, we use $B_t$ to denote the projection of the second component. Sometimes we
+%% omit the subscript if it is clear from context. When we write ``exists'' in this
+%% paper, we always mean chosen structure given by a $\Sigma$-type.
+
+%% We write propositional equality as $t = u$, with $\refl_t$ for reflexivity. We
+%% have equality reflection and uniqueness of identity proofs (UIP) for
+%% $\blank=\blank$. The unit type is $\top$, with inhabitant $\tt$.
+
+%% \todo{compactify cumulativity}
+
+%% \begin{figure}
+%% \begin{mathpar}
+%%   \inferrule*[lab=universe formation]
+%%              {i < j}
+%%              {\Gamma \vdash \Set_i : \Set_{j}}
+
+%%   \inferrule*[lab=function formation]
+%%              {\Gamma \vdash A : \U_i \\ \Gamma,\,x : A \vdash B : \U_j}
+%%              {\Gamma \vdash (x : A)\ra B : \U_{\max(i,j)} }
+
+%%   \inferrule*[lab=$\Sigma$ formation]
+%%              {\Gamma \vdash A : \U_i \\ \Gamma,\,x : A \vdash B : \U_j}
+%%              {\Gamma \vdash (x : A)\times B : \U_{\max(i,j)} }
+
+%%   \inferrule*[lab=unit formation]
+%%              {\\}
+%%              {\Gamma \vdash \top : \Set_i}
+
+%%   \inferrule*[lab=equality formation]
+%%              {\Gamma \vdash A : \Set_i \\ \Gamma \vdash t : A \\ \Gamma \vdash u : A}
+%%              {\Gamma \vdash t = u : \Set_i}
+
+%%   \inferrule*[lab=type lifting]
+%%              {\Gamma \vdash A : \Set_i \\ i \leq j}
+%%              {\Gamma \vdash\,\Lift_j A : \Set_{j}}
+
+%%   \inferrule*[lab=lift composition]{}{\Lift_j (\Lift_i A) \equiv\,\,\Lift_{j} A}
+
+%%   \inferrule*[lab=identity lift]{\Gamma \vdash A : \Set_i}{(\Lift_i A) \equiv\,A}
+
+%%   \inferrule*[lab=term lifting]
+%%              {}
+%%              {\{t\,|\,\Gamma \vdash t : A\} \equiv \{t\,|\,\Gamma \vdash t :\,\,\Lift_i A\}}
+
+%%   \inferrule*[lab=context lifting]
+%%              {}
+%%              {(\Gamma,\,x : A) \equiv (\Gamma,\,x :\,\,\Lift_i A)}
+
+%%   \inferrule*[lab=universe lifting]{}{\Lift_j \Set_i \equiv \Set_i}
+
+%%   \inferrule*[lab=function lifting]
+%%              {}
+%%              {\Lift_i((x : A)\ra B) \equiv (x :\,\Lift_i A)\ra\,\Lift_i B}
+
+%%   \inferrule*[lab=pair lifting]
+%%              {}
+%%              {\Lift_i((x : A)\times B) \equiv (x :\,\Lift_i A)\,\times\Lift_i B}
+
+%%   \inferrule*[lab=unit lifting]
+%%              {}
+%%              {\Lift_i\top \equiv \top}
+
+%%   \inferrule*[lab=equality lifting]
+%%              {}
+%%              {\Lift_i(t = u) \equiv (t = u)}
+%% \end{mathpar}
+
+%% \caption{Some of the rules for lifting.}
+%% \label{lifting}
+%% \end{figure}
+
+%% We also have a lifting operation on types, which introduces cumulativity. In
+%% Figure \ref{lifting}, we include an excerpt of cETT's rules. We change
+%% presentation slightly from Sterling: we use a $\U_{\max(i,j)}$ return type in
+%% functions and pairs, instead of having both $A$ and $B$ types in the same
+%% $\Set_i$, and we define type lifting with $i \leq j$ instead of $i < j$. These
+%% changed rules are all derivable from the original ones. In general, we can
+%% expect that lifting never impedes constructions, because it appropriately
+%% computes out of the way.
+
+%% Of special note is the \emph{term lifting rule}. It is a sort equation, an
+%% equation between sets of terms, expressing that lifted types have exactly the
+%% same terms as unlifted ones. This allows us to have $t :\,\,\Lift_j A$ whenever $t
+%% : A$. Together with the universe lifting rule, this implies $A : \Set_{i+j}$
+%% whenever $A : \Set_i$. Similarly, the lifting rules for functions and pairs give
+%% us cumulativity for $NatAlg$. We derive a notion of subtyping:
+
+%% \begin{definition}[Cumulative subtyping]
+%%   For $A : \Set_i$, $i \leq j$ and $B : \Set_{j}$, we define $A \leq B : \Set_{j}$
+%%   as $B =\,\,\Lift_j A$.  It follows from term lifting and equality reflection
+%%   that whenever $A \leq B$ and $t : A$, then also $t : B$.
+%% \end{definition}
+
+%% \subsection{Universe Polymorphism}
+%% \label{sec:universe_polymorphism}
+
+%% For the sake of simplicity, we do not add this to cETT directly. Instead, we
+%% quantify over levels in an unspecified metatheory \emph{outside} cETT. Hence, a
+%% universe polymorphic cETT term is understood as a $\mathbb{N}$-indexed family of
+%% terms. We reuse the notation of cETT functions for universe polymorphism, e.g.\ as
+%% in the following function:
+%% \[
+%% \lambda\,i.\,\Set_i : (i : \mathbb{N})\ra \Set_{i+1}
+%% \]
+%% In this paper, we do not need to internalize level-polymorphic constructions,
+%% so this setup is sufficient.
 
@@
 \begin{definition}[Notion of model for the theory of signatures]\label{def:tos}
-For levels $i$ and $j$, $\ToS_{i,j} : \Set_{\max(i,\,j)+1}$ is a cETT type whose
+For levels $i$ and $j$, $\ToS_{i,j} : \Set_{\max(i,\,1+j)}$ is a cETT type whose
 elements are ToS models (or ToS-algebras). $\ToS_{i,j}$ is an iterated
@@
 \begin{alignat*}{3}
-  & Inductive : \Con_{\bGamma} \ra \Set_i \\
-  & Inductive\,\Gamma :\equiv (A : \Ty_{\bGamma}\,\Gamma)\ra \Tm_{\bGamma}\,\Gamma\,A
+  & \Inductive : \Con_{\bGamma} \ra \Set_i \\
+  & \Inductive\,\Gamma :\equiv (A : \Ty_{\bGamma}\,\Gamma)\ra \Tm_{\bGamma}\,\Gamma\,A
 \end{alignat*}
@@
 
-In this section we construct QIITs from initial ToS-algebras. From now on, we assume
-the existence of such algebras.
+In this section we construct QIITs from initial ToS-algebras. Before we can assume
+the existence of such algebras. First, we need cumulative subtyping for $\ToS$.
+\todo{foobar}
+
+\begin{lemma}[Cumulativity of $\ToS$] If $j \leq j'$, then $\ToS_{i,j} \leq \ToS_{i, j'}$.
+\end{lemma}
+\begin{proof}
+  The $j$ universe level only appears in function codomains, hence
+  the subtyping rules for functions and $\Sigma$ can be used to get
+  $\ToS$ subtyping.
+\end{proof}
 
@@
 
-\textbf{For each $i$, we assume that $\tos_i : \ToS_{i, 1+i}$ exists}. The reason for the
-$1 + i$ level of $\ToS$ sorts is that $\tos_i$ may contain $A : \Set_i$ types in
-$\Ty$, and since we intend $\tos_i$ to be a freely generated algebra, the sorts
-must be large enough to avoid size paradoxes.
-
-\begin{lemma}[Cumulativity of $\ToS$] If $i \leq i'$ and $j \leq j'$,
-  then $\ToS_{i,j} \leq \ToS_{i', j'}$.
-\end{lemma}
-\begin{proof} $\ToS_{i,j}$ is uniformly defined in $i$, $j$, hence cumulative subtyping extends
-  through dependent pairs and functions to $\ToS$.
-\end{proof}
+%% For each level $i$,
+%% \begin{enumerate}
+%%   \item We assume that $\tos_i : \ToS_{i, 1+i}$ exists.
+%%   \item For each $i'$ such that $(1+i) \leq i'$, we assume $\Induction
+%% \end{enumerate}
+
+%% \textbf{For each $i$, we assume that $\tos_i : \ToS_{i, 1+i}$ exists}. The reason for the
+%% $1 + i$ level of $\ToS$ sorts is that $\tos_i$ may contain $A : \Set_i$ types in
+%% $\Ty$, and since we intend $\tos_i$ to be a freely generated algebra, the sorts
+%% must be large enough to avoid size paradoxes.
+
+%% \textbf{For each $i$ and $j$, we assume $\mathsf{ToSind :
 
@@
 \end{document}
+
+%% B : Set j
+
+%% (A : Set_i) -> B
+
+%% A : Seti
+%% B : Set j
+%% (x : A) -> B : Set(max i j)
+
+%% (Con : Setj)x
+%% (Ty  : Con -> Setj)x
+
+%% (Pie : (A : Seti) -> (A -> Ty Gamma) -> Ty Gamma
+%% : Set(max(suci, j))
+
+%% : Set(max(sucj, max(suci,j)))
+%%    max(1+j, max(1+i, j))
+%%  = max(1+i, 1+j)
+
+%% REVISE LEVELS!!
+
+%% syntax levels:
+
+  %% ToS(i, 1+i) : Set(max(1+i, 2+i))
+
+%% REWRITE metatheory
+
+%% write about limitations!

### 2020-01-07 draft/paper.tex
compact intro

@@
 \newcommand{\Inductive}{\mathsf{Inductive}}
+\newcommand{\Subtype}{\mathsf{Subtype}}
+\newcommand{\subtype}{\mathsf{subtype}}
 
@@
 initial models represent syntax, and initiality corresponds to induction on
-syntax. Several variants have been used, from contextual categories \cite{TODO} and
-comprehension categories \cite{TODO} to categories with families \cite{TODO}, but all of these are
-categories with extra structure.
-
-The main motivation of the current paper is to generalize previously formalized
-\cite{kaposi2019constructing} QIITs so that they accommodate more algebraic
-formulations of type theories, and more aspects of their practical
-metatheory. As a side effect of fulfilling this goal, infinitary QIITs such as
-Cauchy real numbers \cite{TODO} are covered as well.
+syntax. Several variants have been used, from contextual categories \cite{TODO}
+and comprehension categories \cite{TODO} to categories with families
+\cite{TODO}.
+
+Real, surreal \cite{TODO} and ordinal numbers \cite{TODO} are examples of
+\emph{infinitary} QIITs, which are represented as inductive trees which may
+branch infinitely.
+
+%% The main motivation of the current paper is to generalize previously formalized
+%% \cite{kaposi2019constructing} QIITs so that they accommodate more algebraic
+%% formulations of type theories, and more aspects of their practical
+%% metatheory. As a side effect of fulfilling this goal, infinitary QIITs such as
+%% Cauchy real numbers \cite{TODO} are covered as well.
 
 \subsection{Contributions}
-
-We add the following features to QIITs:
+The current paper extends previously formalized QIITs
+\cite{kaposi2019constructing}. We add the following features:
 \begin{enumerate}
@@
   \textbf{Large constructors, large elimination} and algebras at different
-  universe levels. Large elimination is routinely used in the metatheory of type
-  theory, but it has not been presented explicitly in previous works about
+  universe levels. This fills in an important formal gap, since large
+  elimination is routinely used in the metatheory of type theories, but it has
+  not been but it has not been presented explicitly in previous works about
   QIITs.
   \item
-  \textbf{Infinitary constructors}. This allows infinitely branching
-  trees. Also, the theory of QIIT signatures is itself large and infinitary,
-  thus it can ``eat itself'', i.e.\ include its own signature and provide
-  its own metatheory. This was not possible previously in
-  \cite{kaposi2019constructing}, where only finitary QIITs were described. We
-  exploit self-representation to bootstrap the model theory of signatures,
-  without having to assume any pre-existing internal syntax.
+  \textbf{Infinitary constructors}.
+  Of special note here is that the theory of QIIT signatures is itself large and
+  infinitary, thus it can ``eat itself'', i.e.\ include its own signature and
+  provide its own metatheory. This was not possible previously in
+  \cite{kaposi2019constructing}, where only finitary QIITs were described. In
+  this paper we use self-representation to bootstrap the model theory of
+  signatures, without having to assume any pre-existing internal syntax.
   \item
@@
 the paper. In Section \ref{sec:tos}, we introduce the theory of QIIT
-signatures. TODO
+signatures. In Section \ref{sec:semantics} we give categorical semantics to
+signatures. In Section \ref{sec:tossig} we build model theory for the theory
+of QIIT signatures. In Section \ref{sec:termmodels} we give a term model construction
+of QIITs. We discuss related work and conclude in Sections \ref{TODO}
 
 \section{Metatheory}
-
-%% We use extensional type theory as metatheory, extended with two features:
-%% cumulative universes and a limited form of universe polymorphism.
-
-%% The extensions are used to conveniently talk about large elimination and
-%% algebras at different universe levels. For example, consider natural number
-%% algebras at level $i$, given as the $\Sigma$-type $NatAlg_i :\equiv(Nat :
-%% \Set_i)\times Nat\times(Nat \ra Nat)$. The initial such algebra is the set of
-%% natural numbers, which is at level $0$, but in type theory we often want to
-%% eliminate into larger $\Set_i$, for example when computing a $Nat$-indexed
-%% family of types. Hence, we would like to
-
-%% Cumulativity is used to reduce bureaucratic overhead of
-%% handling algebras at different universe levels. For a simple example, consider
-%% natural number algebras at level $i$, given as the $\Sigma$-type $NatAlg_i
-%% :\equiv(Nat : \Set_i)\times Nat\times(Nat \ra Nat)$. The initial such algebra is
-%% the set of natural numbers, which is at level $0$, but in type theory we often
-%% want to eliminate into larger $\Set_i$, for example when computing a
-%% $Nat$-indexed family of types.
-
-%% We are working in a variant of extensional type theory with cumulative
-%% universes, abbreviated as cETT.
-
-%% In this paper we consider QII algebras at arbitrary finite levels, along with
-%% large eliminations, where the initial algebra can be at a different level than
-%% the target algebra. For a simple example, consider natural number algebras at
-%% level $i$, given as the $\Sigma$-type $NatAlg_i :\equiv(Nat : \Set_i)\times
-%% Nat\times(Nat \ra Nat)$. The initial such algebra is the set of natural numbers,
-%% which is at level $0$, but in type theory we often want to eliminate into larger
-%% $\Set_i$, for example when computing a $Nat$-indexed family of types. To support
-%% convenient reasoning about levels in this paper, we need two features:
-%% cumulativity and a way to quantify over finite levels. The former is essential
-%% to reduce bureaucratic overhead, while the latter is required to talk about
-%% arbitrary levels.
-
-%% \subsection{Cumulativity}
-
-%% %% Cumulativity makes handling of algebras and constructions at different levels
-%% %% much easier. For example, given an algebra $\gamma : NatAlg_i$, cumulativity
-%% %% allows us to also have $\gamma : NatAlg_{i+j}$. In a non-cumulative setting such
-%% %% as Agda, we may only have for each type $A : \Set_i$ a type $\mathsf{Lift}_{i\,j}\,A :
-%% %% \Set_{i+j}$ which is isomorphic to $A$ up to definitional equality.
-
-%% In its most basic form, cumulativity requires that whenever $A : \Set_i$, and $i
-%% < j$ then also $A : \Set_j$. We also need cumulativity for $\Sigma$ and function
-%% types, in order to have cumulativity for QII algebras in general. For example,
-%% we want to have $\gamma : NatAlg_j$ whenever $\gamma : NatAlg_i$ and $i < j$.
-
-%% We use Sterling's cumulative algebraic type theory \cite{sterling2019algebraic} as basis for our metatheory. \todo{REASONS, revamp}
-
-%% %% The reason is that it is the only rigorous
-%% %% algebraic definition of cumulativity that we are aware of, and it also
-
-%% %% The reason for this is twofold. First, it
-%% %% supports the right kind of cumulativity for our purposes. Second, it is itself a
-%% %% finitary QIIT with sort equations, hence this paper provides a model theory for
-%% %% it
-
-%% We extend the base theory with an extensional identity type, $\Sigma$-types and
-%% the unit type. In \cite{sterling2019algebraic} a proof of canonicity is provided for the base
-%% theory, which also includes a standard set-theoretic model. It is
-%% straightforward to extend the canonicity proof to cover our additional type
-%% formers.
-
-%% From now on, we refer to this theory as cETT (cumulative extensional type
-%% theory). When working in cETT, we use the following notation. We have
-%% Russell-style universes $\Set_i$ indexed by natural numbers, dependent functions
-%% as $(x : A)\ra B$, and dependent pairs as $(x : A)\times B$ with projections
-%% $\proj1$ and $\proj2$. We sometimes leave parameters implicit in dependent
-%% function types, e.g.\ write $\id : A \ra A$ instead of $\id : (A : \Set_i)\ra A
-%% \ra A$. We also use subscripts as a field projection notation for iterated
-%% pairs. For example, for $t : (A : \Set_i) \times (B : \Set_i) \times (f : A \ra
-%% B)$, we use $B_t$ to denote the projection of the second component. Sometimes we
-%% omit the subscript if it is clear from context. When we write ``exists'' in this
-%% paper, we always mean chosen structure given by a $\Sigma$-type.
-
-%% We write propositional equality as $t = u$, with $\refl_t$ for reflexivity. We
-%% have equality reflection and uniqueness of identity proofs (UIP) for
-%% $\blank=\blank$. The unit type is $\top$, with inhabitant $\tt$.
-
-%% \todo{compactify cumulativity}
-
-%% \begin{figure}
-%% \begin{mathpar}
-%%   \inferrule*[lab=universe formation]
-%%              {i < j}
-%%              {\Gamma \vdash \Set_i : \Set_{j}}
-
-%%   \inferrule*[lab=function formation]
-%%              {\Gamma \vdash A : \U_i \\ \Gamma,\,x : A \vdash B : \U_j}
-%%              {\Gamma \vdash (x : A)\ra B : \U_{\max(i,j)} }
-
-%%   \inferrule*[lab=$\Sigma$ formation]
-%%              {\Gamma \vdash A : \U_i \\ \Gamma,\,x : A \vdash B : \U_j}
-%%              {\Gamma \vdash (x : A)\times B : \U_{\max(i,j)} }
-
-%%   \inferrule*[lab=unit formation]
-%%              {\\}
-%%              {\Gamma \vdash \top : \Set_i}
-
-%%   \inferrule*[lab=equality formation]
-%%              {\Gamma \vdash A : \Set_i \\ \Gamma \vdash t : A \\ \Gamma \vdash u : A}
-%%              {\Gamma \vdash t = u : \Set_i}
-
-%%   \inferrule*[lab=type lifting]
-%%              {\Gamma \vdash A : \Set_i \\ i \leq j}
-%%              {\Gamma \vdash\,\Lift_j A : \Set_{j}}
-
-%%   \inferrule*[lab=lift composition]{}{\Lift_j (\Lift_i A) \equiv\,\,\Lift_{j} A}
-
-%%   \inferrule*[lab=identity lift]{\Gamma \vdash A : \Set_i}{(\Lift_i A) \equiv\,A}
-
-%%   \inferrule*[lab=term lifting]
-%%              {}
-%%              {\{t\,|\,\Gamma \vdash t : A\} \equiv \{t\,|\,\Gamma \vdash t :\,\,\Lift_i A\}}
-
-%%   \inferrule*[lab=context lifting]
-%%              {}
-%%              {(\Gamma,\,x : A) \equiv (\Gamma,\,x :\,\,\Lift_i A)}
-
-%%   \inferrule*[lab=universe lifting]{}{\Lift_j \Set_i \equiv \Set_i}
-
-%%   \inferrule*[lab=function lifting]
-%%              {}
-%%              {\Lift_i((x : A)\ra B) \equiv (x :\,\Lift_i A)\ra\,\Lift_i B}
-
-%%   \inferrule*[lab=pair lifting]
-%%              {}
-%%              {\Lift_i((x : A)\times B) \equiv (x :\,\Lift_i A)\,\times\Lift_i B}
-
-%%   \inferrule*[lab=unit lifting]
-%%              {}
-%%              {\Lift_i\top \equiv \top}
-
-%%   \inferrule*[lab=equality lifting]
-%%              {}
-%%              {\Lift_i(t = u) \equiv (t = u)}
-%% \end{mathpar}
-
-%% \caption{Some of the rules for lifting.}
-%% \label{lifting}
-%% \end{figure}
-
-%% We also have a lifting operation on types, which introduces cumulativity. In
-%% Figure \ref{lifting}, we include an excerpt of cETT's rules. We change
-%% presentation slightly from Sterling: we use a $\U_{\max(i,j)}$ return type in
-%% functions and pairs, instead of having both $A$ and $B$ types in the same
-%% $\Set_i$, and we define type lifting with $i \leq j$ instead of $i < j$. These
-%% changed rules are all derivable from the original ones. In general, we can
-%% expect that lifting never impedes constructions, because it appropriately
-%% computes out of the way.
-
-%% Of special note is the \emph{term lifting rule}. It is a sort equation, an
-%% equation between sets of terms, expressing that lifted types have exactly the
-%% same terms as unlifted ones. This allows us to have $t :\,\,\Lift_j A$ whenever $t
-%% : A$. Together with the universe lifting rule, this implies $A : \Set_{i+j}$
-%% whenever $A : \Set_i$. Similarly, the lifting rules for functions and pairs give
-%% us cumulativity for $NatAlg$. We derive a notion of subtyping:
-
-%% \begin{definition}[Cumulative subtyping]
-%%   For $A : \Set_i$, $i \leq j$ and $B : \Set_{j}$, we define $A \leq B : \Set_{j}$
-%%   as $B =\,\,\Lift_j A$.  It follows from term lifting and equality reflection
-%%   that whenever $A \leq B$ and $t : A$, then also $t : B$.
-%% \end{definition}
-
-%% \subsection{Universe Polymorphism}
-%% \label{sec:universe_polymorphism}
-
-%% For the sake of simplicity, we do not add this to cETT directly. Instead, we
-%% quantify over levels in an unspecified metatheory \emph{outside} cETT. Hence, a
-%% universe polymorphic cETT term is understood as a $\mathbb{N}$-indexed family of
-%% terms. We reuse the notation of cETT functions for universe polymorphism, e.g.\ as
-%% in the following function:
-%% \[
-%% \lambda\,i.\,\Set_i : (i : \mathbb{N})\ra \Set_{i+1}
-%% \]
-%% In this paper, we do not need to internalize level-polymorphic constructions,
-%% so this setup is sufficient.
+\label{sec:metatheory}
+
+The metatheory used in this paper is extensional type theory, extended with a
+form of cumulativity and an external notion of universe polymorphism. We refer
+to this theory as cETT. We review the used features and notations
+in the following.
+
+\subsection{Core Extensional Theory}
+
+We have Russell-style predicative universes $\Set_i$ indexed by natural
+numbers, dependent functions as $(x : A)\ra B$, and dependent pairs as $(x :
+A)\times B$ with projections $\proj1$ and $\proj2$. We sometimes leave
+parameters implicit in dependent function types, e.g.\ write $\id : A \ra A$
+instead of $\id : (A : \Set_i)\ra A \ra A$. We also use subscripts as a field
+projection notation for iterated pairs. For example, for $t : (A : \Set_i)
+\times (B : \Set_i) \times (f : A \ra B)$, we use $B_t$ to denote the projection
+of the second component. Sometimes we omit the subscript if it is clear from
+context. When we write ``exists'' in this paper, we always mean chosen structure
+given by a $\Sigma$-type.
+
+\todo{univ lvl for pi sigma}
+
+%% Both for functions and $\Sigma$, the output universe level is given as the
+%% maximum of the levels of the constituent type.
+
+We write propositional equality as $t = u$, with $\refl_t$ for reflexivity. We
+have equality reflection and uniqueness of identity proofs (UIP). The unit type
+is $\top$, with inhabitant $\tt$.
+
+\subsection{Cumulativity}
+
+We use cumulative universes and cumulative subtyping as described
+in \cite{timany2018cumulative}. Concretely, we have a $\blank\leq\blank$ subtyping
+relation on types, and a corresponding subtyping rule.
+\begin{mathpar}
+  \inferrule*{i \leq j}
+             {\Gamma \vdash \Set_i \leq \Set_j}
+
+  \inferrule*{\Gamma,\,x : A \vdash B \leq B'}
+             {\Gamma \vdash (x : A)\ra B \leq (x : A)\ra B'}
+
+  \inferrule*{\Gamma\vdash A \leq A' \\ \Gamma,\,x : A \vdash B \leq B'}
+             {\Gamma \vdash (x : A)\times B \leq (x : A') \times B'}
+
+  \inferrule*{\\}
+             {\Gamma \vdash A \leq A}
+
+  \inferrule*{\Gamma \vdash A \leq B \\ \Gamma\vdash B \leq C}
+             {\Gamma \vdash A \leq C}
+
+  \inferrule*{\Gamma\vdash A \leq A' \\ \Gamma\vdash t : A}
+             {\Gamma \vdash t : A'}
+\end{mathpar}
+
+Additionally, we have an internal type for $\blank\leq\blank$, analogously to
+$\blank=\blank$ which is an internalization of definitional equality.
+\begin{mathpar}
+  \inferrule*{\Gamma \vdash A : \Set_i \\ \Gamma \vdash B : \Set_j}
+             {\Gamma \vdash \Subtype\,A\,B : \Set_{\max(i, j)}}
+
+  \inferrule*{\Gamma \vdash A \leq B}
+             {\Gamma \vdash \subtype : \Subtype\,A\,B}
+
+  \inferrule*{\Gamma \vdash t : \Subtype\,A\,B}
+             {\Gamma \vdash A \leq B}
+\end{mathpar}
+
+We use cumulativity to reduce bureaucratic overhead when dealing with
+constructions at different universe levels. The internal $\Subtype$ is used in
+Section \ref{sec:termmodels} to prove cumulativity for general QIIT
+algebras. For example, consider natural number algebras at level $i$, given as
+the $\Sigma$-type $NatAlg_i :\equiv(Nat : \Set_i)\times Nat\times(Nat \ra
+Nat)$. It follows from the subtyping rules that $i \leq j$ implies $NatAlg_{i}
+\leq NatAlg_{j}$. However, cumulativity for arbitrary QIIT algebras does not
+follow judgmentally; it can only be proven by induction on signatures, hence
+the need for $\Subtype$.
+
+Internal subtyping is not included in \cite{timany2018cumulative}, but it can be
+justified by the set-theoretic model given there.
+
+\subsection{Universe Polymorphism}
+\label{sec:universe_polymorphism}
+
+We need to talk about constructions at arbitrary universe levels. For the sake
+of simplicity, we do not assume a notion of universe polymorphism in cETT,
+instead we quantify over levels in an unspecified theory outside of cETT. Hence,
+a universe polymorphic cETT term is understood as a $\mathbb{N}$-indexed family
+of cETT terms. This is sufficient in this paper, since we do not need to
+internalize level polymorphism.
+
+We reuse the notation of cETT functions for
+universe polymorphism, e.g.\ as in the following function:
+\[
+\lambda\,i.\,\Set_i : (i : \mathbb{N})\ra \Set_{i+1}
+\]
 
@@
 \Tm\,\Gamma\,(\El\,a)\ra\Tm\,\Gamma\,(\El\,a)\ra \Tm\,\Gamma\,\U$, with $\Refl :
-(t : \Tm\,\Gamma\,(\El\,a))\ra \Id\,a\,t\,t$, equality reflection and UIP.
+(t : \Tm\,\Gamma\,(\El\,a))\ra \Tm\,\Gamma\,(\El\,(\Id\,a\,t\,t))$, equality reflection and UIP.
 
@@
 of $\Pii$ types. This allows, for example, a signature for trees branching with
-arbitrary small sets. This is a signature at level 1, since we have $\Set_0$ as
+arbitrary small types. This is a signature at level 1, since we have $\Set_0$ as
 a $\Pie$ domain type.
@@
 \begin{example}{Recursive equations}. Again, the universe is closed under $\Id$, which allows us to write equations
-in $\Pii$ domains. A minimal example:
+in $\Pii$ domains. A minimal (and trivial) example:
 \begin{alignat*}{3}
@@
 \end{alignat*}
-More interesting (and complicated) examples are boundary conditions in various
-cubical type theories\cite{TODO}. Note that our $\Id$ allows iterated equations
-as well, but these are all trivial in the semantics, where we assume UIP.
+More interesting (and complicated) examples for recursive equations are boundary
+conditions in various cubical type theories \cite{TODO}. Note that our $\Id$
+allows iterated equations as well, but these are all trivial in the semantics,
+where we assume UIP.
 \end{example}
@@
 Our semantics has a ``type-theoretic'' flavor, which is inspired by the cubical
-set model of Martin-Löf type theory by Bezem et al.\cite{cubical}. The core idea
+set model of Martin-Löf type theory by Bezem et al. \cite{cubical}. The core idea
 is to avoid strictness issues by starting from basic ingredients which are already
@@
 \section{Model Theory of the Theory of Signatures}
+\label{sec:tossig}
 
@@
 \section{Term Models of QIITs}
+\label{sec:termmodels}
 

### 2020-01-07 draft/paper.tex
paper

@@
 
-\todo{univ lvl for pi sigma}
-
-%% Both for functions and $\Sigma$, the output universe level is given as the
-%% maximum of the levels of the constituent type.
+Both for functions and $\Sigma$, the output universe level is given as the
+maximum of the levels of the constituent types, e.g.\ $(x : A)\ra B : \Set_{\max(i,j)}$
+when $A : \Set_i$ and $B : \Set_j$.
 
@@
 have equality reflection and uniqueness of identity proofs (UIP). The unit type
-is $\top$, with inhabitant $\tt$.
+is $\top : \Set_0$, with inhabitant $\tt$.
 
@@
 in \cite{timany2018cumulative}. Concretely, we have a $\blank\leq\blank$ subtyping
-relation on types, and a corresponding subtyping rule.
+relation on types, specified by the following rules:
 \begin{mathpar}
@@
 a universe polymorphic cETT term is understood as a $\mathbb{N}$-indexed family
-of cETT terms. This is sufficient in this paper, since we do not need to
-internalize level polymorphism.
-
-We reuse the notation of cETT functions for
+of cETT terms. We reuse the notation of cETT functions for
 universe polymorphism, e.g.\ as in the following function:
@@
 \begin{definition}[Notion of model for the theory of signatures]\label{def:tos}
-For levels $i$ and $j$, $\ToS_{i,j} : \Set_{\max(i,\,1+j)}$ is a cETT type whose
+For levels $i$ and $j$, $\ToS_{i,j} : \Set_{\max(i+1,\,j+1)}$ is a cETT type whose
 elements are ToS models (or ToS-algebras). $\ToS_{i,j}$ is an iterated
@@
 \end{alignat*}
-We abbreviate non-dependent inductive $\Pii$ as $\blank\arri\blank$, and likewise we use $\blank\arre\blank$ and $\blank\arrinf\blank$.
+We abbreviate non-dependent inductive $\Pii$ as $\blank\arri\blank$, and
+likewise we use $\blank\arre\blank$ and $\blank\arrinf\blank$ for non-dependent
+external and infinitary functions.
 
@@
 \begin{definition}[Notion of signature]\label{def:signature}
-A QIIT signature at level $i$ is a context in an arbitrary p$M : \ToS_{i,j}$
+A QIIT signature at level $i$ is a context in an arbitrary $M : \ToS_{i,j}$
 model. We define the type of such signatures as follows:
@@
 In the following, for $i$ and $j$ levels we define a model $\bM_{i,j} : \ToS_{i,
-  1+\max(i,1+j)}$, such than $\Con_{\bM_{i,j}}$ is a type of structured categories (of
+  \max(i,j+1)+1}$, such than $\Con_{\bM_{i,j}}$ is a type of structured categories (of
 algebras). The level $i$ marks the level of all external sets in function
 domains, and $j$ marks the level of all internal sorts in an algebra. Hence,
-every algebra has level $\max(i,1+j)$. The bump is only needed for $j$, since
-algebras only contain elements of $T : \Set_j$ types, while inductive sets are
-themselves elements of $\Set_i$. For example, $NatAlg_i : \Set_{\max(0, 1+i)} :
-\Set_{1+\max(0, 1+i)}$,
+every algebra has level $\max(i,j+1)$. The bump is only needed for $j$, since
+algebras merely contain elements of $A : \Set_i$ types, while inductive sets are
+themselves elements of $\Set_j$. For example, $NatAlg_j : \Set_{\max(0, j+1)} :
+\Set_{\max(0, j+1)+1}$,
 
@@
 
-We define $\bCon : \Set_{1+\max(i,1+j)}$ as $\flCwF_{\max(i,1+j)}$.
+We define $\bCon : \Set_{\max(i,j+1)+1}$ as $\flCwF_{\max(i,j+1)}$.
 %% We define
@@
 
-\todo{segue text?}
+In order to talk about weak structure-preservation in the interpretation of substitutions, we need to specify isomorphisms for contexts and types.
 
@@
 
-\todo{we have a level lift here, decide how to handle}
-
-%% We define $\boldsymbol{\Sub\,\Gamma\,\Delta} : \Set_{1 + \max(i, j+1)}$ as the type
-%% of weak flCwF morphisms from $\boldsymbol{\Gamma}$ to $\boldsymbol{\Delta}$.
-
 \begin{definition}\label{def:weakmorphism} A \emph{weak flCwF morphism} $\boldsymbol{\sigma : \Sub\,\Gamma\,\Delta}$ is a functor between underlying categories, which additionally maps types to types and terms to terms, and satisfies the following:

### 2020-01-08 draft/paper.tex
paper

@@
 \newcommand{\arrinf}{\Rightarrow^{\mathsf{inf}}}
+\newcommand{\syn}{\mathsf{syn}}
+\newcommand{\SynSig}{\mathsf{SynSig}}
 
@@
 %% comma separated list
-\keywords{inductive types, quotient inductive-inductive types}  %% \keywords are mandatory in final camera-ready submission
+\keywords{inductive types, infinitary inductive types, quotient inductive-inductive types, categories with families}
 
@@
 
-\todo{revise metatheory intro, use Timany cumulativity}
-\todo{recheck universe levels, prettify levels}
-\todo{explain the limitations and design choices in univ levels}
-\todo{write the rest}
-\todo{see if we can cut back on flCwF, add more examples instead}
-
 The aim of this work is to provide theoretical underpinning to a general notion
@@
 
-Categories are a good example. Signatures for QIITs allow having multiple sorts,
+Categories are a prime example for a structure which is described by a quotient
+inductive-inductive signature. Signatures for QIITs allow having multiple sorts,
 with later ones indexed over previous ones, and equations as well. We need both
@@
 initial models represent syntax, and initiality corresponds to induction on
-syntax. Several variants have been used, from contextual categories \cite{TODO}
-and comprehension categories \cite{TODO} to categories with families
-\cite{TODO}.
-
-Real, surreal \cite{TODO} and ordinal numbers \cite{TODO} are examples of
-\emph{infinitary} QIITs, which are represented as inductive trees which may
-branch infinitely.
+syntax. Several variants have been used, from contextual categories \cite{gat}
+and comprehension categories \cite{jacobs1993comprehension} to categories with
+families \cite{Dybjer96internaltype}. Additionally, real, surreal \cite{HoTTbook}
+and ordinal numbers \cite{lumsdaineShulman} can be defined as \emph{infinitary}
+QIITs, which are represented as possibly infinitely branching inductive trees.
 
@@
 \subsection{Contributions}
-The current paper extends previously formalized QIITs
-\cite{kaposi2019constructing}. We add the following features:
+The current paper extends the syntax and semantics of QIITs as previously described
+in the literature \cite{kaposi2019constructing, gabe, dijkstra2017quotient}. We add the following features:
 \begin{enumerate}
@@
   \textbf{Large constructors, large elimination} and algebras at different
-  universe levels. This fills in an important formal gap, since large
-  elimination is routinely used in the metatheory of type theories, but it has
-  not been but it has not been presented explicitly in previous works about
-  QIITs.
+  universe levels. This fills in an important formal gap; large models are
+  routinely used in the metatheory of type theories, but they have not been
+  presented explicitly in previous QIIT literature.
   \item
@@
 strict CwF morphisms, we need to consider weak ones, and instead of modeling
-types as displayed CwFs, we need to model them as CwF isofibrations. In a
-nutshell, the new semantics must be given mutually with a form of a ``structure
-identity principle''\cite{TODO}, which says that signature extension respects algebra
-isomorphisms.
+types as displayed CwFs, we need to model them as CwF isofibrations. The latter
+amounts to showing that signature extension respects algebra isomorphisms.
 
@@
 of QIIT signatures. In Section \ref{sec:termmodels} we give a term model construction
-of QIITs. We discuss related work and conclude in Sections \ref{TODO}
+of QIITs. We discuss related work and conclude in Sections \ref{sec:relatedwork}-\ref{sec:conclusion}.
 
@@
 
-Both for functions and $\Sigma$, the output universe level is given as the
+Both for function types and $\Sigma$, the output universe level is given as the
 maximum of the levels of the constituent types, e.g.\ $(x : A)\ra B : \Set_{\max(i,j)}$
@@
 \subsection{Cumulativity}
+\label{sec:cumulativity}
 
@@
 A \textbf{category with families} (CwF), where all four underlying sets (of
-objects, morphisms, types and terms) are in $\Set_j$. Following notation in
-\cite{kaposi2019constructing}, we denote these respectively as $\Con : \Set_j$,
-$\Sub : \Con \ra \Con \ra \Set_j$, $\Ty : \Con \ra \Set_j$ and $\Tm : (\Gamma :
-\Con) \ra \Ty\,\Gamma \ra \Set_j$. We use $\id$ and $\blank\circ\blank$ to
+objects, morphisms, types and terms) are in $\Set_i$. Following notation in
+\cite{kaposi2019constructing}, we denote these respectively as $\Con : \Set_i$,
+$\Sub : \Con \ra \Con \ra \Set_i$, $\Ty : \Con \ra \Set_i$ and $\Tm : (\Gamma :
+\Con) \ra \Ty\,\Gamma \ra \Set_i$. We use $\id$ and $\blank\circ\blank$ to
 denote identity and composition for substitution. We denote the empty context as
@@
 
-\textbf{External function space} $\Pie : (A : \Set_i)\ra(A \ra \Ty\,\Gamma)\ra
+\textbf{External function space} $\Pie : (A : \Set_j)\ra(A \ra \Ty\,\Gamma)\ra
 \Ty\,\Gamma$, with $\appe : \Tm\,\Gamma\,(\Pie\,A\,B)\ra((x : A)\ra
@@
 
-\textbf{Infinitary function space} $\Piinf : (A : \Set_i)\ra(A \ra
+\textbf{Infinitary function space} $\Piinf : (A : \Set_j)\ra(A \ra
 \Tm\,\Gamma\,\U)\ra \Tm\,\Gamma\,\U$, with $\appinf :
@@
 \begin{definition}[Notion of signature]\label{def:signature}
-A QIIT signature at level $i$ is a context in an arbitrary $M : \ToS_{i,j}$
+A QIIT signature at level $j$ is a context in an arbitrary $M : \ToS_{i,j}$
 model. We define the type of such signatures as follows:
 \[
-  \Sig_i :\equiv (j : \mathbb{N})\ra(M : \ToS_{i,j})\ra \Con_M
+  \Sig_j :\equiv (i : \mathbb{N})\ra(M : \ToS_{i,j})\ra \Con_M
 \]
 
-Here, $i$ refers to the level of external types appearing in the signature, in
-the domains of $\Pie$ and $\Piinf$ functions, while the quantified $j$ level is
+Here, $j$ refers to the level of external types appearing in the signature, in
+the domains of $\Pie$ and $\Piinf$ functions, while the quantified $i$ level is
 required to allow interpreting a signature in arbitrary-sized ToS models. Note
-that $\Sig_i$ is universe-polymorphic, so it is a family of cETT types and it is
+that $\Sig_j$ is universe-polymorphic, so it is a family of cETT types and it is
 not in any cETT universe.
@@
   & NatSig : \Sig_0 \\
-  & NatSig :\equiv \lambda (j :\mathbb{N})(M : \ToS_{0,j}).\\
+  & NatSig :\equiv \lambda (i :\mathbb{N})(M : \ToS_{i,0}).\\
   & \hspace{1em}(\emptycon_M\,\ext_M\, (N : \U_M) \,\ext_M\,(zero : \El_M\,N)\\
@@
 signature in an arbitrary ToS model, by applying a signature to a model.
-$\Sig_i$ can be viewed as a precursor to a Church-encoding for the theory of
+$\Sig_j$ can be viewed as a precursor to a Church-encoding for the theory of
 signatures, but we only need contexts encoded in this way, and not other ToS
-components. In functional programming, this is sometimes called
-``finally tagless''\cite{carette2007finally}, and it is used for defining and
-interpreting embedded languages.
+components. In functional programming, this is sometimes called ``finally
+tagless'' \cite{carette2007finally}, and it is used for defining and interpreting
+embedded languages.
 \end{definition}
@@
 More interesting (and complicated) examples for recursive equations are boundary
-conditions in various cubical type theories \cite{TODO}. Note that our $\Id$
+conditions in various cubical type theories \cite{cohen2016cubical, angiuli2016computational, angiuli2018cartesian}. Note that our $\Id$
 allows iterated equations as well, but these are all trivial in the semantics,
@@
 \end{example}
+
+\emph{Remark.} Since signatures are parametrized by a single universe level, all
+external types in constructors must be contained in the same $\Set_j$
+universe. We opted for this setup for the sake of simplicity. Cumulativity helps
+here: it allows us to pick a $j$ level which is large enough to accommodate all
+external types in a signature.
 
@@
 
-In the following, for $i$ and $j$ levels we define a model $\bM_{i,j} : \ToS_{i,
-  \max(i,j+1)+1}$, such than $\Con_{\bM_{i,j}}$ is a type of structured categories (of
-algebras). The level $i$ marks the level of all external sets in function
-domains, and $j$ marks the level of all internal sorts in an algebra. Hence,
-every algebra has level $\max(i,j+1)$. The bump is only needed for $j$, since
-algebras merely contain elements of $A : \Set_i$ types, while inductive sets are
-themselves elements of $\Set_j$. For example, $NatAlg_j : \Set_{\max(0, j+1)} :
-\Set_{\max(0, j+1)+1}$,
+In the following, for given $i$ and $j$ levels, we define a model $\bM_{i,j} :
+\ToS_{\max(i+1,j)+1, j}$ such that $\Con_{\bM_{i,j}}$ is a type of structured
+categories (of algebras). The level $i$ marks the level of all internal sorts in
+an algebra, and the level $j$ marks the level of all external sets in function
+domains. Hence, every algebra has level $\max(i+1,j)$. The bump is only needed
+for $i$, since algebras merely contain elements of $A : \Set_j$ types, while
+inductive sets are themselves elements of $\Set_i$. For example, $NatAlg_i :
+\Set_{\max(i+1,0)} : \Set_{\max(i+1, 0)+1}$.
 
@@
 
-We define $\bCon : \Set_{\max(i,j+1)+1}$ as $\flCwF_{\max(i,j+1)}$.
+We define $\bCon : \Set_{\max(i+1,j)+1}$ as $\flCwF_{\max(i+1,j)}$.
 %% We define
@@
 \begin{definition}[Finite limit CwFs]\label{def:flCwF}
-For each level $i$ we define $\flCwF_i : \Set_{1+i}$ as an iterated $\Sigma$-type
+For each level $i$ we define $\flCwF_i : \Set_{i+1}$ as an iterated $\Sigma$-type
 with the following components:
@@
 
-\begin{theorem}[Equivalence of initiality and induction]
+\begin{theorem}[Equivalence of initiality and induction, c.f. \cite{kaposi2019constructing}]
 An object $\Gamma : \Con_{\bGamma}$ supports induction if and only if it is
 initial. Moreover, induction and initiality are both proof irrelevant
-predicates.
+predicates. \qed
 \end{theorem}
-\begin{proof} By brief internal flCwF reasoning in \cite{kaposi2019constructing}.
-\end{proof}
 
@@
 \subsection{Substitutions}
-
-\begin{definition}\label{def:weakmorphism} A \emph{weak flCwF morphism} $\boldsymbol{\sigma : \Sub\,\Gamma\,\Delta}$ is a functor between underlying categories, which additionally maps types to types and terms to terms, and satisfies the following:
+\label{sec:substitutions}
+
+A \emph{weak flCwF morphism} $\boldsymbol{\sigma : \Sub\,\Gamma\,\Delta}$ is a functor between underlying categories, which additionally maps types to types and terms to terms, and satisfies the following:
   \begin{enumerate}
@@
   \end{enumerate}
-\end{definition}
 
@@
 
+Note that the definition we just gave lives in $\Set_{\max(i+1,j)}$, but by
+cumulativity it is also in $\Set_{\max(i+1,j)+1}$, as required by our $\bM_{i,j}
+: \ToS_{\max(i+1,j)+1, j}$ specification of the model being defined.
+
 \begin{theorem}\label{thm:flpres} Every $\boldsymbol{\sigma : \Sub\,\Gamma\,\Delta}$ preserves fl-structure up to type isomorphism. That is, we have
@@
 
-We define $\boldsymbol{\Ty\,\Gamma} : \Set_{1 + \max(i, j+1)}$ as the type of
-split flCwF-isofibrations over $\bGamma$. We extend Ahrens' and Lumsdaine's
-displayed categories and their definition of isofibrations \cite{displayedcats}.
-We first define displayed flCwFs, then specify iso-cleaving as additional structure
-on top of that.
-
-\begin{definition}[Displayed flCwF] The type of displayed flCwFs at level $i$ is given as the logical predicate interpretation \cite{TODO} of $\flCwF_i$.
+We define $\boldsymbol{\Ty\,\Gamma} : \Set_{\max(i+1, j)+1}$ as the type of
+split flCwF-isofibrations over $\bGamma$, at level $\max(i+1, j)$. We extend
+Ahrens' and Lumsdaine's displayed categories and their definition of
+isofibrations \cite{displayedcats}.  We first define displayed flCwFs, then
+specify iso-cleaving as additional structure on top of that.
+
+\begin{definition}[Displayed flCwF] The type of displayed flCwFs at level $i$ is given as the logical predicate interpretation (see e.g.\ \cite{bernardy12parametricity} or \cite{hiit}) of $\flCwF_i$.
 For each flCwF component in $\bGamma$, there is a component in a displayed flCwF which ``lies over'' it.
 
-In situations where we need to refer to both ``base'' and displayed things, we give
-\ul{underlined} names to contexts, substitutions, types and terms in a base
-flCwF. For example, we may have $\ulGamma : \Con_{\bGamma}$ living in
-$\boldsymbol{\Gamma : \Con}$, and $\Gamma : \Con_{\bA}\,\ulGamma$ living in
-$\boldsymbol{A : \Ty\,\Gamma}$. We only use underlining on cETT variable names,
-and overload flCwF component names for displayed counterparts.
+\emph{Notation.} In situations where we need to refer to both ``base'' and
+displayed things, we give \ul{underlined} names to contexts, substitutions,
+types and terms in a base flCwF. For example, we may have $\ulGamma :
+\Con_{\bGamma}$ living in $\boldsymbol{\Gamma : \Con}$, and $\Gamma :
+\Con_{\bA}\,\ulGamma$ living in a displayed flCwF over $\bGamma$. We only use
+underlining on cETT variable names, and overload flCwF component names for
+displayed counterparts. For example, a $\Con$ component is named the same in
+a base flCwF and a displayed one.
 
@@
 types cannot be interpreted strictly, as we will see in Section
-\ref{TODO}. Strict preservation is expressed with metatheoretic equality, which
+\ref{sec:infinitaryfunction}. Strict preservation is expressed with metatheoretic equality, which
 is respected by all cETT constructions, but isomorphisms are not necessarily
@@
 
-\emph{Remark.} It is not possible to model types as fibrations or opfibrations, because
-we have no restriction on the variance of ToS types. For example, the type which extends
-a pointed set to a $Nat$-signature is invariant in the base signature.
+\emph{Remark.} It is not possible to model types as fibrations or opfibrations,
+because we have no restriction on the variance of ToS types. For example, the
+type which extends a pointed set to a natural number signature, is neither a
+fibration nor an opfibration.
 
@@
 
-\todo{iso-cleaving}
+Iso-cleaving is given iso-cleaving in $\bA$ and the action of $\bsigma$ on
+isomorphisms, e.g.\ we have $\coe_{\boldsymbol{A[\sigma]}}\,\ulsigma\,\Gamma
+:\equiv \coe_{\bA}\,(\bsigma\,\ulsigma)\,\Gamma$.
 
@@
 $\boldsymbol{A[\sigma\circ\delta]} = \boldsymbol{A[\sigma][\delta]}$, follows
-from Lemma \ref{lem:idcomppres} and split cleaving given by $\coe_{\id}$,
-$\coe_{\circ}$, $\coh_{\id}$ and $\coh_{\circ}$ in $\bA$.
+from Lemma \ref{lem:idcomppres} and split cleaving given by $\coe^{\id}$,
+$\coe^{\circ}$, $\coh^{\id}$ and $\coh^{\circ}$ laws in $\bA$.
 
@@
 
-$\boldsymbol{\Tm\,\Gamma\,A} : \Set_{1 + \max(i, 1+j)}$ is defined as the type of
+$\boldsymbol{\Tm\,\Gamma\,A} : \Set_{\max(i+1, j)+1}$ is defined as the type of
 \emph{weak flCwF sections} of $\bA$. The underlying functions of $\bt :
@@
 
-Similarly to Definition \ref{def:weakmorphism}, we denote the evident preservation
+Similarly to Section \ref{sec:substitutions}, we denote the evident preservation
 isomorphisms as $\bt_{\emptycon} : \bt\,\emptycon \simeq_{\id} \emptycon$ and
@@
 We define $\boldsymbol{\U : \Ty\,\Gamma}$ as the isofibration which is
-constantly $\bSet_j$. A constant isofibration does not actually depend on
-the base flCwF, and has trivial iso-cleaving where $\coe$-s are identity
-functions. Hence, we have $\Con_{\bU}\,\ulGamma :\equiv \Set_j$ and $\Sub_{\bU}\,\Gamma\,\Delta\,\ulsigma :\equiv \Gamma \ra \Delta$.
-
-\emph{Remark.} The type $\bTm\,\bGamma\,\bU$ is strictly equal to $\bSub\,\bGamma\,\bSet_j$, so in the following we will think about semantic elements of the universe as weak morphisms from $\bGamma$ to $\bSet_j$.
+constantly $\bSet_i$. A constant isofibration does not actually depend on the
+base flCwF, and has trivial iso-cleaving where $\coe$-s are identity
+functions. Hence, we have $\Con_{\bU}\,\ulGamma :\equiv \Set_i$ and
+$\Sub_{\bU}\,\Gamma\,\Delta\,\ulsigma :\equiv \Gamma \ra \Delta$.
+
+\emph{Remark.} The type $\bTm\,\bGamma\,\bU$ is strictly equal to
+$\bSub\,\bGamma\,\bSet_i$, so it is helpful to think about semantic elements of
+the universe as weak morphisms from $\bGamma$ to $\bSet_i$.
 
@@
 \boldsymbol{:} \bTy\,\bGamma$. We define this as a dependent product of
-isofibrations, indexed by a discrete domain. With a general $\bA \boldsymbol{:}
-\bTy\,\bGamma$ domain, $\bPii$ would not be definable, since variance issues
-preclude $\Pi$-types in the CwF of categories \cite{TODO}.
-
-%% _{(\bPii\,\ba\,\bB)}
+isofibrations, indexed by a discrete domain. The discreteness is essential: with
+a general $\bA \boldsymbol{:} \bTy\,\bGamma$ domain, $\bPii$ would not be
+definable because of variance issues. Indeed, the category of categories is not
+locally cartesian closed and does not support a general $\Pi$ type
+\cite{johnstone2002sketches}.
 
@@
 
-For $A : \Set_i$ and $\bB : A \ra \bTy\,\bGamma$, we define $\bPie\,A\,\bB
+For $A : \Set_j$ and $\bB : A \ra \bTy\,\bGamma$, we define $\bPie\,A\,\bB
 \boldsymbol{:} \bTy\,\bGamma$ as the $A$-indexed direct product of $\bB$. Since
@@
 \subsection{Infinitary Function Space}
-
-For $A : \Set_i$ and $\bb : A \ra \bTm\,\bGamma\,\bU$, we aim to define
+\label{sec:infinitaryfunction}
+
+For $A : \Set_j$ and $\bb : A \ra \bTm\,\bGamma\,\bU$, we aim to define
 $\bPiinf\,A\,\bb \boldsymbol{:} \bTm\,\bGamma\,\bU$. The underlying functions
@@
 
-With this, we have defined the $\bM_{i,j} : \ToS_{i, 1+\max(i,1+j)}$ model that we set out to define
-in Section \ref{sec:overview}.
+With this, we have defined the $\bM_{i,j} : \ToS_{\max(i+1,j)+1, j}$ model that
+we set out to define in Section \ref{sec:overview}.
 
@@
 
-\begin{definition}[Signature for ToS] For each level $i$,
-we define $\ToSSig_i : \Sig_{1+i}$, as the signature for the theory of
-signatures with external sets in $\Set_i$. This is a large and infinitary QIIT
-signature, as we have $\Pie$ and $\Piinf$ abstracting over $A : \Set_i$ and
+\begin{definition}[Signature for ToS] For each level $j$,
+we define $\ToSSig_j : \Sig_{j+i}$, as the signature for the theory of
+signatures with external sets in $\Set_j$. This is a large and infinitary QIIT
+signature, as we have $\Pie$ and $\Piinf$ abstracting over $A : \Set_j$ and
 branching with $A \ra \Ty\,\Gamma$ and $A \ra \Tm\,\Gamma\,\U$ respectively. We
@@
   &           && \ext (\Piinf : \Pii(\Gamma : Con)\\
-  &           && \hspace{4.2em}(\Pie \Set_i (\lambda\,A.\,(A \arrinf \Ty\,\Gamma)\arri\El(\Ty\,\Gamma))))\\
+  &           && \hspace{4.2em}(\Pie \Set_j (\lambda\,A.\,(A \arrinf \Ty\,\Gamma)\arri\El(\Ty\,\Gamma))))\\
   &           && ...
@@
 
-Now, for each $j$, the interpretation of $\ToSSig_i$ in $\bM_{1+i,j}$ yields an
+Now, for each $i$, the interpretation of $\ToSSig_j$ in $\bM_{i,j+1}$ yields an
 flCwF $\bGamma$ such that $\Con_{\bGamma} = \ToS_{i,j}$. In short, we can
@@
 computation of the interpretation and the fact that $\ToSSig$ is precisely the
-internal representation of $\ToS$. Also, since we have an flCwF of
-$\ToS_{i,j}$-algebras, we can use Definition \ref{def:induction} for the notion
-of $\ToS$-induction.
+internal representation of $\ToS$. Hence, we have self-description modulo the
+bumping of the $j$ level. Also, as we get an flCwF of $\ToS_{i,j}$-algebras, we
+can use Definition \ref{def:induction} for the notion of $\ToS$-induction.
 
@@
 
-In this section we construct QIITs from initial ToS-algebras. Before we can assume
-the existence of such algebras. First, we need cumulative subtyping for $\ToS$.
-\todo{foobar}
-
-\begin{lemma}[Cumulativity of $\ToS$] If $j \leq j'$, then $\ToS_{i,j} \leq \ToS_{i, j'}$.
+In this section we construct QIITs from initial ToS-algebras. For this, we need
+to assume the existence of such algebras.
+
+\subsection{Assuming Syntax for the Theory of Signatures}
+
+\begin{lemma}[Cumulativity of $\ToS$]\label{lemma:cumulativity} If $i \leq i'$, then $\ToS_{i,j} \leq \ToS_{i', j}$.
+  This follows from the definition of ToS and the subtyping rules in Section
+  \ref{sec:cumulativity}. \qed
 \end{lemma}
-\begin{proof}
-  The $j$ universe level only appears in function codomains, hence
-  the subtyping rules for functions and $\Sigma$ can be used to get
-  $\ToS$ subtyping.
-\end{proof}
-
-\todo{assumption in environment}
+
+\textbf{Assumption.} For each level $j$ and $j'$ such that $j+1 \leq j'$, we
+assume the existence of $\syn_j : \ToS_{j+1,j}$, and we assume that $\syn_j$,
+considered as an element of $\ToS_{j',j}$ by Lemma \ref{lemma:cumulativity}, is
+inductive in the sense of Definition \ref{def:induction}.
+
+We explain this assumption. The syntax for the theory of signatures is
+postulated at the lowest possible level $\ToS_{j+1,j}$. This is the lowest
+because signatures may contain $A : \Set_j$ types, and since we want to view the
+syntax as freely generated, its inductive sorts must be large enough to contain
+the $A$ types. Otherwise we would run into Russell's paradox. Then, the
+induction assumption says that we have induction at all levels larger than
+$j+1$.
+
+\begin{example} We have $\syn_0 : \ToS_{1, 0}$, which is the syntax of closed QIIT signatures.
+We want to define a function $\mathsf{length} : \Con_{\syn_0} \ra \mathbb{N}$ by induction, which returns the length of a (syntactic) signature as a metatheoretic natural number.
+To this end, we define a displayed ToS over $\syn_0$, where $\Con$ is defined as constantly $\mathbb{N}$, every other sort is defined as constantly $\top$, $\emptycon$ is defined as $0$ and $\Gamma\ext A$ is defined as $\Gamma + 1$.
+By the induction assumption, we get a ToS-section from $\syn_0$ to the displayed model, whose action on contexts is exactly the $\mathsf{length}$ function.
+Note that the induction assumption requires that the displayed model is at least at level 1, but this is not problematic because by cumulativity $\mathbb{N} : \Set_1$.
+\end{example}
+
+With $\syn$ at hand, we can use an alternative, more conventional representation of
+signatures.
+
+\begin{definition} We define $\SynSig_j : \Set_{j+1}$, the type of \emph{syntactic signatures} at $j$, as $\Con_{\syn_j}$.
+\end{definition}
+
+We can convert a signature to a syntactic one by interpreting it in some
+$\syn_j$ model, and we can convert in the other direction by using ToS-induction
+to interpret a $\Gamma : \Con_{\syn_j}$ in an arbitrary ToS model.
+
+%% is inductive
+%% when considered as the element of $\ToS_{j',j}$ by Lemma
+%% \ref{lemma:cumulativity}.
+
+%% In the induction assumption, we are relying on the fact that there is an flCwF
+%% whose objects are $\ToS_{j',j}$ algebras.
+
+%% Above, $\syn$ abbreviates ``syntax'' for $\ToS$. In the assumption of induction,
+%% we are using Definition \ref{def:induction}.
 
@@
 %% \emph{Assumption.}
+
+\section{Related Work}
+\label{sec:relatedwork}
+
+\section{Conclusion}
+\label{sec:conclusion}
 

### 2020-01-09 draft/paper.tex
paper

@@
 \usepackage{scalerel}
-
+\usepackage{stmaryrd}
 \usepackage{tikz-cd}
@@
 \newcommand{\subtype}{\mathsf{subtype}}
+\newcommand{\NatSig}{\mathsf{NatSig}}
 
@@
 \newcommand{\bM}{\boldsymbol{\mathsf{M}}}
+\newcommand{\bT}{\boldsymbol{\mathsf{T}}}
+\newcommand{\bS}{\boldsymbol{\mathsf{S}}}
 
@@
 \newcommand{\ulgamma}{\ul{\gamma}}
+\newcommand{\ulOmega}{\ul{\Omega}}
 \newcommand{\uldelta}{\ul{\delta}}
@@
 \newcommand{\coh}{\mathsf{coh}}
+\newcommand{\llb}{\llbracket}
+\newcommand{\rrb}{\rrbracket}
 
@@
 \begin{alignat*}{3}
-  & NatSig : \Sig_0 \\
-  & NatSig :\equiv \lambda (i :\mathbb{N})(M : \ToS_{i,0}).\\
+  & \NatSig : \Sig_0 \\
+  & \NatSig :\equiv \lambda (i :\mathbb{N})(M : \ToS_{i,0}).\\
   & \hspace{1em}(\emptycon_M\,\ext_M\, (N : \U_M) \,\ext_M\,(zero : \El_M\,N)\\
@@
 \begin{alignat*}{3}
-& TreeSig :\equiv \\
+& \mathsf{TreeSig} :\equiv \\
 & \hspace{1em}\emptycon \ext (Tree : \U)\\
@@
 \begin{alignat*}{3}
-  & RecEqSig :\equiv\\
+  & \mathsf{RecEqSig} :\equiv\\
   & \hspace{0.5em}\emptycon\,\ext (A : \U) \ext (a : \El\,A)\ext (f : \Pii\,(x : A)\,(\Id\,A\,x\,a\arri\El\,A))
@@
 
-\begin{theorem}[Equivalence of initiality and induction, c.f. \cite{kaposi2019constructing}]
+\begin{theorem}[Equivalence of initiality and induction, c.f. \cite{kaposi2019constructing}]\label{thm:initialind}
 An object $\Gamma : \Con_{\bGamma}$ supports induction if and only if it is
-initial. Moreover, induction and initiality are both proof irrelevant
+initial. Moreover, induction and initiality are both proof-irrelevant
 predicates. \qed
@@
 \subsection{Identity and Composition}
+\label{sec:idcomp}
 
@@
   \ra \Tm\,\Gamma\,(A[\sigma])}$ is given similarly to
-$\boldsymbol{\blank\circ\blank}$ in Section \ref{TODO}. Underlying functions
+$\boldsymbol{\blank\circ\blank}$ in Section \ref{sec:idcomp}. Underlying functions
 are given by function composition, and preservation morphisms are also similar:
@@
 \begin{example} We have $\syn_0 : \ToS_{1, 0}$, which is the syntax of closed QIIT signatures.
-We want to define a function $\mathsf{length} : \Con_{\syn_0} \ra \mathbb{N}$ by induction, which returns the length of a (syntactic) signature as a metatheoretic natural number.
+We want to define a function $\mathsf{length} : \Con_{\syn_0} \ra \mathbb{N}$ by induction, which returns the length of a syntactic context as a metatheoretic natural number.
 To this end, we define a displayed ToS over $\syn_0$, where $\Con$ is defined as constantly $\mathbb{N}$, every other sort is defined as constantly $\top$, $\emptycon$ is defined as $0$ and $\Gamma\ext A$ is defined as $\Gamma + 1$.
@@
 
-%% is inductive
-%% when considered as the element of $\ToS_{j',j}$ by Lemma
-%% \ref{lemma:cumulativity}.
-
-%% In the induction assumption, we are relying on the fact that there is an flCwF
-%% whose objects are $\ToS_{j',j}$ algebras.
-
-%% Above, $\syn$ abbreviates ``syntax'' for $\ToS$. In the assumption of induction,
-%% we are using Definition \ref{def:induction}.
-
-%% For each level $i$,
-%% \begin{enumerate}
-%%   \item We assume that $\tos_i : \ToS_{i, 1+i}$ exists.
-%%   \item For each $i'$ such that $(1+i) \leq i'$, we assume $\Induction
-%% \end{enumerate}
-
-%% \textbf{For each $i$, we assume that $\tos_i : \ToS_{i, 1+i}$ exists}. The reason for the
-%% $1 + i$ level of $\ToS$ sorts is that $\tos_i$ may contain $A : \Set_i$ types in
-%% $\Ty$, and since we intend $\tos_i$ to be a freely generated algebra, the sorts
-%% must be large enough to avoid size paradoxes.
-
-%% \textbf{For each $i$ and $j$, we assume $\mathsf{ToSind :
-
-%% \todo{assumption as environment}
-
-%% \emph{Assumption.}
+\subsection{Term Algebras}
+
+The basic idea is that initial algebras can be built from the terms of $\syn_j$.
+For example, consider the syntactic signature for natural numbers:
+\[
+\NatSig :\equiv \emptycon\ext(N : \U)\ext(zero : \El\,N)\ext (suc : N
+\arri \El\,N)
+\]
+The type $\Tm_{\syn}\,\NatSig\,(\El_{\syn}\,N)$ is isomorphic to the usual type
+of natural numbers, since, intuitively, such terms can only be built from
+iterated usage of $zero$ and $suc$. We will build a term algebra for each
+signature in this manner. First, we describe the interpretation of $\syn$ in the
+$\Set$ model.
+
+\begin{definition}[The $\Set$ model of $\ToS$] For each $j$, we have $\bS : \ToS_{j+1, j}$, which can be given by restricting the $\bM$ model
+of Section \ref{sec:semantics} so that we only have the first $\Con$ components
+in the interpretations for contexts, substitution, types, term, and we only have
+actions on contexts in the interpretations of term and substitution
+formers. Hence, we have:
+\begin{alignat*}{3}
+  & \Con_{\bS} && :\equiv \Set_{j+2} \\
+  & \Sub_{\bS}\,\Gamma\,\Delta && :\equiv \Gamma \ra \Delta\\
+  & \Ty_{\bS}\,\Gamma && :\equiv \Gamma \ra \Set_{j+2}\\
+  & \Tm_{\bS}\,\Gamma\,A && :\equiv (\gamma : \Gamma) \ra A\,\gamma
+\end{alignat*}
+\end{definition}
+
+For every $M : \ToS_{j+1,j}$, there is a unique strict $\ToS$-morphism from
+$\syn_j$ to $M$. This follows from the induction assumption on $\syn_j$ and
+Theorem \ref{thm:initialind}. We denote this morphism as
+$\llbracket\blank\rrbracket_M$.  For example, given $\Gamma : \Con_\syn$, we
+have $\llbracket\Gamma\rrbracket_M : \Con_M$.  Also, for every displayed
+$\ToS$-model $M$ over $\syn_j$, there is a strict $\ToS$-section of $M$. We also
+denote this as $\llbracket\blank\rrbracket_M$, so e.g.\ for $\ulGamma :
+\Con_\syn$ we have $\llbracket\ulGamma\rrbracket_M : \Con_M\,\ulGamma$.
+
+For a syntactic signature $\Gamma$, $\llb\Gamma\rrb_{\bS}$ yields exactly the
+type of $\Gamma$-algebras, e.g.\ $\llb\NatSig\rrb_{\bS}$ is a left-nested
+$\Sigma$-type, which is the type of pointed sets with an endofunction.
+
+\begin{definition}[Term algebra construction]
+For each syntactic signature $\ulOmega : \SynSig_j$, we define a displayed $\ToS$
+model over $\syn_j$, named $\bT$. The underlying sets are as follows:
+\begin{alignat*}{3}
+  & \Con\,\ulGamma :\equiv \Sub_{\syn}\,\ulOmega\,\ulGamma \ra \llb\ulGamma\rrb_{\bS}
+\end{alignat*}
+
+\end{definition}
 

### 2020-01-10 draft/paper.tex
list of related work

@@
 
+Cartmell \cite{gat}. Displayed categories Ahrens \cite{displayedcats}.
+
+Paolo paper \cite{gabe}. Finitary case \cite{kaposi2019constructing}.
+
+Direct construction of concrete QIIT: Streicher
+\cite{streicher2012semantics}, Brunerie \cite{brunerie}, Ambroise
+\cite{induction-is-enough}.
+
+Higher inductive types: HIT descriptions \cite{niels,hiit}, HITs in
+cubical type theories\cite{cubicalhits,10.1145/3290314}, semantics for
+subclasses of HITs
+\cite{lumsdaineShulman,10.1145/3209108.3209130,nielsmsc,sojakova,moeneclaey}
+Construction of example IITs without UIP: Jasper \cite{jasper}.
+
 \section{Conclusion}

### 2020-01-10 draft/paper.tex
paper

--- 2020-01-09 draft/paper.tex
@@
 \newcommand{\ulsigma}{\ul{\sigma}}
+\newcommand{\ulnu}{\ul{\nu}}
 \newcommand{\ulepsilon}{\ul{\epsilon}}
@@
 \newcommand{\ulA}{\ul{A}}
+\newcommand{\ula}{\ul{a}}
 \newcommand{\ulB}{\ul{B}}
@@
 \subsection{Universe}
+\label{sec:universe}
 
@@
 \end{example}
+
+For every $M : \ToS_{j+1,j}$, there is a unique strict $\ToS$-morphism from
+$\syn_j$ to $M$. This follows from the induction assumption on $\syn_j$ and
+Theorem \ref{thm:initialind}. We denote this morphism as
+$\llbracket\blank\rrbracket_M$.  For example, given $\Gamma : \Con_\syn$, we
+have $\llbracket\Gamma\rrbracket_M : \Con_M$.  Also, for every displayed
+$\ToS$-model $M$ over $\syn_j$, there is a strict $\ToS$-section of $M$. We also
+denote this as $\llbracket\blank\rrbracket_M$, so e.g.\ for $\ulGamma :
+\Con_\syn$ we have $\llbracket\ulGamma\rrbracket_M : \Con_M\,\ulGamma$.
 
@@
 \end{alignat*}
-\end{definition}
-
-For every $M : \ToS_{j+1,j}$, there is a unique strict $\ToS$-morphism from
-$\syn_j$ to $M$. This follows from the induction assumption on $\syn_j$ and
-Theorem \ref{thm:initialind}. We denote this morphism as
-$\llbracket\blank\rrbracket_M$.  For example, given $\Gamma : \Con_\syn$, we
-have $\llbracket\Gamma\rrbracket_M : \Con_M$.  Also, for every displayed
-$\ToS$-model $M$ over $\syn_j$, there is a strict $\ToS$-section of $M$. We also
-denote this as $\llbracket\blank\rrbracket_M$, so e.g.\ for $\ulGamma :
-\Con_\syn$ we have $\llbracket\ulGamma\rrbracket_M : \Con_M\,\ulGamma$.
-
-For a syntactic signature $\Gamma$, $\llb\Gamma\rrb_{\bS}$ yields exactly the
-type of $\Gamma$-algebras, e.g.\ $\llb\NatSig\rrb_{\bS}$ is a left-nested
-$\Sigma$-type, which is the type of pointed sets with an endofunction.
+Now, for $\Gamma : \Con_{\syn}$, the type of $\Gamma$-algebras is given by
+$\llb\Gamma\rrb_{\bS}$. E.g.\, $\llb\NatSig\rrb_{\bS}$ yields a left-nested
+$\Sigma$-type of pointed sets with an endofunction. Also, $\llb\Gamma\rrb_{\bM}$
+extends $\llb\Gamma\rrb_{\bS}$ to an flCwF of $\Gamma$-algebras.
+\end{definition}
 
@@
 \begin{alignat*}{3}
-  & \Con\,\ulGamma :\equiv \Sub_{\syn}\,\ulOmega\,\ulGamma \ra \llb\ulGamma\rrb_{\bS}
-\end{alignat*}
+  & \Con_{\bT}\,\ulGamma &&:\equiv \Sub\,\ulOmega\,\ulGamma \ra \llb\ulGamma\rrb_{\bS}\\
+  & \Sub_{\bT}\,\Gamma\,\Delta\,\ulsigma &&:\equiv (\ulnu : \Sub\,\ulOmega\,\ulGamma)
+  \ra \Delta\,(\ulsigma\circ\nu) \simeq \llb\ulsigma\rrb_{\bS}\,(\Gamma\,\ulnu)\\
+  & \Ty_{\bT}\,\Gamma\,\ulA &&:\equiv (\ulnu : \Sub\,\ulOmega\,\ulGamma) \ra
+    \Tm\,\ulOmega\,(\ulA[\ulnu]) \ra \llb\ulA\rrb_{\bS}\,(\Gamma\,\ulnu)\\
+  & \Tm_{\bT}\,\Gamma\,A\,\ult &&:\equiv (\ulnu : \Sub\,\ulOmega\,\ulGamma)\ra
+    \llb\ulA\rrb_{\bS}\,\ulnu\,(\ult[\ulnu]) \simeq_{\id} \llb\ult\rrb_{\bS} (\Gamma\,\ulnu)
+\end{alignat*}
+
+Above, the $\simeq$ in the definition of $\Sub_{\bT}$ is a context isomorphism
+in $\llb\ulDelta\rrb_{\bM}$, which is the flCwF of $\ulDelta$-algebras. The
+$\simeq_{\id}$ in $\Ty_{\bT}$ is a vertical context isomorphism in the displayed
+flCwF given by $\llb\ulA\rrb_{\bM}$.
+
+So far, the underlying sets in $\bT$ are similar to what was given in
+\cite{kaposi2019constructing} in the construction of term algebras, but there is
+an important difference: in ibid.\, strict equalities are used instead of
+isomorphisms. In our case, isomorphisms are necessary once again because of
+infinitary functions types and our identity type; we shall see this shortly. The
+universe is interpreted as follows:
+\begin{alignat*}{3}
+  & \U_{\bT} : (\ulnu : \Sub\,\ulOmega\,\ulGamma)(\ult : \Tm\,\ulOmega\,\U)\ra
+              \Set_{j+1}\\
+  & \U_{\bT}\,\ulnu\,\ult :\equiv \Tm\,\ulOmega\,(\El\,\ult)\\
+  & \El_{\bT}\,a : (\ulnu : \Sub\,\ulOmega\,\ulGamma)
+             (\ult : \Tm\,\ulOmega\,(\El\,(\ula[\ulnu])))\ra
+              \llb\ula\rrb_{\bS}\,(\Gamma\,\ulnu)\\
+  & \El_{\bT}\,a\,\ulnu\,\ult :\equiv (a\,\ulnu)\,\ult
+\end{alignat*}
+
+Hence, a syntactic $\ult : \Tm\,\ulOmega\,\U$ is interpreted as a set of terms
+with type $\El\,\ult$. In the interpretation of $\El$, note that
+\[
+  a\,\ulnu : \U_{\bT}\,\ulnu\,(\ula[\ulnu]) \simeq_{\id}
+                 \llb\ula\rrb_{\bS}\,(\Gamma\,\ulnu)
+\]
+hence
+\[
+a\,\ulnu : \Tm\,\ulOmega\,(\El\,(\ula[\ulnu])) \simeq_{\id}
+           \llb\ula\rrb_{\bS}\,(\Gamma\,\ulnu)
+\]
+The $\simeq_{\id}$ above is just an isomorphism of sets, since it lives in
+$\llb\U\rrb_{\bM}$ which was given as the flCwF of sets in Section
+\ref{sec:universe}. This above isomorphism is a good summary of the
+construction: the interpretation of a $\ula : \Tm\,\ulOmega\,\U$ in the term
+algebra is isomorphic to a set of terms.
+
+Inductive functions are interpreted by transport along such isomorphism:
+\[
+  \Pii_{\bT}\,a\,B\,\ulnu\,\ult :\equiv \lambda\,\alpha.\,
+         B\,(\ulnu,\,(a\,\ulnu)^{-1}\,\alpha)\,(\ult\,\appitt\,((a\,\ulnu)^{-1}\,\alpha))
+\]
+
+For the infinitary function space, we need the following, where $\simeq_{\id}$ is
+again set isomorphism.
+\begin{alignat*}{3}
+  & \Piinf_{\bT}\,A\,b\,\ulnu : \Tm\,\ulOmega\,(\El\,(\Piinf\,A\,(\lambda\,\alpha.\,(b\,\alpha)[\ulnu])))\\
+  & \hspace{4.2em}\simeq_{\id} ((\alpha : A) \ra \llb\alpha\rrb_{\bS}\,(\Gamma\,\ulnu))
+\end{alignat*}
+This can be given using the natural isomorphism consisting of $\appinf$ and
+$\laminf$. However, the sides are not strictly equal. For the identity type, we
+build the following isomorphism using equality reflection.
+\begin{alignat*}{3}
+  & \Id_{\bT}\,a\,t\,u\,\ulnu : \Tm\,\ulOmega\,(\El\,(\Id\,(\ula[\ulnu])\,(\ult[\ulnu])\,(\ulu[\ulnu]))\\
+  & \hspace{4.2em}\simeq_{\id} (\llb\ult\rrb_{\bS}\,(\Gamma\,\ulnu) =   \llb\ulu\rrb_{\bS}\,(\Gamma\,\ulnu))
+\end{alignat*}
+
+We omit the rest of the definition of $\bT$. The interpretations of equations in
+the CwF and fl-structure parts become fairly technical, and we also need to
+utilize iso-cleaving to interpret type substitution and substitution
+laws. However, the basic shape of the model remains similar to
+\cite{kaposi2019constructing}.
 
@@
 %% write about limitations!
+p

### 2020-01-10 draft/paper.tex
related work

--- 2020-01-10 draft/paper.tex
@@
 
-Cartmell \cite{gat}. Displayed categories Ahrens \cite{displayedcats}.
-
-Paolo paper \cite{gabe}. Finitary case \cite{kaposi2019constructing}.
-
-Direct construction of concrete QIIT: Streicher
-\cite{streicher2012semantics}, Brunerie \cite{brunerie}, Ambroise
-\cite{induction-is-enough}.
-
-Higher inductive types: HIT descriptions \cite{niels,hiit}, HITs in
-cubical type theories\cite{cubicalhits,10.1145/3290314}, semantics for
-subclasses of HITs
-\cite{lumsdaineShulman,10.1145/3209108.3209130,nielsmsc,sojakova,moeneclaey}
-Construction of example IITs without UIP: Jasper \cite{jasper}.
-
-\section{Conclusion}
+Cartmell \cite{gat} defines generalised algebraic theories (GATs)
+using type theoretic syntax. Compared to our QII signatures, his
+signatures support sort equations but do not cover infinitary
+constructors or recursive equations. A way to encode sort equations in
+our signatures is using isomorphisms instead of equalities. It seems
+that infinitary constructors and sort equations are not compatible in
+a setting with UIP. In contrast to our algebraic definition,
+Cartmell's signatures are given by presyntax, named variables and
+typing relations. He also does not construct categories of algebras
+explicitly, but is concerned with showing that the classifying
+category for GATs is that of contextual categories.
+
+A more semantic approach to QIITs is given by Altenkirch et al.\
+\cite{gabe}. They generalise the usual functorial semantics of
+inductive types to QIITs by considering towers of functors and
+building complete categories of algebras from such towers. Their
+notion of signature does not enforce strict positivity, hence
+describes a larger class of QII signatures. The equivalence of
+initiality and induction is showed, but initial algebras are not
+constructed.
+
+Kaposi et al's work \cite{kaposi2019constructing} is the most direct
+precursor of our work. They don't consider infinitary constructors or
+constructors with recursive equations which make their semantics much
+simpler.
+
+Streicher's initiality proof \cite{streicher2012semantics} is a direct
+construction of a concrete QIIT: that of the syntax of type theory. He
+uses preterms, typing relations and quotients to construct the initial
+model of type theory in a set-theoretic metalanguage. A recent
+formalisation of the same result for a larger type theory was done by
+Brunerie et al.\ \cite{brunerie} in Agda with UIP, function
+extensionality and quotients. Lafont et al.\
+\cite{induction-is-enough} construct all finitary inductive-inductive
+types in Agda using only UIP and function extensionality. Examples of
+inductive-inductive types are constructed without using UIP by Hugunin
+\cite{jasper}.
+
+Higher inductive types (HITs) are generalisations of QIITs in settings
+with proof-relevant identity types, in fact, they were introduced
+before QIITs \cite{HoTTbook}. \cite{niels} is a description of a
+subset of HITs. \cite{hiit} describes a syntax for higher
+inductive-inductive types using a theory of signatures similar to ours,
+however does not construct categories of algebras or show
+initiality. Semantics for different subclasses of HITs are given by
+\cite{lumsdaineShulman,10.1145/3209108.3209130,nielsmsc,sojakova,moeneclaey}. Cubical
+type theories were shown to support HITs in a computational way
+\cite{niels,hiit}.
+
+Our notion of displayed CwF is an extension of displayed category
+\cite{displayedcats} however in a setting with UIP.
+
+\section{Conclusions and further work}
 \label{sec:conclusion}
+
+Existence of the theory of signatures:
+
+\cite[Section 9]{lumsdaineShulman} shows that infinitary QIITs are not
+constructible in type theory without choice, so the existence of the
+theory of signatures is a non-conservative extension of extensional
+type theory. The setoid model supports an example QIT which branches
+over an arbitrary setoid, and this suggests that it might support the
+theory of signatures as well.
 

### 2020-01-10 draft/paper.tex
paper

--- 2020-01-10 draft/paper.tex
@@
 \newcommand{\bS}{\boldsymbol{\mathsf{S}}}
+\newcommand{\bP}{\boldsymbol{\mathsf{P}}}
+\newcommand{\bD}{\boldsymbol{\mathsf{D}}}
 
@@
 
+\subsection{Useful Model Fragments of $\bM$}
+
+First, we need to describe three model fragments of $\bM$, which can be
+respectively used to compute notions of algebras, displayed algebras and
+sections for each syntactic signature. This is a rephrasing of the
+$\blank^{\A}$, $\blank^{\D}$ and $\blank^{\S}$ interpretations in
+\cite{kaposi2019constructing}.
+
+\begin{definition}[The $\Set$ model of $\ToS$] For each $j$, we have $\bA : \ToS_{j+1, j}$, which can be given by restricting the $\bM$ model
+of Section \ref{sec:semantics} so that we only have the first $\Con$ components
+in the interpretations for contexts, substitution, types, term, and we only have
+actions on contexts in the interpretations of term and substitution
+formers. Hence, we have:
+\begin{alignat*}{3}
+  & \Con_{\bA} && :\equiv \Set_{j+2} \\
+  & \Sub_{\bA}\,\Gamma\,\Delta && :\equiv \Gamma \ra \Delta\\
+  & \Ty_{\bA}\,\Gamma && :\equiv \Gamma \ra \Set_{j+2}\\
+  & \Tm_{\bA}\,\Gamma\,A && :\equiv (\gamma : \Gamma) \ra A\,\gamma
+\end{alignat*}
+Now, for $\Gamma : \Con_{\syn}$, the type of $\Gamma$-algebras is given by
+$\llb\Gamma\rrb_{\bA}$. E.g.\, $\llb\NatSig\rrb_{\bA}$ yields a left-nested
+$\Sigma$-type of pointed sets with an endofunction. Also, $\llb\Gamma\rrb_{\bM}$
+extends $\llb\Gamma\rrb_{\bA}$ to an flCwF of $\Gamma$-algebras.
+\end{definition}
+
+\begin{definition}[Logical predicate model of $\ToS$ over the $\Set$ model]
+For each $j$ level we have $\bD$, which is a displayed $\ToS$ model over $\bA$
+at level $j$. This model, analogously to $\bA$, is given by restricting $\bM$ to
+the \emph{third} components everywhere, corresponding to types or actions on
+types. Hence, we have:
+\begin{alignat*}{3}
+  & \Con_{\bD}\,\ulGamma && :\equiv \ulGamma \ra \Set_{j+2} \\
+  & \Sub_{\bD}\,\Gamma\,\Delta\,\ulsigma && :\equiv (\ulgamma : \ulGamma) \ra \Gamma\,\gamma \ra \Delta(\llb\ulsigma\rrb_{\bA}\,\ulgamma)\\
+  & \Ty_{\bD}\,\Gamma\,\ulA && :\equiv (\ulgamma : \ulGamma) \ra \Gamma\,\ulgamma \ra \llb\ulA\rrb_{\bA}\,\ulgamma \ra \Set_{j+2}\\
+  & \Tm_{\bD}\,\Gamma\,A\,\ult && :\equiv (\ulgamma : \ulGamma)(\gamma : \Gamma\,\ulGamma)
+    \ra A\,\ulgamma\,\gamma\,(\llb\ult\rrb_{\bA}\,\ulgamma)
+\end{alignat*}
+\end{definition}
+
+For $\Gamma : \SynSig_j$, the type of displayed $\Gamma$-algebras over some
+$\ulgamma : \llb\Gamma\rrb_{\bA}$ is given by $\llb\Gamma\rrb_{\bD}\,\ulgamma$. In
+other words, $\llb\Gamma\rrb_{\bD}$ yields the notion of types in the flCwF of
+$\Gamma$-algebras given by $\llb\Gamma\rrb_{\bM}$.
+
+\begin{definition}[Displayed algebra section model of $\ToS$] Analogously to
+$\bA$ and $\bD$, we define $\bS$ as a displayed $\ToS$ model over the total
+model of $\bD$, which is given by restricting $\bM$ to the fourth components,
+corresponding to interpretations of terms and actions on terms.
+
+For $\ulgamma : \llb\Gamma\rrb_{\bA}$ and $\gamma : \llb\Gamma\rrb_{\bD}\,\ulgamma$,
+the type of $\Gamma$-sections is computed as $\llb\Gamma\rrb_{\bS}\,\ulgamma\,\gamma$.
+\end{definition}
+
 \subsection{Term Algebras}
@@
 
-\begin{definition}[The $\Set$ model of $\ToS$] For each $j$, we have $\bS : \ToS_{j+1, j}$, which can be given by restricting the $\bM$ model
-of Section \ref{sec:semantics} so that we only have the first $\Con$ components
-in the interpretations for contexts, substitution, types, term, and we only have
-actions on contexts in the interpretations of term and substitution
-formers. Hence, we have:
-\begin{alignat*}{3}
-  & \Con_{\bS} && :\equiv \Set_{j+2} \\
-  & \Sub_{\bS}\,\Gamma\,\Delta && :\equiv \Gamma \ra \Delta\\
-  & \Ty_{\bS}\,\Gamma && :\equiv \Gamma \ra \Set_{j+2}\\
-  & \Tm_{\bS}\,\Gamma\,A && :\equiv (\gamma : \Gamma) \ra A\,\gamma
-\end{alignat*}
-Now, for $\Gamma : \Con_{\syn}$, the type of $\Gamma$-algebras is given by
-$\llb\Gamma\rrb_{\bS}$. E.g.\, $\llb\NatSig\rrb_{\bS}$ yields a left-nested
-$\Sigma$-type of pointed sets with an endofunction. Also, $\llb\Gamma\rrb_{\bM}$
-extends $\llb\Gamma\rrb_{\bS}$ to an flCwF of $\Gamma$-algebras.
-\end{definition}
-
 \begin{definition}[Term algebra construction]
 For each syntactic signature $\ulOmega : \SynSig_j$, we define a displayed $\ToS$
-model over $\syn_j$, named $\bT$. The underlying sets are as follows:
-\begin{alignat*}{3}
-  & \Con_{\bT}\,\ulGamma &&:\equiv \Sub\,\ulOmega\,\ulGamma \ra \llb\ulGamma\rrb_{\bS}\\
-  & \Sub_{\bT}\,\Gamma\,\Delta\,\ulsigma &&:\equiv (\ulnu : \Sub\,\ulOmega\,\ulGamma)
-  \ra \Delta\,(\ulsigma\circ\nu) \simeq \llb\ulsigma\rrb_{\bS}\,(\Gamma\,\ulnu)\\
-  & \Ty_{\bT}\,\Gamma\,\ulA &&:\equiv (\ulnu : \Sub\,\ulOmega\,\ulGamma) \ra
-    \Tm\,\ulOmega\,(\ulA[\ulnu]) \ra \llb\ulA\rrb_{\bS}\,(\Gamma\,\ulnu)\\
-  & \Tm_{\bT}\,\Gamma\,A\,\ult &&:\equiv (\ulnu : \Sub\,\ulOmega\,\ulGamma)\ra
-    \llb\ulA\rrb_{\bS}\,\ulnu\,(\ult[\ulnu]) \simeq_{\id} \llb\ult\rrb_{\bS} (\Gamma\,\ulnu)
-\end{alignat*}
-
-Above, the $\simeq$ in the definition of $\Sub_{\bT}$ is a context isomorphism
+model over $\syn_j$, named $\bT_{\ulOmega}$. The underlying sets are as follows:
+\begin{alignat*}{3}
+  & \Con_{\bT_{\ulOmega}}\,\ulGamma &&:\equiv \Sub\,\ulOmega\,\ulGamma \ra \llb\ulGamma\rrb_{\bA}\\
+  & \Sub_{\bT_{\ulOmega}}\,\Gamma\,\Delta\,\ulsigma &&:\equiv (\ulnu : \Sub\,\ulOmega\,\ulGamma)
+  \ra \Delta\,(\ulsigma\circ\nu) \simeq \llb\ulsigma\rrb_{\bA}\,(\Gamma\,\ulnu)\\
+  & \Ty_{\bT_{\ulOmega}}\,\Gamma\,\ulA &&:\equiv (\ulnu : \Sub\,\ulOmega\,\ulGamma) \ra
+    \Tm\,\ulOmega\,(\ulA[\ulnu]) \ra \llb\ulA\rrb_{\bA}\,(\Gamma\,\ulnu)\\
+  & \Tm_{\bT_{\ulOmega}}\,\Gamma\,A\,\ult &&:\equiv (\ulnu : \Sub\,\ulOmega\,\ulGamma)\ra
+    \llb\ulA\rrb_{\bA}\,\ulnu\,(\ult[\ulnu]) \simeq_{\id} \llb\ult\rrb_{\bA} (\Gamma\,\ulnu)
+\end{alignat*}
+
+Above, the $\simeq$ in the definition of $\Sub_{\bT_{\ulOmega}}$ is a context isomorphism
 in $\llb\ulDelta\rrb_{\bM}$, which is the flCwF of $\ulDelta$-algebras. The
-$\simeq_{\id}$ in $\Ty_{\bT}$ is a vertical context isomorphism in the displayed
+$\simeq_{\id}$ in $\Ty_{\bT_{\ulOmega}}$ is a vertical context isomorphism in the displayed
 flCwF given by $\llb\ulA\rrb_{\bM}$.
 
-So far, the underlying sets in $\bT$ are similar to what was given in
+So far, the underlying sets in $\bT_{\ulOmega}$ are similar to what was given in
 \cite{kaposi2019constructing} in the construction of term algebras, but there is
@@
 \begin{alignat*}{3}
-  & \U_{\bT} : (\ulnu : \Sub\,\ulOmega\,\ulGamma)(\ult : \Tm\,\ulOmega\,\U)\ra
+  & \U_{\bT_{\ulOmega}} : (\ulnu : \Sub\,\ulOmega\,\ulGamma)(\ult : \Tm\,\ulOmega\,\U)\ra
               \Set_{j+1}\\
-  & \U_{\bT}\,\ulnu\,\ult :\equiv \Tm\,\ulOmega\,(\El\,\ult)\\
-  & \El_{\bT}\,a : (\ulnu : \Sub\,\ulOmega\,\ulGamma)
+  & \U_{\bT_{\ulOmega}}\,\ulnu\,\ult :\equiv \Tm\,\ulOmega\,(\El\,\ult)\\
+  & \El_{\bT_{\ulOmega}}\,a : (\ulnu : \Sub\,\ulOmega\,\ulGamma)
              (\ult : \Tm\,\ulOmega\,(\El\,(\ula[\ulnu])))\ra
-              \llb\ula\rrb_{\bS}\,(\Gamma\,\ulnu)\\
-  & \El_{\bT}\,a\,\ulnu\,\ult :\equiv (a\,\ulnu)\,\ult
+              \llb\ula\rrb_{\bA}\,(\Gamma\,\ulnu)\\
+  & \El_{\bT_{\ulOmega}}\,a\,\ulnu\,\ult :\equiv (a\,\ulnu)\,\ult
 \end{alignat*}
@@
 \[
-  a\,\ulnu : \U_{\bT}\,\ulnu\,(\ula[\ulnu]) \simeq_{\id}
-                 \llb\ula\rrb_{\bS}\,(\Gamma\,\ulnu)
+  a\,\ulnu : \U_{\bT_{\ulOmega}}\,\ulnu\,(\ula[\ulnu]) \simeq_{\id}
+                 \llb\ula\rrb_{\bA}\,(\Gamma\,\ulnu)
 \]
@@
 a\,\ulnu : \Tm\,\ulOmega\,(\El\,(\ula[\ulnu])) \simeq_{\id}
-           \llb\ula\rrb_{\bS}\,(\Gamma\,\ulnu)
+           \llb\ula\rrb_{\bA}\,(\Gamma\,\ulnu)
 \]
@@
 \[
-  \Pii_{\bT}\,a\,B\,\ulnu\,\ult :\equiv \lambda\,\alpha.\,
+  \Pii_{\bT_{\ulOmega}}\,a\,B\,\ulnu\,\ult :\equiv \lambda\,\alpha.\,
          B\,(\ulnu,\,(a\,\ulnu)^{-1}\,\alpha)\,(\ult\,\appitt\,((a\,\ulnu)^{-1}\,\alpha))
@@
 \begin{alignat*}{3}
-  & \Piinf_{\bT}\,A\,b\,\ulnu : \Tm\,\ulOmega\,(\El\,(\Piinf\,A\,(\lambda\,\alpha.\,(b\,\alpha)[\ulnu])))\\
-  & \hspace{4.2em}\simeq_{\id} ((\alpha : A) \ra \llb\alpha\rrb_{\bS}\,(\Gamma\,\ulnu))
+  & \Piinf_{\bT_{\ulOmega}}\,A\,b\,\ulnu : \Tm\,\ulOmega\,(\El\,(\Piinf\,A\,(\lambda\,\alpha.\,(b\,\alpha)[\ulnu])))\\
+  & \hspace{4.2em}\simeq_{\id} ((\alpha : A) \ra \llb\alpha\rrb_{\bA}\,(\Gamma\,\ulnu))
 \end{alignat*}
@@
 \begin{alignat*}{3}
-  & \Id_{\bT}\,a\,t\,u\,\ulnu : \Tm\,\ulOmega\,(\El\,(\Id\,(\ula[\ulnu])\,(\ult[\ulnu])\,(\ulu[\ulnu]))\\
-  & \hspace{4.2em}\simeq_{\id} (\llb\ult\rrb_{\bS}\,(\Gamma\,\ulnu) =   \llb\ulu\rrb_{\bS}\,(\Gamma\,\ulnu))
-\end{alignat*}
-
-We omit the rest of the definition of $\bT$. The interpretations of equations in
+  & \Id_{\bT_{\ulOmega}}\,a\,t\,u\,\ulnu : \Tm\,\ulOmega\,(\El\,(\Id\,(\ula[\ulnu])\,(\ult[\ulnu])\,(\ulu[\ulnu]))\\
+  & \hspace{4.2em}\simeq_{\id} (\llb\ult\rrb_{\bA}\,(\Gamma\,\ulnu) =   \llb\ulu\rrb_{\bA}\,(\Gamma\,\ulnu))
+\end{alignat*}
+
+We omit the rest of the definition of $\bT_{\ulOmega}$. The interpretations of equations in
 the CwF and fl-structure parts become fairly technical, and we also need to
@@
 \cite{kaposi2019constructing}.
-
-\end{definition}
+\end{definition}
+
+Now, we can build the term algebra for $\ulOmega$ by taking
+$\llb\ulOmega\rrb_{\bT_{\ulOmega}}\,\id$, which has type $\llb\ulOmega\rrb_{\bA}$.
+
+\emph{Remark.} If we start with a syntactic signature at level $j$, then the
+underlying sets in the term algebra are all in $\Set_{j+1}$. Hence, the term
+algebra for $\NatSig : \SynSig_0$ has an underlying set in $\Set_{j+1}$. This is
+a bit inconvenient, since normally we would have natural number in $\Set_0$. Our
+current term model construction cannot avoid this level bump, since $\syn_j$ is
+necessarily a large type, and we do not have a way to construct a small set from
+a large set of terms. Perhaps this would be possible with a \emph{resizing rule}
+\cite{voevodsky2011resizing}. Also, if we only consider closed finitary QIITs,
+with no possibility of referring to external types in signatures, then we can
+modify the current term model construction so that we always build sets in
+$\Set_0$. This would cover natural numbers and most dependent type theories.
+
+\subsection{Term Algebras Support Induction}
+
+%% We assume $\ulOmega : \SynSig_j$.
+%% We abbreviate the term algebra
+%% $\llb\ulOmega\rrb_{\bT_{\ulOmega}}\,\id$ as
+
+%% $\llb\ulOmega\rrb_{\bT_{\ulOmega}}\,\id$ as b
+%% $\llb\ulOmega\rrb_{\bT_{\ulOmega}}\,\id$ as
 
@@
 %% write about limitations!
-p

### 2020-01-10 draft/paper.tex
paper

@@
 \ref{lem:idcomppres}. We do not have to show functoriality of term substitution
-here, since that is derivable in any CwF \cite{TODO}.
+here, since that is derivable in any CwF, see e.g. \cite{kaposi2019constructing}.
 
@@
 
-First, we need to describe three model fragments of $\bM$, which can be
-respectively used to compute notions of algebras, displayed algebras and
-sections for each syntactic signature. This is a rephrasing of the
-$\blank^{\A}$, $\blank^{\D}$ and $\blank^{\S}$ interpretations in
+We will need in the following three model fragments of $\bM$, which can be used
+to compute notions of algebras, displayed algebras and sections respectively for
+each syntactic signature. This is a rephrasing of the $\blank^{\A}$,
+$\blank^{\D}$ and $\blank^{\S}$ interpretations in
 \cite{kaposi2019constructing}.
@@
 
-We omit the rest of the definition of $\bT_{\ulOmega}$. The interpretations of equations in
-the CwF and fl-structure parts become fairly technical, and we also need to
-utilize iso-cleaving to interpret type substitution and substitution
+We omit the rest of the definition of $\bT_{\ulOmega}$. The interpretations of
+equations in the CwF and the type formers are fairly technical, and we also
+need to utilize iso-cleaving to interpret type substitution and substitution
 laws. However, the basic shape of the model remains similar to

### 2020-01-10 draft/paper.tex

@@
 
-Cartmell \cite{gat}. Displayed categories Ahrens \cite{displayedcats}.
-
-Paolo paper \cite{gabe}. Finitary case \cite{kaposi2019constructing}.
-
-Direct construction of concrete QIIT: Streicher
-\cite{streicher2012semantics}, Brunerie \cite{brunerie}, Ambroise
-\cite{induction-is-enough}.
-
-Higher inductive types: HIT descriptions \cite{niels,hiit}, HITs in
-cubical type theories\cite{cubicalhits,10.1145/3290314}, semantics for
-subclasses of HITs
-\cite{lumsdaineShulman,10.1145/3209108.3209130,nielsmsc,sojakova,moeneclaey}
-Construction of example IITs without UIP: Jasper \cite{jasper}.
-
-\section{Conclusion}
+Cartmell \cite{gat} defines generalised algebraic theories (GATs)
+using type theoretic syntax. Compared to our QII signatures, his
+signatures support sort equations but do not cover infinitary
+constructors or recursive equations. A way to encode sort equations in
+our signatures is using isomorphisms instead of equalities. It seems
+that infinitary constructors and sort equations are not compatible in
+a setting with UIP. In contrast to our algebraic definition,
+Cartmell's signatures are given by presyntax, named variables and
+typing relations. He also does not construct categories of algebras
+explicitly, but is concerned with showing that the classifying
+category for GATs is that of contextual categories.
+
+A more semantic approach to QIITs is given by Altenkirch et al.\
+\cite{gabe}. They generalise the usual functorial semantics of
+inductive types to QIITs by considering towers of functors and
+building complete categories of algebras from such towers. Their
+notion of signature does not enforce strict positivity, hence
+describes a larger class of QII signatures. The equivalence of
+initiality and induction is showed, but initial algebras are not
+constructed.
+
+Kaposi et al's work \cite{kaposi2019constructing} is the most direct
+precursor of our work. They don't consider infinitary constructors or
+constructors with recursive equations which make their semantics much
+simpler.
+
+Streicher's initiality proof \cite{streicher2012semantics} is a direct
+construction of a concrete QIIT: that of the syntax of type theory. He
+uses preterms, typing relations and quotients to construct the initial
+model of type theory in a set-theoretic metalanguage. A recent
+formalisation of the same result for a larger type theory was done by
+Brunerie et al.\ \cite{brunerie} in Agda with UIP, function
+extensionality and quotients. Lafont et al.\
+\cite{induction-is-enough} construct all finitary inductive-inductive
+types in Agda using only UIP and function extensionality. Examples of
+inductive-inductive types are constructed without using UIP by Hugunin
+\cite{jasper}.
+
+Higher inductive types (HITs) are generalisations of QIITs in settings
+with proof-relevant identity types, in fact, they were introduced
+before QIITs \cite{HoTTbook}. \cite{niels} is a description of a
+subset of HITs. \cite{hiit} describes a syntax for higher
+inductive-inductive types using a theory of signatures similar to ours,
+however does not construct categories of algebras or show
+initiality. Semantics for different subclasses of HITs are given by
+\cite{lumsdaineShulman,10.1145/3209108.3209130,nielsmsc,sojakova,moeneclaey}. Cubical
+type theories were shown to support HITs in a computational way
+\cite{niels,hiit}.
+
+Our notion of displayed CwF is an extension of displayed category
+\cite{displayedcats} however in a setting with UIP.
+
+\section{Conclusions and further work}
+>>>>>>> eba72f32c6607b337c6315a124afe03f6f3ad92d
 \label{sec:conclusion}
+
+Existence of the theory of signatures:
+
+\cite[Section 9]{lumsdaineShulman} shows that infinitary QIITs are not
+constructible in type theory without choice, so the existence of the
+theory of signatures is a non-conservative extension of extensional
+type theory. The setoid model supports an example QIT which branches
+over an arbitrary setoid, and this suggests that it might support the
+theory of signatures as well.
 

### 2020-01-10 draft/paper.tex
paper
examples in intro

@@
 
-Categories are a prime example for a structure which is described by a quotient
-inductive-inductive signature. Signatures for QIITs allow having multiple sorts,
-with later ones indexed over previous ones, and equations as well. We need both
-features in order to write down the signature of categories.
+Categories are a prime example for a structure which is described by a
+quotient inductive-inductive signature. Signatures for QIITs allow
+having multiple sorts, with later ones indexed over previous ones, and
+equations as well. We need both features in order to write down the
+following signature of categories.
+\begin{alignat*}{3}
+  & Ob  && : \U \\
+  & Mor && : Ob\ra Ob\ra\U \\
+  & \blank\circ\blank && : Mor\,J\,K\ra Mor\,I\,J\ra Mor\,I\,K \\
+  & id && : Mor\,I\,I \\
+  & ass && : (f\circ g)\circ h = f\circ(g\circ h) \\
+  & idl && : id\circ f = f \\
+  & idr && : f\circ id = f
+\end{alignat*}
+The sort $Mor$ is indexed twice over the sort $Ob$, $ass$, $idl$ and
+$idr$ express equations between the constructors $\blank\circ\blank$
+and $id$.
 
@@
 and comprehension categories \cite{jacobs1993comprehension} to categories with
-families \cite{Dybjer96internaltype}. Additionally, real, surreal \cite{HoTTbook}
-and ordinal numbers \cite{lumsdaineShulman} can be defined as \emph{infinitary}
-QIITs, which are represented as possibly infinitely branching inductive trees.
+families \cite{Dybjer96internaltype} (CwF).
+%
+Whereas the initial category can be given by first defining the type
+of objects (the empty type) and then defining morphisms as a separate
+inductive family, for the initial CwF one has to define $Ob$, $Mor$
+and two extra sorts $Ty$ and $Tm$ \emph{inductive-inductively}, at the
+same time. This is because there is a constructor of $Ob$ which refers
+to $Ty$, constructors of $Ty$ refer to $Ob$ and $Sub$, and so on. This
+is illustrated by the following excerpt of the signature for CwFs
+extending the above signature for categories.
+\begin{alignat*}{3}
+  & Ty && : Ob\ra \U \\
+  & empty  && : Ob \\
+  & ext && : (\Gamma:Ob)\ra Ty\,\Gamma\ra Ob \\
+  & \blank[\blank] && : Ty\,\Delta\ra Mor\,\Gamma\,\Delta\ra Ty\,\Gamma
+\end{alignat*}
+
+While signatures for type theory usually only contain finitary
+constructors, important examples of QIITs have infinitary
+constructors, such as the following signature for infinitaly branching
+trees where the order of subtrees does not matter.
+\begin{alignat*}{3}
+  & Tree && : \U \\
+  & leaf  && : Tree \\
+  & node && : (\N\ra Tree)\ra Tree \\
+  & perm && : (g :\N\ra Tree)(f : \N\ra\N)\ra \mathsf{isIso}\,f\ra \\
+  & && \hspace{0.8em} node\,g = node (g\circ f)
+\end{alignat*}
+Here $node$ and $perm$ are infinitary constructurs, $perm$ expresses
+that reordering subtrees by a permutation (a $\N\ra\N$ function which
+is an isomorphism) does not matter.
+
+Real-world examples examples of infinitary QIITs include the real, surreal numbers \cite{HoTTbook}, ordinal numbers \cite{lumsdaineShulman} and the partiality monad \cite{partiality}.
 
@@
 \section{Conclusions and further work}
->>>>>>> eba72f32c6607b337c6315a124afe03f6f3ad92d
 \label{sec:conclusion}

### 2020-01-10 draft/paper.tex
rename metavars

@@
   & empty  && : Ob \\
-  & ext && : (\Gamma:Ob)\ra Ty\,\Gamma\ra Ob \\
-  & \blank[\blank] && : Ty\,\Delta\ra Mor\,\Gamma\,\Delta\ra Ty\,\Gamma
+  & ext && : (I:Ob)\ra Ty\,I\ra Ob \\
+  & \blank[\blank] && : Ty\,J\ra Mor\,I\,J\ra Ty\,I
 \end{alignat*}

### 2020-01-10 draft/paper.tex
paper

--- 2020-01-10 draft/paper.tex
@@
 
-\begin{definition}[The $\Set$ model of $\ToS$] For each $j$, we have $\bA : \ToS_{j+1, j}$, which can be given by restricting the $\bM$ model
+\begin{definition}[The $\Set$ model of $\ToS$] For each $i$ and $j$, we have $\bA : \ToS_{\max(i+1, j)+1, j}$, which can be given by restricting the $\bM_{i,j}$ model
 of Section \ref{sec:semantics} so that we only have the first $\Con$ components
@@
 \begin{alignat*}{3}
-  & \Con_{\bA} && :\equiv \Set_{j+2} \\
+  & \Con_{\bA} && :\equiv \Set_{\max(i+1, j)} \\
   & \Sub_{\bA}\,\Gamma\,\Delta && :\equiv \Gamma \ra \Delta\\
-  & \Ty_{\bA}\,\Gamma && :\equiv \Gamma \ra \Set_{j+2}\\
+  & \Ty_{\bA}\,\Gamma && :\equiv \Gamma \ra \Set_{\max(i+1, j)}\\
   & \Tm_{\bA}\,\Gamma\,A && :\equiv (\gamma : \Gamma) \ra A\,\gamma
 \end{alignat*}
-Now, for $\Gamma : \Con_{\syn}$, the type of $\Gamma$-algebras is given by
-$\llb\Gamma\rrb_{\bA}$. E.g.\, $\llb\NatSig\rrb_{\bA}$ yields a left-nested
-$\Sigma$-type of pointed sets with an endofunction. Also, $\llb\Gamma\rrb_{\bM}$
-extends $\llb\Gamma\rrb_{\bA}$ to an flCwF of $\Gamma$-algebras.
+Now, for some $\Gamma : \Con_{\syn_j}$, the type of $\Gamma$-algebras at level
+$i$ is given by $\llb\Gamma\rrb_{\bA}$, where we implicitly lift $\syn_j :
+\ToS_{j+1, j}$ to $\ToS_{\max(i+1,j)+1, j}$. E.g.\, $\llb\NatSig\rrb_{\bA}$
+yields a left-nested $\Sigma$-type of pointed sets with an endofunction. Also,
+$\llb\Gamma\rrb_{\bM}$ extends $\llb\Gamma\rrb_{\bA}$ to an flCwF of
+$\Gamma$-algebras.
 \end{definition}
@@
 \begin{definition}[Logical predicate model of $\ToS$ over the $\Set$ model]
-For each $j$ level we have $\bD$, which is a displayed $\ToS$ model over $\bA$
-at level $j$. This model, analogously to $\bA$, is given by restricting $\bM$ to
+For each $i$ and $j$ level we have $\bD$, which is a displayed $\ToS$ model over
+$\bA$. This model, analogously to $\bA$, is given by restricting $\bM_{i,j}$ to
 the \emph{third} components everywhere, corresponding to types or actions on
@@
 \begin{alignat*}{3}
-  & \Con_{\bD}\,\ulGamma && :\equiv \ulGamma \ra \Set_{j+2} \\
+  & \Con_{\bD}\,\ulGamma && :\equiv \ulGamma \ra \Set_{\max(i+1, j)} \\
   & \Sub_{\bD}\,\Gamma\,\Delta\,\ulsigma && :\equiv (\ulgamma : \ulGamma) \ra \Gamma\,\gamma \ra \Delta(\llb\ulsigma\rrb_{\bA}\,\ulgamma)\\
-  & \Ty_{\bD}\,\Gamma\,\ulA && :\equiv (\ulgamma : \ulGamma) \ra \Gamma\,\ulgamma \ra \llb\ulA\rrb_{\bA}\,\ulgamma \ra \Set_{j+2}\\
+  & \Ty_{\bD}\,\Gamma\,\ulA && :\equiv (\ulgamma : \ulGamma) \ra \Gamma\,\ulgamma \ra \llb\ulA\rrb_{\bA}\,\ulgamma \ra \Set_{\max(i+1, j)}\\
   & \Tm_{\bD}\,\Gamma\,A\,\ult && :\equiv (\ulgamma : \ulGamma)(\gamma : \Gamma\,\ulGamma)
@@
 
-For $\Gamma : \SynSig_j$, the type of displayed $\Gamma$-algebras over some
-$\ulgamma : \llb\Gamma\rrb_{\bA}$ is given by $\llb\Gamma\rrb_{\bD}\,\ulgamma$. In
-other words, $\llb\Gamma\rrb_{\bD}$ yields the notion of types in the flCwF of
-$\Gamma$-algebras given by $\llb\Gamma\rrb_{\bM}$.
+For $\Gamma : \SynSig_j$, the type of displayed $\Gamma$-algebras at level $i$
+over some $\ulgamma : \llb\Gamma\rrb_{\bA}$ is given by
+$\llb\Gamma\rrb_{\bD}\,\ulgamma$. Here, we also implicitly lift $\Gamma$ to live
+in the appropriately sized $\syn$. In other words, $\llb\Gamma\rrb_{\bD}$ yields
+the notion of types in the flCwF of $\Gamma$-algebras given by
+$\llb\Gamma\rrb_{\bM}$.
 
 \begin{definition}[Displayed algebra section model of $\ToS$] Analogously to
-$\bA$ and $\bD$, we define $\bS$ as a displayed $\ToS$ model over the total
-model of $\bD$, which is given by restricting $\bM$ to the fourth components,
-corresponding to interpretations of terms and actions on terms.
-
-For $\ulgamma : \llb\Gamma\rrb_{\bA}$ and $\gamma : \llb\Gamma\rrb_{\bD}\,\ulgamma$,
-the type of $\Gamma$-sections is computed as $\llb\Gamma\rrb_{\bS}\,\ulgamma\,\gamma$.
+$\bA$ and $\bD$, for each $i$ and $j$ levels we define $\bS$ as a displayed
+  $\ToS$ model over the total model of $\bD$, which is given by restricting
+  $\bM_{i,j}$ to the fourth components, corresponding to interpretations of terms and
+  actions on terms.
+
+For $\ulgamma : \llb\Gamma\rrb_{\bA}$ and $\gamma :
+\llb\Gamma\rrb_{\bD}\,\ulgamma$, the type of $\Gamma$-sections at level $i$ is
+computed as $\llb\Gamma\rrb_{\bS}\,\ulgamma\,\gamma$.
 \end{definition}
@@
 current term model construction cannot avoid this level bump, since $\syn_j$ is
-necessarily a large type, and we do not have a way to construct a small set from
+necessarily large, and we do not have a way to construct a small set from
 a large set of terms. Perhaps this would be possible with a \emph{resizing rule}
@@
 
+\subsection{Cumulativity of Algebras}
+
+We would like to show that term algebras are initial, but we want to do this
+\emph{on all universe levels}, i.e.\ that terms algebras are initial when lifted
+to any higher level. This requires showing that $\Gamma$-algebras are
+cumulative. We do this by induction on syntactic signatures.
+
+\begin{definition}[Cumulativity model] We assume $j$, $k$ and $l$ levels such that $j+1 \leq k$, $j+1 \leq l$ and $k \leq l$. We define a displayed model over $\syn_j : \ToS_{j+1, j}$ lifted to
+$\ToS_{l, j}$. In the following, we notate the level of algebras computed by
+  $\llb\blank\rrb_{\bA}$ as an extra index, as in
+  $\llb\ulGamma\rrb_{\bA_k}$. The underlying sets of the model are as follows.
+\begin{alignat*}{3}
+  & \Con\,\ulGamma && :\equiv \Subtype\,\llb\ulGamma\rrb_{\bA_k}\,\llb\ulGamma\rrb_{\bA_l}\\
+  & \Sub\,\Gamma\,\Delta\,\ulsigma && :\equiv (\gamma : \llb\ulGamma\rrb_{\bA_k})\ra
+  \llb\ulsigma\rrb_{\bA_k}\,\gamma = \llb\ulsigma\rrb_{\bA_l}\,\gamma\\
+  & \Ty\,\Gamma\,\ulA &&:\equiv (\gamma : \llb\ulGamma\rrb_{\bA_k})\ra
+      \Subtype\,(\llb\ulA\rrb_{\bA_k}\,\gamma)\,(\llb\ulA\rrb_{\bA_l}\,\gamma)\\
+  & \Tm\,\Gamma\,A\,\ult &&:\equiv (\gamma : \llb\ulGamma\rrb_{\bA_k})\ra
+     \llb\ult\rrb_{\bA_k}\,\gamma = \llb\ult\rrb_{\bA_l}\,\gamma
+\end{alignat*}
+The rest of the model is straightforward to define. Now, it follows from the
+induction assumption for $\syn$ and the reflection rule for $\Subtype$ in
+Section \ref{sec:cumulativity}, that $\llb\ulGamma\rrb_{\bA_k} \leq
+\llb\ulGamma\rrb_{\bA_l}$.
+
+\end{definition}
+
 \subsection{Term Algebras Support Induction}
-
-%% We assume $\ulOmega : \SynSig_j$.
-%% We abbreviate the term algebra
-%% $\llb\ulOmega\rrb_{\bT_{\ulOmega}}\,\id$ as
-
-%% $\llb\ulOmega\rrb_{\bT_{\ulOmega}}\,\id$ as b
-%% $\llb\ulOmega\rrb_{\bT_{\ulOmega}}\,\id$ as
 
@@
 
-Cartmell \cite{gat} defines generalised algebraic theories (GATs)
-using type theoretic syntax. Compared to our QII signatures, his
-signatures support sort equations but do not cover infinitary
-constructors or recursive equations. A way to encode sort equations in
-our signatures is using isomorphisms instead of equalities. It seems
-that infinitary constructors and sort equations are not compatible in
-a setting with UIP. In contrast to our algebraic definition,
-Cartmell's signatures are given by presyntax, named variables and
-typing relations. He also does not construct categories of algebras
-explicitly, but is concerned with showing that the classifying
-category for GATs is that of contextual categories.
+Cartmell \cite{gat} defines generalized algebraic theories (GATs) using type
+theoretic syntax. Compared to our QII signatures, his signatures support sort
+equations but do not cover infinitary constructors or recursive equations. A way
+to encode sort equations in our signatures is using isomorphisms instead of
+equalities. In contrast to our algebraic definition, Cartmell's signatures are
+given by presyntax, named variables and typing relations, there is no
+explicit model theory provided for signatures, and no explicit term model
+construction. Cartmell focuses instead on showing that contextual categories
+serve as classifying categories for GATs.
 
 A more semantic approach to QIITs is given by Altenkirch et al.\
-\cite{gabe}. They generalise the usual functorial semantics of
+\cite{gabe}. They generalize the initial algebra semantics of
 inductive types to QIITs by considering towers of functors and
@@
 
-Kaposi et al's work \cite{kaposi2019constructing} is the most direct
-precursor of our work. They don't consider infinitary constructors or
-constructors with recursive equations which make their semantics much
-simpler.
+Kaposi et al's work \cite{kaposi2019constructing} is the direct precursor of our
+work. They don not consider infinitary constructors or constructors with
+recursive equations, which make their semantics considerably simpler. They also
+do not provide a model theory of signatures, instead they assume signatures
+as an ad-hoc QIIT.
 
 Streicher's initiality proof \cite{streicher2012semantics} is a direct
-construction of a concrete QIIT: that of the syntax of type theory. He
-uses preterms, typing relations and quotients to construct the initial
-model of type theory in a set-theoretic metalanguage. A recent
-formalisation of the same result for a larger type theory was done by
-Brunerie et al.\ \cite{brunerie} in Agda with UIP, function
-extensionality and quotients. Lafont et al.\
-\cite{induction-is-enough} construct all finitary inductive-inductive
-types in Agda using only UIP and function extensionality. Examples of
+construction of a concrete QIIT: that of the syntax of type theory. He uses
+preterms, typing relations and quotients to construct the initial model of type
+theory in a set-theoretic metalanguage. A recent formalisation of the same
+result for a larger type theory was done by Brunerie et al.\ \cite{brunerie} in
+Agda with UIP, function extensionality and quotients. Lafont et
+al.\ \cite{induction-is-enough} construct all finitary inductive-inductive types
+in Agda using only UIP and function extensionality. Examples of
 inductive-inductive types are constructed without using UIP by Hugunin
@@
 
-Higher inductive types (HITs) are generalisations of QIITs in settings
-with proof-relevant identity types, in fact, they were introduced
-before QIITs \cite{HoTTbook}. \cite{niels} is a description of a
-subset of HITs. \cite{hiit} describes a syntax for higher
-inductive-inductive types using a theory of signatures similar to ours,
-however does not construct categories of algebras or show
-initiality. Semantics for different subclasses of HITs are given by
+Higher inductive types (HITs) are generalisations of QIITs in settings with
+proof-relevant identity types. They were introduced before QIITs
+\cite{HoTTbook}. \cite{niels} is a description of a subset of HITs. \cite{hiit}
+describes a syntax for higher inductive-inductive types using a theory of
+signatures similar to ours, however does not construct categories of algebras or
+show initiality. Semantics for different subclasses of HITs are given by
 \cite{lumsdaineShulman,10.1145/3209108.3209130,nielsmsc,sojakova,moeneclaey}. Cubical
@@
 
-Our notion of displayed CwF is an extension of displayed category
-\cite{displayedcats} however in a setting with UIP.
+Our notion of displayed CwF is an extension of displayed categories
+\cite{displayedcats}, although in a setting with UIP.
 
 \section{Conclusions and further work}
->>>>>>> eba72f32c6607b337c6315a124afe03f6f3ad92d
 \label{sec:conclusion}

### 2020-01-10 draft/paper.tex

@@
 and comprehension categories \cite{jacobs1993comprehension} to categories with
-families \cite{Dybjer96internaltype}. Additionally, real, surreal \cite{HoTTbook}
-and ordinal numbers \cite{lumsdaineShulman} can be defined as \emph{infinitary}
-QIITs, which are represented as possibly infinitely branching inductive trees.
+families \cite{Dybjer96internaltype} (CwF).
+%
+Whereas the initial category can be given by first defining the type
+of objects (the empty type) and then defining morphisms as a separate
+inductive family, for the initial CwF one has to define $Ob$, $Mor$
+and two extra sorts $Ty$ and $Tm$ \emph{inductive-inductively}, at the
+same time. This is because there is a constructor of $Ob$ which refers
+to $Ty$, constructors of $Ty$ refer to $Ob$ and $Sub$, and so on. This
+is illustrated by the following excerpt of the signature for CwFs
+extending the above signature for categories.
+\begin{alignat*}{3}
+  & Ty && : Ob\ra \U \\
+  & empty  && : Ob \\
+  & ext && : (I:Ob)\ra Ty\,I\ra Ob \\
+  & \blank[\blank] && : Ty\,J\ra Mor\,I\,J\ra Ty\,I
+\end{alignat*}
+
+While signatures for type theory usually only contain finitary
+constructors, important examples of QIITs have infinitary
+constructors, such as the following signature for infinitaly branching
+trees where the order of subtrees does not matter.
+\begin{alignat*}{3}
+  & Tree && : \U \\
+  & leaf  && : Tree \\
+  & node && : (\N\ra Tree)\ra Tree \\
+  & perm && : (g :\N\ra Tree)(f : \N\ra\N)\ra \mathsf{isIso}\,f\ra \\
+  & && \hspace{0.8em} node\,g = node (g\circ f)
+\end{alignat*}
+Here $node$ and $perm$ are infinitary constructurs, $perm$ expresses
+that reordering subtrees by a permutation (a $\N\ra\N$ function which
+is an isomorphism) does not matter.
+
+Real-world examples examples of infinitary QIITs include the real, surreal numbers \cite{HoTTbook}, ordinal numbers \cite{lumsdaineShulman} and the partiality monad \cite{partiality}.
 

### 2020-01-10 draft/paper.tex
paper

@@
 \newcommand{\bD}{\boldsymbol{\mathsf{D}}}
+\newcommand{\bI}{\boldsymbol{\mathsf{I}}}
 
@@
 \end{alignat*}
-The sort $Mor$ is indexed twice over the sort $Ob$, $ass$, $idl$ and
-$idr$ express equations between the constructors $\blank\circ\blank$
-and $id$.
-
-The benefit of having a QIIT signature is getting a model theory ``for free'',
+The benefit of having a QII signature is getting a model theory ``for free'',
 from the metatheory of QIITs. This model theory includes a category of algebras
@@
 \end{alignat*}
-
-While signatures for type theory usually only contain finitary
-constructors, important examples of QIITs have infinitary
-constructors, such as the following signature for infinitaly branching
-trees where the order of subtrees does not matter.
+While signatures for type theory usually only contain finitary constructors,
+some QIITs have infinitary constructors, such as the following signature for
+countable branching trees where the order of subtrees does not matter.
 \begin{alignat*}{3}
@@
 Real-world examples examples of infinitary QIITs include the real, surreal numbers \cite{HoTTbook}, ordinal numbers \cite{lumsdaineShulman} and the partiality monad \cite{partiality}.
-
-%% The main motivation of the current paper is to generalize previously formalized
-%% \cite{kaposi2019constructing} QIITs so that they accommodate more algebraic
-%% formulations of type theories, and more aspects of their practical
-%% metatheory. As a side effect of fulfilling this goal, infinitary QIITs such as
-%% Cauchy real numbers \cite{TODO} are covered as well.
 
@@
   of constructors. These have occurred previously in syntaxes of cubical type
-  theories, as boundary conditions\cite{TODO}.
+  theories, as boundary conditions \cite{cohen2016cubical, angiuli2016computational, angiuli2018cartesian}.
 \end{enumerate}
@@
 
-\textbf{Assumption.} For each level $j$ and $j'$ such that $j+1 \leq j'$, we
+\textbf{Assumption.} For each level $j$ and $k$ such that $j+1 \leq k$, we
 assume the existence of $\syn_j : \ToS_{j+1,j}$, and we assume that $\syn_j$,
-considered as an element of $\ToS_{j',j}$ by Lemma \ref{lemma:cumulativity}, is
+considered as an element of $\ToS_{k,j}$ by Lemma \ref{lemma:cumulativity}, is
 inductive in the sense of Definition \ref{def:induction}.
@@
 $\llb\Gamma\rrb_{\bM}$ extends $\llb\Gamma\rrb_{\bA}$ to an flCwF of
-$\Gamma$-algebras.
+$\Gamma$-algebras, and $\llb\Gamma\rrb_{\bA} = \Con_{\llb\Gamma\rrb_{\bM}}$.
 \end{definition}
@@
 the notion of types in the flCwF of $\Gamma$-algebras given by
-$\llb\Gamma\rrb_{\bM}$.
+$\llb\Gamma\rrb_{\bM}$, so we have $\llb\Gamma\rrb_{\bD} = \Ty_{\llb\Gamma\rrb_{\bM}}$.
 
@@
 \llb\Gamma\rrb_{\bD}\,\ulgamma$, the type of $\Gamma$-sections at level $i$ is
-computed as $\llb\Gamma\rrb_{\bS}\,\ulgamma\,\gamma$.
+computed as $\llb\Gamma\rrb_{\bS}\,\ulgamma\,\gamma$, and we have
+$\llb\Gamma\rrb_{\bS} = \Tm_{\llb\Gamma\rrb_{\bM}}$.
 \end{definition}
@@
 underlying sets in the term algebra are all in $\Set_{j+1}$. Hence, the term
-algebra for $\NatSig : \SynSig_0$ has an underlying set in $\Set_{j+1}$. This is
-a bit inconvenient, since normally we would have natural number in $\Set_0$. Our
+algebra for $\NatSig : \SynSig_0$ has an underlying set in $\Set_1$. This is
+a bit inconvenient, since normally we would have natural numbers in $\Set_0$. Our
 current term model construction cannot avoid this level bump, since $\syn_j$ is
@@
 \emph{on all universe levels}, i.e.\ that terms algebras are initial when lifted
-to any higher level. This requires showing that $\Gamma$-algebras are
+to any higher level. This requires showing that QII algebras are
 cumulative. We do this by induction on syntactic signatures.
@@
 $\ToS_{l, j}$. In the following, we notate the level of algebras computed by
-  $\llb\blank\rrb_{\bA}$ as an extra index, as in
+  $\llb\blank\rrb_{\bA}$ with an extra index, as in
   $\llb\ulGamma\rrb_{\bA_k}$. The underlying sets of the model are as follows.
@@
 \llb\ulGamma\rrb_{\bA_l}$.
-
 \end{definition}
@@
 \subsection{Term Algebras Support Induction}
+
+\begin{definition} We assume $j$ and $k$ such that $j + 1 \leq k$, and we also assume
+$\ulOmega : \SynSig_j$ and $\gamma :
+\llb\ulOmega\rrb_{\bD_k}\,(\llb\ulOmega\rrb_{\bT_{\ulOmega}}\,\id)$.  Hence, $\gamma$ is
+a displayed $\ulOmega$-algebra over the term algebra, at level $k$. We are using
+the cumulativity of $\ulOmega$ here to lift the term algebra appropriately.
+We aim to show that $\gamma$ has a section. We define a displayed model over $\syn_j$ lifted
+to $\ToS_{k, j}$, which we name $\bI_{\ulOmega}$. The underlying sets are:
+\begin{alignat*}{3}
+  & \Con_{\bI_{\ulOmega}}\,\ulGamma &&:\equiv (\ulnu : \Sub\,\ulOmega\,\ulGamma)\ra
+  \llb\ulGamma\rrb_{\bS}\,(\llb\ulnu\rrb_{\bA}\,(\llb\ulOmega\rrb_{\bT_{\ulOmega}}\,\id))\,\gamma\\
+  & \Sub_{\bI_{\ulOmega}}\,\Gamma\,\Delta\,\ulsigma && :\equiv (\ulnu : \Sub\,\ulOmega\,\ulGamma)\ra
+    \Delta\,(\ulsigma\circ\ulnu) = \llb\ulsigma\rrb_{\bS}\,(\Gamma\,\ulnu)\\
+  & \Ty_{\bI_{\ulOmega}}\,\Gamma\,\ulA &&:\equiv
+  (\ulnu : \Sub\,\ulOmega\,\ulGamma)(t : \Tm\,\ulOmega\,(\ulA[\ulnu]))\\
+  & && \hspace{1.2em}\ra
+  \llb\ulA\rrb_{\bS}\,(\llb\ult\rrb_{\bA}\,(\llb\ulOmega\rrb_{\bT_{\ulOmega}}\,\id))\,
+  (\llb\ult\rrb_{\bD}\,\gamma)\,(\Gamma\,\ulnu)\\
+  & \Tm_{\bI_{\ulOmega}}\,\Gamma\,A\,\ult &&:\equiv A\,\ulnu\,(\ult[\ulnu]) = \llb\ult\rrb_{\bS}\,(\Gamma\,\ulnu)
+\end{alignat*}
+
+Here, there is no essential change compared to \cite{kaposi2019constructing},
+and we follow ibid. in the definition of $\bI_{\ulOmega}$. The reason is that although we
+have weakened strict algebra equality to isomorphism, in the current
+construction we only have to show equalities of substitutions and terms, which
+we do not need to weaken (and they cannot be sensibly weakened anyway).
+\end{definition}
+
+\begin{theorem}[Initiality of term algebras] For each $j$ and $k$ such that $j + 1 \leq k$, and $\ulOmega : \SynSig_j$, the term algebra given by $\llb\ulOmega\rrb_{\bT_{\ulOmega}}\,\id$ is initial at level $k$.
+\end{theorem}
+\begin{proof}
+For each $\gamma :
+\llb\ulOmega\rrb_{\bD_k}\,(\llb\ulOmega\rrb_{\bT_{\ulOmega}}\,\id)$, we have
+$\llb\ulOmega\rrb_{\bI_{\ulOmega}}\,\id :
+\llb\ulGamma\rrb_{\bS}\,(\llb\ulOmega\rrb_{\bT_{\ulOmega}}\,\id)\,\gamma$. Hence,
+term algebras are inductive in the sense of Definition \ref{def:induction}, and
+by Theorem \ref{thm:initialind} they are also initial.
+\end{proof}
 
@@
 
-Kaposi et al's work \cite{kaposi2019constructing} is the direct precursor of our
-work. They don not consider infinitary constructors or constructors with
-recursive equations, which make their semantics considerably simpler. They also
-do not provide a model theory of signatures, instead they assume signatures
-as an ad-hoc QIIT.
+The work of Kaposi et al. \cite{kaposi2019constructing} is the direct precursor
+of our work. They do not consider infinitary constructors or constructors with
+recursive equations, which makes their semantics considerably simpler. They also
+do not provide a model theory of signatures, instead they assume signatures as
+an ad-hoc QIIT.
 
 Streicher's initiality proof \cite{streicher2012semantics} is a direct
-construction of a concrete QIIT: that of the syntax of type theory. He uses
+construction of a concrete QIIT: that of the syntax of a type theory. He uses
 preterms, typing relations and quotients to construct the initial model of type
-theory in a set-theoretic metalanguage. A recent formalisation of the same
+theory in a set-theoretic metalanguage. A recent formalization of the same
 result for a larger type theory was done by Brunerie et al.\ \cite{brunerie} in
@@
 
-Higher inductive types (HITs) are generalisations of QIITs in settings with
+Higher inductive types (HITs) are generalizations of QIITs in settings with
 proof-relevant identity types. They were introduced before QIITs
@@
 \end{document}
-
-%% B : Set j
-
-%% (A : Set_i) -> B
-
-%% A : Seti
-%% B : Set j
-%% (x : A) -> B : Set(max i j)
-
-%% (Con : Setj)x
-%% (Ty  : Con -> Setj)x
-
-%% (Pie : (A : Seti) -> (A -> Ty Gamma) -> Ty Gamma
-%% : Set(max(suci, j))
-
-%% : Set(max(sucj, max(suci,j)))
-%%    max(1+j, max(1+i, j))
-%%  = max(1+i, 1+j)
-
-%% REVISE LEVELS!!
-
-%% syntax levels:
-
-  %% ToS(i, 1+i) : Set(max(1+i, 2+i))
-
-%% REWRITE metatheory
-
-%% write about limitations!

### 2020-01-10 draft/paper.tex
paper

@@
 
-We can convert a signature to a syntactic one by interpreting it in some
-$\syn_j$ model, and we can convert in the other direction by using ToS-induction
-to interpret a $\Gamma : \Con_{\syn_j}$ in an arbitrary ToS model.
+We can convert a signature to a syntactic one by interpreting it in $\syn_j$,
+and we can convert in the other direction by using ToS-induction to interpret a
+$\Gamma : \Con_{\syn_j}$ in an arbitrary ToS model.
 

### 2020-01-10 draft/paper.tex
= instead of \equiv

@@
 algebras. For example, consider natural number algebras at level $i$, given as
-the $\Sigma$-type $NatAlg_i :\equiv(Nat : \Set_i)\times Nat\times(Nat \ra
+the $\Sigma$-type $NatAlg_i :=(Nat : \Set_i)\times Nat\times(Nat \ra
 Nat)$. It follows from the subtyping rules that $i \leq j$ implies $NatAlg_{i}
@@
       (u : \Tm\,\Gamma\,(\El\,a)) \ra \Tm\,\Gamma\,(B[\id, u])\\
-  & t\appitt u && :\equiv (\appi\,t)[\id, u]
+  & t\appitt u && := (\appi\,t)[\id, u]
 \end{alignat*}
@@
 %% \begin{alignat*}{3}
-%%   & A\arri B && :\equiv \Pi\,A\,(B[\p])\\
-%%   & A \arre\,B &&:\equiv \Pie\,A\,(\lambda\,\_.\,B) \\
-%%   & A \arrinf\,B &&:\equiv \Piinf\,A\,(\lambda\,\_.\,B)
-%% $A\arri B :\equiv \Pi\,A\,(B[\p])$, and $A \arre\,B :\equiv \Pie\,A\,(\lambda\,\_.\,B)$, and likewise
-%% $A \arrinf\,B :\equiv \Piinf\,A\,(\lambda\,\_.\,B)$.
+%%   & A\arri B && := \Pi\,A\,(B[\p])\\
+%%   & A \arre\,B &&:= \Pie\,A\,(\lambda\,\_.\,B) \\
+%%   & A \arrinf\,B &&:= \Piinf\,A\,(\lambda\,\_.\,B)
+%% $A\arri B := \Pi\,A\,(B[\p])$, and $A \arre\,B := \Pie\,A\,(\lambda\,\_.\,B)$, and likewise
+%% $A \arrinf\,B := \Piinf\,A\,(\lambda\,\_.\,B)$.
 
@@
 \[
-  \Sig_j :\equiv (i : \mathbb{N})\ra(M : \ToS_{i,j})\ra \Con_M
+  \Sig_j := (i : \mathbb{N})\ra(M : \ToS_{i,j})\ra \Con_M
 \]
@@
   & \NatSig : \Sig_0 \\
-  & \NatSig :\equiv \lambda (i :\mathbb{N})(M : \ToS_{i,0}).\\
+  & \NatSig := \lambda (i :\mathbb{N})(M : \ToS_{i,0}).\\
   & \hspace{1em}(\emptycon_M\,\ext_M\, (N : \U_M) \,\ext_M\,(zero : \El_M\,N)\\
@@
 \begin{alignat*}{3}
-& \mathsf{TreeSig} :\equiv \\
+& \mathsf{TreeSig} := \\
 & \hspace{1em}\emptycon \ext (Tree : \U)\\
@@
 %% \begin{alignat*}{3}
-%%   & cETTSig :\equiv \\
+%%   & cETTSig := \\
 %%   & ...\,\ext\,Tm{\Lift} : \Pii\,(\Gamma : Con)\,(\Pii\,(A : Ty{\appitt}\Gamma)((\IdU\,(Tm{\appitt}\Gamma{\appitt}A)\,(Tm{\appitt}\Gamma{\appitt}(\Lift_j{\appitt}A))))
@@
 \begin{alignat*}{3}
-  & \mathsf{RecEqSig} :\equiv\\
+  & \mathsf{RecEqSig} :=\\
   & \hspace{0.5em}\emptycon\,\ext (A : \U) \ext (a : \El\,A)\ext (f : \Pii\,(x : A)\,(\Id\,A\,x\,a\arri\El\,A))
@@
   & \Inductive : \Con_{\bGamma} \ra \Set_i \\
-  & \Inductive\,\Gamma :\equiv (A : \Ty_{\bGamma}\,\Gamma)\ra \Tm_{\bGamma}\,\Gamma\,A
+  & \Inductive\,\Gamma := (A : \Ty_{\bGamma}\,\Gamma)\ra \Tm_{\bGamma}\,\Gamma\,A
 \end{alignat*}
@@
 \begin{alignat*}{3}
-  & (\boldsymbol{\sigma \circ \delta})_{\emptycon}^{-1} :\equiv
+  & (\boldsymbol{\sigma \circ \delta})_{\emptycon}^{-1} :=
     \bsigma\,\bdelta_{\emptycon}^{-1} \circ \bdelta_{\emptycon}^{-1} \\
-  & (\boldsymbol{\sigma \circ \delta})_{\ext}^{-1} :\equiv
+  & (\boldsymbol{\sigma \circ \delta})_{\ext}^{-1} :=
     \bsigma\,\bdelta_{\ext}^{-1} \circ \bdelta_{\ext}^{-1}
@@
 \begin{alignat*}{3}
-  & \Con_{\boldsymbol{A[\sigma]}}\,\ulGamma && :\equiv \Con_{\bA}\,(\bsigma\,\ulGamma)\\
-  & \Sub_{\boldsymbol{A[\sigma]}}\,\Gamma\,\Delta\,\ulsigma && :\equiv
+  & \Con_{\boldsymbol{A[\sigma]}}\,\ulGamma && := \Con_{\bA}\,(\bsigma\,\ulGamma)\\
+  & \Sub_{\boldsymbol{A[\sigma]}}\,\Gamma\,\Delta\,\ulsigma && :=
     \Sub_{\bA}\,\Gamma\,\Delta\,(\bsigma\,\ulsigma)\\
-  & \Ty_{\boldsymbol{A[\sigma]}}\,\Gamma\,\ulA && :\equiv
+  & \Ty_{\boldsymbol{A[\sigma]}}\,\Gamma\,\ulA && :=
       \Ty_{\bA}\,\Gamma\,(\bsigma\,\ulA)\\
-  & \Tm_{\boldsymbol{A[\sigma]}}\,\Gamma\,A\,\ult && :\equiv
+  & \Tm_{\boldsymbol{A[\sigma]}}\,\Gamma\,A\,\ult && :=
       \Tm_{\bA}\,\Gamma\,A\,(\bsigma\,\ult)
@@
 
-Moreover, $\id_{\boldsymbol{A[\sigma]}} :\equiv \id_{\bA}$,
-$\sigma\circ_{\boldsymbol{A[\sigma]}}\delta :\equiv \sigma\circ_{\bA}\delta$,
+Moreover, $\id_{\boldsymbol{A[\sigma]}} := \id_{\bA}$,
+$\sigma\circ_{\boldsymbol{A[\sigma]}}\delta := \sigma\circ_{\bA}\delta$,
 and likewise substitution components are given by corresponding substitution
@@
 \begin{alignat*}{3}
-  &\emptycon_{\boldsymbol{A[\sigma]}} && :\equiv
+  &\emptycon_{\boldsymbol{A[\sigma]}} && :=
     \coe\,\bsigma_{\emptycon}^{-1}\,\emptycon_{\bA}\\
-  &\Gamma\ext_{\boldsymbol{A[\sigma]}}A && :\equiv
+  &\Gamma\ext_{\boldsymbol{A[\sigma]}}A && :=
     \coe\,\bsigma_{\ext}^{-1}\,(\Gamma\ext_{\bA} A)\\
-  &\Id_{\boldsymbol{A[\sigma]}}\,t\,u && :\equiv
+  &\Id_{\boldsymbol{A[\sigma]}}\,t\,u && :=
     \coe\,\bsigma_{\Id}^{-1}\,(\Id_{\bA}\,t\,u)
@@
 \begin{alignat*}{3}
-  & \epsilon_{\boldsymbol{A[\sigma]}} && :\equiv
+  & \epsilon_{\boldsymbol{A[\sigma]}} && :=
     \coh\,\bsigma_{\emptycon}^{-1}\,\emptycon_{\bA} \circ \epsilon_{\bA}\\
-  & \p_{\boldsymbol{A[\sigma]}} && :\equiv
+  & \p_{\boldsymbol{A[\sigma]}} && :=
     \p_{\bA} \circ (\coh\,\bsigma_{\ext}^{-1}\,(\Gamma\ext A))^{-1}\\
-  & \appK_{\boldsymbol{A[\sigma]}}\,t && :\equiv
+  & \appK_{\boldsymbol{A[\sigma]}}\,t && :=
     \appK_{\bA}\,((\coh\,\bsigma_{\K}\,(\K\,\Delta))^{-1}\circ t)
@@
 isomorphisms, e.g.\ we have $\coe_{\boldsymbol{A[\sigma]}}\,\ulsigma\,\Gamma
-:\equiv \coe_{\bA}\,(\bsigma\,\ulsigma)\,\Gamma$.
+:= \coe_{\bA}\,(\bsigma\,\ulsigma)\,\Gamma$.
 
@@
 \begin{alignat*}{3}
-  & (\boldsymbol{t[\sigma]})_{\emptycon}^{-1} :\equiv
+  & (\boldsymbol{t[\sigma]})_{\emptycon}^{-1} :=
     \bt\,\bsigma_{\emptycon}^{-1} \circ \bt_{\emptycon}^{-1} \\
-  & (\boldsymbol{t[\sigma]})_{\ext}^{-1} :\equiv
+  & (\boldsymbol{t[\sigma]})_{\ext}^{-1} :=
     \bt\,\bsigma_{\ext}^{-1} \circ \bt_{\ext}^{-1}
@@
   A)\,(A[\p])}$ is likewise a strict flCwF section given by second projections.
-Substitution extension $\boldsymbol{\sigma,\,t}$ is given by pointwise combining $\bsigma$ and $\bt$ with metatheoretic $\Sigma$ pairing, e.g.\ $\Con_{\boldsymbol{(\sigma,t)}}\,\ulGamma :\equiv (\bsigma\,\ulGamma,\,\bt\,\ulGamma)$.
+Substitution extension $\boldsymbol{\sigma,\,t}$ is given by pointwise combining $\bsigma$ and $\bt$ with metatheoretic $\Sigma$ pairing, e.g.\ $\Con_{\boldsymbol{(\sigma,t)}}\,\ulGamma := (\bsigma\,\ulGamma,\,\bt\,\ulGamma)$.
 
@@
 
-\begin{definition} For a level $i$, we write $\bSet_i$ for the flCwF of sets where $\Con_{\bSet_i} :\equiv \Set_i$ and $\Sub_{\bSet_i}\,\Gamma\,\Delta :\equiv \Gamma \ra \Delta$.
+\begin{definition} For a level $i$, we write $\bSet_i$ for the flCwF of sets where $\Con_{\bSet_i} := \Set_i$ and $\Sub_{\bSet_i}\,\Gamma\,\Delta := \Gamma \ra \Delta$.
 \end{definition}
@@
 base flCwF, and has trivial iso-cleaving where $\coe$-s are identity
-functions. Hence, we have $\Con_{\bU}\,\ulGamma :\equiv \Set_i$ and
-$\Sub_{\bU}\,\Gamma\,\Delta\,\ulsigma :\equiv \Gamma \ra \Delta$.
+functions. Hence, we have $\Con_{\bU}\,\ulGamma := \Set_i$ and
+$\Sub_{\bU}\,\Gamma\,\Delta\,\ulsigma := \Gamma \ra \Delta$.
 
@@
 \begin{alignat*}{3}
-  & \Con_{\bEl\,\ba}\,\ulGamma && :\equiv \ba\,\ulGamma\\
-  & \Sub_{\bEl\,\ba}\,\Gamma\,\Delta\,\ulsigma && :\equiv \ba\,\ulsigma\,\Gamma = \Delta\\
-  & \Ty_{\bEl\,\ba}\,\Gamma\,\ulA && :\equiv \ba\,\ulA\,\Gamma\\
-  & \Tm_{\bEl\,\ba}\,\Gamma\,A\,\ult && :\equiv \ba\,\ult\,\Gamma = A
+  & \Con_{\bEl\,\ba}\,\ulGamma && := \ba\,\ulGamma\\
+  & \Sub_{\bEl\,\ba}\,\Gamma\,\Delta\,\ulsigma && := \ba\,\ulsigma\,\Gamma = \Delta\\
+  & \Ty_{\bEl\,\ba}\,\Gamma\,\ulA && := \ba\,\ulA\,\Gamma\\
+  & \Tm_{\bEl\,\ba}\,\Gamma\,A\,\ult && := \ba\,\ult\,\Gamma = A
 \end{alignat*}
@@
 \begin{alignat*}{3}
-  & \emptycon_{\bEl\,\ba} && :\equiv \ba_{\emptycon}^{-1}\,\tt\\
-  & (\Gamma\ext_{\bEl\,\ba} A) && :\equiv \ba_{\ext}^{-1}\,(\Gamma,\,A)
+  & \emptycon_{\bEl\,\ba} && := \ba_{\emptycon}^{-1}\,\tt\\
+  & (\Gamma\ext_{\bEl\,\ba} A) && := \ba_{\ext}^{-1}\,(\Gamma,\,A)
 \end{alignat*}
 We likewise use preservation isomorphisms to define $\K$, $\Id$ and $\Sigma$.
-Context coercion is $\coe\,\ulsigma\,\Gamma :\equiv \ba\,\ulsigma\,\Gamma$. Type coercion, for $A : \ba\,\ulA\,\Gamma$ is given as $\coe\,\ult\,A :\equiv \ba\,\ult\,(\ba_{\ext}^{-1}\,(\Gamma,\,A))$.
+Context coercion is $\coe\,\ulsigma\,\Gamma := \ba\,\ulsigma\,\Gamma$. Type coercion, for $A : \ba\,\ulA\,\Gamma$ is given as $\coe\,\ult\,A := \ba\,\ult\,(\ba_{\ext}^{-1}\,(\Gamma,\,A))$.
 
@@
 \begin{alignat*}{3}
-  & \Con_{(\bPii\,\ba\,\bB)}\,\ulGamma &&:\equiv (\gamma : \ba\,\ulGamma)\ra \Con_{\bB}\,(\ulGamma, \gamma)\\
-  & \Ty_{(\bPii\,\ba\,\bB)}\,\Gamma\,\ulA &&:\equiv (\gamma : \ba\,\ulGamma)(a : \ba\,\ulA\,\gamma)\ra \Ty_{\bB}\,(\Gamma\,\gamma)\,(\ulA, a)
+  & \Con_{(\bPii\,\ba\,\bB)}\,\ulGamma &&:= (\gamma : \ba\,\ulGamma)\ra \Con_{\bB}\,(\ulGamma, \gamma)\\
+  & \Ty_{(\bPii\,\ba\,\bB)}\,\Gamma\,\ulA &&:= (\gamma : \ba\,\ulGamma)(a : \ba\,\ulA\,\gamma)\ra \Ty_{\bB}\,(\Gamma\,\gamma)\,(\ulA, a)
 \end{alignat*}
@@
 \begin{alignat*}{3}
-  & \Sub_{(\bPii\,\ba\,\bB)}\,\Gamma\,\Delta\,\ulsigma :\equiv
+  & \Sub_{(\bPii\,\ba\,\bB)}\,\Gamma\,\Delta\,\ulsigma :=
   (\gamma : \ba\,\ulGamma)(\delta : \ba\,\ulDelta)
@@
 \begin{alignat*}{3}
-  & \Sub_{(\bPii\,\ba\,\bB)}\,\Gamma\,\Delta\,\ulsigma :\equiv
+  & \Sub_{(\bPii\,\ba\,\bB)}\,\Gamma\,\Delta\,\ulsigma :=
   (\gamma : \ba\,\ulGamma)\ra\Sub_{\bB}\,(\Gamma\,\gamma)\,(\Delta\,(\ba\,\ulsigma\,\gamma)\,(\ulsigma, \refl)
@@
 \begin{alignat*}{3}
-  & \coe\,\ulsigma\,\Gamma &&:\equiv
+  & \coe\,\ulsigma\,\Gamma &&:=
     \lambda\,\gamma.\,\coe_{\bB}\,(\ulsigma,\refl)\,(\Gamma\,(\ba\,(\ulsigma^{-1})\,\gamma))\\
-  & \coe\,\ult\,A &&:\equiv
+  & \coe\,\ult\,A &&:=
     \lambda\,\gamma\,a.\,\coe_{\bB}\,(\ult,\refl)\,(A\,(\ba\,(\ult^{-1})\,(\ba_{\ext}^{-1}(\gamma,a))))
@@
 \begin{alignat*}{3}
-  & (\bPiinf\,A\,\bb)\,\ulGamma    &&:\equiv (a : A)\ra \bb\,a\,\ulGamma\\
-  & (\bPiinf\,A\,\bb)\,\ulsigma    &&:\equiv \lambda\,a.\, \bb\,a\,\ulsigma\\
-  & (\bPiinf\,A\,\bb)\,\ulA\       &&:\equiv \lambda\,\Gamma.\,(a : A)\ra \bb\,a\,\ulA\,(\Gamma\, a)\\
-  & (\bPiinf\,A\,\bb)\,\ult        &&:\equiv \lambda\,a.\, \bb\,a\,\ult
+  & (\bPiinf\,A\,\bb)\,\ulGamma    &&:= (a : A)\ra \bb\,a\,\ulGamma\\
+  & (\bPiinf\,A\,\bb)\,\ulsigma    &&:= \lambda\,a.\, \bb\,a\,\ulsigma\\
+  & (\bPiinf\,A\,\bb)\,\ulA\       &&:= \lambda\,\Gamma.\,(a : A)\ra \bb\,a\,\ulA\,(\Gamma\, a)\\
+  & (\bPiinf\,A\,\bb)\,\ult        &&:= \lambda\,a.\, \bb\,a\,\ult
 \end{alignat*}
@@
   %% &                               && : \top\ra (a : A) \ra \bb\,a\,\emptycon\\
-  &(\bPiinf\,A\,\bb)_{\emptycon}^{-1} && :\equiv \lambda\,\_\,a.\,(\bb\,a)_{\emptycon}^{-1}\,\tt\\
+  &(\bPiinf\,A\,\bb)_{\emptycon}^{-1} && := \lambda\,\_\,a.\,(\bb\,a)_{\emptycon}^{-1}\,\tt\\
   &(\bPiinf\,A\,\bb)_{\ext}^{-1} && : (\Gamma : (\bPiinf\,A\,\bb)\,\ulGamma)\times((\bPiinf\,A\,\bb)\,\ulA\,\Gamma)\\
   & && \hspace{0.5em}\ra (\bPiinf\,A\,\bb)\,(\ulGamma \ext \ulA)\\
-  & (\bPiinf\,A\,\bb)_{\ext}^{-1} && :\equiv \lambda\,(\Gamma,A)\,a.\,(\bb\,a)_{\ext}^{-1}(\Gamma\,a,\,A\,a)
+  & (\bPiinf\,A\,\bb)_{\ext}^{-1} && := \lambda\,(\Gamma,A)\,a.\,(\bb\,a)_{\ext}^{-1}(\Gamma\,a,\,A\,a)
 \end{alignat*}
@@
 \begin{alignat*}{3}
-& (\bId\,\bt\,\bu)\,\ulGamma &&:\equiv \bt\,\ulGamma = \bu\,\ulGamma\\
-& (\bId\,\bt\,\bu)\,\ulA     && :\equiv \lambda\,e.\, \bt\,\ulA = \bu\,\ulA
+& (\bId\,\bt\,\bu)\,\ulGamma &&:= \bt\,\ulGamma = \bu\,\ulGamma\\
+& (\bId\,\bt\,\bu)\,\ulA     && := \lambda\,e.\, \bt\,\ulA = \bu\,\ulA
 \end{alignat*}
@@
 \begin{alignat*}{3}
-  & \Con_{\bA} && :\equiv \Set_{\max(i+1, j)} \\
-  & \Sub_{\bA}\,\Gamma\,\Delta && :\equiv \Gamma \ra \Delta\\
-  & \Ty_{\bA}\,\Gamma && :\equiv \Gamma \ra \Set_{\max(i+1, j)}\\
-  & \Tm_{\bA}\,\Gamma\,A && :\equiv (\gamma : \Gamma) \ra A\,\gamma
+  & \Con_{\bA} && := \Set_{\max(i+1, j)} \\
+  & \Sub_{\bA}\,\Gamma\,\Delta && := \Gamma \ra \Delta\\
+  & \Ty_{\bA}\,\Gamma && := \Gamma \ra \Set_{\max(i+1, j)}\\
+  & \Tm_{\bA}\,\Gamma\,A && := (\gamma : \Gamma) \ra A\,\gamma
 \end{alignat*}
@@
 \begin{alignat*}{3}
-  & \Con_{\bD}\,\ulGamma && :\equiv \ulGamma \ra \Set_{\max(i+1, j)} \\
-  & \Sub_{\bD}\,\Gamma\,\Delta\,\ulsigma && :\equiv (\ulgamma : \ulGamma) \ra \Gamma\,\gamma \ra \Delta(\llb\ulsigma\rrb_{\bA}\,\ulgamma)\\
-  & \Ty_{\bD}\,\Gamma\,\ulA && :\equiv (\ulgamma : \ulGamma) \ra \Gamma\,\ulgamma \ra \llb\ulA\rrb_{\bA}\,\ulgamma \ra \Set_{\max(i+1, j)}\\
-  & \Tm_{\bD}\,\Gamma\,A\,\ult && :\equiv (\ulgamma : \ulGamma)(\gamma : \Gamma\,\ulGamma)
+  & \Con_{\bD}\,\ulGamma && := \ulGamma \ra \Set_{\max(i+1, j)} \\
+  & \Sub_{\bD}\,\Gamma\,\Delta\,\ulsigma && := (\ulgamma : \ulGamma) \ra \Gamma\,\gamma \ra \Delta(\llb\ulsigma\rrb_{\bA}\,\ulgamma)\\
+  & \Ty_{\bD}\,\Gamma\,\ulA && := (\ulgamma : \ulGamma) \ra \Gamma\,\ulgamma \ra \llb\ulA\rrb_{\bA}\,\ulgamma \ra \Set_{\max(i+1, j)}\\
+  & \Tm_{\bD}\,\Gamma\,A\,\ult && := (\ulgamma : \ulGamma)(\gamma : \Gamma\,\ulGamma)
     \ra A\,\ulgamma\,\gamma\,(\llb\ult\rrb_{\bA}\,\ulgamma)
@@
 \[
-\NatSig :\equiv \emptycon\ext(N : \U)\ext(zero : \El\,N)\ext (suc : N
+\NatSig := \emptycon\ext(N : \U)\ext(zero : \El\,N)\ext (suc : N
 \arri \El\,N)
@@
 \begin{alignat*}{3}
-  & \Con_{\bT_{\ulOmega}}\,\ulGamma &&:\equiv \Sub\,\ulOmega\,\ulGamma \ra \llb\ulGamma\rrb_{\bA}\\
-  & \Sub_{\bT_{\ulOmega}}\,\Gamma\,\Delta\,\ulsigma &&:\equiv (\ulnu : \Sub\,\ulOmega\,\ulGamma)
+  & \Con_{\bT_{\ulOmega}}\,\ulGamma &&:= \Sub\,\ulOmega\,\ulGamma \ra \llb\ulGamma\rrb_{\bA}\\
+  & \Sub_{\bT_{\ulOmega}}\,\Gamma\,\Delta\,\ulsigma &&:= (\ulnu : \Sub\,\ulOmega\,\ulGamma)
   \ra \Delta\,(\ulsigma\circ\nu) \simeq \llb\ulsigma\rrb_{\bA}\,(\Gamma\,\ulnu)\\
-  & \Ty_{\bT_{\ulOmega}}\,\Gamma\,\ulA &&:\equiv (\ulnu : \Sub\,\ulOmega\,\ulGamma) \ra
+  & \Ty_{\bT_{\ulOmega}}\,\Gamma\,\ulA &&:= (\ulnu : \Sub\,\ulOmega\,\ulGamma) \ra
     \Tm\,\ulOmega\,(\ulA[\ulnu]) \ra \llb\ulA\rrb_{\bA}\,(\Gamma\,\ulnu)\\
-  & \Tm_{\bT_{\ulOmega}}\,\Gamma\,A\,\ult &&:\equiv (\ulnu : \Sub\,\ulOmega\,\ulGamma)\ra
+  & \Tm_{\bT_{\ulOmega}}\,\Gamma\,A\,\ult &&:= (\ulnu : \Sub\,\ulOmega\,\ulGamma)\ra
     \llb\ulA\rrb_{\bA}\,\ulnu\,(\ult[\ulnu]) \simeq_{\id} \llb\ult\rrb_{\bA} (\Gamma\,\ulnu)
@@
               \Set_{j+1}\\
-  & \U_{\bT_{\ulOmega}}\,\ulnu\,\ult :\equiv \Tm\,\ulOmega\,(\El\,\ult)\\
+  & \U_{\bT_{\ulOmega}}\,\ulnu\,\ult := \Tm\,\ulOmega\,(\El\,\ult)\\
   & \El_{\bT_{\ulOmega}}\,a : (\ulnu : \Sub\,\ulOmega\,\ulGamma)
@@
               \llb\ula\rrb_{\bA}\,(\Gamma\,\ulnu)\\
-  & \El_{\bT_{\ulOmega}}\,a\,\ulnu\,\ult :\equiv (a\,\ulnu)\,\ult
+  & \El_{\bT_{\ulOmega}}\,a\,\ulnu\,\ult := (a\,\ulnu)\,\ult
 \end{alignat*}
@@
 \[
-  \Pii_{\bT_{\ulOmega}}\,a\,B\,\ulnu\,\ult :\equiv \lambda\,\alpha.\,
+  \Pii_{\bT_{\ulOmega}}\,a\,B\,\ulnu\,\ult := \lambda\,\alpha.\,
          B\,(\ulnu,\,(a\,\ulnu)^{-1}\,\alpha)\,(\ult\,\appitt\,((a\,\ulnu)^{-1}\,\alpha))
@@
 \begin{alignat*}{3}
-  & \Con\,\ulGamma && :\equiv \Subtype\,\llb\ulGamma\rrb_{\bA_k}\,\llb\ulGamma\rrb_{\bA_l}\\
-  & \Sub\,\Gamma\,\Delta\,\ulsigma && :\equiv (\gamma : \llb\ulGamma\rrb_{\bA_k})\ra
+  & \Con\,\ulGamma && := \Subtype\,\llb\ulGamma\rrb_{\bA_k}\,\llb\ulGamma\rrb_{\bA_l}\\
+  & \Sub\,\Gamma\,\Delta\,\ulsigma && := (\gamma : \llb\ulGamma\rrb_{\bA_k})\ra
   \llb\ulsigma\rrb_{\bA_k}\,\gamma = \llb\ulsigma\rrb_{\bA_l}\,\gamma\\
-  & \Ty\,\Gamma\,\ulA &&:\equiv (\gamma : \llb\ulGamma\rrb_{\bA_k})\ra
+  & \Ty\,\Gamma\,\ulA &&:= (\gamma : \llb\ulGamma\rrb_{\bA_k})\ra
       \Subtype\,(\llb\ulA\rrb_{\bA_k}\,\gamma)\,(\llb\ulA\rrb_{\bA_l}\,\gamma)\\
-  & \Tm\,\Gamma\,A\,\ult &&:\equiv (\gamma : \llb\ulGamma\rrb_{\bA_k})\ra
+  & \Tm\,\Gamma\,A\,\ult &&:= (\gamma : \llb\ulGamma\rrb_{\bA_k})\ra
      \llb\ult\rrb_{\bA_k}\,\gamma = \llb\ult\rrb_{\bA_l}\,\gamma
@@
 \begin{alignat*}{3}
-  & \Con_{\bI_{\ulOmega}}\,\ulGamma &&:\equiv (\ulnu : \Sub\,\ulOmega\,\ulGamma)\ra
+  & \Con_{\bI_{\ulOmega}}\,\ulGamma &&:= (\ulnu : \Sub\,\ulOmega\,\ulGamma)\ra
   \llb\ulGamma\rrb_{\bS}\,(\llb\ulnu\rrb_{\bA}\,(\llb\ulOmega\rrb_{\bT_{\ulOmega}}\,\id))\,\gamma\\
-  & \Sub_{\bI_{\ulOmega}}\,\Gamma\,\Delta\,\ulsigma && :\equiv (\ulnu : \Sub\,\ulOmega\,\ulGamma)\ra
+  & \Sub_{\bI_{\ulOmega}}\,\Gamma\,\Delta\,\ulsigma && := (\ulnu : \Sub\,\ulOmega\,\ulGamma)\ra
     \Delta\,(\ulsigma\circ\ulnu) = \llb\ulsigma\rrb_{\bS}\,(\Gamma\,\ulnu)\\
-  & \Ty_{\bI_{\ulOmega}}\,\Gamma\,\ulA &&:\equiv
+  & \Ty_{\bI_{\ulOmega}}\,\Gamma\,\ulA &&:=
   (\ulnu : \Sub\,\ulOmega\,\ulGamma)(t : \Tm\,\ulOmega\,(\ulA[\ulnu]))\\
@@
   (\llb\ult\rrb_{\bD}\,\gamma)\,(\Gamma\,\ulnu)\\
-  & \Tm_{\bI_{\ulOmega}}\,\Gamma\,A\,\ult &&:\equiv A\,\ulnu\,(\ult[\ulnu]) = \llb\ult\rrb_{\bS}\,(\Gamma\,\ulnu)
+  & \Tm_{\bI_{\ulOmega}}\,\Gamma\,A\,\ult &&:= A\,\ulnu\,(\ult[\ulnu]) = \llb\ult\rrb_{\bS}\,(\Gamma\,\ulnu)
 \end{alignat*}

### 2020-01-10 draft/paper.tex
= instead of := when just unfolding definitions
s
not necessarily third, fourth, hence use Ty, Tm

@@
 \begin{alignat*}{3}
-  & \Con_{\bA} && := \Set_{\max(i+1, j)} \\
-  & \Sub_{\bA}\,\Gamma\,\Delta && := \Gamma \ra \Delta\\
-  & \Ty_{\bA}\,\Gamma && := \Gamma \ra \Set_{\max(i+1, j)}\\
-  & \Tm_{\bA}\,\Gamma\,A && := (\gamma : \Gamma) \ra A\,\gamma
+  & \Con_{\bA} && = \Set_{\max(i+1, j)} \\
+  & \Sub_{\bA}\,\Gamma\,\Delta && = \Gamma \ra \Delta\\
+  & \Ty_{\bA}\,\Gamma && = \Gamma \ra \Set_{\max(i+1, j)}\\
+  & \Tm_{\bA}\,\Gamma\,A && = (\gamma : \Gamma) \ra A\,\gamma
 \end{alignat*}
@@
 $\bA$. This model, analogously to $\bA$, is given by restricting $\bM_{i,j}$ to
-the \emph{third} components everywhere, corresponding to types or actions on
+the $\Ty$ components everywhere, corresponding to types or actions on
 types. Hence, we have:
 \begin{alignat*}{3}
-  & \Con_{\bD}\,\ulGamma && := \ulGamma \ra \Set_{\max(i+1, j)} \\
-  & \Sub_{\bD}\,\Gamma\,\Delta\,\ulsigma && := (\ulgamma : \ulGamma) \ra \Gamma\,\gamma \ra \Delta(\llb\ulsigma\rrb_{\bA}\,\ulgamma)\\
-  & \Ty_{\bD}\,\Gamma\,\ulA && := (\ulgamma : \ulGamma) \ra \Gamma\,\ulgamma \ra \llb\ulA\rrb_{\bA}\,\ulgamma \ra \Set_{\max(i+1, j)}\\
-  & \Tm_{\bD}\,\Gamma\,A\,\ult && := (\ulgamma : \ulGamma)(\gamma : \Gamma\,\ulGamma)
-    \ra A\,\ulgamma\,\gamma\,(\llb\ult\rrb_{\bA}\,\ulgamma)
+  & \Con_{\bD}\,\ulGamma && = \ulGamma \ra \Set_{\max(i+1, j)} \\
+  & \Sub_{\bD}\,\Gamma\,\Delta\,\ulsigma && = (\ulgamma : \ulGamma) \ra \Gamma\,\gamma \ra \Delta\,(\ulsigma\,\ulgamma)\\
+  & \Ty_{\bD}\,\Gamma\,\ulA && = (\ulgamma : \ulGamma) \ra \Gamma\,\ulgamma \ra \ulA\,\ulgamma \ra \Set_{\max(i+1, j)}\\
+  & \Tm_{\bD}\,\Gamma\,A\,\ult && = (\ulgamma : \ulGamma)(\gamma : \Gamma\,\ulGamma)
+    \ra A\,\ulgamma\,\gamma\,(\ult\,\ulgamma)
 \end{alignat*}
@@
 $\ToS$ model over the total model of $\bD$, which is given by restricting
-$\bM_{i,j}$ to the fourth components, corresponding to interpretations of terms and
+$\bM_{i,j}$ to the $\Tm$ components, corresponding to interpretations of terms and
 actions on terms.

### 2020-01-10 draft/paper.tex
brackets for readability
spaces

@@
 \begin{alignat*}{3}
-& (\bId\,\bt\,\bu)\,\ulGamma &&:= \bt\,\ulGamma = \bu\,\ulGamma\\
-& (\bId\,\bt\,\bu)\,\ulA     && := \lambda\,e.\, \bt\,\ulA = \bu\,\ulA
+& (\bId\,\bt\,\bu)\,\ulGamma &&:= (\bt\,\ulGamma = \bu\,\ulGamma)\\
+& (\bId\,\bt\,\bu)\,\ulA     && := \lambda\,e.\, (\bt\,\ulA = \bu\,\ulA)
 \end{alignat*}
@@
 \begin{alignat*}{3}
-  (\bId\,\bt\,\bu)\,(\ulsigma : \Sub\,\ulGamma\,\ulDelta) : \bt\,\ulGamma = \bu\,\ulGamma \ra
-       \bt\,\ulDelta = \bu\,\ulDelta
+  (\bId\,\bt\,\bu)\,(\ulsigma : \Sub\,\ulGamma\,\ulDelta) : (\bt\,\ulGamma = \bu\,\ulGamma) \ra
+       (\bt\,\ulDelta = \bu\,\ulDelta)
 \end{alignat*}
@@
 \cite{kaposi2019constructing} in the construction of term algebras, but there is
-an important difference: in ibid.\, strict equalities are used instead of
+an important difference: in ibid.\ strict equalities are used instead of
 isomorphisms. In our case, isomorphisms are necessary once again because of
@@
 Here, there is no essential change compared to \cite{kaposi2019constructing},
-and we follow ibid. in the definition of $\bI_{\ulOmega}$. The reason is that although we
+and we follow ibid.\ in the definition of $\bI_{\ulOmega}$. The reason is that although we
 have weakened strict algebra equality to isomorphism, in the current

### 2020-01-10 draft/paper.tex
typo
paper

--- 2020-01-10 draft/paper.tex
@@
 algebras. For example, consider natural number algebras at level $i$, given as
-the $\Sigma$-type $NatAlg_i :=(Nat : \Set_i)\times Nat\times(Nat \ra
+the $\Sigma$-type $NatAlg_i :\equiv(Nat : \Set_i)\times Nat\times(Nat \ra
 Nat)$. It follows from the subtyping rules that $i \leq j$ implies $NatAlg_{i}
@@
       (u : \Tm\,\Gamma\,(\El\,a)) \ra \Tm\,\Gamma\,(B[\id, u])\\
-  & t\appitt u && := (\appi\,t)[\id, u]
+  & t\appitt u && :\equiv (\appi\,t)[\id, u]
 \end{alignat*}
@@
 %% \begin{alignat*}{3}
-%%   & A\arri B && := \Pi\,A\,(B[\p])\\
-%%   & A \arre\,B &&:= \Pie\,A\,(\lambda\,\_.\,B) \\
-%%   & A \arrinf\,B &&:= \Piinf\,A\,(\lambda\,\_.\,B)
-%% $A\arri B := \Pi\,A\,(B[\p])$, and $A \arre\,B := \Pie\,A\,(\lambda\,\_.\,B)$, and likewise
-%% $A \arrinf\,B := \Piinf\,A\,(\lambda\,\_.\,B)$.
+%%   & A\arri B && :\equiv \Pi\,A\,(B[\p])\\
+%%   & A \arre\,B &&:\equiv \Pie\,A\,(\lambda\,\_.\,B) \\
+%%   & A \arrinf\,B &&:\equiv \Piinf\,A\,(\lambda\,\_.\,B)
+%% $A\arri B :\equiv \Pi\,A\,(B[\p])$, and $A \arre\,B :\equiv \Pie\,A\,(\lambda\,\_.\,B)$, and likewise
+%% $A \arrinf\,B :\equiv \Piinf\,A\,(\lambda\,\_.\,B)$.
 
@@
 \[
-  \Sig_j := (i : \mathbb{N})\ra(M : \ToS_{i,j})\ra \Con_M
+  \Sig_j :\equiv (i : \mathbb{N})\ra(M : \ToS_{i,j})\ra \Con_M
 \]
@@
   & \NatSig : \Sig_0 \\
-  & \NatSig := \lambda (i :\mathbb{N})(M : \ToS_{i,0}).\\
+  & \NatSig :\equiv \lambda (i :\mathbb{N})(M : \ToS_{i,0}).\\
   & \hspace{1em}(\emptycon_M\,\ext_M\, (N : \U_M) \,\ext_M\,(zero : \El_M\,N)\\
@@
 \begin{alignat*}{3}
-& \mathsf{TreeSig} := \\
+& \mathsf{TreeSig} :\equiv \\
 & \hspace{1em}\emptycon \ext (Tree : \U)\\
@@
 %% \begin{alignat*}{3}
-%%   & cETTSig := \\
+%%   & cETTSig :\equiv \\
 %%   & ...\,\ext\,Tm{\Lift} : \Pii\,(\Gamma : Con)\,(\Pii\,(A : Ty{\appitt}\Gamma)((\IdU\,(Tm{\appitt}\Gamma{\appitt}A)\,(Tm{\appitt}\Gamma{\appitt}(\Lift_j{\appitt}A))))
@@
 \begin{alignat*}{3}
-  & \mathsf{RecEqSig} :=\\
+  & \mathsf{RecEqSig} :\equiv\\
   & \hspace{0.5em}\emptycon\,\ext (A : \U) \ext (a : \El\,A)\ext (f : \Pii\,(x : A)\,(\Id\,A\,x\,a\arri\El\,A))
@@
   & \Inductive : \Con_{\bGamma} \ra \Set_i \\
-  & \Inductive\,\Gamma := (A : \Ty_{\bGamma}\,\Gamma)\ra \Tm_{\bGamma}\,\Gamma\,A
+  & \Inductive\,\Gamma :\equiv (A : \Ty_{\bGamma}\,\Gamma)\ra \Tm_{\bGamma}\,\Gamma\,A
 \end{alignat*}
@@
 \begin{alignat*}{3}
-  & (\boldsymbol{\sigma \circ \delta})_{\emptycon}^{-1} :=
+  & (\boldsymbol{\sigma \circ \delta})_{\emptycon}^{-1} :\equiv
     \bsigma\,\bdelta_{\emptycon}^{-1} \circ \bdelta_{\emptycon}^{-1} \\
-  & (\boldsymbol{\sigma \circ \delta})_{\ext}^{-1} :=
+  & (\boldsymbol{\sigma \circ \delta})_{\ext}^{-1} :\equiv
     \bsigma\,\bdelta_{\ext}^{-1} \circ \bdelta_{\ext}^{-1}
@@
 \begin{alignat*}{3}
-  & \Con_{\boldsymbol{A[\sigma]}}\,\ulGamma && := \Con_{\bA}\,(\bsigma\,\ulGamma)\\
-  & \Sub_{\boldsymbol{A[\sigma]}}\,\Gamma\,\Delta\,\ulsigma && :=
+  & \Con_{\boldsymbol{A[\sigma]}}\,\ulGamma && :\equiv \Con_{\bA}\,(\bsigma\,\ulGamma)\\
+  & \Sub_{\boldsymbol{A[\sigma]}}\,\Gamma\,\Delta\,\ulsigma && :\equiv
     \Sub_{\bA}\,\Gamma\,\Delta\,(\bsigma\,\ulsigma)\\
-  & \Ty_{\boldsymbol{A[\sigma]}}\,\Gamma\,\ulA && :=
+  & \Ty_{\boldsymbol{A[\sigma]}}\,\Gamma\,\ulA && :\equiv
       \Ty_{\bA}\,\Gamma\,(\bsigma\,\ulA)\\
-  & \Tm_{\boldsymbol{A[\sigma]}}\,\Gamma\,A\,\ult && :=
+  & \Tm_{\boldsymbol{A[\sigma]}}\,\Gamma\,A\,\ult && :\equiv
       \Tm_{\bA}\,\Gamma\,A\,(\bsigma\,\ult)
@@
 
-Moreover, $\id_{\boldsymbol{A[\sigma]}} := \id_{\bA}$,
-$\sigma\circ_{\boldsymbol{A[\sigma]}}\delta := \sigma\circ_{\bA}\delta$,
+Moreover, $\id_{\boldsymbol{A[\sigma]}} :\equiv \id_{\bA}$,
+$\sigma\circ_{\boldsymbol{A[\sigma]}}\delta :\equiv \sigma\circ_{\bA}\delta$,
 and likewise substitution components are given by corresponding substitution
@@
 \begin{alignat*}{3}
-  &\emptycon_{\boldsymbol{A[\sigma]}} && :=
+  &\emptycon_{\boldsymbol{A[\sigma]}} && :\equiv
     \coe\,\bsigma_{\emptycon}^{-1}\,\emptycon_{\bA}\\
-  &\Gamma\ext_{\boldsymbol{A[\sigma]}}A && :=
+  &\Gamma\ext_{\boldsymbol{A[\sigma]}}A && :\equiv
     \coe\,\bsigma_{\ext}^{-1}\,(\Gamma\ext_{\bA} A)\\
-  &\Id_{\boldsymbol{A[\sigma]}}\,t\,u && :=
+  &\Id_{\boldsymbol{A[\sigma]}}\,t\,u && :\equiv
     \coe\,\bsigma_{\Id}^{-1}\,(\Id_{\bA}\,t\,u)
@@
 \begin{alignat*}{3}
-  & \epsilon_{\boldsymbol{A[\sigma]}} && :=
+  & \epsilon_{\boldsymbol{A[\sigma]}} && :\equiv
     \coh\,\bsigma_{\emptycon}^{-1}\,\emptycon_{\bA} \circ \epsilon_{\bA}\\
-  & \p_{\boldsymbol{A[\sigma]}} && :=
+  & \p_{\boldsymbol{A[\sigma]}} && :\equiv
     \p_{\bA} \circ (\coh\,\bsigma_{\ext}^{-1}\,(\Gamma\ext A))^{-1}\\
-  & \appK_{\boldsymbol{A[\sigma]}}\,t && :=
+  & \appK_{\boldsymbol{A[\sigma]}}\,t && :\equiv
     \appK_{\bA}\,((\coh\,\bsigma_{\K}\,(\K\,\Delta))^{-1}\circ t)
@@
 isomorphisms, e.g.\ we have $\coe_{\boldsymbol{A[\sigma]}}\,\ulsigma\,\Gamma
-:= \coe_{\bA}\,(\bsigma\,\ulsigma)\,\Gamma$.
+:\equiv \coe_{\bA}\,(\bsigma\,\ulsigma)\,\Gamma$.
 
@@
 \begin{alignat*}{3}
-  & (\boldsymbol{t[\sigma]})_{\emptycon}^{-1} :=
+  & (\boldsymbol{t[\sigma]})_{\emptycon}^{-1} :\equiv
     \bt\,\bsigma_{\emptycon}^{-1} \circ \bt_{\emptycon}^{-1} \\
-  & (\boldsymbol{t[\sigma]})_{\ext}^{-1} :=
+  & (\boldsymbol{t[\sigma]})_{\ext}^{-1} :\equiv
     \bt\,\bsigma_{\ext}^{-1} \circ \bt_{\ext}^{-1}
@@
   A)\,(A[\p])}$ is likewise a strict flCwF section given by second projections.
-Substitution extension $\boldsymbol{\sigma,\,t}$ is given by pointwise combining $\bsigma$ and $\bt$ with metatheoretic $\Sigma$ pairing, e.g.\ $\Con_{\boldsymbol{(\sigma,t)}}\,\ulGamma := (\bsigma\,\ulGamma,\,\bt\,\ulGamma)$.
+Substitution extension $\boldsymbol{\sigma,\,t}$ is given by pointwise combining $\bsigma$ and $\bt$ with metatheoretic $\Sigma$ pairing, e.g.\ $\Con_{\boldsymbol{(\sigma,t)}}\,\ulGamma :\equiv (\bsigma\,\ulGamma,\,\bt\,\ulGamma)$.
 
@@
 
-\begin{definition} For a level $i$, we write $\bSet_i$ for the flCwF of sets where $\Con_{\bSet_i} := \Set_i$ and $\Sub_{\bSet_i}\,\Gamma\,\Delta := \Gamma \ra \Delta$.
+\begin{definition} For a level $i$, we write $\bSet_i$ for the flCwF of sets where $\Con_{\bSet_i} :\equiv \Set_i$ and $\Sub_{\bSet_i}\,\Gamma\,\Delta :\equiv \Gamma \ra \Delta$.
 \end{definition}
@@
 base flCwF, and has trivial iso-cleaving where $\coe$-s are identity
-functions. Hence, we have $\Con_{\bU}\,\ulGamma := \Set_i$ and
-$\Sub_{\bU}\,\Gamma\,\Delta\,\ulsigma := \Gamma \ra \Delta$.
+functions. Hence, we have $\Con_{\bU}\,\ulGamma :\equiv \Set_i$ and
+$\Sub_{\bU}\,\Gamma\,\Delta\,\ulsigma :\equiv \Gamma \ra \Delta$.
 
@@
 \begin{alignat*}{3}
-  & \Con_{\bEl\,\ba}\,\ulGamma && := \ba\,\ulGamma\\
-  & \Sub_{\bEl\,\ba}\,\Gamma\,\Delta\,\ulsigma && := \ba\,\ulsigma\,\Gamma = \Delta\\
-  & \Ty_{\bEl\,\ba}\,\Gamma\,\ulA && := \ba\,\ulA\,\Gamma\\
-  & \Tm_{\bEl\,\ba}\,\Gamma\,A\,\ult && := \ba\,\ult\,\Gamma = A
+  & \Con_{\bEl\,\ba}\,\ulGamma && :\equiv \ba\,\ulGamma\\
+  & \Sub_{\bEl\,\ba}\,\Gamma\,\Delta\,\ulsigma && :\equiv \ba\,\ulsigma\,\Gamma = \Delta\\
+  & \Ty_{\bEl\,\ba}\,\Gamma\,\ulA && :\equiv \ba\,\ulA\,\Gamma\\
+  & \Tm_{\bEl\,\ba}\,\Gamma\,A\,\ult && :\equiv \ba\,\ult\,\Gamma = A
 \end{alignat*}
@@
 \begin{alignat*}{3}
-  & \emptycon_{\bEl\,\ba} && := \ba_{\emptycon}^{-1}\,\tt\\
-  & (\Gamma\ext_{\bEl\,\ba} A) && := \ba_{\ext}^{-1}\,(\Gamma,\,A)
+  & \emptycon_{\bEl\,\ba} && :\equiv \ba_{\emptycon}^{-1}\,\tt\\
+  & (\Gamma\ext_{\bEl\,\ba} A) && :\equiv \ba_{\ext}^{-1}\,(\Gamma,\,A)
 \end{alignat*}
 We likewise use preservation isomorphisms to define $\K$, $\Id$ and $\Sigma$.
-Context coercion is $\coe\,\ulsigma\,\Gamma := \ba\,\ulsigma\,\Gamma$. Type coercion, for $A : \ba\,\ulA\,\Gamma$ is given as $\coe\,\ult\,A := \ba\,\ult\,(\ba_{\ext}^{-1}\,(\Gamma,\,A))$.
+Context coercion is $\coe\,\ulsigma\,\Gamma :\equiv \ba\,\ulsigma\,\Gamma$. Type coercion, for $A : \ba\,\ulA\,\Gamma$ is given as $\coe\,\ult\,A :\equiv \ba\,\ult\,(\ba_{\ext}^{-1}\,(\Gamma,\,A))$.
 
@@
 \begin{alignat*}{3}
-  & \Con_{(\bPii\,\ba\,\bB)}\,\ulGamma &&:= (\gamma : \ba\,\ulGamma)\ra \Con_{\bB}\,(\ulGamma, \gamma)\\
-  & \Ty_{(\bPii\,\ba\,\bB)}\,\Gamma\,\ulA &&:= (\gamma : \ba\,\ulGamma)(a : \ba\,\ulA\,\gamma)\ra \Ty_{\bB}\,(\Gamma\,\gamma)\,(\ulA, a)
+  & \Con_{(\bPii\,\ba\,\bB)}\,\ulGamma &&:\equiv (\gamma : \ba\,\ulGamma)\ra \Con_{\bB}\,(\ulGamma, \gamma)\\
+  & \Ty_{(\bPii\,\ba\,\bB)}\,\Gamma\,\ulA &&:\equiv (\gamma : \ba\,\ulGamma)(a : \ba\,\ulA\,\gamma)\ra \Ty_{\bB}\,(\Gamma\,\gamma)\,(\ulA, a)
 \end{alignat*}
@@
 \begin{alignat*}{3}
-  & \Sub_{(\bPii\,\ba\,\bB)}\,\Gamma\,\Delta\,\ulsigma :=
+  & \Sub_{(\bPii\,\ba\,\bB)}\,\Gamma\,\Delta\,\ulsigma :\equiv
   (\gamma : \ba\,\ulGamma)(\delta : \ba\,\ulDelta)
@@
 \begin{alignat*}{3}
-  & \Sub_{(\bPii\,\ba\,\bB)}\,\Gamma\,\Delta\,\ulsigma :=
+  & \Sub_{(\bPii\,\ba\,\bB)}\,\Gamma\,\Delta\,\ulsigma :\equiv
   (\gamma : \ba\,\ulGamma)\ra\Sub_{\bB}\,(\Gamma\,\gamma)\,(\Delta\,(\ba\,\ulsigma\,\gamma)\,(\ulsigma, \refl)
@@
 \begin{alignat*}{3}
-  & \coe\,\ulsigma\,\Gamma &&:=
+  & \coe\,\ulsigma\,\Gamma &&:\equiv
     \lambda\,\gamma.\,\coe_{\bB}\,(\ulsigma,\refl)\,(\Gamma\,(\ba\,(\ulsigma^{-1})\,\gamma))\\
-  & \coe\,\ult\,A &&:=
+  & \coe\,\ult\,A &&:\equiv
     \lambda\,\gamma\,a.\,\coe_{\bB}\,(\ult,\refl)\,(A\,(\ba\,(\ult^{-1})\,(\ba_{\ext}^{-1}(\gamma,a))))
@@
 \begin{alignat*}{3}
-  & (\bPiinf\,A\,\bb)\,\ulGamma    &&:= (a : A)\ra \bb\,a\,\ulGamma\\
-  & (\bPiinf\,A\,\bb)\,\ulsigma    &&:= \lambda\,a.\, \bb\,a\,\ulsigma\\
-  & (\bPiinf\,A\,\bb)\,\ulA\       &&:= \lambda\,\Gamma.\,(a : A)\ra \bb\,a\,\ulA\,(\Gamma\, a)\\
-  & (\bPiinf\,A\,\bb)\,\ult        &&:= \lambda\,a.\, \bb\,a\,\ult
+  & (\bPiinf\,A\,\bb)\,\ulGamma    &&:\equiv (a : A)\ra \bb\,a\,\ulGamma\\
+  & (\bPiinf\,A\,\bb)\,\ulsigma    &&:\equiv \lambda\,a.\, \bb\,a\,\ulsigma\\
+  & (\bPiinf\,A\,\bb)\,\ulA\       &&:\equiv \lambda\,\Gamma.\,(a : A)\ra \bb\,a\,\ulA\,(\Gamma\, a)\\
+  & (\bPiinf\,A\,\bb)\,\ult        &&:\equiv \lambda\,a.\, \bb\,a\,\ult
 \end{alignat*}
@@
   %% &                               && : \top\ra (a : A) \ra \bb\,a\,\emptycon\\
-  &(\bPiinf\,A\,\bb)_{\emptycon}^{-1} && := \lambda\,\_\,a.\,(\bb\,a)_{\emptycon}^{-1}\,\tt\\
+  &(\bPiinf\,A\,\bb)_{\emptycon}^{-1} && :\equiv \lambda\,\_\,a.\,(\bb\,a)_{\emptycon}^{-1}\,\tt\\
   &(\bPiinf\,A\,\bb)_{\ext}^{-1} && : (\Gamma : (\bPiinf\,A\,\bb)\,\ulGamma)\times((\bPiinf\,A\,\bb)\,\ulA\,\Gamma)\\
   & && \hspace{0.5em}\ra (\bPiinf\,A\,\bb)\,(\ulGamma \ext \ulA)\\
-  & (\bPiinf\,A\,\bb)_{\ext}^{-1} && := \lambda\,(\Gamma,A)\,a.\,(\bb\,a)_{\ext}^{-1}(\Gamma\,a,\,A\,a)
+  & (\bPiinf\,A\,\bb)_{\ext}^{-1} && :\equiv \lambda\,(\Gamma,A)\,a.\,(\bb\,a)_{\ext}^{-1}(\Gamma\,a,\,A\,a)
 \end{alignat*}
@@
 \begin{alignat*}{3}
-& (\bId\,\bt\,\bu)\,\ulGamma &&:= \bt\,\ulGamma = \bu\,\ulGamma\\
-& (\bId\,\bt\,\bu)\,\ulA     && := \lambda\,e.\, \bt\,\ulA = \bu\,\ulA
+& (\bId\,\bt\,\bu)\,\ulGamma &&:\equiv \bt\,\ulGamma = \bu\,\ulGamma\\
+& (\bId\,\bt\,\bu)\,\ulA     && :\equiv \lambda\,e.\, \bt\,\ulA = \bu\,\ulA
 \end{alignat*}
@@
 of Section \ref{sec:semantics} so that we only have the first $\Con$ components
-in the interpretations for contexts, substitution, types, term, and we only have
-actions on contexts in the interpretations of term and substitution
+in the interpretations for contexts, substitutions, types, terms, and we only
+have actions on contexts in the interpretations of term and substitution
 formers. Hence, we have:
 \begin{alignat*}{3}
-  & \Con_{\bA} && := \Set_{\max(i+1, j)} \\
-  & \Sub_{\bA}\,\Gamma\,\Delta && := \Gamma \ra \Delta\\
-  & \Ty_{\bA}\,\Gamma && := \Gamma \ra \Set_{\max(i+1, j)}\\
-  & \Tm_{\bA}\,\Gamma\,A && := (\gamma : \Gamma) \ra A\,\gamma
+  & \Con_{\bA} && = \Set_{\max(i+1, j)} \\
+  & \Sub_{\bA}\,\Gamma\,\Delta && = \Gamma \ra \Delta\\
+  & \Ty_{\bA}\,\Gamma && = \Gamma \ra \Set_{\max(i+1, j)}\\
+  & \Tm_{\bA}\,\Gamma\,A && = (\gamma : \Gamma) \ra A\,\gamma
 \end{alignat*}
@@
 \begin{alignat*}{3}
-  & \Con_{\bD}\,\ulGamma && := \ulGamma \ra \Set_{\max(i+1, j)} \\
-  & \Sub_{\bD}\,\Gamma\,\Delta\,\ulsigma && := (\ulgamma : \ulGamma) \ra \Gamma\,\gamma \ra \Delta(\llb\ulsigma\rrb_{\bA}\,\ulgamma)\\
-  & \Ty_{\bD}\,\Gamma\,\ulA && := (\ulgamma : \ulGamma) \ra \Gamma\,\ulgamma \ra \llb\ulA\rrb_{\bA}\,\ulgamma \ra \Set_{\max(i+1, j)}\\
-  & \Tm_{\bD}\,\Gamma\,A\,\ult && := (\ulgamma : \ulGamma)(\gamma : \Gamma\,\ulGamma)
+  & \Con_{\bD}\,\ulGamma && = \ulGamma \ra \Set_{\max(i+1, j)} \\
+  & \Sub_{\bD}\,\Gamma\,\Delta\,\ulsigma && = (\ulgamma : \ulGamma) \ra \Gamma\,\gamma \ra \Delta(\llb\ulsigma\rrb_{\bA}\,\ulgamma)\\
+  & \Ty_{\bD}\,\Gamma\,\ulA && = (\ulgamma : \ulGamma) \ra \Gamma\,\ulgamma \ra \llb\ulA\rrb_{\bA}\,\ulgamma \ra \Set_{\max(i+1, j)}\\
+  & \Tm_{\bD}\,\Gamma\,A\,\ult && = (\ulgamma : \ulGamma)(\gamma : \Gamma\,\ulGamma)
     \ra A\,\ulgamma\,\gamma\,(\llb\ult\rrb_{\bA}\,\ulgamma)
@@
 \[
-\NatSig := \emptycon\ext(N : \U)\ext(zero : \El\,N)\ext (suc : N
+\NatSig :\equiv \emptycon\ext(N : \U)\ext(zero : \El\,N)\ext (suc : N
 \arri \El\,N)
@@
 \begin{alignat*}{3}
-  & \Con_{\bT_{\ulOmega}}\,\ulGamma &&:= \Sub\,\ulOmega\,\ulGamma \ra \llb\ulGamma\rrb_{\bA}\\
-  & \Sub_{\bT_{\ulOmega}}\,\Gamma\,\Delta\,\ulsigma &&:= (\ulnu : \Sub\,\ulOmega\,\ulGamma)
+  & \Con_{\bT_{\ulOmega}}\,\ulGamma &&:\equiv \Sub\,\ulOmega\,\ulGamma \ra \llb\ulGamma\rrb_{\bA}\\
+  & \Sub_{\bT_{\ulOmega}}\,\Gamma\,\Delta\,\ulsigma &&:\equiv (\ulnu : \Sub\,\ulOmega\,\ulGamma)
   \ra \Delta\,(\ulsigma\circ\nu) \simeq \llb\ulsigma\rrb_{\bA}\,(\Gamma\,\ulnu)\\
-  & \Ty_{\bT_{\ulOmega}}\,\Gamma\,\ulA &&:= (\ulnu : \Sub\,\ulOmega\,\ulGamma) \ra
+  & \Ty_{\bT_{\ulOmega}}\,\Gamma\,\ulA &&:\equiv (\ulnu : \Sub\,\ulOmega\,\ulGamma) \ra
     \Tm\,\ulOmega\,(\ulA[\ulnu]) \ra \llb\ulA\rrb_{\bA}\,(\Gamma\,\ulnu)\\
-  & \Tm_{\bT_{\ulOmega}}\,\Gamma\,A\,\ult &&:= (\ulnu : \Sub\,\ulOmega\,\ulGamma)\ra
+  & \Tm_{\bT_{\ulOmega}}\,\Gamma\,A\,\ult &&:\equiv (\ulnu : \Sub\,\ulOmega\,\ulGamma)\ra
     \llb\ulA\rrb_{\bA}\,\ulnu\,(\ult[\ulnu]) \simeq_{\id} \llb\ult\rrb_{\bA} (\Gamma\,\ulnu)
@@
               \Set_{j+1}\\
-  & \U_{\bT_{\ulOmega}}\,\ulnu\,\ult := \Tm\,\ulOmega\,(\El\,\ult)\\
+  & \U_{\bT_{\ulOmega}}\,\ulnu\,\ult :\equiv \Tm\,\ulOmega\,(\El\,\ult)\\
   & \El_{\bT_{\ulOmega}}\,a : (\ulnu : \Sub\,\ulOmega\,\ulGamma)
@@
               \llb\ula\rrb_{\bA}\,(\Gamma\,\ulnu)\\
-  & \El_{\bT_{\ulOmega}}\,a\,\ulnu\,\ult := (a\,\ulnu)\,\ult
+  & \El_{\bT_{\ulOmega}}\,a\,\ulnu\,\ult :\equiv (a\,\ulnu)\,\ult
 \end{alignat*}
@@
 \[
-  \Pii_{\bT_{\ulOmega}}\,a\,B\,\ulnu\,\ult := \lambda\,\alpha.\,
+  \Pii_{\bT_{\ulOmega}}\,a\,B\,\ulnu\,\ult :\equiv \lambda\,\alpha.\,
          B\,(\ulnu,\,(a\,\ulnu)^{-1}\,\alpha)\,(\ult\,\appitt\,((a\,\ulnu)^{-1}\,\alpha))
@@
 \begin{alignat*}{3}
-  & \Con\,\ulGamma && := \Subtype\,\llb\ulGamma\rrb_{\bA_k}\,\llb\ulGamma\rrb_{\bA_l}\\
-  & \Sub\,\Gamma\,\Delta\,\ulsigma && := (\gamma : \llb\ulGamma\rrb_{\bA_k})\ra
+  & \Con\,\ulGamma && :\equiv \Subtype\,\llb\ulGamma\rrb_{\bA_k}\,\llb\ulGamma\rrb_{\bA_l}\\
+  & \Sub\,\Gamma\,\Delta\,\ulsigma && :\equiv (\gamma : \llb\ulGamma\rrb_{\bA_k})\ra
   \llb\ulsigma\rrb_{\bA_k}\,\gamma = \llb\ulsigma\rrb_{\bA_l}\,\gamma\\
-  & \Ty\,\Gamma\,\ulA &&:= (\gamma : \llb\ulGamma\rrb_{\bA_k})\ra
+  & \Ty\,\Gamma\,\ulA &&:\equiv (\gamma : \llb\ulGamma\rrb_{\bA_k})\ra
       \Subtype\,(\llb\ulA\rrb_{\bA_k}\,\gamma)\,(\llb\ulA\rrb_{\bA_l}\,\gamma)\\
-  & \Tm\,\Gamma\,A\,\ult &&:= (\gamma : \llb\ulGamma\rrb_{\bA_k})\ra
+  & \Tm\,\Gamma\,A\,\ult &&:\equiv (\gamma : \llb\ulGamma\rrb_{\bA_k})\ra
      \llb\ult\rrb_{\bA_k}\,\gamma = \llb\ult\rrb_{\bA_l}\,\gamma
@@
 \begin{alignat*}{3}
-  & \Con_{\bI_{\ulOmega}}\,\ulGamma &&:= (\ulnu : \Sub\,\ulOmega\,\ulGamma)\ra
+  & \Con_{\bI_{\ulOmega}}\,\ulGamma &&:\equiv (\ulnu : \Sub\,\ulOmega\,\ulGamma)\ra
   \llb\ulGamma\rrb_{\bS}\,(\llb\ulnu\rrb_{\bA}\,(\llb\ulOmega\rrb_{\bT_{\ulOmega}}\,\id))\,\gamma\\
-  & \Sub_{\bI_{\ulOmega}}\,\Gamma\,\Delta\,\ulsigma && := (\ulnu : \Sub\,\ulOmega\,\ulGamma)\ra
+  & \Sub_{\bI_{\ulOmega}}\,\Gamma\,\Delta\,\ulsigma && :\equiv (\ulnu : \Sub\,\ulOmega\,\ulGamma)\ra
     \Delta\,(\ulsigma\circ\ulnu) = \llb\ulsigma\rrb_{\bS}\,(\Gamma\,\ulnu)\\
-  & \Ty_{\bI_{\ulOmega}}\,\Gamma\,\ulA &&:=
+  & \Ty_{\bI_{\ulOmega}}\,\Gamma\,\ulA &&:\equiv
   (\ulnu : \Sub\,\ulOmega\,\ulGamma)(t : \Tm\,\ulOmega\,(\ulA[\ulnu]))\\
@@
   (\llb\ult\rrb_{\bD}\,\gamma)\,(\Gamma\,\ulnu)\\
-  & \Tm_{\bI_{\ulOmega}}\,\Gamma\,A\,\ult &&:= A\,\ulnu\,(\ult[\ulnu]) = \llb\ult\rrb_{\bS}\,(\Gamma\,\ulnu)
+  & \Tm_{\bI_{\ulOmega}}\,\Gamma\,A\,\ult &&:\equiv A\,\ulnu\,(\ult[\ulnu]) = \llb\ult\rrb_{\bS}\,(\Gamma\,\ulnu)
 \end{alignat*}
@@
 equalities. In contrast to our algebraic definition, Cartmell's signatures are
-given by presyntax, named variables and typing relations, there is no
-explicit model theory provided for signatures, and no explicit term model
-construction. Cartmell focuses instead on showing that contextual categories
-serve as classifying categories for GATs.
+given by presyntax, named variables and typing relations, there is no explicit
+model theory provided for signatures, and no explicit term model construction is
+given. Cartmell focuses instead on showing that contextual categories serve as
+classifying categories for GATs.
 
@@
 
-Streicher's initiality proof \cite{streicher2012semantics} is a direct
-construction of a concrete QIIT: that of the syntax of a type theory. He uses
-preterms, typing relations and quotients to construct the initial model of type
-theory in a set-theoretic metalanguage. A recent formalization of the same
-result for a larger type theory was done by Brunerie et al.\ \cite{brunerie} in
-Agda with UIP, function extensionality and quotients. Lafont et
-al.\ \cite{induction-is-enough} construct all finitary inductive-inductive types
-in Agda using only UIP and function extensionality. Examples of
-inductive-inductive types are constructed without using UIP by Hugunin
-\cite{jasper}.
+%% Streicher's initiality proof \cite{streicher2012semantics} is a direct
+%% construction of a concrete QIIT: that of the syntax of a type theory. He uses
+%% preterms, typing relations and quotients to construct the initial model of type
+%% theory in a set-theoretic metalanguage. A recent formalization of the same
+%% result for a larger type theory was done by Brunerie et al.\ \cite{brunerie} in
+%% Agda with UIP, function extensionality and quotients. Lafont et
+%% al.\ \cite{induction-is-enough} construct all finitary inductive-inductive types
+%% in Agda using only UIP and function extensionality. Examples of
+%% inductive-inductive types are constructed without using UIP by Hugunin
+%% \cite{jasper}.
 
@@
 
-Existence of the theory of signatures:
-
-\cite[Section 9]{lumsdaineShulman} shows that infinitary QIITs are not
-constructible in type theory without choice, so the existence of the
-theory of signatures is a non-conservative extension of extensional
-type theory. The setoid model supports an example QIT which branches
-over an arbitrary setoid, and this suggests that it might support the
-theory of signatures as well.
+An important motivation of the current work was to use QIITs as a framework for
+algebraic theories, with the metatheory of type theories in mind as a key
+application. We would prefer QIITs to
+\begin{itemize}
+\item Be formally precise.
+\item Not gloss over issues of size.
+\item Be rich enough to cover most type theories in the wild, including the theory
+      of QIIT signatures.
+\item Be direct enough, so that signatures for type theories can
+      be written out without excessive encoding overhead.
+\item Be suitable for practical implementation in proof assistants.
+\item Be reducible to a minimal set of basic type formers.
+\end{itemize}
+
+With the current work, we have improved the state of QIITs with respect to
+the above criteria. However, a number of open research problems remain.
+
+With regards to the expressiveness of QIIT signatures, we do not yet
+support sort equations, i.e. equations of elements of $\Tm\,\Gamma\,U$ in
+signatures. Sort equations are included in Cartmell's generalized algebraic
+theories \cite{gat}, and they appear to be highly useful for giving an algebraic
+representation for Russell-style universes and cumulative universes
+\cite{sterling2019algebraic}. We leave this for future work, but we note that
+the current isofibration-based semantics does not work in the presence of strict
+sort equations, since they are not invariant under isomorphism; instead, sort
+equations are compatible with the stricter semantics of
+\cite{kaposi2019constructing}.
+
+While we have made an effort to shape the syntax and semantics of QIITs to
+be amenable to implementation in proof assistants, much needs to be done before
+we can have a practical implementation. For one, we would need to consider QIITs
+in a type theory where transports along equality proofs compute, and would need
+to work out computing transports for QIITs. Cubical Agda has recently made
+strides in implementing HITs \cite{vezzosicubical}, but as of now it does not
+support computing transports on indexed inductive types.
+
+With regards to the reduction of QIITs to simple type formers, the
+reduction of infinitary QIITs appears to be more challenging than the finitary
+case. \cite[Section 9]{lumsdaineShulman} shows that infinitary QIITs are not
+constructible in type theory without choice, so such QIITs are a
+non-conservative extension of extensional type theory. In the finitary case, a
+generalization of the approach in \cite{induction-is-enough} seems promising ;
+this amounts to a Streicher-style initial algebra construction
+\cite{streicher2012semantics} for the theory of finitary QIIT signatures. In
+particular, Brunerie et al. \cite{brunerie} has formalized in Agda this
+construction for a comparable type theory, using UIP, function extensionality,
+propositional extensionality and simple quotient types.
+
+Another line of possible future work would be to explore a more general
+functorial style of semantics for QIITs. So far, we considered set-based
+1-categorical semantics, which is what we need when we want to reason
+inductively about syntaxes of type theories. However, it would be fruitful to
+consider algebras in structured categories other than the category of sets.
 

### 2020-01-10 draft/paper.tex

--- 2020-01-10 draft/paper.tex
@@
 of Section \ref{sec:semantics} so that we only have the first $\Con$ components
-in the interpretations for contexts, substitution, types, term, and we only have
-actions on contexts in the interpretations of term and substitution
+in the interpretations for contexts, substitutions, types, terms, and we only
+have actions on contexts in the interpretations of term and substitution
 formers. Hence, we have:
@@
 We would like to show that term algebras are initial, but we want to do this
-\emph{on all universe levels}, i.e.\ that terms algebras are initial when lifted
+\emph{on all universe levels}, i.e.\ that term algebras are initial when lifted
 to any higher level. This requires showing that QII algebras are

### 2020-01-10 draft/paper.tex
paper
\U instead of U
missing word

@@
 substitutions are appropriate structure-preserving functors. Then, each
-signature can applied to this model, yielding an interpretation of the signature
+signature can be applied to this model, yielding an interpretation of the signature
 as a structured category of algebras.
@@
 
-\section{Conclusions and further work}
+\section{Conclusions and Further Work}
 \label{sec:conclusion}
@@
 With regards to the expressiveness of QIIT signatures, we do not yet
-support sort equations, i.e. equations of elements of $\Tm\,\Gamma\,U$ in
+support sort equations, i.e. equations of elements of $\Tm\,\Gamma\,\U$ in
 signatures. Sort equations are included in Cartmell's generalized algebraic

### 2020-01-10 draft/paper.tex
small fixes

@@
   \item A CwF with underlying sets all in $\Set_i$. We reuse the component
-    notations from Definition \ref{def:tos}.
-  \item $\Sigma$-type $\Sg : (A : \Ty\,\Gamma)\ra \Ty\,(\Gamma \ext A)
+    notations $\Con$, $\Sub$, $\Ty$, etc.\ from Definition \ref{def:tos}.
+  \item $\Sigma$-types $\Sg : (A : \Ty\,\Gamma)\ra \Ty\,(\Gamma \ext A)
     \ra \Ty\,\Gamma$, with term formers $\proj1$, $\proj2$ and $\blank,\blank$.
@@
 whose objects are types $A : \Ty\,\Gamma$, and morphisms from $A$ to $B$ are
-terms $t : \Tm\,(\Gamma\,\ext\,A)\,B[\p]$. Identity morphisms are given by $\q :
-\Tm\,(\Gamma\,\ext\,A)\,A[\p]$, and composition $t \circ u$ by $t[\p, u]$. The
+terms $t : \Tm\,(\Gamma\,\ext\,A)\,(B[\p])$. Identity morphisms are given by $\q :
+\Tm\,(\Gamma\,\ext\,A)\,(A[\p])$, and composition $t \circ u$ by $t[\p, u]$. The
 assignment of type categories to contexts extends to a split indexed category. For
@@
 $\Ty\,\Gamma$, which sends $A$ to $A[\sigma]$ and $t :
-\Tm\,(\Gamma\,\ext\,A)\,B[\p]$ to $t[\sigma\circ \p, \q]$.
+\Tm\,(\Gamma\,\ext\,A)\,(B[\p])$ to $t[\sigma\circ \p, \q]$.
 \end{definition}

### 2020-01-10 draft/paper.tex
small fixes

@@
 This isomorphism is the identity on $\bsigma\,\Gamma$, hence we can extract the
-desired $\bsigma_{\Sigma} : \bsigma\,\Sigma\,A\,B) \simeq
+desired $\bsigma_{\Sigma} : \bsigma\,(\Sigma\,A\,B) \simeq
 \Sigma\,(\bsigma\,A)\,((\bsigma\,B)[\bsigma_{\ext}^{-1}])$ from it.
@@
 \end{alignat*}
-This yields a type isomorphism $\bsigma\,(\K\,\Delta)) \simeq
+This yields a type isomorphism $\bsigma\,(\K\,\Delta) \simeq
 \K\,(\bsigma\,\Delta)$ in the empty context, and we use the functorial action of

### 2020-01-10 draft/paper.tex
paper

--- 2020-01-10 draft/paper.tex
@@
 is illustrated by the following excerpt of the signature for CwFs
-extending the above signature for categories.
+extending the previous signature for categories.
 \begin{alignat*}{3}
@@
 not easily readable. We use instead a nameful notation for binders in context
-extension, $\Pii$ and $\lami$, e.g.\ as $(\emptycon\,\ext (a : U) \ext (t :
+extension, $\Pii$ and $\lami$, e.g.\ as $(\emptycon\,\ext (a : \U) \ext (t :
 \El\,a))$. We also define a type-theoretic flavor of $\appi$ for convenience:
@@
 signatures, but we only need contexts encoded in this way, and not other ToS
-components. In functional programming, this is sometimes called ``finally
-tagless'' \cite{carette2007finally}, and it is used for defining and interpreting
-embedded languages.
+components. In functional programming, this representation is sometimes called
+``finally tagless'' \cite{carette2007finally}, and it is used for defining and
+interpreting embedded languages.
 \end{definition}
@@
 substitutions are appropriate structure-preserving functors. Then, each
-signature can be applied to this model, yielding an interpretation of the signature
-as a structured category of algebras.
+signature can be applied to this model, yielding an interpretation of the
+signature as a structured category of algebras.
 
@@
 structured CwFs instead of lex categories. The reason here is that CwFs allow us
-to compute induction principles in strictly the same way as one would write them
-down in type theory, since we have $\Ty$ and $\Tm$ for a primitive notion of
-dependent objects and morphisms. In contrast, dependent objects in lex
-categories is a derived notion, and the induction principles we get are only up
-to isomorphism. This issue is perhaps not relevant from a purely categorical
+to compute induction principles in strictly the same way as one would write in
+type theory, since we have $\Ty$ and $\Tm$ for a primitive notion of dependent
+objects and morphisms. In contrast, dependent objects in lex categories is a
+derived notion, and the induction principles we get are only up to
+isomorphism. This issue is perhaps not relevant from a purely categorical
 perspective, but we are concerned with eventually implementing QIITs in proof
@@
 In contrast to \cite{kaposi2019constructing}, it is not sufficient to model
-types with displayed flCwFs. In ibid.\ substitutions are modeled as strict
+types as displayed flCwFs. In ibid.\ substitutions are modeled as strict
 morphisms, which we cannot do because our infinitary function types and identity
@@
 Moreover, $\id_{\boldsymbol{A[\sigma]}} := \id_{\bA}$,
-$\sigma\circ_{\boldsymbol{A[\sigma]}}\delta := \sigma\circ_{\bA}\delta$,
-and likewise substitution components are given by corresponding substitution
-components in $\bA$.
+$\sigma\circ_{\boldsymbol{A[\sigma]}}\delta := \sigma\circ_{\bA}\delta$, and
+likewise components for substitution are given by corresponding components in
+$\bA$.
 
@@
 
-We will need in the following three model fragments of $\bM$, which can be used
+In the following, We will need three model fragments of $\bM$, which can be used
 to compute notions of algebras, displayed algebras and sections respectively for
@@
 of natural numbers, since, intuitively, such terms can only be built from
-iterated usage of $zero$ and $suc$. We will build a term algebra for each
-signature in this manner. First, we describe the interpretation of $\syn$ in the
-$\Set$ model.
+iterated usage of $zero$ and $suc$. We build a term algebra for each
+signature in this manner.
 
@@
 \end{alignat*}
-
 Hence, a syntactic $\ult : \Tm\,\ulOmega\,\U$ is interpreted as a set of terms

### 2020-01-10 draft/paper.tex

@@
 where contexts are categories supporting the above requirements and
-substitutions are appropriate structure-preserving functors. Then, each
-signature can be applied to this model, yielding an interpretation of the
-signature as a structured category of algebras.
+substitutions are appropriate structure-preserving functors. Then, each signature can be applied to this model, yielding an interpretation of the signature
+as a structured category of algebras.
 

### 2020-01-10 draft/paper.tex
paper
missing word

--- 2020-01-10 draft/paper.tex
@@
 
-Iso-cleaving is given iso-cleaving in $\bA$ and the action of $\bsigma$ on
+Iso-cleaving is given by iso-cleaving in $\bA$ and the action of $\bsigma$ on
 isomorphisms, e.g.\ we have $\coe_{\boldsymbol{A[\sigma]}}\,\ulsigma\,\Gamma

### 2020-01-19 types2020abstract/abstract.tex
types20 abstract

\documentclass{easychair}

\usepackage{doc}

\newcommand{\easychair}{\textsf{easychair}}
\newcommand{\miktex}{MiK{\TeX}}
\newcommand{\texniccenter}{{\TeX}nicCenter}
\newcommand{\makefile}{\texttt{Makefile}}
\newcommand{\latexeditor}{LEd}

\newcommand{\ToS}{\mathsf{ToS}}
\newcommand{\U}{\mathsf{U}}

\title{Generalizations of Quotient Inductive-Inductive Types %
  \thanks{Both authors author were supported by the European Union, co-financed
    by the European Social Fund (EFOP-3.6.3-VEKOP-16-2017-00002). The second
    author was supported by by the National Research, Development and Innovation
    Fund of Hungary, financed under the Thematic Excellence Programme funding
    scheme, Project no.\ ED\_18-1-2019-0030 (Application-specific highly
    reliable IT solutions). }}

\author{
Andr\'as Kov\'acs
\and
Ambrus Kaposi
}

% Institutes for affiliations are also joined by \and,
\institute{
  E\"otv\"os Lor\'and University,
  Budapest, Hungary \\
  \email{kovacsandras|akaposi@inf.elte.hu}
}

\authorrunning{Kov\'acs, Kaposi}
\titlerunning{Generalizations of Quotient Inductive-Inductive Types}
\pagenumbering{gobble}
\begin{document}

\maketitle

Quotient inductive-inductive types (QIITs) are the most general class of
inductive types studied thus far in a set-truncated setting, i.e.\ in the
presence of uniqueness of identity proofs (UIP). In the current work, we develop
QIITs further, focusing on applications in practical metatheory of type
theories. We extend previous work on QIITs \cite{kaposi2019constructing} with
the following:
\begin{enumerate}
  \item
  \textbf{Large constructors, large elimination} and algebras at different
  universe levels. This fills in an important formal gap; large models are
  routinely used in the metatheory of type theories, but they have not been
  presented explicitly in previous QIIT literature.
  \item
  \textbf{Infinitary constructors}.
  This covers real, surreal \cite{HoTTbook} and ordinal numbers. Additionally,
  the theory which describes QII signatures is itself a large and infinitary
  QIIT, which allows the theory of signatures to describe its own signature
  (modulo universe levels), and provide its own model theory. This was not
  possible previously in \cite{kaposi2019constructing}, where only finitary
  QIITs were described.
  \item
  \textbf{Recursive equations}, i.e. equations appearing as assumptions of
  constructors. These have occurred previously in syntaxes of cubical type
  theories, as boundary conditions \cite{cohen2016cubical,
    angiuli2016computational, angiuli2018cartesian}.
  \item
  \textbf{Sort equations}. Sort equations were included in Cartmell's
  generalized algebraic theories (GATs) \cite{gat}, which overlap significantly
  with finitary QIITs. Sort equations appear to be useful for algebraic
  presentations of Russell-style and cumulative universes
  \cite{sterling2019algebraic}.
\end{enumerate}

\textbf{Self-describing signatures}. In the current work, we
would also like to streamline and make more rigorous the specification of
signatures. Previous descriptions of GATs \cite{gat, sterling2019algebraic} used
raw syntax with well-formedness relations to describe signatures, which is
rather unwieldy to formally handle. Also, the precursor of the current work
\cite{kaposi2019constructing} used an ad-hoc QIIT to describe signatures, which
did not have a model theory worked out, and its existence was simply assumed.

In contrast, equipped with large elimination and self-description, we are able
to specify signatures and develop a model theory for signatures, without ever
using raw syntax or assuming the existence of a particular QIIT. We do the
following in order.
\begin{enumerate}
  \item
  We specify a notion of model for the theory of signatures (ToS); this is a
  category with family (CwF) extended with several type formers, allowing to
  represent a signature as a typing context, with types specifying various
  constructors.
  \item We say that a signature is a context in an \emph{arbitrary} model, i.e.
  a function with type $(M : \mathsf{ToS})\to\mathsf{Con}_M$. This can be
  viewed as a fragment of a Church-encoding; here we do not care about
  encoding the whole syntax of ToS, nor the initiality of the syntax, we only
  need a representation of signatures and the ability to interpret a signature in
  a $\mathsf{ToS}$-model.
  \item We give a semantics for signatures, as a particular $M : \mathsf{ToS}$
    model which interprets each CwF context $\Gamma$ as a structured category of
    $\Gamma$-algebras.
  \item We give a (large, infinitary) signature for $\mathsf{ToS}$ itself, such
    that interpreting the signature in the semantic model yields a structured
    category of $\mathsf{ToS}$-algebras.
\end{enumerate}
In the above construction, everything is appropriately indexed with universe
levels (we omit the details), and there is a ``bump'' of levels at every
instance of self-description.

\textbf{Extending semantics to infinitary constructors and sort equations.}
Previously in \cite{kaposi2019constructing}, contexts in $\ToS$ were interpreted
as CwFs of algebras with extra structure, substitutions as strict morphisms of
such CwFs, and types as displayed CwFs. Infinitary constructors force a major
change: substitutions are interpreted as weak CwF morphisms, and types as CwF
isofibrations, which are displayed CwFs with an additional lifting structure of
isomorphisms. In short, this means that the semantics of infinitary constructors
can be only given mutually with a form of invariance under algebra
isomorphisms. Recursive equations similarly require this kind of semantics.

However, strict sort equations are not invariant under isomorphisms. For
example, if we have an isomorphism in $\mathbf{Set}\times\mathbf{Set}$ between
$(A, B)$ and $(A', B')$, and we also know that $A = B$ strictly, then it is not
necessarily the case that $A' = B'$. This means that a strict semantics for sort
equations is incompatible with the isofibration semantics for infinitary
constructors. Our current solution is to simply keep the troublesome features
apart. Hence we have
\begin{enumerate}
\item A theory of signatures supporting recursive equations and infinitary
  constructors, but no sort equations. This $\ToS$ can describe itself, and by a
  term model construction we can reduce all described QIITs to an assumed syntax
  of $\ToS$. The term model construction is more involved than in
  \cite{kaposi2019constructing}.
\item A theory of signatures supporting sort equations, but no recursive
  equations and infinitary constructors. This $\ToS$ is infinitary and has no
  sort equations, so we can give it a model theory as an infinitary QIIT. The
  QIITs described here support a stricter semantics which is not invariant under
  isomorphisms, and we also have a term model construction. Here, the semantics
  and the term models are straightforward extensions of
  \cite{kaposi2019constructing}.
\end{enumerate}

\bibliographystyle{plain}
\bibliography{references}

\end{document}

### 2020-01-19 draft/paper.tex

--- 2020-01-10 draft/paper.tex
@@
 following signature of categories.
+\begingroup
+\allowdisplaybreaks
 \begin{alignat*}{3}
@@
 \end{alignat*}
+\endgroup
 The benefit of having a QII signature is getting a model theory ``for free'',

### 2020-01-20 types2020abstract/abstract.tex
small fixes

@@
 raw syntax with well-formedness relations to describe signatures, which is
-rather unwieldy to formally handle. Also, the precursor of the current work
+rather unwieldy to handle formally. Also, the precursor of the current work
 \cite{kaposi2019constructing} used an ad-hoc QIIT to describe signatures, which
@@
   \item
-  We specify a notion of model for the theory of signatures (ToS); this is a
+  We specify a notion of model for the theory of signatures ($\ToS$); this is a
   category with family (CwF) extended with several type formers, allowing to
@@
   viewed as a fragment of a Church-encoding; here we do not care about
-  encoding the whole syntax of ToS, nor the initiality of the syntax, we only
+  encoding the whole syntax of $\ToS$, nor the initiality of the syntax, we only
   need a representation of signatures and the ability to interpret a signature in

### 2020-02-15 types2020abstract/abstract.tex
update TYPES abstract

--- 2020-01-19 types2020abstract/abstract.tex
@@
 \usepackage{doc}
+\usepackage{scalerel}
+\usepackage{amsfonts}
 
@@
 \newcommand{\latexeditor}{LEd}
+\newcommand{\emptycon}{\scaleobj{.75}\bullet}
+\newcommand{\ext}{\triangleright}
+\newcommand{\arri}{\Rightarrow}
 
@@
 
-\textbf{Self-describing signatures}. In the current work, we
+\subsubsection*{Self-describing signatures} In the current work, we
 would also like to streamline and make more rigorous the specification of
@@
     encoding the whole syntax of ToS, nor the initiality of the syntax, we only
-  need a representation of signatures and the ability to interpret a signature in
-  a $\mathsf{ToS}$-model.
+    need a representation of signatures and the ability to interpret a signature
+    in a $\mathsf{ToS}$-model. For example, the signature for natural number
+    algebras is a function
+    \[
+    \lambda(M : \mathsf{ToS}).\,(\emptycon_M\,\ext_M\,(N : \mathsf{U}_M)\,\ext_M\,(zero : \mathsf{El}_M\,N)\,\ext_M\,(suc : N\arri_M\,\mathsf{El}_M\,N))
+    \]
+    which maps every model $M$ to a typing context in $M$, consisting of the
+    declaration of a sort and two constructors.
+
   \item We give a semantics for signatures, as a particular $M : \mathsf{ToS}$
     model which interprets each CwF context $\Gamma$ as a structured category of
-    $\Gamma$-algebras.
+    $\Gamma$-algebras. E.g.\ for the signature of natural numbers, we get a
+    structured category of $\mathbb{N}$-algebras, with
+    $\mathbb{N}$-homomorphisms as morphisms.
   \item We give a (large, infinitary) signature for $\mathsf{ToS}$ itself, such
     that interpreting the signature in the semantic model yields a structured
-    category of $\mathsf{ToS}$-algebras.
+    category of $\mathsf{ToS}$-algebras. From this, we acquire notions of
+    \emph{recursion} and \emph{induction}, hence we gain the ability to define
+    further constructions by induction on an assumed initial model of the theory
+    of signatures.
 \end{enumerate}
@@
 
-\textbf{Extending semantics to infinitary constructors and sort equations.}
+\subsubsection*{Extending semantics to infinitary constructors and sort equations.}
 Previously in \cite{kaposi2019constructing}, contexts in $\ToS$ were interpreted
@@
 such CwFs, and types as displayed CwFs. Infinitary constructors force a major
-change: substitutions are interpreted as weak CwF morphisms, and types as CwF
-isofibrations, which are displayed CwFs with an additional lifting structure of
-isomorphisms. In short, this means that the semantics of infinitary constructors
-can be only given mutually with a form of invariance under algebra
+change: substitutions must be interpreted as weak CwF morphisms, and types as
+CwF isofibrations, which are displayed CwFs with an additional lifting structure
+for isomorphisms. In short, this means that the semantics of infinitary
+constructors can be only given mutually with a form of invariance under algebra
 isomorphisms. Recursive equations similarly require this kind of semantics.
@@
   term model construction we can reduce all described QIITs to an assumed syntax
-  of $\ToS$. The term model construction is more involved than in
-  \cite{kaposi2019constructing}.
+  of the same $\ToS$. This term model construction is also weakened (i.e.\ it is
+  up to algebra isomorphisms), hence it is significantly more complicated than
+  in \cite{kaposi2019constructing}.
 \item A theory of signatures supporting sort equations, but no recursive
   equations and infinitary constructors. This $\ToS$ is infinitary and has no
-  sort equations, so we can give it a model theory as an infinitary QIIT. The
-  QIITs described here support a stricter semantics which is not invariant under
+  sort equations, so we can give it a model theory as an infinitary QIIT. This
+  $\ToS$ supports a stricter semantics which is not invariant under
   isomorphisms, and we also have a term model construction. Here, the semantics

### 2020-03-18 draft/paper.tex
some response edits

@@
 families \cite{Dybjer96internaltype} (CwF).
-%
-Whereas the initial category can be given by first defining the type
-of objects (the empty type) and then defining morphisms as a separate
-inductive family, for the initial CwF one has to define $Ob$, $Mor$
-and two extra sorts $Ty$ and $Tm$ \emph{inductive-inductively}, at the
-same time. This is because there is a constructor of $Ob$ which refers
-to $Ty$, constructors of $Ty$ refer to $Ob$ and $Sub$, and so on. This
-is illustrated by the following excerpt of the signature for CwFs
-extending the previous signature for categories.
+
+Whereas the initial category can be given by first defining the type of objects
+(the empty type) and then defining morphisms separately, for an initial model of
+a type theory given as a CwF one has to define $Ob$, $Mor$ and two extra sorts
+$Ty$ and $Tm$ \emph{inductive-inductively}, at the same time. This is because
+there is a constructor of $Ob$ which refers to $Ty$, constructors of $Ty$ refer
+to $Ob$ and $Sub$, and so on. This is illustrated by the following excerpt of
+the signature for CwFs extending the previous signature for categories.
 \begin{alignat*}{3}
@@
 \end{alignat*}
-While signatures for type theory usually only contain finitary constructors,
+While signatures for type theories usually only contain finitary constructors,
 some QIITs have infinitary constructors, such as the following signature for
-countable branching trees where the order of subtrees does not matter.
+countably branching trees where the order of subtrees does not matter,
+i.e.\ trees with permuted subtrees in a branch are equal.
 \begin{alignat*}{3}
@@
   & perm && : (g :\N\ra Tree)(f : \N\ra\N)\ra \mathsf{isIso}\,f\ra \\
-  & && \hspace{0.8em} node\,g = node (g\circ f)
-\end{alignat*}
-Here $node$ and $perm$ are infinitary constructurs, $perm$ expresses
-that reordering subtrees by a permutation (a $\N\ra\N$ function which
-is an isomorphism) does not matter.
-
-Real-world examples examples of infinitary QIITs include the real, surreal numbers \cite{HoTTbook}, ordinal numbers \cite{lumsdaineShulman} and the partiality monad \cite{partiality}.
+  & && \hspace{0.8em} node\,g = node\,(g\circ f)
+\end{alignat*}
+Additional examples for infinitary QIITs are real, surreal numbers
+\cite{HoTTbook}, ordinal numbers \cite{lumsdaineShulman} and the partiality
+monad \cite{partiality}.
 
@@
   & \Inductive : \Con_{\bGamma} \ra \Set_i \\
-  & \Inductive\,\Gamma := (A : \Ty_{\bGamma}\,\Gamma)\ra \Tm_{\bGamma}\,\Gamma\,A
+  & \Inductive\,\Gamma := (A : \Ty_{\boldsymbol{\Gamma}}\,\Gamma)\ra \Tm_{\bGamma}\,\Gamma\,A
 \end{alignat*}
@@
 
-A \emph{weak flCwF morphism} $\boldsymbol{\sigma : \Sub\,\Gamma\,\Delta}$ is a functor between underlying categories, which additionally maps types to types and terms to terms, and satisfies the following:
+A \emph{weak flCwF morphism} $\boldsymbol{\sigma : \Sub\,\Gamma\,\Delta}$ is a functor between underlying categories, which additionally maps types to types and terms to terms, and satisfies the following mere property:
   \begin{enumerate}
@@
 
-In short, $\bsigma$ preserves substitution strictly and preserves
-empty context and context extension up to isomorphism. We notate the evident
-isomorphisms as $\bsigma_{\emptycon} :
-\bsigma\,\emptycon \simeq \emptycon$ and $\bsigma_{\ext}
-:
-\bsigma\,(\Gamma\,\ext\,A)\,\simeq\,\bsigma\,\Gamma\,\ext\,\bsigma\,A$. Our notion of weak morphism is the same as in \cite{birkedal2018modal}, when restricted to CwFs.
+In short, $\bsigma$ preserves substitution strictly and preserves empty context
+and context extension up to isomorphism. We notate the evident isomorphisms as
+$\bsigma_{\emptycon} : \bsigma\,\emptycon \simeq \emptycon$ and $\bsigma_{\ext}
+: \bsigma\,(\Gamma\,\ext\,A)\,\simeq\,\bsigma\,\Gamma\,\ext\,\bsigma\,A$. Our
+notion of weak morphism is the same as in \cite{birkedal2018modal}, when
+restricted to CwFs.
 
@@
 likewise components for substitution are given by corresponding components in
-$\bA$.
-
-Context and type formers are given by coercing $\bA$
+$\bA$. Context and type formers are given by coercing $\bA$
 structures along $\bsigma$ preservation isomorphisms. For example:
@@
 \subsection{Term Substitution}
-
 $\boldsymbol{\blank[\blank] : \Tm\,\Delta\,A \ra (\sigma : \Sub\,\Gamma\,\Delta)
@@
 \end{alignat*}
-
 We also have the same decomposition of derived isomorphisms as in Lemma

### 2020-03-18 draft/paper.tex
cleanup references

@@
 \newcommand{\NatSig}{\mathsf{NatSig}}
+\newcommand{\mi}[1]{\mathit{#1}}
 
@@
 \begin{alignat*}{3}
-  & Ob  && : \U \\
-  & Mor && : Ob\ra Ob\ra\U \\
-  & \blank\circ\blank && : Mor\,J\,K\ra Mor\,I\,J\ra Mor\,I\,K \\
-  & id && : Mor\,I\,I \\
-  & ass && : (f\circ g)\circ h = f\circ(g\circ h) \\
-  & idl && : id\circ f = f \\
-  & idr && : f\circ id = f
+  & \mi{Ob}  && : \U \\
+  & \mi{Mor} && : \mi{Ob}\ra \mi{Ob}\ra\U \\
+  & \blank\circ\blank && : \mi{Mor}\,J\,K\ra \mi{Mor}\,I\,J\ra \mi{Mor}\,I\,K \\
+  & \mi{id} && : \mi{Mor}\,I\,I \\
+  & \mi{ass} && : (f\circ g)\circ h = f\circ(g\circ h) \\
+  & \mi{idl} && : \mi{id}\circ f = f \\
+  & \mi{idr} && : f\circ \mi{id} = f
 \end{alignat*}
@@
 and comprehension categories \cite{jacobs1993comprehension} to categories with
-families \cite{Dybjer96internaltype} (CwF).
+families \cite{internaltt} (CwF).
 
@@
 (the empty type) and then defining morphisms separately, for an initial model of
-a type theory given as a CwF one has to define $Ob$, $Mor$ and two extra sorts
-$Ty$ and $Tm$ \emph{inductive-inductively}, at the same time. This is because
-there is a constructor of $Ob$ which refers to $Ty$, constructors of $Ty$ refer
-to $Ob$ and $Sub$, and so on. This is illustrated by the following excerpt of
-the signature for CwFs extending the previous signature for categories.
-\begin{alignat*}{3}
-  & Ty && : Ob\ra \U \\
-  & empty  && : Ob \\
-  & ext && : (I:Ob)\ra Ty\,I\ra Ob \\
-  & \blank[\blank] && : Ty\,J\ra Mor\,I\,J\ra Ty\,I
+a type theory given as a CwF one has to define $\mi{Ob}$, $\mi{Mor}$ and two
+extra sorts $\mi{Ty}$ and $\mi{Tm}$ \emph{inductive-inductively}, at the same
+time. This is because there is a constructor of $\mi{Ob}$ which refers to
+$\mi{Ty}$, constructors of $\mi{Ty}$ refer to $\mi{Ob}$ and $\mi{Sub}$, and so
+on. This is illustrated by the following excerpt of the signature for CwFs
+extending the previous signature for categories.
+\begin{alignat*}{3}
+  & \mi{Ty}        && : \mi{Ob}\ra \U \\
+  & \mi{empty}     && : \mi{Ob} \\
+  & \mi{ext}       && : (I:\mi{Ob})\ra \mi{Ty}\,I\ra \mi{Ob} \\
+  & \blank[\blank] && : \mi{Ty}\,J\ra \mi{Mor}\,I\,J\ra \mi{Ty}\,I
 \end{alignat*}
@@
 \begin{alignat*}{3}
-  & Tree && : \U \\
-  & leaf  && : Tree \\
-  & node && : (\N\ra Tree)\ra Tree \\
-  & perm && : (g :\N\ra Tree)(f : \N\ra\N)\ra \mathsf{isIso}\,f\ra \\
-  & && \hspace{0.8em} node\,g = node\,(g\circ f)
+  & \mi{Tree} && : \U \\
+  & \mi{leaf}  && : \mi{Tree} \\
+  & \mi{node} && : (\N\ra \mi{Tree})\ra \mi{Tree} \\
+  & \mi{perm} && : (g :\N\ra \mi{Tree})(f : \N\ra\N)\ra \mathsf{isIso}\,f\ra \\
+  & && \hspace{0.8em} \mi{node}\,g = \mi{node}\,(g\circ f)
 \end{alignat*}
@@
 The current paper extends the syntax and semantics of QIITs as previously described
-in the literature \cite{kaposi2019constructing, gabe, dijkstra2017quotient}. We add the following features:
+in the literature \cite{kaposi2019constructing, qiits, dijkstra2017quotient}. We add the following features:
 \begin{enumerate}
@@
   of constructors. These have occurred previously in syntaxes of cubical type
-  theories, as boundary conditions \cite{cohen2016cubical, angiuli2016computational, angiuli2018cartesian}.
+  theories, as boundary conditions \cite{cchm, angiuli2016computational, angiuli2018cartesian}.
 \end{enumerate}
@@
 
-We have Russell-style predicative universes $\Set_i$ indexed by natural
-numbers, dependent functions as $(x : A)\ra B$, and dependent pairs as $(x :
-A)\times B$ with projections $\proj1$ and $\proj2$. We sometimes leave
-parameters implicit in dependent function types, e.g.\ write $\id : A \ra A$
-instead of $\id : (A : \Set_i)\ra A \ra A$. We also use subscripts as a field
-projection notation for iterated pairs. For example, for $t : (A : \Set_i)
-\times (B : \Set_i) \times (f : A \ra B)$, we use $B_t$ to denote the projection
-of the second component. Sometimes we omit the subscript if it is clear from
-context. When we write ``exists'' in this paper, we always mean chosen structure
-given by a $\Sigma$-type.
+We have Russell-style predicative universes $\Set_i$ indexed by natural numbers,
+dependent functions as $(x : A)\ra B$, and dependent pairs as $(x : A)\times
+B$. We sometimes leave parameters implicit in dependent function types,
+e.g.\ write $\id : A \ra A$ instead of $\id : (A : \Set_i)\ra A \ra A$. We also
+use subscripts as a field projection notation for iterated pairs. For example,
+for $t : (A : \Set_i) \times (B : \Set_i) \times (f : A \ra B)$, we use $B_t$ to
+denote the projection of the second component. Sometimes we omit the subscript
+if it is clear from context. When we write ``exists'' in this paper, we always
+mean chosen structure given by a $\Sigma$-type.
 
@@
 algebras. For example, consider natural number algebras at level $i$, given as
-the $\Sigma$-type $NatAlg_i :=(Nat : \Set_i)\times Nat\times(Nat \ra
-Nat)$. It follows from the subtyping rules that $i \leq j$ implies $NatAlg_{i}
-\leq NatAlg_{j}$. However, cumulativity for arbitrary QIIT algebras does not
-follow judgmentally; it can only be proven by induction on signatures, hence
-the need for $\Subtype$.
+the $\Sigma$-type $\mi{NatAlg}_i :=(\mi{Nat} : \Set_i)\times
+\mi{Nat}\times(\mi{Nat} \ra \mi{Nat})$. It follows from the subtyping rules that
+$i \leq j$ implies $NatAlg_{i} \leq \mi{NatAlg}_{j}$. However, cumulativity for
+arbitrary QIIT algebras does not follow judgmentally; it can only be proven by
+induction on signatures, hence the need for $\Subtype$.
 
@@
 \Tm\,\Gamma\,\U)\ra \Tm\,\Gamma\,\U$, with $\appinf :
-\Tm\,\Gamma\,(\El\,(\Pie\,A\,b))\ra((x : A)\ra \Tm\,\Gamma\,(\El\,(b\,x)))$ and
+\Tm\,\Gamma\,(\El\,(\Piinf\,A\,b))\ra((x : A)\ra \Tm\,\Gamma\,(\El\,(b\,x)))$ and
 its inverse $\laminf$.
@@
   & \NatSig := \lambda (i :\mathbb{N})(M : \ToS_{i,0}).\\
-  & \hspace{1em}(\emptycon_M\,\ext_M\, (N : \U_M) \,\ext_M\,(zero : \El_M\,N)\\
-  & \hspace{2.67em}\ext_M (suc : N\arri_M\El_M\,N))
+  & \hspace{1em}(\emptycon_M\,\ext_M\, (N : \U_M) \,\ext_M\,(\mi{zero} : \El_M\,N)\\
+  & \hspace{2.67em}\ext_M (\mi{suc} : N\arri_M\El_M\,N))
 \end{alignat*}
@@
 & \mathsf{TreeSig} := \\
-& \hspace{1em}\emptycon \ext (Tree : \U)\\
-& \hspace{1.35em}\ext (node : \Pie\,\Set_0\,(\lambda A.\,(A\,\arrinf\,Tree)\arri\El\,Tree))
+& \hspace{1em}\emptycon \ext (\mi{Tree} : \U)\\
+& \hspace{1.35em}\ext (\mi{node} : \Pie\,\Set_0\,(\lambda A.\,(A\,\arrinf\,\mi{Tree})\arri\El\,\mi{Tree}))
 \end{alignat*}
@@
 More interesting (and complicated) examples for recursive equations are boundary
-conditions in various cubical type theories \cite{cohen2016cubical, angiuli2016computational, angiuli2018cartesian}. Note that our $\Id$
+conditions in various cubical type theories \cite{cchm, angiuli2016computational, angiuli2018cartesian}. Note that our $\Id$
 allows iterated equations as well, but these are all trivial in the semantics,
@@
 Our semantics has a ``type-theoretic'' flavor, which is inspired by the cubical
-set model of Martin-Löf type theory by Bezem et al. \cite{cubical}. The core idea
+set model of Martin-Löf type theory by Bezem et al. \cite{cubicalmodel}. The core idea
 is to avoid strictness issues by starting from basic ingredients which are already
@@
 for $i$, since algebras merely contain elements of $A : \Set_j$ types, while
-inductive sets are themselves elements of $\Set_i$. For example, $NatAlg_i :
+inductive sets are themselves elements of $\Set_i$. For example, $\mi{NatAlg}_i :
 \Set_{\max(i+1,0)} : \Set_{\max(i+1, 0)+1}$.
@@
 
-A \emph{weak flCwF morphism} $\boldsymbol{\sigma : \Sub\,\Gamma\,\Delta}$ is a functor between underlying categories, which additionally maps types to types and terms to terms, and satisfies the following mere property:
+A \emph{weak flCwF morphism} $\boldsymbol{\sigma : \Sub\,\Gamma\,\Delta}$ is a
+functor between underlying categories, which also maps types to types and terms
+to terms, and satisfies the following mere properties:
   \begin{enumerate}
@@
 : \bsigma\,(\Gamma\,\ext\,A)\,\simeq\,\bsigma\,\Gamma\,\ext\,\bsigma\,A$. Our
-notion of weak morphism is the same as in \cite{birkedal2018modal}, when
+notion of weak morphism is the same as in \cite{birkedal2020modal}, when
 restricted to CwFs.
@@
 locally cartesian closed and does not support a general $\Pi$ type
-\cite{johnstone2002sketches}.
+\cite[Section A1.5]{johnstone2002sketches}.
 
@@
 and we can convert in the other direction by using ToS-induction to interpret a
-$\Gamma : \Con_{\syn_j}$ in an arbitrary ToS model.
+$\Gamma : \Con_{\syn_j}$ in an arbitrary ToS model. This is merely a logical
+equivalence, external to cETT (because of unvierse polymorphism), and not an
+isomorphism.
 
@@
 
-In the following, We will need three model fragments of $\bM$, which can be used
+In the following, we will need three model fragments of $\bM$, which can be used
 to compute notions of algebras, displayed algebras and sections respectively for
@@
   & \Con_{\bD}\,\ulGamma && = \ulGamma \ra \Set_{\max(i+1, j)} \\
-  & \Sub_{\bD}\,\Gamma\,\Delta\,\ulsigma && = (\ulgamma : \ulGamma) \ra \Gamma\,\gamma \ra \Delta\,(\ulsigma\,\ulgamma)\\
+  & \Sub_{\bD}\,\Gamma\,\Delta\,\ulsigma && = (\ulgamma : \ulGamma) \ra \Gamma\,\ulgamma \ra \Delta\,(\ulsigma\,\ulgamma)\\
   & \Ty_{\bD}\,\Gamma\,\ulA && = (\ulgamma : \ulGamma) \ra \Gamma\,\ulgamma \ra \ulA\,\ulgamma \ra \Set_{\max(i+1, j)}\\
-  & \Tm_{\bD}\,\Gamma\,A\,\ult && = (\ulgamma : \ulGamma)(\gamma : \Gamma\,\ulGamma)
+  & \Tm_{\bD}\,\Gamma\,A\,\ult && = (\ulgamma : \ulGamma)(\gamma : \Gamma\,\ulgamma)
     \ra A\,\ulgamma\,\gamma\,(\ult\,\ulgamma)
@@
 \[
-\NatSig := \emptycon\ext(N : \U)\ext(zero : \El\,N)\ext (suc : N
+\NatSig := \emptycon\ext(N : \U)\ext(\mi{zero} : \El\,N)\ext (\mi{suc} : N
 \arri \El\,N)
@@
 of natural numbers, since, intuitively, such terms can only be built from
-iterated usage of $zero$ and $suc$. We build a term algebra for each
+iterated usage of $\mi{zero}$ and $\mi{suc}$. We build a term algebra for each
 signature in this manner.
@@
   & \Sub_{\bT_{\ulOmega}}\,\Gamma\,\Delta\,\ulsigma &&:= (\ulnu : \Sub\,\ulOmega\,\ulGamma)
-  \ra \Delta\,(\ulsigma\circ\nu) \simeq \llb\ulsigma\rrb_{\bA}\,(\Gamma\,\ulnu)\\
+  \ra \Delta\,(\ulsigma\circ\ulnu) \simeq \llb\ulsigma\rrb_{\bA}\,(\Gamma\,\ulnu)\\
   & \Ty_{\bT_{\ulOmega}}\,\Gamma\,\ulA &&:= (\ulnu : \Sub\,\ulOmega\,\ulGamma) \ra
@@
 proof-relevant identity types. They were introduced before QIITs
-\cite{HoTTbook}. \cite{niels} is a description of a subset of HITs. \cite{hiit}
+\cite{HoTTbook}. \cite{hitsinprogramming} is a description of a subset of HITs. \cite{hiit}
 describes a syntax for higher inductive-inductive types using a theory of
@@
 
-With regards to the reduction of QIITs to simple type formers, the
-reduction of infinitary QIITs appears to be more challenging than the finitary
+With regards to the reduction of QIITs to simple type formers, the reduction of
+infinitary QIITs appears to be more challenging than the finitary
 case. \cite[Section 9]{lumsdaineShulman} shows that infinitary QIITs are not
-constructible in type theory without choice, so such QIITs are a
-non-conservative extension of extensional type theory. In the finitary case, a
-generalization of the approach in \cite{induction-is-enough} seems promising ;
-this amounts to a Streicher-style initial algebra construction
+constructible from inductive types and simple quotients with relations. In the
+finitary case, a generalization of the approach in \cite{induction-is-enough}
+seems promising ; this amounts to a Streicher-style initial algebra construction
 \cite{streicher2012semantics} for the theory of finitary QIIT signatures. In
-particular, Brunerie et al. \cite{brunerie} has formalized in Agda this
+particular, Brunerie et al. \cite{brunerie} have formalized in Agda this
 construction for a comparable type theory, using UIP, function extensionality,

### 2020-03-18 draft/response.txt
response draft

Response to review 1:

- "it would improve the paper to give more of an intuitive understanding"

We agree and will try to find space for this. A possible intuitive explanation
is that in KKA the universe U was empty, and all substitutions were "neutral",
i.e. semantic subsitutions were functors which permute, duplicate or forget
components of algebras, or freely reinterpret components, and all such functors
strictly preserve limits. In the current paper, U is not empty, it's closed
under identity and infinitary function types, which don't just simply reshuffle
structure and hence preserve limits only weakly.

- "As stated this is only a "logical equivalence,""

We will note this. Isomorphism is probably not feasible, while impredicative
Church codings work to some extent (see "Impredicative encodings of HITs" by
Awodey et al.), they don't support large elimination, and we need interpretation of
signatures in large models.

Response to review 2:

- "it would be good to have a story how to work with QIITs in practice"

We didn't elaborate QIIT usage and applications because of space constraints.
However, we don't think "escaping from quotients" is an issue in practice. We
can print quotients as a propositionally truncated strings, which work just as
well as printing untruncated strings. Likewise with first digits of real
numbers. We agree though that working with quotients requires at least some
deliberation and rephrasing.

- "El (Π^{inf} A b) = Π^{ext} A (El ∘ b)"

This equation plausibly holds in the M model, we have not included/checked it
because it is not really important with respect to practical usage of signatures.

- (A : Ty_Γ₀ Γ) → Tm_Γ₀ Γ A

We'll try to add more illustration. Here probably the best would be to note the
following:

  Con_Γ        type of Γ-algebras
  Sub_Γ Γ Δ    type of homomorphisms from Γ to Δ
  Ty_Γ Γ       type of displayed algebras over Γ
  Tm_Γ Γ A     type of dependent morphisms from Γ to A

If Γ₀ is the CwF of Nat-algebras, then Γ : Con_Γ₀ is a particular Nat-algebra
(N, z, s), A has type (Nᴾ : N → Set) × (zᴾ : Nᴾ z) × (sᴾ : (n : N) → Nᴾ n → Nᵖ
(s n)) and Tm_Γ₀ Γ A is a type which is equal to (Nˢ : (n : N) → Nᴾ n) × (Nˢ z =
zᴾ) × ((n : N) → Nˢ (s n) = sᴾ n (Nˢ n)). It could be illustrative as well to
see how "initiality" looks like in the same setting:

  Initial Γ := (Δ : Con_Γ) → isContr (Sub_Γ Γ Δ)

"Inductive" is a dependently typed analogue of "Initial" in the sense that there
is a dependent morphism (Tm Γ A) from an inductive object Γ to any A dependent
object over it.

Response to review 3:

- "Bootstrapping how exactly?"

The benefit of bootstrapping is that the only ToS-related thing which we have to
manually invent is the notion of ToS model in def 3.1, and notions of
ToS-induction, ToS-morphism, etc. are computed by the semantics of ToSSig. The
induction principle for ToS is rather huge (it says that every displayed ToS
algebra has a section), and it was a bit unsatisfactory in the previous
"Constructing QIITs" paper that it was ad-hoc assumed/invented.

- "the signature ToSSig helps you
   with the notion of ToS-induction, but what is the concrete link with
   the assumed initial model syn in Section 6?"

The concrete link is that we need a notion of initiality before we can assume an
*initial* ToS model. The M interpretation of ToSSig computes (among other
things) this. ToSSig is defined in a way so that the Con part of its M
interpretation yields exactly the notion of ToS algebra from def 3.1, so that we
get an flCwF where objects are ToS algebras and morphisms are strict ToS
morphisms.

- "Why finite limit CwFs?"

The equivalence of induction and initiality (which we reference from
"Constructing QIITs") requires some (not all) components of CwF + identity +
constant families. We add all CwF+Id+K+Sigma components so that we get the
structure which is (bi)equivalent to finitely complete categories. There are also
many potentially useful/interesting constructions on such CwFs in the works of
Dybjer, Clairambault & Castellan, so we think that it's worth to support this
structure. We'll try to make this motivation clearer

- "Making sense of Section 6"
- "but the flCwF structure seems to be missing"

There is no fl-structure involved here, we define three (displayed) models of
ToS, the set model, the logical predicate model and the section model. We agree
that it is hard to parse what is happening here, as ToS is a CwF with extra
structure (def 3.1), but in M, the interpretation of each context is also a CwF
with extra structure (flCwF), so we have "a structured CwF of structured
CwFs". This part is explained and illustrated in detail in the previous
"Constructing QIITs" work, we should direct the reader there more clearly in the
text (sadly we don't have space for similar exposition here)

- "That size issue again"

We can actually eliminate into any type from the (Nat : Set₁) given by the term
model: while the induction principle from Thm 6.10 requires as input a displayed
Nat-algebra at level 1, by cumulativity it is enough to provide a disp. algebra at
level 0. This should work for any particular QIIT. We only
included the general cumulativity proof for plain algebras in def 6.8 and omitted it
for displayed algebras. The displayed case is a bit more fiddly, because we need to handle
cases where the base algebra is at a different level than the displayed algebra itself.
Levels are rather painful, we made an effort to keep the universe setup as simple as
possible, and we made a point to avoid "heterogeneously" leveled constructions such as this.
We'll try to note this in the paper.

- "The category example in Section 1 is not very convincing"

The Cat signature itself requires (finitary) QII features to specify, and it's a
very prominent structure. It happens to be the case that the initial Cat is easy
to construct, but we don't think that detracts from its value as an example; also,
our M semantics provides a bunch of structure unrelated to initial algebras.

- "Smaller comments & typos"

We will try to improve based on these comments

- "Could the usual parametricity arguments be used to show that your
  "Church-encoded" definition of signatures in fact is initial?"

Definitely not internally to cETT, because of the universe polymorphism, but
even without that it's probably not feasible; see "Impredicative encodings of
HITs" by Awodey et al. We don't really need the Church-encoding to be nice and
well-behaved in this paper, we only need a way to write down signatures and
interpret them in a model, similarly to how programmers use "finally tagless"
encodings.

--------------------------------------------------------------------------------

TODO:
- intuitive expl for isofib semantics necessity
- directing readers to CQIITs in sec 6
- improve based on review 3 "Smaller comments"
- response to displayed algebra cumulativity

### 2020-04-03 draft/response.txt
typo fix

@@
 However, we don't think "escaping from quotients" is an issue in practice. We
-can print quotients as a propositionally truncated strings, which work just as
-well as printing untruncated strings. Likewise with first digits of real
-numbers. We agree though that working with quotients requires at least some
-deliberation and rephrasing.
+can print quotients as propositionally truncated strings, which works just as
+well as printing untruncated strings. We agree though that working with
+quotients requires at least some deliberation and rephrasing.
 
@@
 zᴾ) × ((n : N) → Nˢ (s n) = sᴾ n (Nˢ n)). It could be illustrative as well to
-see how "initiality" looks like in the same setting:
+see how initiality looks like in the same setting:
 
@@
 ToS-induction, ToS-morphism, etc. are computed by the semantics of ToSSig. The
-induction principle for ToS is rather huge (it says that every displayed ToS
-algebra has a section), and it was a bit unsatisfactory in the previous
-"Constructing QIITs" paper that it was ad-hoc assumed/invented.
+induction principle for ToS is rather huge, and it was a bit unsatisfactory in
+the previous "Constructing QIITs" paper that it was ad-hoc assumed/invented.
 
@@
 Dybjer, Clairambault & Castellan, so we think that it's worth to support this
-structure. We'll try to make this motivation clearer
+structure. We'll try to make this motivation clearer.
 
@@
 model: while the induction principle from Thm 6.10 requires as input a displayed
-Nat-algebra at level 1, by cumulativity it is enough to provide a disp. algebra at
-level 0. This should work for any particular QIIT. We only
-included the general cumulativity proof for plain algebras in def 6.8 and omitted it
-for displayed algebras. The displayed case is a bit more fiddly, because we need to handle
-cases where the base algebra is at a different level than the displayed algebra itself.
-Levels are rather painful, we made an effort to keep the universe setup as simple as
-possible, and we made a point to avoid "heterogeneously" leveled constructions such as this.
-We'll try to note this in the paper.
+Nat-algebra at level 1 or more, by cumulativity it is enough to make the
+induction motive return in Set₀. This should work for any particular QIIT. We
+only included the general cumulativity proof for plain algebras in def 6.8 and
+omitted it for displayed algebras. The displayed case is a bit more fiddly,
+because we need to handle cases where the base algebra is at a different level
+than the displayed algebra itself.  Levels are rather painful, we made an effort
+to keep the universe setup as simple as possible, and we purposefully avoided
+"heterogeneously" leveled constructions such as this.  We'll try to note this in
+the paper.
 
@@
 - intuitive expl for isofib semantics necessity
+- flCwf motivation clearer
 - directing readers to CQIITs in sec 6
@@
 - response to displayed algebra cumulativity
+
+Set₀

### 2020-04-11 draft/paper.tex
typo fix
cut fluff from intro to make space for review edits

@@
 \begin{alignat*}{3}
-  & \mi{Ob}  && : \U \\
-  & \mi{Mor} && : \mi{Ob}\ra \mi{Ob}\ra\U \\
+  & \mi{Ob}  && : \Set \\
+  & \mi{Mor} && : \mi{Ob}\ra \mi{Ob}\ra\Set \\
   & \blank\circ\blank && : \mi{Mor}\,J\,K\ra \mi{Mor}\,I\,J\ra \mi{Mor}\,I\,K \\
@@
 
-Whereas the initial category can be given by first defining the type of objects
-(the empty type) and then defining morphisms separately, for an initial model of
-a type theory given as a CwF one has to define $\mi{Ob}$, $\mi{Mor}$ and two
-extra sorts $\mi{Ty}$ and $\mi{Tm}$ \emph{inductive-inductively}, at the same
-time. This is because there is a constructor of $\mi{Ob}$ which refers to
-$\mi{Ty}$, constructors of $\mi{Ty}$ refer to $\mi{Ob}$ and $\mi{Sub}$, and so
-on. This is illustrated by the following excerpt of the signature for CwFs
-extending the previous signature for categories.
-\begin{alignat*}{3}
-  & \mi{Ty}        && : \mi{Ob}\ra \U \\
-  & \mi{empty}     && : \mi{Ob} \\
-  & \mi{ext}       && : (I:\mi{Ob})\ra \mi{Ty}\,I\ra \mi{Ob} \\
-  & \blank[\blank] && : \mi{Ty}\,J\ra \mi{Mor}\,I\,J\ra \mi{Ty}\,I
-\end{alignat*}
-While signatures for type theories usually only contain finitary constructors,
-some QIITs have infinitary constructors, such as the following signature for
-countably branching trees where the order of subtrees does not matter,
-i.e.\ trees with permuted subtrees in a branch are equal.
-\begin{alignat*}{3}
-  & \mi{Tree} && : \U \\
-  & \mi{leaf}  && : \mi{Tree} \\
-  & \mi{node} && : (\N\ra \mi{Tree})\ra \mi{Tree} \\
-  & \mi{perm} && : (g :\N\ra \mi{Tree})(f : \N\ra\N)\ra \mathsf{isIso}\,f\ra \\
-  & && \hspace{0.8em} \mi{node}\,g = \mi{node}\,(g\circ f)
-\end{alignat*}
-Additional examples for infinitary QIITs are real, surreal numbers
-\cite{HoTTbook}, ordinal numbers \cite{lumsdaineShulman} and the partiality
-monad \cite{partiality}.
-
-\subsection{Contributions}
-The current paper extends the syntax and semantics of QIITs as previously described
-in the literature \cite{kaposi2019constructing, qiits, dijkstra2017quotient}. We add the following features:
+A prime motivation of the current work is to further develop QIITs as a
+framework for the metatheory of type theories, to cover more theories and
+support more applications. To this end, we extend the syntax and semantics of
+QIITs as previously described in the literature \cite{kaposi2019constructing,
+  qiits, dijkstra2017quotient}.
+
+%% Whereas the initial category can be given by first defining the type of objects
+%% (the empty type) and then defining morphisms separately, for an initial model of
+%% a type theory given as a CwF, one has to define $\mi{Ob}$, $\mi{Mor}$ and two
+%% extra sorts $\mi{Ty}$ and $\mi{Tm}$ \emph{inductive-inductively}, at the same
+%% time. This is illustrated by the following excerpt of the signature for CwFs
+%% extending the previous signature for categories.
+%% \begin{alignat*}{3}
+%%   & \mi{Ty}        && : \mi{Ob}\ra \Set \\
+%%   & \mi{U}         && : \mi{Ty\,i} \\
+%%   & \mi{empty}     && : \mi{Ob} \\
+%%   & \mi{ext}       && : (I:\mi{Ob})\ra \mi{Ty}\,I\ra \mi{Ob} \\
+%%   & \blank[\blank] && : \mi{Ty}\,J\ra \mi{Mor}\,I\,J\ra \mi{Ty}\,I
+%% \end{alignat*}
+%% While signatures for type theories usually only contain finitary constructors,
+%% some QIITs have infinitary constructors, such as the following signature for
+%% countably branching trees where the order of the subtrees does not matter,
+%% i.e.\ permuting subtrees yields an equal tree.
+%% \begin{alignat*}{3}
+%%   & \mi{Tree} && : \U \\
+%%   & \mi{leaf}  && : \mi{Tree} \\
+%%   & \mi{node} && : (\N\ra \mi{Tree})\ra \mi{Tree} \\
+%%   & \mi{perm} && : (g :\N\ra \mi{Tree})(f : \N\ra\N)\ra \mathsf{isIso}\,f\ra \\
+%%   & && \hspace{0.8em} \mi{node}\,g = \mi{node}\,(g\circ f)
+%% \end{alignat*}
+%% Additional examples for infinitary QIITs are real, surreal numbers
+%% \cite{HoTTbook}, ordinal numbers \cite{lumsdaineShulman} and the partiality
+%% monad \cite{partiality}.
+
+%% \subsection{Contributions}
+%% The current paper extends the syntax and semantics of QIITs as previously described
+%% in the literature \cite{kaposi2019constructing, qiits, dijkstra2017quotient}. We add the following features:
+
 \begin{enumerate}
@@
   routinely used in the metatheory of type theories, but they have not been
-  presented explicitly in previous QIIT literature.
+  presented explicitly in previous QIIT literature. For example, interpreting
+  syntactic contexts as sets already requires a notion of large models.
   \item
   \textbf{Infinitary constructors}.
-  Of special note here is that the theory of QIIT signatures is itself large and
-  infinitary, thus it can ``eat itself'', i.e.\ include its own signature and
-  provide its own metatheory. This was not possible previously in
-  \cite{kaposi2019constructing}, where only finitary QIITs were described. In
-  this paper we use self-representation to bootstrap the model theory of
-  signatures, without having to assume any pre-existing internal syntax.
+  This allows specification of infinitely branching trees. Examples of
+  infinitary QIITs in previous works include real, surreal numbers
+  \cite{HoTTbook}, ordinal numbers \cite{lumsdaineShulman} and a partiality
+  monad \cite{partiality}. Of special note here is that the theory of QIIT
+  signatures is itself large and infinitary, thus it can ``eat itself'',
+  i.e.\ include its own signature and provide its own metatheory. This was not
+  possible previously in \cite{kaposi2019constructing}, where only finitary
+  QIITs were described. In this paper we use self-representation to bootstrap
+  the model theory of signatures, without having to assume any pre-existing
+  internal syntax.
   \item
@@
 & (\bsigma\,\Gamma\,\ext\,\bsigma\,(\Sigma\,A\,B)) \simeq
-  (\bsigma\,\Gamma\,\ext\,\bsigma\,A\,\ext\,(\Sigma\,B)[\bsigma_{\ext}^{-1}]) \\
-& \simeq (\bsigma\,\Gamma\,\ext\,\Sigma\,(\bsigma\,A)\,((\Sigma\,B)[\bsigma_{\ext}^{-1}]))
+  (\bsigma\,\Gamma\,\ext\,\bsigma\,A\,\ext\,(\bsigma\,B)[\bsigma_{\ext}^{-1}]) \\
+& \simeq (\bsigma\,\Gamma\,\ext\,\Sigma\,(\bsigma\,A)\,((\bsigma\,B)[\bsigma_{\ext}^{-1}]))
 \end{alignat*}

### 2020-04-11 draft/paper.tex
fixes

@@
 
-Cartmell \cite{gat} defines generalized algebraic theories (GATs) using type
-theoretic syntax. Compared to our QII signatures, his signatures support sort
-equations but do not cover infinitary constructors or recursive equations. A way
-to encode sort equations in our signatures is using isomorphisms instead of
-equalities. In contrast to our algebraic definition, Cartmell's signatures are
-given by presyntax, named variables and typing relations, there is no explicit
-model theory provided for signatures, and no explicit term model construction is
-given. Cartmell focuses instead on showing that contextual categories serve as
-classifying categories for GATs.
-
-A more semantic approach to QIITs is given by Altenkirch et al.\
-\cite{gabe}. They generalize the initial algebra semantics of
-inductive types to QIITs by considering towers of functors and
-building complete categories of algebras from such towers. Their
-notion of signature does not enforce strict positivity, hence
-describes a larger class of QII signatures. The equivalence of
-initiality and induction is showed, but initial algebras are not
-constructed.
+Cartmell \cite{gat} defines generalized algebraic theories (GATs) using
+type-theoretic syntax. Compared to our QII signatures, his signatures support
+sort equations but do not cover infinitary constructors or recursive
+equations. A way to encode sort equations in our signatures is using
+isomorphisms instead of equalities. In contrast to our algebraic definition,
+Cartmell's signatures are given by presyntax, named variables and typing
+relations, there is no explicit model theory provided for signatures, and no
+explicit term model construction is given. Cartmell focuses instead on showing
+that contextual categories serve as classifying categories for GATs.
+
+A more semantic approach to QIITs is given by Altenkirch et
+al.\ \cite{qiits}. They generalize the initial algebra semantics of inductive
+types to QIITs by considering towers of functors and building complete
+categories of algebras from them. Their notion of signature does not enforce
+strict positivity, hence describes a larger class of QII signatures. They show
+equivalence of initiality and induction, but do not construct initial algebras,
+nor would that be possible because of the lack of a positivity restriction.
 
@@
 proof-relevant identity types. They were introduced before QIITs
-\cite{HoTTbook}. \cite{hitsinprogramming} is a description of a subset of HITs. \cite{hiit}
-describes a syntax for higher inductive-inductive types using a theory of
-signatures similar to ours, however does not construct categories of algebras or
-show initiality. Semantics for different subclasses of HITs are given by
+\cite{HoTTbook}. \cite{hiit} describes a syntax for higher inductive-inductive
+types using a theory of signatures similar to ours, but it does not construct
+categories of algebras and initial algebras. Semantics for different subclasses
+of HITs are given by
 \cite{lumsdaineShulman,10.1145/3209108.3209130,nielsmsc,sojakova,moeneclaey}. Cubical
-type theories were shown to support HITs in a computational way
-\cite{niels,hiit}.
+type theories were shown to support some HITs in a computational way
+\cite{cubicalhits,10.1145/3290314}.
 
@@
 finitary case, a generalization of the approach in \cite{induction-is-enough}
-seems promising ; this amounts to a Streicher-style initial algebra construction
+seems promising; this amounts to a Streicher-style initial algebra construction
 \cite{streicher2012semantics} for the theory of finitary QIIT signatures. In

### 2020-04-12 draft/paper.tex
fmt change & induction response

@@
-%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-\documentclass[sigplan,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%% For double-blind review submission, w/ CCS and ACM Reference
-%\documentclass[sigplan,review,anonymous]{acmart}\settopmatter{printfolios=true}
-%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
-%\documentclass[sigplan,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%% For single-blind review submission, w/ CCS and ACM Reference
-%\documentclass[sigplan,review]{acmart}\settopmatter{printfolios=true}
-%% For final camera-ready submission, w/ required CCS and ACM Reference
-%\documentclass[sigplan]{acmart}\settopmatter{}
+\documentclass[sigplan,screen]{acmart}
 
@@
 \affiliation{
-  \department{Department of Programming Languages and Compilers}
+  %% \department{Department of Programming Languages and Compilers}
   \institution{E{\"o}tv{\"o}s Lor{\'a}nd University}
   \city{Budapest}
-  \country{Hungary}                   %% \country is recommended
+  \country{Hungary}
 }
-\email{kovacsandras@inf.elte.hu}         %% \email is recommended
+\email{kovacsandras@inf.elte.hu}
+
+\author{Ambrus Kaposi}
+\orcid{0000-0001-9897-8936}
+\affiliation{
+  %% \department{Department of Programming Languages and Compilers}
+  \institution{E{\"o}tv{\"o}s Lor{\'a}nd University}
+  \city{Budapest}
+  \country{Hungary}
+}
+\email{akaposi@inf.elte.hu}
 
@@
 QIITs as previously described in the literature \cite{kaposi2019constructing,
-  qiits, dijkstra2017quotient}.
+  qiits, dijkstra2017quotient}, with the following features.
 
@@
 \end{enumerate}
-\end{definition}
-
-\begin{definition} We abbreviate the additional structure on CwFs consisting of $\Sigma$, $\Id$ and
+We abbreviate the additional structure on CwFs consisting of $\Sigma$, $\Id$ and
 $\Kfam$ as \emph{fl-structure}.
@@
 
-In the categorical semantics for a concrete signature, this informally means
-that an algebra supports induction if for any dependent algebra over it (which
-is a bundle of induction motives and methods), there is a dependent morphism
-into it (which is a bundle of eliminator functions and their $\beta$-rules).
+For an example, if we interpret $\mathsf{\NatSig}$ in the $\bM$ model, we get an
+flCwF of natural number algebras, where $\Con$ is the type of algebras and
+$\Sub\,\Gamma\,\Delta$ is the type of homomorphisms between $\Gamma$ and
+$\Delta$ algebras. $\Ty$ is the type of displayed algebras, and $\Tm$ is the
+type of their sections:
+\begin{alignat*}{3}
+  & \Ty\,(N,\,z,\,s) \equiv(N^D : N \to \Set)\\ & \hspace{2em}\times (N^D\,z)
+  \times ((n : N)\to N^D\,n \to N^D(s\,n))\\ &
+  \Tm\,(N,\,z,\,s)\,(N^D,\,z^D,\,s^D) \equiv (N^S : (n : N)\to
+  N^D\,n)\\ & \hspace{2em}\times (N^S\,z = z^D)\times ((n : N) \to N^S\,(s\,n) =
+  s^D\,n\,(N^S\,n))
+\end{alignat*}
+Thus, for natural number algebras, $\Inductive$ is exactly the predicate which
+holds when an algebra supports induction.
+
+%% $\Inductive$ can be viewed as a
+%% dependently typed analogue of initiality: an algebra $\Gamma$ is inductive if
+%% for every dependent algebra $A : \Ty\,\Gamma$ over it there is a dependent
+%% morphism from $\Gamma$ to $A$, given as an element of $\Tm\,\Gamma\,A$.
+
+%% For some $(N,\,z,\,s)$ algebra,
+%% $\Ty\,(N,\,z,\,s)$ is defined as $(N^P : N \to \Set)\times N^P\,z \times ((n :
+%% N)\to N^P\,n \to N^P(s\,n))$, the type of displayed algebras, and
+%% $\Tm\,(N,\,z,\,s)\,(N^P,\,z^P,\,s^P)$ is $(N^S : (n : N)\to N^P\,n) \times (N^S\,z = z^P)
+%% \times ((n : N) \to N^S\,(s\,n) = s^P\,(N^S\,n))$.
+
+%% Here, $\Con_{\bGamma}$ is the type of algebras for some signature, e.g.\ the
+%% type of natural number algebras in the case of
+%% $\mathsf{NatSig}$. $\Ty_{\bGamma}\,\Gamma$ is a displayed algebra over a
+%% concrete algebra
+
+%% In the categorical semantics for a concrete signature, this informally means
+%% that an algebra supports induction if for any dependent algebra over it (which
+%% is a bundle of induction motives and methods), there is a dependent morphism
+%% into it (which is a bundle of eliminator functions and their $\beta$-rules).
 
@@
 $\blank^{\D}$ and $\blank^{\S}$ interpretations in
-\cite{kaposi2019constructing}.
+\cite{kaposi2019constructing}, where they are discussed at more length.
 

### 2020-04-12 draft/paper.tex
acks

@@
 \begin{acks}
-This work was supported by the European Union, co-financed by the
-European Social Fund (EFOP-3.6.3-VEKOP-16-2017-00002) and COST Action
-EUTypes CA15123.
+The first author was supported by the European Union, co-financed by the
+European Social Fund (EFOP-3.6.3-VEKOP-16-2017-00002). The second author was
+supported by the National Research, Development and Innovation Fund of Hungary,
+financed under the Thematic Excellence Programme funding scheme, Project
+no. ED18-1-2019-0030 (Application-specific highly reliable IT solutions), by the
+New National Excellence Program of the Ministry for Innovation and Technology,
+Project no. \'UNKP-19-4-ELTE-874, and by the Bolyai Fellowship of the Hungarian
+Academy of Sciences, Project no. BO/00659/19/3.
 \end{acks}

### 2020-05-01 draft/paper.tex
bug
finalization

@@
 
-%% Conference information
-%% Supplied to authors by publisher for camera-ready submission;
-%% use defaults for review submission.
-\acmConference[PL'18]{ACM SIGPLAN Conference on Programming Languages}{January 01--03, 2018}{New York, NY, USA}
-\acmYear{2018}
-\acmISBN{} % \acmISBN{978-x-xxxx-xxxx-x/YY/MM}
-\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
+\copyrightyear{2020}
+\acmYear{2020}
+\setcopyright{acmlicensed}
+\acmConference[LICS '20]{Proceedings of the 35th Annual ACM/IEEE Symposium on Logic
+  in Computer Science (LICS)}{July 8--11, 2020}{Saarbr\"ucken, Germany}
+\acmBooktitle{Proceedings of the 35th Annual ACM/IEEE Symposium on Logic in
+  Computer Science (LICS '20), July 8--11, 2020, Saarbr\"ucken, Germany}
+\acmPrice{15.00}
+\acmDOI{10.1145/3373718.3394770}
+\acmISBN{978-1-4503-7104-9/20/07}
 \startPage{1}
-
-%% Copyright information
-%% Supplied to authors (based on authors' rights management selection;
-%% see authors.acm.org) by publisher for camera-ready submission;
-%% use 'none' for review submission.
-\setcopyright{none}
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-%\setcopyright{rightsretained}
-%\copyrightyear{2018}           %% If different from \acmYear
 
@@
 
-%% ------------------------------------------------------------
+%% --------------------------------------------------------------------------------
+
 \usepackage[utf8]{inputenc}
-\usepackage{todonotes}
-\presetkeys{todonotes}{inline}{}
 \usepackage{amsmath}
@@
 \usepackage{stmaryrd}
-\usepackage{tikz-cd}
-
-\newtheorem{notation}{Notation}
-
-\input{abbrevs.tex}
+
+%% --------------------------------------------------------------------------------
+
+\newcommand{\refl}{\mathsf{refl}}
+\newcommand{\id}{\mathsf{id}}
+\newcommand{\Con}{\mathsf{Con}}
+\newcommand{\Sub}{\mathsf{Sub}}
+\newcommand{\Tm}{\mathsf{Tm}}
+\newcommand{\Ty}{\mathsf{Ty}}
+\renewcommand{\U}{\mathsf{U}}
+\newcommand{\El}{\mathsf{El}}
+\newcommand{\Id}{\mathsf{Id}}
+\newcommand{\proj}{\mathsf{proj}}
+\renewcommand{\tt}{\mathsf{tt}}
+\newcommand{\blank}{\mathord{\hspace{1pt}\text{--}\hspace{1pt}}}
+\newcommand{\ra}{\rightarrow}
+\newcommand{\Set}{\mathsf{Set}}
 \newcommand{\Lift}{\Uparrow}
@@
 \newcommand{\mi}[1]{\mathit{#1}}
-
 \newcommand{\Sg}{\Sigma}
@@
 \newcommand{\q}{\mathsf{q}}
+\newcommand{\K}{\mathsf{K}}
 \newcommand{\lamK}{\mathsf{lam}^{\K}}
 \newcommand{\appK}{\mathsf{app}^{\K}}
-
+\newcommand{\A}{\mathsf{A}}
+\newcommand{\D}{\mathsf{D}}
+\renewcommand{\S}{\mathsf{S}}
 \newcommand{\arri}{\Rightarrow}
@@
 \newcommand{\SynSig}{\mathsf{SynSig}}
-
 \newcommand{\bCon}{\boldsymbol{\Con}}
@@
 \newcommand{\bI}{\boldsymbol{\mathsf{I}}}
-
 \newcommand{\ul}[1]{\underline{#1}}
@@
 \newcommand{\tos}{\mathsf{tos}}
-
 \newcommand{\coe}{\mathsf{coe}}
@@
 
-%% ------------------------------------------------------------
+%% --------------------------------------------------------------------------------
 
@@
 
-%% Whereas the initial category can be given by first defining the type of objects
-%% (the empty type) and then defining morphisms separately, for an initial model of
-%% a type theory given as a CwF, one has to define $\mi{Ob}$, $\mi{Mor}$ and two
-%% extra sorts $\mi{Ty}$ and $\mi{Tm}$ \emph{inductive-inductively}, at the same
-%% time. This is illustrated by the following excerpt of the signature for CwFs
-%% extending the previous signature for categories.
-%% \begin{alignat*}{3}
-%%   & \mi{Ty}        && : \mi{Ob}\ra \Set \\
-%%   & \mi{U}         && : \mi{Ty\,i} \\
-%%   & \mi{empty}     && : \mi{Ob} \\
-%%   & \mi{ext}       && : (I:\mi{Ob})\ra \mi{Ty}\,I\ra \mi{Ob} \\
-%%   & \blank[\blank] && : \mi{Ty}\,J\ra \mi{Mor}\,I\,J\ra \mi{Ty}\,I
-%% \end{alignat*}
-%% While signatures for type theories usually only contain finitary constructors,
-%% some QIITs have infinitary constructors, such as the following signature for
-%% countably branching trees where the order of the subtrees does not matter,
-%% i.e.\ permuting subtrees yields an equal tree.
-%% \begin{alignat*}{3}
-%%   & \mi{Tree} && : \U \\
-%%   & \mi{leaf}  && : \mi{Tree} \\
-%%   & \mi{node} && : (\N\ra \mi{Tree})\ra \mi{Tree} \\
-%%   & \mi{perm} && : (g :\N\ra \mi{Tree})(f : \N\ra\N)\ra \mathsf{isIso}\,f\ra \\
-%%   & && \hspace{0.8em} \mi{node}\,g = \mi{node}\,(g\circ f)
-%% \end{alignat*}
-%% Additional examples for infinitary QIITs are real, surreal numbers
-%% \cite{HoTTbook}, ordinal numbers \cite{lumsdaineShulman} and the partiality
-%% monad \cite{partiality}.
-
-%% \subsection{Contributions}
-%% The current paper extends the syntax and semantics of QIITs as previously described
-%% in the literature \cite{kaposi2019constructing, qiits, dijkstra2017quotient}. We add the following features:
-
 \begin{enumerate}
@@
   \item
-  \textbf{Recursive equations}, i.e. equations appearing as assumptions
-  of constructors. These have occurred previously in syntaxes of cubical type
-  theories, as boundary conditions \cite{cchm, angiuli2016computational, angiuli2018cartesian}.
+  \textbf{Recursive equations}, i.e. equations appearing as assumptions of
+  constructors. These have occurred previously in syntaxes of cubical type
+  theories, as boundary conditions \cite{cchm, angiuli2016computational,
+    angiuli2018cartesian}.
 \end{enumerate}
@@
 
-Additionally, we have an internal type for $\blank\leq\blank$, analogously to
-$\blank=\blank$ which is an internalization of definitional equality.
+Additionally, we have an internal $\Subtype$ type, which internalizes subtyping,
+analogously to how $\blank=\blank$ internalizes definitional equality. Hence, we
+have analogous reflection and uniqueness rules.
 \begin{mathpar}
@@
              {\Gamma \vdash A \leq B}
+
+  \inferrule*{\Gamma \vdash t : \Subtype\,A\,B \\ \Gamma \vdash u : \Subtype\,A\,B}
+             {\Gamma \vdash t \equiv u}
 \end{mathpar}
@@
 infinitary QIIT itself, and we would like to define ToS and a notion of
-signature without referring to QIITs, only using features present in cETT. As a
-first step, we define a notion of model.
+signature without referring to QIITs, only using features present in cETT.
+
+In previous works by Cartmell \cite{gat} and Sterling
+\cite{sterling2019algebraic}, signatures for generalized algebraic theories are
+defined using raw syntax together with well-formedness relations. In this way,
+signatures can be specified without already assuming the existence of GATs or
+QIITs. However, raw syntax is notoriously difficult to work with, and we prefer
+to avoid it altogether.
+
+At this point, we do not actually need \emph{syntactic} signatures, which make
+it possible to do induction on signatures. We only need a way to write down
+well-formed signatures, and interpret them in arbitrary models of ToS. For this,
+a weak Church-like encoding suffices, where a signature is given as a typing
+context in an arbitrary model of ToS. For this, we first need to specify the
+notion of ToS models. However, this is the \emph{only} piece of information
+about ToS which we need to manually provide. Other concepts such as
+homomorphisms of ToS models and ToS-induction, will be derived from the
+semantics of signatures and self-description in Section \ref{sec:tossig}.
 
@@
 
-%% \begin{example}{Sort equations}. An interesting example is the term lifting
-%% rule for cETT (omitting everything except term lifting):
-%% \begin{alignat*}{3}
-%%   & cETTSig := \\
-%%   & ...\,\ext\,Tm{\Lift} : \Pii\,(\Gamma : Con)\,(\Pii\,(A : Ty{\appitt}\Gamma)((\IdU\,(Tm{\appitt}\Gamma{\appitt}A)\,(Tm{\appitt}\Gamma{\appitt}(\Lift_j{\appitt}A))))
-%% \end{alignat*}
-%% \end{example}
-
 \begin{example}{Recursive equations}. Again, the universe is closed under $\Id$, which allows us to write equations
@@
 More interesting (and complicated) examples for recursive equations are boundary
-conditions in various cubical type theories \cite{cchm, angiuli2016computational, angiuli2018cartesian}. Note that our $\Id$
-allows iterated equations as well, but these are all trivial in the semantics,
-where we assume UIP.
+conditions in various cubical type theories \cite{cchm,
+  angiuli2016computational, angiuli2018cartesian}. Note that our $\Id$ allows
+iterated equations as well, but these are all trivial in the semantics, where we
+assume UIP.
 \end{example}
@@
 perspective, but we are concerned with eventually implementing QIITs in proof
-assistants, so we prefer if our semantics computes strictly.
+assistants, so we prefer if our semantics computes strictly. This was
+demonstrated previously in \cite{hiit}, where we provided a program which
+computed types of induction principles from signatures of higher
+inductive-inductive types, and we believe that the same could be achieved for
+the signatures and semantics described in this paper.
 
@@
 We define $\bCon : \Set_{\max(i+1,j)+1}$ as $\flCwF_{\max(i+1,j)}$.
-%% We define
-%% $\flCwF$ as follows.
 
@@
 
-%% We abbreviate the additional structure on CwFs consisting of $\Sigma$, $\Id$ and
-%% $\Kfam$ as fl-structure. We use $\Ty$ in an flCwF to represent ``dependent
-%% objects'', and $\Tm$ for ``dependent morphisms''. We use these to define
-%% induction.
-
 \begin{definition}[Notion of induction in an flCwF]\label{def:induction}
@@
 
-%% For some $(N,\,z,\,s)$ algebra,
-%% $\Ty\,(N,\,z,\,s)$ is defined as $(N^P : N \to \Set)\times N^P\,z \times ((n :
-%% N)\to N^P\,n \to N^P(s\,n))$, the type of displayed algebras, and
-%% $\Tm\,(N,\,z,\,s)\,(N^P,\,z^P,\,s^P)$ is $(N^S : (n : N)\to N^P\,n) \times (N^S\,z = z^P)
-%% \times ((n : N) \to N^S\,(s\,n) = s^P\,(N^S\,n))$.
-
-%% Here, $\Con_{\bGamma}$ is the type of algebras for some signature, e.g.\ the
-%% type of natural number algebras in the case of
-%% $\mathsf{NatSig}$. $\Ty_{\bGamma}\,\Gamma$ is a displayed algebra over a
-%% concrete algebra
-
-%% In the categorical semantics for a concrete signature, this informally means
-%% that an algebra supports induction if for any dependent algebra over it (which
-%% is a bundle of induction motives and methods), there is a dependent morphism
-%% into it (which is a bundle of eliminator functions and their $\beta$-rules).
-
-\begin{theorem}[Equivalence of initiality and induction, c.f. \cite{kaposi2019constructing}]\label{thm:initialind}
-An object $\Gamma : \Con_{\bGamma}$ supports induction if and only if it is
+\begin{theorem}[Equivalence of initiality and induction, c.f. \cite{kaposi2019constructing}]
+\label{thm:initialind} An object $\Gamma : \Con_{\bGamma}$ supports induction if and only if it is
 initial. Moreover, induction and initiality are both proof-irrelevant
@@
 and the equalizer of $\sigma$ and $\delta$ as $\Gamma\ext
-\Id\,(\K\,\Delta)\,(\lamK\,\sigma)\,(\lamK\,\delta)$.
-
-In order to talk about weak structure-preservation in the interpretation of substitutions, we need to specify isomorphisms for contexts and types.
-
-\begin{definition} A \emph{context isomorphism} is an invertible morphism $\sigma : \Sub\,\Gamma\,\Delta$.
-We note the inverse as $\sigma^{-1}$.
-We also use the notation $\sigma : \Gamma \simeq \Delta$.
-\end{definition}
-
-\begin{definition}[Type categories, c.f.\ \cite{clairambault2014biequivalence}]\label{def:type_categories} For each $\Gamma : \Con$, there is a category
+\Id\,(\K\,\Delta)\,(\lamK\,\sigma)\,(\lamK\,\delta)$. While showing equivalence
+of initiality and induction does not need all flCwF components (e.g.\ $\Sigma$
+is not needed), we build the full flCwF semantics in order to connect to
+Clairambault's and Dybjer's results.
+
+In order to talk about weak structure-preservation in the interpretation of
+substitutions, we need to specify isomorphisms for contexts and types.
+
+\begin{definition}
+A \emph{context isomorphism} is an invertible morphism $\sigma :
+\Sub\,\Gamma\,\Delta$. We note the inverse as $\sigma^{-1}$. We also use the
+notation $\sigma : \Gamma \simeq \Delta$.
+\end{definition}
+
+\begin{definition}[Type categories, c.f.\ \cite{clairambault2014biequivalence}]
+\label{def:type_categories} For each $\Gamma : \Con$, there is a category
 whose objects are types $A : \Ty\,\Gamma$, and morphisms from $A$ to $B$ are
-terms $t : \Tm\,(\Gamma\,\ext\,A)\,(B[\p])$. Identity morphisms are given by $\q :
-\Tm\,(\Gamma\,\ext\,A)\,(A[\p])$, and composition $t \circ u$ by $t[\p, u]$. The
-assignment of type categories to contexts extends to a split indexed category. For
-each $\sigma : \Sub\,\Gamma\,\Delta$, there is a functor from $\Ty\,\Delta$ to
-$\Ty\,\Gamma$, which sends $A$ to $A[\sigma]$ and $t :
+terms $t : \Tm\,(\Gamma\,\ext\,A)\,(B[\p])$. Identity morphisms are given by $\q
+: \Tm\,(\Gamma\,\ext\,A)\,(A[\p])$, and composition $t \circ u$ by $t[\p,
+  u]$. The assignment of type categories to contexts extends to a split indexed
+category. For each $\sigma : \Sub\,\Gamma\,\Delta$, there is a functor from
+$\Ty\,\Delta$ to $\Ty\,\Gamma$, which sends $A$ to $A[\sigma]$ and $t :
 \Tm\,(\Gamma\,\ext\,A)\,(B[\p])$ to $t[\sigma\circ \p, \q]$.
@@
 
-\begin{theorem}\label{thm:flpres} Every $\boldsymbol{\sigma : \Sub\,\Gamma\,\Delta}$ preserves fl-structure up to type isomorphism. That is, we have
+\begin{theorem}\label{thm:flpres}
+Every $\boldsymbol{\sigma : \Sub\,\Gamma\,\Delta}$ preserves fl-structure up to
+type isomorphism. That is, we have
 \begin{alignat*}{3}
@@
 
-It is easy to verify the left and right identity laws and associativity for $\boldsymbol{\blank\circ\blank}$.
-
-\begin{lemma}\label{lem:idcomppres} The derived preservation isomorphisms for the fl-structure can be decomposed analogously; all derived isomorphisms in $\bid$ are identities, and we have
+It is easy to verify the left and right identity laws and associativity for
+$\boldsymbol{\blank\circ\blank}$.
+
+\begin{lemma}\label{lem:idcomppres}
+The derived preservation isomorphisms for the fl-structure can be decomposed
+analogously; all derived isomorphisms in $\bid$ are identities, and we have
 \begin{alignat*}{3}
@@
 \end{lemma}
-\begin{proof} In the case of $\Id$, the equations hold immediately by UIP. For $\Sigma$ and $\K$, we prove by flCwF computation and straightforward unfolding of definitions.
+\begin{proof}
+In the case of $\Id$, the equations hold immediately by UIP. For $\Sigma$ and
+$\K$, we prove by flCwF computation and straightforward unfolding of
+definitions.
 \end{proof}
@@
 
-\begin{definition}[Displayed flCwF] The type of displayed flCwFs at level $i$ is given as the logical predicate interpretation (see e.g.\ \cite{bernardy12parametricity} or \cite{hiit}) of $\flCwF_i$.
-For each flCwF component in $\bGamma$, there is a component in a displayed flCwF which ``lies over'' it.
+\begin{definition}[Displayed flCwF]
+The type of displayed flCwFs at level $i$ is given as the logical predicate
+interpretation (see e.g.\ \cite{bernardy12parametricity} or \cite{hiit}) of
+$\flCwF_i$. For each flCwF component in $\bGamma$, there is a component in a
+displayed flCwF which ``lies over'' it.
 
@@
 
-Concretely, a displayed flCwF $\bA$ over $\bGamma$ has the following underlying sets, which we call displayed contexts, substitutions, types and terms respectively.
+Concretely, a displayed flCwF $\bA$ over $\bGamma$ has the following underlying
+sets, which we call displayed contexts, substitutions, types and terms
+respectively.
 \begin{alignat*}{3}
@@
 In the following we will often omit $_{\bGamma}$ and $_{\bA}$ subscripts on
-components; for example, in the type $\Con_{\bA}\,\emptycon$, the $\emptycon$ is clearly
-a base component in $\bGamma$.
-
-We also need displayed counterparts to the previously defined derived notions on flCwFs;
-these are again given as logical predicate interpretations of the non-displayed definitions.
-
-\begin{definition}[Displayed type categories] For each $\Gamma : \Con_{\bA}\,\ulGamma$, there
-is a displayed category over the type category $\Ty_{\bGamma}\,\ulGamma$, whose objects
-over $\ulA : \Ty_{\bGamma}\,\ulGamma$ are elements of $\Ty_{\bA}\,\Gamma\,\ulA$, and displayed morphisms over $\ult : \Tm_{\bGamma}\,(\ulGamma \ext \ulA)\,(\ulB[\p])$ are elements of $\Tm_{\bA}\,(\Gamma \ext A)\,(B[\p])\,\ult$.
-The identity morphism is given by $\q_{\bA}$, and the composition of $t$ and $u$ is $t[\p_{\bA},u]$.
-Analogously to Definition \ref{def:type_categories}, this extends to a displayed split indexed category.
-\end{definition}
-
-\begin{definition}[Displayed isomorphisms] A \emph{displayed context isomorphism} over $\ulsigma : \ulGamma \simeq \ulDelta$, notated $\sigma : \Gamma \simeq_{\ulsigma} \Delta$, is an invertible displayed morphism $\sigma : \Sub_{\bA}\,\Gamma\,\Delta\,\ulsigma$, with inverse $\sigma^{-1} : \Sub_{\bA}\,\Delta\,\Gamma\,\ulsigma^{-1}$.
-A \emph{displayed type isomorphism} over $\ult : \ulA \simeq \ulB$, notated $t : A \simeq_{\ult} B$, is an isomorphism in a displayed type category.
-\end{definition}
-
-\begin{definition} A \emph{vertical morphism} lies over an identity morphism. We use this definition for context morphisms (substitutions) and type morphisms as well.
+components; for example, in the type $\Con_{\bA}\,\emptycon$, the $\emptycon$ is
+clearly a base component in $\bGamma$.
+
+We also need displayed counterparts to the previously defined derived notions on
+flCwFs; these are again given as logical predicate interpretations of the
+non-displayed definitions.
+
+\begin{definition}[Displayed type categories]
+For each $\Gamma : \Con_{\bA}\,\ulGamma$, there is a displayed category over the
+type category $\Ty_{\bGamma}\,\ulGamma$, whose objects over $\ulA :
+\Ty_{\bGamma}\,\ulGamma$ are elements of $\Ty_{\bA}\,\Gamma\,\ulA$, and
+displayed morphisms over $\ult : \Tm_{\bGamma}\,(\ulGamma \ext
+\ulA)\,(\ulB[\p])$ are elements of $\Tm_{\bA}\,(\Gamma \ext
+A)\,(B[\p])\,\ult$. The identity morphism is given by $\q_{\bA}$, and the
+composition of $t$ and $u$ is $t[\p_{\bA},u]$. Analogously to Definition
+\ref{def:type_categories}, this extends to a displayed split indexed category.
+\end{definition}
+
+\begin{definition}[Displayed isomorphisms]
+A \emph{displayed context isomorphism} over $\ulsigma : \ulGamma \simeq
+\ulDelta$, notated $\sigma : \Gamma \simeq_{\ulsigma} \Delta$, is an invertible
+displayed morphism $\sigma : \Sub_{\bA}\,\Gamma\,\Delta\,\ulsigma$, with inverse
+$\sigma^{-1} : \Sub_{\bA}\,\Delta\,\Gamma\,\ulsigma^{-1}$. A \emph{displayed
+  type isomorphism} over $\ult : \ulA \simeq \ulB$, notated $t : A \simeq_{\ult}
+B$, is an isomorphism in a displayed type category.
+\end{definition}
+
+\begin{definition}
+A \emph{vertical morphism} lies over an identity morphism. We use this
+definition for context morphisms (substitutions) and type morphisms as well.
 \end{definition}
@@
 In contrast to \cite{kaposi2019constructing}, it is not sufficient to model
-types as displayed flCwFs. In ibid.\ substitutions are modeled as strict
-morphisms, which we cannot do because our infinitary function types and identity
-types cannot be interpreted strictly, as we will see in Section
-\ref{sec:infinitaryfunction}. Strict preservation is expressed with metatheoretic equality, which
-is respected by all cETT constructions, but isomorphisms are not necessarily
-respected. Hence, we need to add additional structure to displayed flCwFs which
-expresses preservation of base isomorphisms.
+types as displayed flCwFs. In ibid.\ the universe $\U$ in ToS was empty, and all
+substitutions were ``neutral'', i.e.\ semantic subsitutions were functors which
+may permute, duplicate or forget components of algebras, or freely reinterpret
+components, and it is easy to see that all such functors strictly preserve
+limits. In contrast, the current $\U$ is not empty: it is closed under identity
+and infinitary function types. Hence, substitutions and terms are not neutral
+anymore, as they can contain canonical type codes in $\U$. Semantically, these
+canonical type codes do not merely reshuffle structure, hence they preserve
+limits only weakly. We will return to this in Section
+\ref{sec:infinitaryfunction}. We are forced to use a weaker semantics where
+fl-structure is not preserved strictly, and we also need to add additional
+structure to displayed flCwFs which expresses preservation of base isomorphisms.
 
@@
 \end{theorem}
-\begin{proof} The construction of isomorphisms is the same as in Theorem \ref{thm:flpres}. Indeed, every construction there has a displayed counterpart which we can use here.
+\begin{proof}
+The construction of isomorphisms is the same as in Theorem
+\ref{thm:flpres}. Indeed, every construction there has a displayed counterpart
+which we can use here.
 \end{proof}
@@
 taking a first projection for each component. $\boldsymbol{\q : \Tm\,(\Gamma\ext
-  A)\,(A[\p])}$ is likewise a strict flCwF section given by second projections.
-Substitution extension $\boldsymbol{\sigma,\,t}$ is given by pointwise combining $\bsigma$ and $\bt$ with metatheoretic $\Sigma$ pairing, e.g.\ $\Con_{\boldsymbol{(\sigma,t)}}\,\ulGamma := (\bsigma\,\ulGamma,\,\bt\,\ulGamma)$.
+  A)\,(A[\p])}$ is likewise a strict flCwF section given by second
+projections. Substitution extension $\boldsymbol{\sigma,\,t}$ is given by
+pointwise combining $\bsigma$ and $\bt$ with metatheoretic $\Sigma$ pairing,
+e.g.\ $\Con_{\boldsymbol{(\sigma,t)}}\,\ulGamma :=
+(\bsigma\,\ulGamma,\,\bt\,\ulGamma)$.
 
@@
 
-\begin{definition} For a level $i$, we write $\bSet_i$ for the flCwF of sets where $\Con_{\bSet_i} := \Set_i$ and $\Sub_{\bSet_i}\,\Gamma\,\Delta := \Gamma \ra \Delta$.
+\begin{definition}
+For a level $i$, we write $\bSet_i$ for the flCwF of sets where $\Con_{\bSet_i}
+:= \Set_i$ and $\Sub_{\bSet_i}\,\Gamma\,\Delta := \Gamma \ra \Delta$.
 \end{definition}
@@
 We likewise use preservation isomorphisms to define $\K$, $\Id$ and $\Sigma$.
-Context coercion is $\coe\,\ulsigma\,\Gamma := \ba\,\ulsigma\,\Gamma$. Type coercion, for $A : \ba\,\ulA\,\Gamma$ is given as $\coe\,\ult\,A := \ba\,\ult\,(\ba_{\ext}^{-1}\,(\Gamma,\,A))$.
+Context coercion is $\coe\,\ulsigma\,\Gamma := \ba\,\ulsigma\,\Gamma$. Type
+coercion, for $A : \ba\,\ulA\,\Gamma$ is given as $\coe\,\ult\,A :=
+\ba\,\ult\,(\ba_{\ext}^{-1}\,(\Gamma,\,A))$.
 
@@
 branching with $A \ra \Ty\,\Gamma$ and $A \ra \Tm\,\Gamma\,\U$ respectively. We
-present an excerpt from $\ToSSig_i$ below.
+present an excerpt from $\ToSSig_j$ below.
 \begin{alignat*}{3}
@@
 
-\begin{lemma}[Cumulativity of $\ToS$]\label{lemma:cumulativity} If $i \leq i'$, then $\ToS_{i,j} \leq \ToS_{i', j}$.
-  This follows from the definition of ToS and the subtyping rules in Section
+\begin{lemma}[Cumulativity of $\ToS$]\label{lemma:cumulativity}
+If $i \leq i'$, then $\ToS_{i,j} \leq \ToS_{i', j}$.  This follows from the
+definition of ToS and the subtyping rules in Section
 \ref{sec:cumulativity}. \qed
@@
 
-\begin{example} We have $\syn_0 : \ToS_{1, 0}$, which is the syntax of closed QIIT signatures.
-We want to define a function $\mathsf{length} : \Con_{\syn_0} \ra \mathbb{N}$ by induction, which returns the length of a syntactic context as a metatheoretic natural number.
-To this end, we define a displayed ToS over $\syn_0$, where $\Con$ is defined as constantly $\mathbb{N}$, every other sort is defined as constantly $\top$, $\emptycon$ is defined as $0$ and $\Gamma\ext A$ is defined as $\Gamma + 1$.
-By the induction assumption, we get a ToS-section from $\syn_0$ to the displayed model, whose action on contexts is exactly the $\mathsf{length}$ function.
-Note that the induction assumption requires that the displayed model is at least at level 1, but this is not problematic because by cumulativity $\mathbb{N} : \Set_1$.
+\begin{example}
+We have $\syn_0 : \ToS_{1, 0}$, which is the syntax of closed QIIT
+signatures. We want to define a function $\mathsf{length} : \Con_{\syn_0} \ra
+\mathbb{N}$ by induction, which returns the length of a syntactic context as a
+metatheoretic natural number. To this end, we define a displayed ToS over
+$\syn_0$, where $\Con$ is defined as constantly $\mathbb{N}$, every other sort
+is defined as constantly $\top$, $\emptycon$ is defined as $0$ and $\Gamma\ext
+A$ is defined as $\Gamma + 1$. By the induction assumption, we get a ToS-section
+from $\syn_0$ to the displayed model, whose action on contexts is exactly the
+$\mathsf{length}$ function. Note that the induction assumption requires that the
+displayed model is at least at level 1, but this is not problematic because by
+cumulativity $\mathbb{N} : \Set_1$.
 \end{example}
@@
 
-\begin{definition} We define $\SynSig_j : \Set_{j+1}$, the type of \emph{syntactic signatures} at $j$, as $\Con_{\syn_j}$.
+\begin{definition}
+We define $\SynSig_j : \Set_{j+1}$, the type of \emph{syntactic signatures} at
+$j$, as $\Con_{\syn_j}$.
 \end{definition}
@@
 $\Gamma : \Con_{\syn_j}$ in an arbitrary ToS model. This is merely a logical
-equivalence, external to cETT (because of unvierse polymorphism), and not an
+equivalence, external to cETT (because of universe polymorphism), and not an
 isomorphism.
@@
 
-\begin{definition}[Cumulativity model] We assume $j$, $k$ and $l$ levels such that $j+1 \leq k$, $j+1 \leq l$ and $k \leq l$. We define a displayed model over $\syn_j : \ToS_{j+1, j}$ lifted to
+\begin{definition}[Cumulativity model]
+We assume $j$, $k$ and $l$ levels such that $j+1 \leq k$, $j+1 \leq l$ and $k
+\leq l$. We define a displayed model over $\syn_j : \ToS_{j+1, j}$ lifted to
 $\ToS_{l, j}$. In the following, we notate the level of algebras computed by
@@
 
-\begin{definition} We assume $j$ and $k$ such that $j + 1 \leq k$, and we also assume
-$\ulOmega : \SynSig_j$ and $\gamma :
-\llb\ulOmega\rrb_{\bD_k}\,(\llb\ulOmega\rrb_{\bT_{\ulOmega}}\,\id)$.  Hence, $\gamma$ is
-a displayed $\ulOmega$-algebra over the term algebra, at level $k$. We are using
-the cumulativity of $\ulOmega$ here to lift the term algebra appropriately.
-We aim to show that $\gamma$ has a section. We define a displayed model over $\syn_j$ lifted
-to $\ToS_{k, j}$, which we name $\bI_{\ulOmega}$. The underlying sets are:
+\begin{definition}
+We assume $j$ and $k$ such that $j + 1 \leq k$, and we also assume $\ulOmega :
+\SynSig_j$ and $\gamma :
+\llb\ulOmega\rrb_{\bD_k}\,(\llb\ulOmega\rrb_{\bT_{\ulOmega}}\,\id)$.  Hence,
+$\gamma$ is a displayed $\ulOmega$-algebra over the term algebra, at level
+$k$. We are using the cumulativity of $\ulOmega$ here to lift the term algebra
+appropriately.  We aim to show that $\gamma$ has a section. We define a
+displayed model over $\syn_j$ lifted to $\ToS_{k, j}$, which we name
+$\bI_{\ulOmega}$. The underlying sets are:
 \begin{alignat*}{3}
@@
 Here, there is no essential change compared to \cite{kaposi2019constructing},
-and we follow ibid.\ in the definition of $\bI_{\ulOmega}$. The reason is that although we
-have weakened strict algebra equality to isomorphism, in the current
+and we follow ibid.\ in the definition of $\bI_{\ulOmega}$. The reason is that
+although we have weakened strict algebra equality to isomorphism, in the current
 construction we only have to show equalities of substitutions and terms, which
@@
 
-\begin{theorem}[Initiality of term algebras] For each $j$ and $k$ such that $j + 1 \leq k$, and $\ulOmega : \SynSig_j$, the term algebra given by $\llb\ulOmega\rrb_{\bT_{\ulOmega}}\,\id$ is initial at level $k$.
+\begin{theorem}[Initiality of term algebras]
+For each $j$ and $k$ such that $j + 1 \leq k$, and $\ulOmega : \SynSig_j$, the
+term algebra given by $\llb\ulOmega\rrb_{\bT_{\ulOmega}}\,\id$ is initial at
+level $k$.
 \end{theorem}
@@
 Cartmell \cite{gat} defines generalized algebraic theories (GATs) using
-type-theoretic syntax. Compared to our QII signatures, his signatures support
-sort equations but do not cover infinitary constructors or recursive
-equations. A way to encode sort equations in our signatures is using
+type-theoretic syntax. Compared to our QII signatures, he supports infinite
+signatures and sort equations but does not cover infinitary constructors or
+recursive equations. A way to encode sort equations in our system is using
 isomorphisms instead of equalities. In contrast to our algebraic definition,
@@
 strict positivity, hence describes a larger class of QII signatures. They show
-equivalence of initiality and induction, but do not construct initial algebras,
-nor would that be possible because of the lack of a positivity restriction.
+equivalence of initiality and induction, but the lack of a positivity
+restriction prevents construction of initial algebras.
 
@@
 an ad-hoc QIIT.
-
-%% Streicher's initiality proof \cite{streicher2012semantics} is a direct
-%% construction of a concrete QIIT: that of the syntax of a type theory. He uses
-%% preterms, typing relations and quotients to construct the initial model of type
-%% theory in a set-theoretic metalanguage. A recent formalization of the same
-%% result for a larger type theory was done by Brunerie et al.\ \cite{brunerie} in
-%% Agda with UIP, function extensionality and quotients. Lafont et
-%% al.\ \cite{induction-is-enough} construct all finitary inductive-inductive types
-%% in Agda using only UIP and function extensionality. Examples of
-%% inductive-inductive types are constructed without using UIP by Hugunin
-%% \cite{jasper}.
 

### 2020-05-03 draft/paper.tex
fix

@@
   &           && \ext (Ty  : Con\arri\U)\\
-  &           && \ext (Tm  : \Pii(\Gamma : Con)(\Ty\,\Gamma \arri \U))\\
+  &           && \ext (Tm  : \Pii(\Gamma : Con)(\Ty\appitt\Gamma \arri \U))\\
   &           && ...\\
   &           && \ext (\Piinf : \Pii(\Gamma : Con)\\
-  &           && \hspace{4.2em}(\Pie \Set_j (\lambda\,A.\,(A \arrinf \Ty\,\Gamma)\arri\El(\Ty\,\Gamma))))\\
+  &           && \hspace{4.2em}(\Pie \Set_j (\lambda\,A.\,(A \arrinf \Ty\appitt\Gamma)\arri\El(\Ty\appitt\Gamma))))\\
   &           && ...

### 2020-06-23 draft/paper.tex
Arxiv version!

@@
-\documentclass[sigplan,screen]{acmart}
-
-\copyrightyear{2020}
-\acmYear{2020}
-\setcopyright{acmlicensed}
-\acmConference[LICS '20]{Proceedings of the 35th Annual ACM/IEEE Symposium on Logic
-  in Computer Science (LICS)}{July 8--11, 2020}{Saarbr\"ucken, Germany}
-\acmBooktitle{Proceedings of the 35th Annual ACM/IEEE Symposium on Logic in
-  Computer Science (LICS '20), July 8--11, 2020, Saarbr\"ucken, Germany}
-\acmPrice{15.00}
-\acmDOI{10.1145/3373718.3394770}
-\acmISBN{978-1-4503-7104-9/20/07}
-\startPage{1}
-
-%% Bibliography style
-\bibliographystyle{ACM-Reference-Format}
-%% Citation style
-%\citestyle{acmauthoryear}  %% For author/year citations
-%\citestyle{acmnumeric}     %% For numeric citations
-%\setcitestyle{nosort}      %% With 'acmnumeric', to disable automatic
-                            %% sorting of references within a single citation;
-                            %% e.g., \cite{Smith99,Carpenter05,Baker12}
-                            %% rendered as [14,5,2] rather than [2,5,14].
-%\setcitesyle{nocompress}   %% With 'acmnumeric', to disable automatic
-                            %% compression of sequential references within a
-                            %% single citation;
-                            %% e.g., \cite{Baker12,Baker14,Baker16}
-                            %% rendered as [2,3,4] rather than [2-4].
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Note: Authors migrating a paper from traditional SIGPLAN
-%% proceedings format to PACMPL format must update the
-%% '\documentclass' and topmatter commands above; see
-%% 'acmart-pacmpl-template.tex'.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% Some recommended packages.
-\usepackage{booktabs}   %% For formal tables:
-                        %% http://ctan.org/pkg/booktabs
-\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
-                        %% http://ctan.org/pkg/subcaption
-
-%% --------------------------------------------------------------------------------
+\documentclass{article}
 
@@
 \usepackage{stmaryrd}
+\usepackage{authblk}
+
+\bibliographystyle{acm}
+
+\theoremstyle{definition}
+\newtheorem{definition}{Definition}
+\newtheorem{example}{Example}
+
+\theoremstyle{theorem}
+\newtheorem{theorem}{Theorem}
+\newtheorem{lemma}{Lemma}
 
@@
 \newcommand{\Ty}{\mathsf{Ty}}
-\renewcommand{\U}{\mathsf{U}}
+\newcommand{\U}{\mathsf{U}}
 \newcommand{\El}{\mathsf{El}}
@@
 
+\title{Large and Infinitary Quotient Inductive-Inductive Types}
+
+%% \date{2020 June}
+\author{Andr{\'a}s Kov{\'a}cs}
+\author{Ambrus Kaposi}
+\affil{E\"otv\"os Lor\'and University}
+
+%% --------------------------------------------------------------------------------
+
 \begin{document}
 
-%% Title information
-\title{Large and Infinitary Quotient Inductive-Inductive Types} %% [Short Title] is optional;
-                                        %% when present, will be used in
-                                        %% header instead of Full Title.
-%% \titlenote{with title note}             %% \titlenote is optional;
-%%                                         %% can be repeated if necessary;
-%%                                         %% contents suppressed with 'anonymous'
-%% \subtitle{Subtitle}                     %% \subtitle is optional
-%% \subtitlenote{with subtitle note}       %% \subtitlenote is optional;
-%%                                         %% can be repeated if necessary;
-%%                                         %% contents suppressed with 'anonymous'
-
-%% Author information
-%% Contents and number of authors suppressed with 'anonymous'.
-%% Each author should be introduced by \author, followed by
-%% \authornote (optional), \orcid (optional), \affiliation, and
-%% \email.
-%% An author may have multiple affiliations and/or emails; repeat the
-%% appropriate command.
-%% Many elements are not rendered, but should be provided for metadata
-%% extraction tools.
-
-\author{Andr{\'a}s Kov{\'a}cs}
-\orcid{0000-0002-6375-9781}
-\affiliation{
-  %% \department{Department of Programming Languages and Compilers}
-  \institution{E{\"o}tv{\"o}s Lor{\'a}nd University}
-  \city{Budapest}
-  \country{Hungary}
-}
-\email{kovacsandras@inf.elte.hu}
-
-\author{Ambrus Kaposi}
-\orcid{0000-0001-9897-8936}
-\affiliation{
-  %% \department{Department of Programming Languages and Compilers}
-  \institution{E{\"o}tv{\"o}s Lor{\'a}nd University}
-  \city{Budapest}
-  \country{Hungary}
-}
-\email{akaposi@inf.elte.hu}
-
-%% Abstract
-%% Note: \begin{abstract}...\end{abstract} environment must come
-%% before \maketitle command
+\maketitle
+
 \begin{abstract}
@@
 \end{abstract}
-
-%% 2012 ACM Computing Classification System (CSS) concepts
-%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
-\begin{CCSXML}
-<ccs2012>
-<concept>
-<concept_id>10003752.10003790.10011740</concept_id>
-<concept_desc>Theory of computation~Type theory</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-</ccs2012>
-\end{CCSXML}
-
-\ccsdesc[500]{Theory of computation~Type theory}
-%% End of generated code
-
-%% Keywords
-%% comma separated list
-\keywords{inductive types, infinitary inductive types, quotient inductive-inductive types, categories with families}
-
-%% \maketitle
-%% Note: \maketitle command must come after title commands, author
-%% commands, abstract environment, Computing Classification System
-%% environment and commands, and keywords command.
-\maketitle
 
@@
 
-\textbf{An identity type} $\Id : (a : \Tm\,\Gamma\,\U)\ra
+An \textbf{identity type} $\Id : (a : \Tm\,\Gamma\,\U)\ra
 \Tm\,\Gamma\,(\El\,a)\ra\Tm\,\Gamma\,(\El\,a)\ra \Tm\,\Gamma\,\U$, with $\Refl :
@@
 
-%% Acknowledgments
-\begin{acks}
-The first author was supported by the European Union, co-financed by the
-European Social Fund (EFOP-3.6.3-VEKOP-16-2017-00002). The second author was
-supported by the National Research, Development and Innovation Fund of Hungary,
-financed under the Thematic Excellence Programme funding scheme, Project
-no. ED18-1-2019-0030 (Application-specific highly reliable IT solutions), by the
-New National Excellence Program of the Ministry for Innovation and Technology,
-Project no. \'UNKP-19-4-ELTE-874, and by the Bolyai Fellowship of the Hungarian
-Academy of Sciences, Project no. BO/00659/19/3.
-\end{acks}
-
-%% Bibliography
+\textbf{Acknowledgments.} The first author was supported by the European Union,
+co-financed by the European Social Fund (EFOP-3.6.3-VEKOP-16-2017-00002). The
+second author was supported by the National Research, Development and Innovation
+Fund of Hungary, financed under the Thematic Excellence Programme funding
+scheme, Project no. ED18-1-2019-0030 (Application-specific highly reliable IT
+solutions), by the New National Excellence Program of the Ministry for
+Innovation and Technology, Project no. \'UNKP-19-4-ELTE-874, and by the Bolyai
+Fellowship of the Hungarian Academy of Sciences, Project no. BO/00659/19/3.
+
 \bibliography{references}

### 2020-06-23 draft/paper.tex
add hyperref
nicer refs

@@
 \usepackage[utf8]{inputenc}
+\usepackage[hidelinks]{hyperref}
 \usepackage{amsmath}
@@
 
-\bibliographystyle{acm}
+\bibliographystyle{IEEEtran}
 
@@
 
-\textbf{Acknowledgments.} The first author was supported by the European Union,
+\paragraph{Acknowledgments.}The first author was supported by the European Union,
 co-financed by the European Social Fund (EFOP-3.6.3-VEKOP-16-2017-00002). The
